### 2017-01-22 gassolarpaper/gassolarpaper.tex
plot solar flux; plot energy fits

% `advanced_example.tex', an advanced example employing the AIAA class
% plus other third-party LaTeX packages.
%
% For a bare-bones usage, see `template.tex'.
%
% Typical processing for PostScript (PS) output:
%
%  latex advanced_example
%  bibtex advanced_example  (bibliography)
%  makeindex -s nomencl.ist -o advanced_example.gls advanced_example.glo
%                            (nomenclature)
%  latex advanced_example   (repeat as needed to resolve references)
%
%  xdvi advanced_example    (onscreen draft display)
%  dvips advanced_example   (postscript)
%  gv advanced_example.ps   (onscreen display)
%  lpr advanced_example.ps  (hardcopy)
%
% With the above, only Encapsulated PostScript (EPS) images can be used.
%
%
%  pdflatex advanced_example
%  bibtex advanced_example    (bibliography)
%  makeindex -s nomencl.ist -o advanced_example.gls advanced_example.glo
%                              (nomenclature)
%  pdflatex advanced_example  (repeat as needed to resolve references)
%
%  acroread advanced_example.pdf  (onscreen display)
%
% If you have EPS figures, you will need to use the epstopdf script
% to convert them to PDF because PDF is a limmited subset of EPS.
% pdflatex accepts a variety of other image formats such as JPG, TIFF,
% PNG, and so forth -- check the documentation for your version.
%
% If you do *not* specify suffixes when using the graphicx package's
% \includegraphics command, latex and pdflatex will automatically select
% the appropriate figure format from those available.  This allows you
% to produce PS and PDF output from the same LaTeX source file.
%
% To generate a large format (e.g., 11"x17") PostScript copy for editing
% purposes, use
%
%  dvips -x 1467 -O -0.65in,0.85in -t tabloid advanced_example
%
% For further details and support, read the Users Manual, aiaa.pdf.

\documentclass[]{aiaa-tc}% insert '[draft]' option to show overfull boxes

 \usepackage{varioref}%  smart page, figure, table, and equation referencing
 \usepackage{wrapfig}%   wrap figures/tables in text (i.e., Di Vinci style)
 \usepackage{threeparttable}% tables with footnotes
 \usepackage{dcolumn}%   decimal-aligned tabular math columns
  \newcolumntype{d}{D{.}{.}{-1}}
 \usepackage{nomencl}%   nomenclature generation via makeindex
  \makeglossary
 \usepackage{amssymb,amsmath}
 \usepackage{subfigure}% subcaptions for subfigures
 \usepackage{subfigmat}% matrices of similar subfigures, aka small mulitples
 \usepackage{fancyvrb}%  extended verbatim environments
 \fvset{fontsize=\footnotesize,xleftmargin=2em}
 \usepackage{lettrine}%  dropped capital letter at beginning of paragraph
%  \usepackage[dvips]{dropping}% alternative dropped capital package
 \usepackage[colorlinks]{hyperref}%  hyperlinks [must be loaded after dropping]
 \usepackage{float}
 \usepackage{longtable,booktabs,tabularx}
 \restylefloat{table}
 \usepackage{graphicx}
 \usepackage{caption}
 \usepackage{siunitx}
 \usepackage{multicol}
 \usepackage{indentfirst}
 \usepackage[labelfont=bf]{caption}

 \title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}

 \author{
  Michael Burton \thanks{Master's Candidate, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
  \ and Warren Hoburg\thanks{Professor, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
  {\normalsize\itshape
   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
 }

 % Data used by 'handcarry' option
 \AIAApapernumber{YEAR-NUMBER}
 \AIAAconference{Conference Name, Date, and Location}
 \AIAAcopyright{\AIAAcopyrightD{YEAR}}

 % Define commands to assure consistent treatment throughout document
 \newcommand{\eqnref}[1]{(\ref{#1})}
 \newcommand{\class}[1]{\texttt{#1}}
 \newcommand{\package}[1]{\texttt{#1}}
 \newcommand{\file}[1]{\texttt{#1}}
 \newcommand{\BibTeX}{\textsc{Bib}\TeX}
 \usepackage{hyperref}
 \hypersetup{citecolor = blue}

\begin{document}

\maketitle

\begin{abstract}
    Using geometric programming, a comparison of solar-electric and gas powered, long-endurance aircraft is accomplished.
    Long-endurance aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, velocity and coverage footprint.
    Geometric programming, a form of convex optimization, can reliably solve problems with thousands of variables in seconds and is used as an optimization tool to evaluate the design trade offs between architectures and requirements.
    Using this approach, a gas powered and a solar-electric powered aircraft were considered.  
    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by limited sunlight during the winter and wind speeds at higher latitudes.
    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
\end{abstract}

\section*{Nomenclature}

\begin{multicols}{2}
\small

\begin{tabbing}
  XXXXXXX \= \kill% this line sets tab stop
$A$ \> wing aspect ratio \\
$\bar{A}$ \> non-dimensional cross sectional area \\
$b$ \> wing span [ft] \\
$BSFC$ \> brake specific fuel consumption [lb/hr/hp] \\
$c$ \> wing chord [m] \\
$C_D$ \> drag coefficient \\
$C_{d_0}$ \> non-wing drag coefficient \\
$c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
$c_{d_p}$ \> wing profile drag coefficient \\
$c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
$C_f$ \> skin friction coefficient \\
$C_L$ \> lift coefficient \\
$c_{l_{\alpha}}$ \> lift slope coefficient \\
$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
$d$ \> tail boom diameter [m] \\
$D_{\text{boom}}$ \> tail boom drag [N] \\
$\Delta W$ \> wing section weight [N] \\
$\Delta y$ \> wing section length [m] \\
$DOY$ \> day of the year \\
$e$ \> Oswald efficiency factor \\
$E$ \> Young's Modulus [Pa] \\
$E_{\text{battery}}$ \> energy stored in battery [J] \\
$E_{\text{sun}}$ \> total solar energy available [Whr] \\
$(E/S)_{\text{sun}}$ \> available solar energy per unit area [Whr/m$^2$] \\
$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
$f_{\text{structural}}$ \> fractional structural weight\\
$g$ \> gravitational constant [m/s$^2$] \\
$h$ \> flight altitude [ft] \\
$h_{\text{batt}}$ \> battery energy density [Whr/kg] \\
$h_{\text{cap}}$ \> spar cap separation [m] \\
$I$ \> cap spar moment of inertia [m$^4$] \\
$I_0$ \> tail boom root moment of inertia [m$^4$] \\
$k$ \> tail boom taper index \\
$K_q$ \> wing loading constant [N/m$^2$] \\
$L$ \> atmospheric lapse rate [K/m] \\
$L'$ \> local lifting force [N/m] \\
$L_\text{h}$ \> horizontal tail lift [N] \\
$l_\text{fuse}$ \> fuselage length [m] \\
$l_\text{h}$ \> horizontal tail moment arm [m] \\
$l_\text{v}$ \> vertical tail moment arm [m] \\
$m$ \> tail boom mass [kg] \\
$m'$ \> local wing mass [kg/m] \\
$m_{\text{h}}$ \> modified horizontal tail lift coefficient \\
$m_{\text{motor}}$ \> electric motor mass \\
$m_{\text{w}}$ \> modified wing lift coefficient \\
$\dot{m}_{\text{fuel}}$ \> fuel flow rate [kg/s] \\
$\mathcal{M}$ \> wing bending moment [Nm] \\
$MTOW$ \> max take-off weight [N] \\
$n$ \> number of wing segments \\
$N$ \> number of flight segments \\
$N_{\text{max}}$ \> load factor\\
$p$ \> air pressure [Pa]\\
$p_{11}$ \> air pressure at 11,000 meters [Pa]\\
$p_{\text{wind}}$ \> percentile wind speed \\
$P_{0}$ \> magnitude of available solar power [W/m$^2$] \\
$P_{\text{avionics}}$ \> avionics power [hp] \\
$P_{\text{oper}}$ \> aircraft operating power [hp] \\
$P_{\text{shaft}}$ \> engine/motor shaft power [hp] \\
$P_{\text{sun}}$ \> useable solar power per unit area [W/m$^2$] \\
$P_{\text{sun surface}}$ \> power emitted at the sun's surface [W/m$^2$] \\
$r_0$ \> average distance from earth to sun \\
$R$ \> aircraft range [nmi] \\
$R_{\text{earth orbit}}$ \> distance from earth to sun \\
$R_{\text{fuse}}$ \> fuselage radius [m] \\
$R_{\text{spec}}$ \> specific gas constant [J/k/K] \\
$R_{\text{sun}}$ \> radius of the sun \\
$q$ \> distributed wing loading [N/m] \\
$\bar{q}$ \> normalized distributed wing loading \\
$Re$ \> aircraft Reynolds number \\
$Re_{\text{boom}}$ \> tail boom Reynolds number \\
$Re_{\text{h}}$ \> horizontal tail Reynolds number \\
$Re_{\text{v}}$ \> vertical tail Reynolds number \\
$S$ \> wing planform area [m$^2$]\\
$\mathcal{S}$ \> shear forces [N] \\
$S_{\text{fuse}}$ \> fuselage wetted surface area [m$^2$]\\
$S_{\text{h}}$ \> horizontal tail planform area [m$^2$]\\
$S_{\text{solar}}$ \> solar cell area [m$^2$]\\
$S_{\text{v}}$ \> vertical tail planform area [m$^2$]\\
$t$ \> flight time [days] \\
$t_0$ \> tail boom root wall thickness [m] \\
$t_{\text{cap}}$ \> spar cap thickness [m] \\
$t_{\text{day}}$ \> hours of daylight [hr] \\
$t_{\text{night}}$ \> hours of darkness [hr] \\
$t_{\text{sun rise}}$ \> time of sun rise [hr] \\
$t_{\text{sun set}}$ \> time of sun set [hr] \\
$T$ \> aircraft thrust [N] \\
$T_{11}$ \> air temperature at 11,000 meters [K] \\
$T_{\text{atm}}$ \> air temperature at flight altitude [K] \\
$T_{\text{sl}}$ \> air temperature at sea level [K] \\
$V$ \> true airspeed [m/s] \\
$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume [m$^3$] \\
$V_{\text{gust}}$ \>  gust velocity [m/s] \\
$V_{\text{h}}$ \> horizontal tail lift coefficient \\
$V_{\text{ref}}$ \>  reference wind airspeed [m/s] \\
$V_{\text{wind}}$ \>  wind speed [m/s] \\
$V_{\text{v}}$ \> vertical tail lift coefficient \\
$w$ \> wing deflection [m] \\
$W$ \> aircraft weight [N] \\
$W_{\text{ave}}$ \> average weight over a flight segment [N] \\
$W_{\text{battery}}$ \> battery weight [N] \\
$w_{\text{cap}}$ \> spar cap width [m] \\
$W_{\text{cent}}$ \> center of aircraft weight [N] \\
$W_{\text{empennage}}$ \> empennage weight [N] \\
$W_{\text{engine}}$ \> engine weight [N] \\
$W_{\text{final}}$ \> end of flight segment aircraft weight [N] \\
$W_{\text{fuel}}$ \> fuel weight [N] \\
$W_{\text{fuselage}}$ \> fuselage weight [N] \\
$W_{\text{h}}$ \> horizontal tail weight [N] \\
$W_{\text{initial}}$ \> start of flight segment aircraft weight [N] \\
$w_{\text{max}}$ \> maximum deflection limit [m] \\
$W_{\text{payload}}$ \> payload weight [N] \\
$W_{\text{skin}}$ \> wing skin weight [N] \\
$W_{\text{solar}}$ \> solar cell weight [N] \\
$W_{\text{spar}}$ \> wing spar weight [N] \\
$W_{\text{structural}}$ \> structural weight [N] \\
$W_{\text{v}}$ \> vertical tail weight [N] \\
$W_{\text{wing}}$ \> wing weight [N] \\
$y$ \> distance from wing root along span [m] \\
$z_{\text{bre}}$ \> Breguet Range helper variable \\
$\alpha$ \> angle of attack \\
$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities [rad] \\
$\Delta$ \> solar declination angle \\
$\eta_{\text{charge}}$ \> battery charging efficiency \\
$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
$\eta_{\text{prop}}$ \> propulsive efficiency \\
$\eta_{\text{solar}}$ \> solar cell efficiency \\
$\mu$ \> air viscosity [kg/m/s] \\
$\mu_0$ \> reference air viscosity [kg/m/s] \\
$\phi$ \> latitude \\
$\rho$ \> air density [kg/m$^3$] \\
$\rho_{\text{carbon fiber}}$ \> density of carbon fiber [kg/m$^3$] \\
$\rho_{\text{foam}}$ \> density of foam [lbf/ft$^3$] \\
$\rho_{\text{solar}}$ \> solar cell density [kg/m$^2$] \\
$\sigma_{\text{carbon fiber}}$ \> maximum carbon fiber stress [Pa] \\
$\tau_{\text{h}}$ \> horiztonal tail thickness to chord ratio \\
$\tau_t$ \> wing thickness to chord ratio \\
$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
$\tau_w$ \> cap spar width to chord ratio \\
$\theta$ \> angle normal to surface \\
$\theta_{\text{boom}}$ \> tail boom deflection angle \\
$\Theta$ \> bending deflection angle 
 \end{tabbing}

\end{multicols}
% \printglossary% creates nomenclature section produced by MakeIndex

\section{Introduction}

The design problem of a long-endurance, station-keeping aircraft is considered.
Solar electric and gas powered architectures are proposed as solutions to this problem.
Because sizing long-endurance aircraft is so multifaceted and the interaction between aerodynamics, structural weight, energy and other disciplines is so intertwined, a systematic and rapid approach is needed to explore this design space.
Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.

The driving condition that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
To achieve a feasible design, solar-electric powered aircraft must carry enough solar cells and batteries required to fly during the day while storing enough energy to fly through the night.\cite{solartech}
If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then it can theoretically fly during any time of the year for long durations. \cite{solartech}
This becomes more difficult to achieve at higher latitudes as the solar flux during the winter solstice decreases.  
Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.  
Figure~\ref{f:solarstart} shows that in order to operate at higher latitudes during the winter a larger aircraft is required.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{mtowvslatsolar.pdf}
    \caption{ \textbf{ Trade study showing how solar-electric powered aircraft size increases with latitude operating capacity.  }}
	\label{f:solarstart}
	\end{center}
\end{figure}

The key driving requirement for a gas powered aircraft is endurance.  
As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
Therefore, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitude only depends on wind speed. 
If a gas-powered aircraft can fly at the latitude with the worst wind speed in can theoretically fly anywhere else in the world.  
 (Figure~\ref{f:mtowvsendurance})

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
    \caption{ \textbf{ Trade study showing how gas powered aircraft weight increases with endurance.  }}
	\label{f:mtowvsendurance}
	\end{center}
\end{figure}

To capture the key trade offs like those presented in Figures~\ref{f:solarstart} and ~\ref{f:mtowvsendurance} a geometric programm is built that models for both gas and solar-electric powered, long-endurance aircraft.  
Equations governing the underlying physics of aerodynamic performance, solar energy collection and storage, and fuel burn are written in a geometric programming form, which can then be solved as an optimization problem.
Then by changing the design parameters, such as endurance, latitude, and percentile wind speeds, multiple solutions can be rapidly generated to observe how the aircraft size and weight is affected by different requirements.

Using this methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
Similar trade offs are calculated for other requirements and performance metrics to determine under what conditions which platform is best suited to meet a set of requirements.

\section{Geometric Programming\cite{gp}}

Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions. GPs have the form

\begin{align} 
\label{e:gpform}
\text{minimize } f_0(\bold{x}) & \nonumber \\
\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,...,m \\
g_i (\bold{x}) &= 1, i = 1,...,p \nonumber 
\end{align}

where the functions $f_i$ must be \emph{monomial} functions and the functions $g_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms

\begin{align}
 \label{e:mon}
g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
\label{e:pos}
f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
\end{align}

The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
Additionally, geometric programs can be solved rapidly with existing algorithms.  
Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}  
To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} \\

\section{Aircraft Sizing Models}

To evaluate the size and performance of both gas and solar-electric powered aircraft underlying physics models are used.  
Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar insolation and aircraft models governing the performance, aerodynamics and weight breakdown are used in the optimization model.
Equations from these models are then expressed in a GP-compatible form to enable their use in the optimization. 
Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
To obtain higher fidelity results from the optimization, higher fidelity models can be implemented.

\subsection{Station Keeping}

The design problem considered is a long-endurance, station keeping aircraft or an aircraft capable of remaining in one location in the air for long periods of time. 
Long-endurance, station keeping aircraft are generally required to operate at specific latitudes, altitudes, and percentile wind speeds.\cite{zephyr} 
Both wind speeds and atmospheric quantities depend on the station keeping requirements.

\subsubsection{Wind Speeds}

To station keep, an aircraft must fly faster than the wind speed.  
Therefore, the aircraft speed is constrained for both the gas and solar-electric powered cases,

\begin{equation}
    \label{e:availreq}
    V \geq V_{\text{wind}}.
\end{equation}

The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,

\begin{equation}
    \label{e:windspeed}
    V_{\text{wind}} = f(\phi, h, DOY, p_{\text{wind}}).
    \end{equation}

Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
It is assumed that the wind speeds can be approximated by using wind data across all longitudes.

An analysis was done of how wind speeds vary from month to month to determine during which seasons or months the wind speeds are higher. Figure~\ref{f:windvsmonth} reveals that wind speeds are significantly worse during the winter months. 
Because the constraining case for solar-electric powered aircraft is the winter solstice and because the wind speeds are highest during the winter months only the wind speed data from December will be used for aircraft the sizing and performance analysis presented in this paper.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{windvsmonth.pdf}
    \caption{ \textbf{ Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively. }}
	\label{f:windvsmonth}
	\end{center}
\end{figure}

Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere. 
Because the gas-powered aircraft is assumed to have a naturally aspirated engine, it will fly between 5,000 and 25,000 ft depending on payload requirement. 
The solar-electric powered aircraft will operate around 60,000 ft to avoid cloud coverage.  
Variations of wind speed with latitude are shown for altitudes at both 15,000 ft at and 60,000 ft.  \\

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{latvswind.pdf}
    \caption{ \textbf{ Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively. Wind data includes winds during December for the year 2005-2015.}}
	\label{f:latvswind}
	\end{center}
\end{figure}

Wind speeds also vary significantly with altitude. (Figure~\ref{f:altvswind})
The high winds around 30,000-40,000 ft makes it difficult for both solar-electric and gas powered long-endurance aircraft to fly in that range of altitudes.  
Solar-electric powered aircraft, that do not have aspirated engines, fly around 60,000 ft take advantage of lower wind speeds and limited cloud coverage.
Gas powered aircraft with aspirated engines can have difficulty operating at 60,000 ft and generally perform better in the lower wind speeds below 25,000 ft.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{altvswind.pdf}
    \caption{ \textbf{ Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.  Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively. Wind data includes winds during December for the year 2005-2015.}}
	\label{f:altvswind}
	\end{center}
\end{figure}

To understand what a certain percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
One lag is equivalent to 6 hours.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics{Boston_dec_2015auto.pdf}}
     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics{Boston_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015. One lag is equivalent to 6 hours. }}
 \label{f:bostonwinds}
\end{figure}

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics{Rome_dec_2015auto.pdf}}
     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics{Rome_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015. One lag is equivalent to 6 hours. }}
 \label{f:romewinds}
\end{figure}

This analysis shows high wind speeds can last for days at a time.  
An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to relocate or land.

\subsubsection{Optimizing Altitude for Solar-Electric Powered Aircraft}

There exists an optimum flight altitude for solar-electric powered aircraft.  
The local minimum in wind speed around 63,000 ft and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
To accomplish this, constraints relating wind speed an altitude are imposed to optimize altitude. 
This approach agrees from an operational perspective as a solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.

Because wind speeds also vary with latitude several constraints are needed to optimize altitude. 
Operationally, aircraft, such as the solar-electric powered aircraft, will fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
For example, an aircraft optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude. 
This is imposed in the optimization model as multiple constraints. For the 35 degree latitude case, 70 separate wind and solar flux constraints that must be satisfied, one for each latitude.

Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg\cite{fitting}.
Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
Altitude can be post-calculated from the optimized air density. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude no constraints for latitudes below 20 degrees are actually included in the optimization. 
Additionally, only wind speed data from the Northern Hemisphere are considered.
Therefore, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
The constraint of wind speed and air density at the 35 degree latitude is

\begin{align}
    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.213} &\geq 2918\rho^{7.587}p_{\text{wind}}^{9.37} + \num{1.58d-10}\rho^{-6.074}p_{\text{wind}}^{60.421} \nonumber \\
    \label{e:windfit35}
    &+ \num{4.05d-16}\rho^{-9.879}p_{\text{wind}}^{13.204} + 1048\rho^{6.33}p_{\text{wind}}^{46.232}
\end{align}

where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed. The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{windfitl35.pdf}
    \caption{ \textbf{ Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}. (Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude. }}
	\label{f:windfitl35}
	\end{center}
\end{figure}

For the gas powered aircraft, the altitude is not optimized.  
Gas-powered aircraft, which are naturally aspirated will be less efficient at higher altitudes.  
Additionally, wind speeds increase nearly linearly up to 30,000 ft.  
Therefore, the altitude for gas-powered aircraft is determined as a requirement of the payload and the minimum altitude necessary for the payload to be effective.\cite{orion}
For this sizing study, the altitude requirement for the gas powered aircraft is $h=15,000$ ft.

\subsubsection{Atmospheric Quantities}

For the solar aircraft air density is being optimized and the operating altitude is computed after the fact using equations specific to the tropopause,\cite{isaatm}

\begin{align}
    \label{e:tropopress}
    p &= \rho R_{\text{spec}}T_{11} \\
    \label{e:tropoalt}
    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}} 
\end{align}

According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.  Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$. \\

Because altitude is a requirement for the gas powered aircraft, density is pre-computed prior to solving the optimization using atmospheric equations specific to the troposphere,\cite{isaatm}

\begin{align}
    \label{e:Talt}
    T_{\text{atm}} = T_{\text{sl}} - Lh \\
    \label{e:rhot}
    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
\end{align}

Air viscosity is also computer prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}

\begin{equation}
    \label{e:sutherland}
    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
\end{equation}

where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K].

\subsection{Solar-Electric Power}

The driving equations for the solar-electric powered aircraft are determined by the energy required to fly throughout the day while charging the batteries enough to last through the night.  
Equations~\ref{e:solarreq} and~\ref{e:solarbatt} capture the solar requirement by enforcing that the total solar energy harvested by the solar cells during the day must be greater than the energy used to operate the aircraft during the day while charging enough batteries to fly during the night.\cite{solartech}
Under these assumptions, if the aircraft can fly for one day and one night, then it could theoretically fly for many days under similar circumstances.

    \begin{align}
        \label{e:solarreq}
        (E/S)_{\text{sun}} \eta_{\text{solar}} S_{\text{solar}} &\geq P_{\text{oper}}t_{\text{day}} + E_{\text{battery}}/\eta_{\text{charge}} \\
        \label{e:solarbatt}
        E_{\text{battery}} &\geq P_{\text{oper}}t_{\text{night}}/\eta_{\text{discharge}}
    \end{align}

    It is assumed that the charge and discharge efficiencies are constant, $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$. 
    It is also assumed that the solar cell efficiency is constant, $\eta_{\text{solar}} = 0.3$. 
    A battery with an energy density of $h_{\text{batt}} = 350$ [Whr/kg], is selected for the solar-powered aircraft.

    To enforce these constraints, two variables were created: $P_{\text{oper}}$ and $S_{\text{solar}}$. 
    $P_{\text{oper}}$ is defined as the average power required to operate the aircraft. 
    $S_{\text{solar}}$ is defined as the amount of area on the aircraft covered by solar cells.

    \begin{align}
        \label{e:solarpoper}
        P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}} \\
        \label{e:solarssolar}
        S_{\text{solar}} &\leq f_{\text{solar}}S
    \end{align}

    For the purposes of this design study, it is assumed that the solar cells are only placed on the wing and that the total solar cell area is limited to some fraction of the wing planform area $f_{\text{solar}}$, to account for control surfaces, leading edge, and other accessories.

    Because solar aircraft requirements are typically given in terms of latitude and yearly availability it is useful to express the total available solar energy per unit area per day as a function of latitude, $\phi$, and the day of the year, $DOY$,

    \begin{equation}
        \label{e:solarfunc}
        (E/S)_{\text{sun}} = f(\phi, DOY).
    \end{equation}

    For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(E/S) = E$). 
    The total available solar energy per day is an integral of the available solar power during the course of the day,

    \begin{equation}
        \label{e:solares}
        E_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} P_{\text{sun}} dt.
    \end{equation}

    Figure~\ref{f:lat45} shows a sample distribution of the available solar power during the spring equinox at 45 degrees latitude.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{lat30.pdf}
    \caption{ \textbf{ The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.  The area under this curve is the total energy available per unit area for the entire day.} }
	\label{f:lat45}
	\end{center}
\end{figure}

The shape of this curve is governed by the angle, $\theta$, between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}

\begin{equation}
    \label{e:solarp}
    P_{\text{sun}} = P_0 \cos{\theta}
\end{equation}

The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}

    \begin{equation}
        \label{e:solartheta}
        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
    \end{equation}

 The declination angle $\Delta$, can be found using the relation\cite{solar}

    \begin{align}
        \label{e:solardelta}
        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
    \end{align}

    where $\beta = 2\pi (DOY-1)/365$.
    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation Equation~\eqref{e:solartheta}, \cite{solar}

    \begin{align}
        \label{e:solartday}
        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
        \label{e:solarsunrise}
        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
        \label{e:solartday2}
        t_{\text{day}} &= 2t_{\text{sun rise}} \\
        \label{e:solartnight}
        t_{\text{night}} &= 24 - t_{\text{day}}
    \end{align}

    where noon is $t=0$. Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.

    The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,

    \begin{align}
        \label{e:solarp0}
        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
        \label{e:solareo}
        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{DOY-93}{365}\right)} \right],
    \end{align}

    where

    \[ \begin{array}{lcl}
        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
        R_{\text{sun}} & : & \text{Radius of the sun} \\
        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
        r_0 & : & \text{Average distance from earth to sun} \\
        0.017 & : & \text{Eccentricity of earth-sun orbit}.
    \end{array} \]

    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year. Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night is calculated from the latitude and the day of the year prior to an optimization solve.

\subsection{Breguet Range}

A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,

\begin{equation}
    \label{e:breguetendurance}
    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}BSFCg} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
\end{equation}

The derivation begins with the differential form of Breguet Range,\cite{br2}

\begin{equation}
    \label{e:breguetdiff}
    -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
\end{equation}

Using the definition of $BSFC$

\begin{equation}
    \label{e:brBSFC}
    BSFC = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
\end{equation}

Equation~\eqref{e:breguetdiff} can be written as

\begin{equation}
    \label{e:brdiff2}
    -dW = g P_{\text{shaft}} BSFC dt.
\end{equation}

It is assumed that $BSFC$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
A constant power to weight ratio assumes constant velocity and constant lift coefficient.\cite{br2}
Dividing by $W$,

\begin{equation}
    \label{e:brdiff2}
    -\frac{dW}{W} = \frac{g P_{\text{shaft}}BSFC }{W} dt,
\end{equation}

and integrating, the Breguet Range equation can be expressed as

\begin{equation}
    \label{e:be1}
    \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}BSFC}{W_{\text{ave}}} t,
\end{equation}

where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.  $W_{\text{ave}}$ is assumed to be the geometric mean, defined as

\begin{equation}
    \label{e:gpmean}
    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
\end{equation}

    To make Equation~\eqref{e:be1} GP compatible, a Taylor expansion of the exponentiated right hand side is used,\cite{hoburgthesis}

\begin{align}
    \label{e:brzbre}
    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t BSFC g}{W}\\
    \label{e:brtaylor}
    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
\end{align}

    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynmial respectively and therefore GP compatible. \\

    For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.  Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized for more accurate results.

\begin{align}
    \label{e:slfweightd}
    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
    \label{e:brzbred}
    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i BSFC g}{\sqrt{W_i W_{i+1}}}\\
    \label{e:brtaylord}
    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24} 
    \end{align}

    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ is used. Additionally, it is assumed that the break specific fuel consumption is constant, $BSFC = 0.6$ [lb/hp/hr].\cite{bsfcperf}

\subsection{Steady Level Flight}

Both the gas and solar powered aircraft are assumed to be in steady level flight.  The balance of the main forces of lift, drag, weight and thrust are modeled as\cite{hoburgthesis}

\begin{align}
    \label{e:slfthrust}
    T &\geq \frac{1}{2} \rho V^2 C_D S\\
    \label{e:slfweight}
    W &= \frac{1}{2} \rho V^2 C_L S . 
\end{align}

The shaft power produced by the engine or motor and can expressed as

\begin{equation}
    \label{e:slfpower}
    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
    \end{equation}

    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.75$. Equations~\eqref{e:slfthrust},~\eqref{e:slfweight}, and~\eqref{e:slfpower} are all monomials and therefore GP-compatible.

\subsection{Aerodynamics}

The aircraft aerodynamics is modeled as a drag build up of wing profile drag, induced drag, and non-wing drag coefficients,

\begin{equation}
    \label{e:aerodragb}
    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L}{\pi e A}.
    \end{equation}

where the Oswald efficiency factor is assumed to be constant, $e=0.9$. 
The non-wing drag coefficient $C_{d_0}$, is estimated as a constant.  
For the gas powered aircraft, $C_{d_0} = 0.01$.  
For the solar-electric powered aircraft, $C_{d_0} = 0.002$.  
It is assumed that the solar powered aircraft will store its batteries in the wings. 
The gas powered aircraft will use a fuselage to store its fuel, thus having a larger non-lifting surface and a higher drag coefficient.\cite{raymer}

    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the \emph{JH01} airfoil. (See Appendix B) 
    Drag polars were produced using XFOIL at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg.\cite{fitting}
    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.

    \begin{align}
        \label{e:aerodragprof}
        c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
                       &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
        Re &= \frac{\rho V S/b}{\mu}
    \end{align}

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{JH01polar.pdf}
    \caption{ \textbf{ The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as a solid line.  The open circles are the calculated drag polars from XFOIL. (Log space RMS error = 0.00489.)} }
	\label{f:JH01polar}
	\end{center}
\end{figure}

As a preliminary sizing study the aspect ratio is assumed to be a constant, $A = 30$.  A more detailed sizing study with a wing structural model, discussed later, will allow the aspect ratio to be a free variable.

\subsection{Weight Breakdown}

The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizer, and engine. 
The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.

\begin{align}
    \label{e:weightmtow}
    MTOW &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} \\
\end{align}

A similiar weight breakdown is done for the solar-electric powered aircraft,

\begin{align}
    \label{e:weightsmtow}
    MTOW &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} \\
    W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
    W_{\text{battery}} &\geq \frac{E_{\text{battery}}}{h_{\text{battery}}} g
\end{align}

where the solar cell density and the battery energy density are assumed to be $\rho_{\text{solar}} = 0.27$ [kg/m$^2$] and $h_{\text{batt}} = 350$ [Whr/kg].\cite{solartech}\cite{solarparam}

% \subsection{Comparison to Existing Platforms}
% 
% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
% Known pramaters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}. The model is then solved by maximizing endurance.
% 
% \begin{longtable}{lcccc}
%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
%      \toprule
%      \toprule
%      \label{t:existingparams}
%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
%      \midrule
%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
%      \bottomrule
%  \end{longtable}
% 
%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
% 
% \begin{longtable}{ll}
%     \caption{Fixed Optimization Parameters} \\
%      \toprule
%      \toprule
%      Variable &  Value \\
%      \midrule
%      $\eta_{prop}$        & 0.65   \\
%      $A$                 & 25     \\
%      $f_{\text{structural}}$     & 0.35   \\
%      $BSFC$ [lb/hp/hr]    & 0.6    \\
%      \bottomrule
%      \label{t:compparams}
%  \end{longtable}
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics{mtowvehiclecomp.pdf}}
%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics{bvehiclecomp.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Comparison of optimization results to existing platforms.  Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
%  \label{f:vehiclecomp}
% \end{figure}
%  
% \begin{longtable}{lcccc}
% \caption{Optimization Results Comparison} \\
% \label{t:existingplatforms} \\
%      \toprule
%      \toprule
%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
%      \midrule
%      Orion      & 5 & 4.4   & 132 & 146.8\\
%      Predator   & 1  & 1.7  & 55 & 39.0 \\
%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
%      \bottomrule
%  \end{longtable}
% 
% 
% Because of limited information about these aircraft the optimization predictions are not exact. However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.

\section{Long-endurance Aircraft Trade Space}

One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight,

\begin{equation}
    \label{e:structualfraction}
    W_{\text{structural}} \geq f_{\text{structural}}MTOW.
\end{equation}

where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$. Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.

% \begin{longtable}{llll}
% \caption{Fixed Optimization Parameters} \\
% \toprule
% \toprule
% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
% \midrule
% Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
% Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
% Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
% $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
% $A$                 & 25     & $A$                 & 25              \\
% $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
% $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
% $BSFC$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
% Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
%                      &        & $h_{\text{battery}}$ [Whr/kg]  & 350 \\
%                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
%                      &        & Day of the Year      & 355 (Dec 21st)\\
% \bottomrule
% \label{t:gassolarparams}
%  \end{longtable}

\subsection{Sizing for Different Latitude Bands}

An analysis of the max take off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
This analysis is achieved by minimizing the max take off weight in both the gas and solar-electric powered optimization models at different latitudes. 
For this study, the gas-powered aircraft is constrained to fly for 8 days.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics{mtowvslatgassimple.pdf}}
     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics{mtowvslatsolarsimple.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{ Trade study of max take off weight and latitude. Plots show required maximum take off weight required to operate between a band of latitudes.   A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds. }}
 \label{f:latvsmtowtrade}
\end{figure}

One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.

\subsection{Gas Powered Endurance}

While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
The gas powered optimization model was set to minimize max take off weight for different endurance requirements to understand what endurance limits are possible.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{mtowvsendsimple.pdf}
    \caption{ \textbf{ Trade study between endurance and max take off weight for a gas powered archicture. }}
	\label{f:mtowvsendsimple}
	\end{center}
\end{figure}

For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.

\subsection{Sensitivity Analysis}

When a GP is solved, the sensitivity of the objective function with respect to each constraint is also returned.  From this information, the sensitivity to each inputted variable with respect to the objective function can be extracted.\cite{hoburgthesis} 
For example, if the objective function were weight and the sensitivity to solar cell efficiency were -0.5, then a 1\% increase in the solar cell efficiecy would result in a -0.5\% decrease in weight.  
From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 
To achieve a more accurate numbers for these variable they can be replaced with higher fidelity models.  
For example, the structural weight fraction $f_{\text{structrual}}$, which has a high sensitivity, could be replaced by a structural model of the wing, fuselage and empennage. 
Table~\ref{t:sens} shows the varialbes with the highest sensitivities for both the solar-electric powered aircraft optimized at the $\pm$40 degree latitude requirement minimizing weight.

\begin{longtable}{ll}
\caption{Sensitivities} \\
\toprule
\toprule
 Varialbe & Sensitivities \\
\midrule
$h_{\text{batt}}$         & -3.475 \\
$\eta_{\text{solar}}$     & -1.268 \\
$f_{\text{structural}}$   & 2.445  \\
$\eta_{\text{discharge}}$ & -5.068 \\
$\eta_{\text{prop}}$      & -4.743 \\
$c_{d_0}$                 & 0.9935 \\
\bottomrule
\label{t:sens}
 \end{longtable}

% \subsection{Solar-electric Powered Trade Space Exploration}
% 
% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.  (Figure~\ref{f:solarcontours}) 
% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{3}% number of columns
%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl40a80.pdf}}
%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl40a85.pdf}}
%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl40a90.pdf}}
%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl35a80.pdf}}
%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl35a85.pdf}}
%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl35a90.pdf}}
%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl30a80.pdf}}
%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl30a85.pdf}}
%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl30a90.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Matrix of contour plots of battery energy density and structural weight fraction vs wing span for solar-electric powered aircraft. Each point on the graph corresponds to a unique design.} }
%  \label{f:solarcontours}
% \end{figure}

\section{Detailed Models}

To improve the fidelity of the optimization, higher fidelity models can be implemented.  
A spar and structural wing model are added to both the gas and solar-electric powered optimization models.  
An empennage model is added to account for the additional weight and drag of horizontal and vertical tails and a tail boom.  
Additional models were added to the gas powered aircraft optimization model to account for engine weight and performance. 
These models, similar to those described in section III, are a set of constraints derived from governing first principles put in a GP-compatible form. 
This results in a more comprehensive and detailed optimization model for aircraft sizing.   \\

\subsection{Wing Spar Model}

A wing spar is modeled as one of the main structural elements of both aircraft. 
A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.

\subsubsection{Loading Case for Gas Powered Aircraft}

It is assumed that the key spar sizing case for the gas powered aircraft is standard wing bending during steady level flight, fully fueled. The distributed load is a combination of the lifting distribution and weight distribution along the wing

\begin{equation}
    \label{e:wloading}
    q(y) = L'(y) - N_{\text{max}}gm'(y)
\end{equation}

where $q(y)$ is the spanwise distributed load. 
The wing loading distribution, $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}

\begin{equation}
    \label{e:wingloading}
    q(y) \approx K_q c(y) 
\end{equation}

where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as

\begin{equation}
    \label{e:kq}
    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
\end{equation}

where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$. The local chord for a constant tapered wing\cite{bending} is defined as

\begin{align}
    \label{e:localchord}
    c(y) &= \frac{S}{b} \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right) \\
    \bar{c}(y) \equiv \frac{c(y)}{S/b} &= \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
\end{align}

where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load

\begin{align}
    \label{e:qbar}
    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right) \\
    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} &= \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
\end{align}

is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is

\begin{equation}
    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{empennage}}.
\end{equation}

\subsubsection{Loading Case for Solar-electric Powered Aircraft}

The key sizing case for the solar-electric aircraft spar is a gust load that varies across the span.  
It is assumed that the solar aircraft is span loaded, meaning the majority of the weight lies along the wings,

\begin{align}
    W_{\text{wing}} &\geq W_{\text{cap}} + W_{\text{skin}} + W_{\text{battery}} + W_{\text{solar}}\\
    W_{\text{cent}} &\geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{empennage}}.
\end{align}

For steady level flight, the same loading distribution as for the gas powered aircraft is assumed,

\begin{equation}
    \label{e:qbar}
    q(y) = \frac{N_{\text{max}}W_{\text{cent}}}{b}\frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right).
\end{equation}

Because of the wing's large moment of inertia, caused by the weight of the batteries inside the wings, the solar-electric aircraft is considered a highly flexible aircraft. 
To account for this flexiblity and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the loading distribution

\begin{align}
    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
\end{align}

where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.

Because the gust velocity is assumed to be vertical to the flight path, the local angle of attack is approximated by

\begin{equation}
    \label{e:gustalpha}
    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
\end{equation}

Because the $\arctan$ function is not GP compatible, a monomial approximation was calculating using techniques described by Hoburg\cite{fitting}.
This approximation is graphically shown in Figure~\ref{f:arctanfit} for the expected range of $V_{\text{gust}}/V \in [0, 0.7]$.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
    \caption{ \textbf{ Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.044) }}
	\label{f:arctanfit}
	\end{center}
\end{figure}

The gust velocity has a profile along the wing\cite{acgust},

\begin{equation}
    \label{e:gustwind}
    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
\end{equation}

where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.

% \begin{figure}[H]
% 	\begin{center}
% 	\includegraphics[width=0.6\textwidth]{gustvschord.pdf}
%     \caption{ \textbf{ Graphical representation of assumption that the local chord is equal to the local lift generated from the elliptical distribution. Assumes the center to wing weight ratio, $\frac{W_{\text{cent}}}{W_{wing}} = 0.14$. }}
% 	\label{f:gustvschord}
% 	\end{center}
% \end{figure}

\subsubsection{Discretized Beam Model}

Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes, the shear forces and moments can be computed from the distributed loads $q(y)$, with boundary conditions of zero shear forces and moments at the wing tips.\cite{bending}

\begin{align}
    \label{e:shear}
    \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
    \label{e:moment}
    M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
    \label{e:shearboundary}
    \mathcal{S}_n &= 0 \\
    \label{e:momentboundary}
    M_n &= 0
\end{align}

Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}

\begin{align}
    \label{e:angle}
    \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
    \label{e:deflection}
    w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
    \label{e:angleboundary}
    \Theta_0 &= 0 \\
    \label{e:defboundary}
    w_0 &= 0 
\end{align}

Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as

\begin{align}
    \label{e:sheargp}
    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
    \label{e:momentgp}
    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
    \label{e:anglegp}
    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
    \label{e:deflection}
    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y 
\end{align}

where the Young's Modulous of carbon fiber is $E = 20$ [MPa].

\subsubsection{Cap Spar for Bending Loads}

For this sizing study a cap spar is considered.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.8\textwidth]{capspar2.jpg}
    \caption{ \textbf{ Cross sectional view of a cap spar.  The two caps are made of carbon fiber, separated by a foam core. A thin shear web is wrapped around the caps and foam to prevent shearing.}}
	\label{f:capspar}
	\end{center}
\end{figure}

The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
The equation for the moment of inertia\cite{bending} of a cap spar is

\begin{equation}
    \label{e:moispar}
    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2
\end{equation}

This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,

\begin{equation}
    \label{e:moispar}
    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2
\end{equation}

There are also geometric constraints imposed on the width and thickness.  The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.  The width of the spar cap is assumed by no greater than 30\% of chord, $\tau_w = 0.3$.

\begin{align}
    \label{e:thickness}
    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
    \label{e:width}
    c(y)\tau_w &\geq w_{\text{cap}} 
    \end{align}

To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.

\begin{align}
    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
    c(y)\tau_w &\geq w_{\text{cap}_i} 
\end{align}

The wing spar at the root must be strong enough to withstand the bending root moment and stiff enough to not exceed some deflection limit.  Both constraints are imposed in the optimization model as

\begin{align}
    \label{e:stresscont}
    \sigma_{\text{carbon fiber}} &\geq \mathcal{M}_0 \frac{h_{\text{cap}}+t_{\text{cap}}}{I}\\
    \label{e:defcont}
    w_n &\leq w_{\text{max}}.
\end{align}

The maximum stress for carbon fiber is, $\sigma_{\text{carbon fiber}} = 475$ [MPa].\cite{carbonfiber}
The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.

Finally, the weight of the spar cap is computed as

\begin{align}
    \label{e:sparmass}
    \Delta W_i &\geq \rho_{\text{carbon fiber}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
    \label{e:sparmasssum}
    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
\end{align}

where $\rho_{\text{carbon fiber}} = 1.4$ [g/cm$^3$].\cite{carbonfiber}

\subsection{Wing Skin Weight}

It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is

\begin{equation}
    \label{e:wingskinweight}
    W_{\text{skin}} \geq 2 \rho_{\text{carbon fiber}} S g 
\end{equation}

where $\rho_{\text{carbon fiber}} = 0.043$ [g/cm$^2$].\cite{carbonfiber}

\subsection{Empennage}

An empennage model with a single tail consisting of a tail boom, horizontal and vertical tail, is added to both the solar-electric and gas powered aircraft.  
The empennage adds both weight and drag to the each aircraft.

The tail boom has a diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E$, and density $\rho_{\text{carbon fiber}} = 1.4$ [g/cm$^3$], and length $l_{\text{h}}$. 
The total mass and root bending inertia are imposed in the optimization model as

\begin{align}
    m &\geq \pi \rho_{\text{carbon fiber}} t_0 d l_{\text{h}} \left( 1 + \frac{1}{2} k\right) \\
    I_0 &\leq \pi t_0 d^3/8
\end{align}

where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.  For both the solar-electric and gas powered aircraft $k=0.8$.  When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis.

\begin{align}
    \label{e:boomdefl}
    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}
\end{align}

The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition, $V_{\text{h}} = 0.45$.\cite{aircraftrules}

\begin{equation}
    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
\end{equation}

% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
% 
% \begin{equation}
%     \label{e:deltacg}
%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
% \end{equation}
% 
% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g. travel range,
% 
% \begin{align}
%     \label{e:smcorr}
%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
%     \label{e:htv}
%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
% \end{align}
% 
% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
% 
% \begin{align}
%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
%     \label{e:smcorrmon}
%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
% \end{align}
% 
% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.

The vertical tail is sized to meet a conservative volume tail coefficient, $V_{\text{v}}= 0.02$,\cite{aircraftrules}

\begin{equation}
    \label{e:vtv}
    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
\end{equation}

where $l_{\text{v}}$ is the vertical tail moment arm and assume to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.

Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{\text{carbon fiber}} = 0.1$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$]. 
The weight of the horizontal and vertical tails is

\begin{align}
    \label{e:htweight}
    W_{\text{h}} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{\text{carbon fiber}} S_{\text{h}} \\
    \label{e:vtweight}
    W_{\text{v}} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{\text{carbon fiber}} S_{\text{v}}
\end{align}

where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.

The drag of the empennage was considered as three separate parts with no interference drag.  The drag of the tail boom was calculated using a turbulent flat plate model,

\begin{align}
    \label{e:boomdrag}
    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
\end{align}

The drag of the horizontal and vertical tails was computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,

\begin{align}
    D_{\text{h}} &\geq \frac{1}{2} c_{d_{\text{h}}} \rho V^2 S_{\text{h}} \\
    D_{\text{v}} &\geq \frac{1}{2} c_{d_{\text{v}}} \rho V^2 S_{\text{v}} \\
    \label{e:taildrag}
    c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
                 & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
                 &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
    Re_{\text{v}} &= \frac{V\rho S_{\text{v}}/b_{\text{v}}}{\mu} \\
    Re_{\text{h}} &= \frac{V\rho S_{\text{h}}/b_{\text{h}}}{\mu} 
\end{align}

where $\tau_{\text{(v,h)}} = 0.08$ is the airfoil thickness. Assuming steady level flight where neither surface is generating lift, the XFOIL data was generated for a zero angle of attack. A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{taildragpolar.pdf}
    \caption{ \textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
	\label{f:taildragpolar}
	\end{center}
\end{figure}

\subsection{Elliptcal Fuselage for Gas Powered Aircraft}

For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.  The fuselage will increase the overall weight and drag of the aircraft.  The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a smal fuselage whose effects will be ignored.

The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,

\begin{equation}
    \label{e:fusevol}
    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
\end{equation}

where the fuel is assumed to be Jet A fuel with density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The volume of the fuselage is constrained by

\begin{equation}
    \label{e:fusevol2}
    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
\end{equation}

where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius. Using the length and radius the surface area can be calculated using Thomsen's approxmiation,\cite{ellipsoidSA}

\begin{equation}
    \label{e:fusesa}
    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + ((2R_{\text{fuse}})^2)^{1.6075}.
\end{equation}

The weight of the fuselage is based off of the surface area

\begin{equation}
    \label{e:fuseweight}
    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{\text{carbon fiber}} g
\end{equation}

where $\rho_{\text{carbon fiber}} = 0.085$ [g/cm$^2$].\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,

\begin{align}
    \label{e:fusedrag}
    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
    C_f &\geq \frac{0.455}{Re^{0.3}}
\end{align}

where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}

\begin{equation}
    \label{e:fuseform}
    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
\end{equation}

\subsection{Engine Weight}

The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}

\begin{equation}
    \label{e:electricengine}
    P_{\text{max}} = B_{PM} m_{\text{motor}}
\end{equation}

where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ration is $B_{PM} = 4140.8$ [W/kg]. \\

The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}

\begin{equation}
    \label{e:powerlaw}
    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
\end{equation}

where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].  Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg\cite{fitting}. Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{powervsweightfit.pdf}
    \caption{ \textbf{ Comparison of power law equation and engine data.  Engine data taken from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

\subsection{Gas Powered Engine Performance}

Two characteristics of gas engines affect the performance of long-endurance aircraft.  
The first is break specific fuel consumption, $BSFC$; a lower $BSFC$ will result in increased endurance.  
The second is the lapse rate.  
Assuming a propeller driven aircraft and a naturally aispirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
RCV Engines Ltd manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
It is assumed that engines of a similar class will perform similarly to the DF70 engine.  
Thus, this engine performance model is only valid for IC engines.

\subsubsection{Break Specific Fuel Consumption Constraint}

A GP-compatible curve was generated from the DF70 data to relate $BSFC$ and shaft power,

\begin{equation}
    \label{e:rpmtobsfc}
    \left( \frac{BSFC}{BSFC_{\text{min}}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
\end{equation}

where $BSFC_{\text{min}} = 0.3162$ [lb/hp/hr].\cite{rcvengines}
Using this approximations the required shaft power determines the $BSFC$.
A comparison of the GP-compatible approximations to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}. A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $BSFC$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of a similar class.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{powertobsfcfit.pdf}
 \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.  Function is representative of engine performance for engines of similar size.  The log-space RMS error is 0.007. }}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

\subsubsection{Engine Lapse Rate Calculation}

The lapse rate $L_{\text{eng}} \leq 1$, is assumed to affect the maximum power output,

\begin{equation}
    \label{e:lapse}
    L_{\text{eng}} = \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
\end{equation}

The lapse is calculated from the required flight altitude $h=15,000$ [ft], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 1000 ft,\cite{enginelapse}

\begin{equation}
    \label{e:lapsefit}
    L_{\text{eng}} = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
\end{equation}

\subsection{Climb Constraints}

Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must be big enough to provide the necessary power to meet the minimum climb rate,

\begin{equation}
    \label{e:climbrate}
    \dot{h} \geq 100 \text{ [ft/min]}.
\end{equation}

The climb rate affects the required thrust, and therefore the required power during climb,

\begin{equation}
    \label{e:climb}
    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}}{V}
\end{equation}

where $W$, is the weight of the aircraft during climb.

\section{Results}

With more detailed structural and performance models the accuracy of the optimization results improves. 
Figure~\ref{f:latvsmtowtrade} was recalculated with the additional constraints included from Section V and is presented in Figure~\ref{f:latvsmtowtradet}. 
The higher fidelity models reveal a more realistic limit to the size of the solar-electirc powered aircraft and actually improve the performance of the gas powered aircraft.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Gas Powered\label{f:mtowvslatgas}]{\includegraphics{mtowvslatgas.pdf}}
     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolar}]{\includegraphics{mtowvslatsolar.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{ Recalulation of trade study of max take off weight and latitude with more detailed structural and performance models.   }}
 \label{f:latvsmtowtradet}
\end{figure}

One way in which adding higher fidelity models increases the accuracy of the optimization is that higher fidelity models allow new varibles to be solved by the optimizer instead of being inputs. 
For example, the aspect ratio and non-wing drag coefficient, which were previously inputs, can now be optimized with more detailed models.  
At 31 degrees latitude and 90th percentile winds the optimum aspect ratio is $A=32.6$ and the the non-wing drag coefficient is $C_{d_0} = 0.0043$ for the solar-electric powered aircraft.

Additional models also improve the optimization accuracy because of new variable dependencies. 
By adding in a wing structural model, air density has a larger effect on the operating altitude of the solar-electric powered aircraft.  
It might be assumed that because the wind speeds are lowest at 67,000 ft at 31 degrees latitude, that the aircraft will always fly at 67,000 ft.  
Figure~\ref{f:altopersimple} confirms that without a structural model the aircraft will want to fly almost exactly at 67,000 ft.  
However, by adding a structural model larger wings now have a weight penalty and according to Equation~\eqref{e:slfweight}, for the same $C_L$, with a smaller wing, higher air density will generate more lift. 
Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density. 
Figure~\ref{f:altoper} shows that with a wing structural model the optimum operating point is not at the minimum wind speed.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Structural weight fraction\label{f:altopersimple}]{\includegraphics{windaltopersimple.pdf}}
     \subfigure[Detailed structural model\label{f:wstruct}]{\includegraphics{windaltoper.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.  With only a structural fraction as the structural model the optimum point is at the minimum wind speed.  With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed. }}
 \label{f:altoper}
\end{figure}

Higher fidelity results can also been seen using detailed structural and performance models in the gas powered aircraft. 
Interesting, a direct comparison between the simple and detailed optimization models shows a cross over around an endurance of 6.5 days, Figure~\ref{f:simpdetcomp}. The comparison was run at a latitude of 38 degrees and 90th percentile winds.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Max take off weight comparison\label{f:simpdetcomp}]{\includegraphics{simpdetcomp.pdf}}
     \subfigure[Non-wing drag coefficient comparison\label{f:nonwingdrag}]{\includegraphics{nonwingdrag.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Comparison of gas powered optimization with and without detailed structural and performance model.  Non-wing drag variation due to Reynolds number effects cause of cross over in Figure~\ref{f:simpdetcomp}. }}
 \label{f:gasdetcomp}
\end{figure}

An analysis of the non-wing drag coefficienct, Figure~\ref{f:nonwingdrag}, reveals that cross over between the static non-wing drag coefficient $C_{d_0}=0.01$, and the more detailed component drag breakdown is one cause of the discrepancy between the detailed and non-detailed optimization models. 
The decrease in non-wing drag, even as the size of the aircraft increases, is due to a Reynolds number effect.  
For a constant airspeed, air density and viscosity, as the aircraft increases in size, Reynolds number will increase which will decrease the non-wing drag coefficient.

\subsection{Solar-electric Powered Trade Space Exploration}

One way analyze the trade space of the solar-electirc powered aircraft it is to change the objective function.  
High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
Altering the objective function to solar cell area can show difference configurations that meet the same requirements.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{solarobjcomp.pdf}
    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$. }}
	\label{f:solarobjcomp}
	\end{center}
\end{figure}

For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, solar cell efficiencies, altitudes, and availabilities.  (Figure~\ref{f:solarcontours}) 
The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.

 \begin{figure}[H]
 \begin{subfigmatrix}{3}% number of columns
  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl35a80.pdf}}
  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl35a85.pdf}}
  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl35a90.pdf}}
  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl30a80.pdf}}
  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl30a85.pdf}}
  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl30a90.pdf}}
  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl25a80.pdf}}
  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl25a85.pdf}}
  \subfigure[25h Latitude, 90\% Wind Speed]{\includegraphics{bcontourl25a90.pdf}}
 \end{subfigmatrix}
 \caption{ \textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft. Each point on the graph corresponds to a unique design.} }
 \label{f:solarcontours}
\end{figure}

\section{Conclusion}

Using geometric programming as a form of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.  
One high level discovery is the important of wind speed.  
If station keeping at all times during the flight is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of then time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery energy density and solar cell efficiency.  Using higher energy density batteries can result in significant weight and performance savings.  
The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between two options.
While the results of these findings are dependent on the accuracy of the models used in the optimization, it is shown that higher fidelity models can be incorporated to increase accuracy.

% produces the bibliography section when processed by BibTeX
\bibliography{biblibrary}
\bibliographystyle{aiaa}

\section*{Appendix A}

\subsection{Wind Speed Constraints}

The each wind speed constraint follows the following form

\begin{equation}
    \label{e:windspeedgen}
    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
\end{equation}

Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.

\tiny
\begin{longtable}{lccccccccccccc} \\
\label{t:windvals} \\
    \toprule
    \toprule
    \caption{Wind Speed Function Coefficients and Exponents} \\
    Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
    \midrule
20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
\bottomrule
\end{longtable}
\normalsize

\section*{Appendix B}

\subsection{Discussion on the use of the \emph{JH01} airfoil}

The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=3e5$.
The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela who helped in the redesign.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics{cpmin.pdf}}
     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics{cpmax.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{ Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.  }}
 \label{f:jhcps}
\end{figure}

% \section*{Appendix C}
% 
% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
% 
% \begin{align}
%     \label{e:mcenter}
%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
%     \label{e:eq6}
%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
% \end{align}
% 
% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
% 
% Using moment and lift approximations
% 
% \begin{align}
%     C_{m_{\text{w}}} &= constant \\
%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
%     \label{e:eq10}
%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
% \end{align}
% 
% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
% 
% \begin{equation}
%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
% \end{equation}
% 
% so that \eqref{e:eq10} can be written as
% 
% \begin{equation}
%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) 
% \end{equation}
% 
% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
% 
% \begin{align}
%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
% \end{align}
% 
% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
% 
% \begin{align}
%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
% \end{align}
% 
% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
% 
% \begin{equation}
%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
% \end{equation}
% 
% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
% 
% \begin{equation}
%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
% \end{equation}
% 
% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
% 
% \begin{equation}
%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
% \end{equation}
% 
% The pitch authority requirement is that at the forward-most c.g. position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$. Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation 
% 
% \begin{align}
%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
% \end{align}
% 
% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.

\end{document}

% - Release $Name:  $ -

### 2017-01-22 gassolar/solar/solar_irradiance.py

@@ -6,7 +6,7 @@ from gpfit.fit import fit
 import pandas as pd
 plt.rcParams.update({'font.size':15})
 
-def get_Eirr(latitude, day):
+def get_Eirr(latitude, day, N=50.0):
     """
     day is juilian day, measured from Jan 1st
     latitude is in degrees
@@ -26,7 +26,7 @@ def get_Eirr(latitude, day):
              0.002697*cos(3*beta) + 0.00148*sin(3*beta))
     tstart = 12/np.pi*arccos(-tan(delta)*tan(lat))
     tend = -tstart
-    t = np.linspace(tstart, tend, 50)
+    t = np.linspace(tstart, tend, N)
     costhsun = sin(delta)*sin(lat) + cos(delta)*cos(lat)*cos(2*np.pi*t/24)
 
     r0 = 149.597e6 # avg distance from earth to sun km
@@ -36,23 +36,55 @@ def get_Eirr(latitude, day):
     P0 = Psun*4*np.pi*Rsun**2/4/np.pi/Reo**2
     tau = np.exp(-0.175/costhsun)
     P = P0*costhsun# *tau
-    E = np.trapz(P)*(abs(tend-tstart))/50.0
+    E = np.trapz(P)*(abs(tend-tstart))/N
     tday = tstart*2
     tnight = 24-tstart*2
     plot = [P, t]
     return E, tday, tnight, plot
 
 if __name__ == "__main__":
-    ES, td, tn, p = get_Eirr(45, 31+28+21)
+    ES, td, tn, p = get_Eirr(30, 355, N=1000)
     fig, ax = plt.subplots()
-    ax.plot(p[1], p[0])
+    ax.fill_between([-12, -td/2], 0, 80, alpha=0.5, facecolor="b",
+                    linewidth=2, color="b")
+    ax.fill_between([td/2, 12], 0, 80, alpha=0.5, facecolor="b", linewidth=2,
+                    color="b")
+    ax.fill_between(p[1], p[0], 80, where=p[0] < 80, linewidth=2, color="b",
+                    facecolor="m", alpha=0.5)
+    ax.fill_between(p[1], 0, p[0], hatch="xx", color="r", linewidth=2,
+                    facecolor="none")
+    newp = []
+    for pp in p[0]:
+        if pp < 80:
+            newp.append(pp)
+        else:
+            newp.append(80)
[…]

### 2017-01-24 gassolarpaper/gassolarpaper.tex
restructure with requirements section

@@
 	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
-    \caption{ \textbf{ Trade study showing how gas powered aircraft weight increases with endurance.  }}
+    \caption{\textbf{Trade study showing how gas powered aircraft weight increases with endurance.}}
 	\label{f:mtowvsendurance}
@@
 
-To capture the key trade offs like those presented in Figures~\ref{f:solarstart} and ~\ref{f:mtowvsendurance} a geometric programm is built that models for both gas and solar-electric powered, long-endurance aircraft.
+To capture the key trade offs like those presented in Figures~\ref{f:solarstart} and~\ref{f:mtowvsendurance} a geometric programm is built that models for both gas and solar-electric powered, long-endurance aircraft.
 Equations governing the underlying physics of aerodynamic performance, solar energy collection and storage, and fuel burn are written in a geometric programming form, which can then be solved as an optimization problem.
@@
 
-\section{Geometric Programming\cite{gp}}
-
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions. GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,...,m \\
-g_i (\bold{x}) &= 1, i = 1,...,p \nonumber 
-\end{align}
-
-where the functions $f_i$ must be \emph{monomial} functions and the functions $g_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}  
-To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} \\
-
-\section{Aircraft Sizing Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft underlying physics models are used.  
-Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar insolation and aircraft models governing the performance, aerodynamics and weight breakdown are used in the optimization model.
-Equations from these models are then expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, higher fidelity models can be implemented.
-
-\subsection{Station Keeping}
-
-The design problem considered is a long-endurance, station keeping aircraft or an aircraft capable of remaining in one location in the air for long periods of time. 
-Long-endurance, station keeping aircraft are generally required to operate at specific latitudes, altitudes, and percentile wind speeds.\cite{zephyr} 
-Both wind speeds and atmospheric quantities depend on the station keeping requirements.
-
-\subsubsection{Wind Speeds}
-
-To station keep, an aircraft must fly faster than the wind speed.  
-Therefore, the aircraft speed is constrained for both the gas and solar-electric powered cases,
+\section{Requirements Analysis}
+
+Design requirements to operate at specific altitudes, latitudes, and percentile wind speeds each of which will affect the performance and design of the aircraft.\cite{zephyr} 
+To station keep an aircraft must fly faster than the wind speed,
 
@@
 
-An analysis was done of how wind speeds vary from month to month to determine during which seasons or months the wind speeds are higher.
-Figure~\ref{f:windvsmonth} reveals that wind speeds are significantly worse during the winter months.
-Because the constraining case for solar-electric powered aircraft is the winter solstice and because the wind speeds are highest during the winter months only the wind speed data from December will be used for aircraft the sizing and performance analysis presented in this paper.
+\subsection{Endurance}
+
+Long-endurance gas powered aircraft, the endurance requirement the amount of time the aircraft needs to remain in the air without refueling.  
+If solar-electric powered aircraft can fly during the shortest day of the year, December 21st, then it can theortically fly at any time the year.  
+Therefore, the endurance requirement for solar-electric powere aircaft is assumed to be a requirement to fly during the winter solstice.  
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section , and with higher wind speeds as shown in Figure~\ref{f:windvsmonth}.
 
@@
 	\includegraphics[width=0.6\textwidth]{windvsmonth.pdf}
-    \caption{ \textbf{ Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
-}}
+    \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
+Lower bound, upper bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.}}
 	\label{f:windvsmonth}
@@
 
+\subsection{Altitude}
+
+Gas powered aircraft generally fly at a given altitude to meet a payload requirement.  
+A naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} mean that gas powered aircraft will generally be more efficiect at lower altitude and generally fly between 5,000 at 25,000 ft depending on the payload requirement.
+For this design study, it is assume d that the aircraft is flying at 15,000 ft.
+
+Solar-electric powered aircraft tend to fly around 60,000 ft to meet payload requirements, avoid cloud coverage and fly in lower wind speeds which have a minimum around 60,000 shown in Figure~\ref{f:altvswind}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth]{altvswind.pdf}
+    \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
+Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
+Wind data includes winds during December for the year 2005-2015.}}
+	\label{f:altvswind}
+	\end{center}
+\end{figure}
+
+The altitude also determines atmospheric properties such as air density and viscosity.  For the solar aircraft air density is calculated using equations specific to the tropopause,\cite{isaatm}
+
+\begin{align}
+    \label{e:tropopress}
+    p &= \rho R_{\text{spec}}T_{11} \\
+    \label{e:tropoalt}
+    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}} 
+\end{align}
+
+According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+\\
+
+For the gas powered aircraft air density is calculated using atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computer prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K].
+
+\subsection{Latitude}
+
+Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
 Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere.
-Because the gas-powered aircraft is assumed to have a naturally aspirated engine, it will fly between 5,000 and 25,000 ft depending on payload requirement. 
-The solar-electric powered aircraft will operate around 60,000 ft to avoid cloud coverage.  
-Variations of wind speed with latitude are shown for altitudes at both 15,000 ft at and 60,000 ft.  \\
 
@@
 	\includegraphics[width=0.6\textwidth]{latvswind.pdf}
-    \caption{ \textbf{ Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
+    \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
 Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
@@
 
-Wind speeds also vary significantly with altitude. (Figure~\ref{f:altvswind})
-The high winds around 30,000-40,000 ft makes it difficult for both solar-electric and gas powered long-endurance aircraft to fly in that range of altitudes.  
-Solar-electric powered aircraft, that do not have aspirated engines, fly around 60,000 ft take advantage of lower wind speeds and limited cloud coverage.
-Gas powered aircraft with aspirated engines can have difficulty operating at 60,000 ft and generally perform better in the lower wind speeds below 25,000 ft.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth]{altvswind.pdf}
-    \caption{ \textbf{ Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
-Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
-Wind data includes winds during December for the year 2005-2015.}}
-	\label{f:altvswind}
-	\end{center}
-\end{figure}
-
+Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
+This is further discussed in Section.
+
+Operationally, long-endurance aircraft will fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
+For example, an aircraft optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude. 
+This can be imposed in the optimization model as multiple constraints. For the 35 degree latitude case, 70 separate wind and solar flux constraints that must be satisfied, one for each latitude.
+
+\subsection{Percentile Wind Speed}
+
+For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data in December across all longitudes and spanning the years 2005-2015.  
 To understand what a certain percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
 Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
-One lag is equivalent to 6 hours.
 
@@
 
-\subsubsection{Optimizing Altitude for Solar-Electric Powered Aircraft}
+\section{Geometric Programming\cite{gp}}
+
+Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions. GPs have the form
+
+\begin{align} 
+\label{e:gpform}
+\text{minimize } f_0(\bold{x}) & \nonumber \\
+\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
+g_i (\bold{x}) &= 1, i = 1,\ldots,p \nonumber 
+\end{align}
+
+where the functions $f_i$ must be \emph{monomial} functions and the functions $g_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
+
+\begin{align}
+ \label{e:mon}
+g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
+\label{e:pos}
+f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
+\end{align}
+
+The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
+Additionally, geometric programs can be solved rapidly with existing algorithms.  
+Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}  
+To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} \\
+
+\section{Aircraft Sizing Models}
+
+To evaluate the size and performance of both gas and solar-electric powered aircraft underlying physics models are used.  
+Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar insolation and aircraft models governing the performance, aerodynamics and weight breakdown are used in the optimization model.
+Equations from these models are then expressed in a GP-compatible form to enable their use in the optimization. 
+Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
+To obtain higher fidelity results from the optimization, higher fidelity models can be implemented.
+
+\subsection{Wind Speeds}
 
@@
 Because wind speeds also vary with latitude several constraints are needed to optimize altitude. 
-Operationally, aircraft, such as the solar-electric powered aircraft, will fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
-For example, an aircraft optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude. 
-This is imposed in the optimization model as multiple constraints. For the 35 degree latitude case, 70 separate wind and solar flux constraints that must be satisfied, one for each latitude.
-
 Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg\cite{fitting}.
@@
 	\includegraphics[width=0.6\textwidth]{windfitl35.pdf}
-    \caption{ \textbf{ Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}.
-(Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.
-}}
+    \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}.
+(Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.}}
 	\label{f:windfitl35}
@@
 
-\subsubsection{Atmospheric Quantities}
-
-For the solar aircraft air density is being optimized and the operating altitude is computed after the fact using equations specific to the tropopause,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}} 
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-Because altitude is a requirement for the gas powered aircraft, density is pre-computed prior to solving the optimization using atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computer prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K].
-
 \subsection{Solar-Electric Power}
 
-The driving equations for the solar-electric powered aircraft are determined by the energy required to fly throughout the day while charging the batteries enough to last through the night.  
-Equations~\ref{e:solarreq} and~\ref{e:solarbatt} capture the solar requirement by enforcing that the total solar energy harvested by the solar cells during the day must be greater than the energy used to operate the aircraft during the day while charging enough batteries to fly during the night.\cite{solartech}
-Under these assumptions, if the aircraft can fly for one day and one night, then it could theoretically fly for many days under similar circumstances.
+At a given time of year DOY, and latitude $\phi$, there exists an availble solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day
+
+\begin{equation}
+    \label{e:solarfunc}
+    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
+\end{equation}
+
+To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$ and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
 
@@
     \label{e:solarreq}
-        (E/S)_{\text{sun}} \eta_{\text{solar}} S_{\text{solar}} &\geq P_{\text{oper}}t_{\text{day}} + E_{\text{battery}}/\eta_{\text{charge}} \\
+    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{battery}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
     \label{e:solarbatt}
-        E_{\text{battery}} &\geq P_{\text{oper}}t_{\text{night}}/\eta_{\text{discharge}}
-    \end{align}
-
-    It is assumed that the charge and discharge efficiencies are constant, $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$. 
-    It is also assumed that the solar cell efficiency is constant, $\eta_{\text{solar}} = 0.3$. 
-    A battery with an energy density of $h_{\text{batt}} = 350$ [Whr/kg], is selected for the solar-powered aircraft.
-
-    To enforce these constraints, two variables were created: $P_{\text{oper}}$ and $S_{\text{solar}}$. 
-    $P_{\text{oper}}$ is defined as the average power required to operate the aircraft. 
-    $S_{\text{solar}}$ is defined as the amount of area on the aircraft covered by solar cells.
-
-    \begin{align}
-        \label{e:solarpoper}
-        P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}} \\
-        \label{e:solarssolar}
-        S_{\text{solar}} &\leq f_{\text{solar}}S
-    \end{align}
-
-    For the purposes of this design study, it is assumed that the solar cells are only placed on the wing and that the total solar cell area is limited to some fraction of the wing planform area $f_{\text{solar}}$, to account for control surfaces, leading edge, and other accessories.
-
-    Because solar aircraft requirements are typically given in terms of latitude and yearly availability it is useful to express the total available solar energy per unit area per day as a function of latitude, $\phi$, and the day of the year, $DOY$,
-
-    \begin{equation}
-        \label{e:solarfunc}
-        (E/S)_{\text{sun}} = f(\phi, DOY).
-    \end{equation}
-
-    For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(E/S) = E$). 
-    The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        E_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} P_{\text{sun}} dt.
-    \end{equation}
-
-    Figure~\ref{f:lat45} shows a sample distribution of the available solar power during the spring equinox at 45 degrees latitude.
+    E_{\text{battery}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_C \eta_{\text{solar}} S_{\text{solar}}
+\end{align}
+
+where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
+An additional energy term $(E/S)_C$ is included in Equation~\ref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on battery power.
+Equations~\ref{e:solarreq} and~\ref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
 
@@
 	\includegraphics[width=0.6\textwidth]{lat30.pdf}
-    \caption{ \textbf{ The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.
-The area under this curve is the total energy available per unit area for the entire day.} }
-	\label{f:lat45}
+    \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.
+For a solar-electirc powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
+	\label{f:lat30}
 	\end{center}
@@
 
-The shape of this curve is governed by the angle, $\theta$, between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
+In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
+
+\begin{align}
+    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
+    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}. \\
+\end{align}
+
+where $\eta_{\text{motor}} = 0.95$.
+Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_C$ are functions of $(P/S)_{\text{min}}$.
+The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions $(P/S)_{\text{min}}$, one for each latitude between 20 and 60 degrees latitude.
+Figure~\ref{f:energyapprox} sample monomial approximations at 30 degrees latitude.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics{Benergy.pdf}}
+     \subfigure[Monomial Approximation for $(E/S)_C$\label{f:Cenergy}]{\includegraphics{Cenergy.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude.}}
+ \label{f:energyapprox}
+\end{figure}
+
+    For the purposes of this design study, it is assumed that the solar cells are can be placed on the wings and the horizontal tail but are not placed on the vertical tail.  
+    The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are placed on the horizontal tail,
+
+\begin{equation}
+    \label{e:solarssolar}
+    S_{\text{solar}} \leq f_{\text{solar}}S.
+\end{equation}
+
+The total available solar energy per day is an integral of the available solar power during the course of the day,
+
+    \begin{equation}
+        \label{e:solares}
+        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
+    \end{equation}
+
+For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
+The solar irradiated power $P_{\text{sun}}$ a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
 

### 2017-01-24 gassolarpaper/gassolarpaper.tex
comment out simple model section

@@
 
-\section{Long-endurance Aircraft Trade Space}
-
-One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight,
-
-\begin{equation}
-    \label{e:structualfraction}
-    W_{\text{structural}} \geq f_{\text{structural}}MTOW.
-\end{equation}
-
-where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
+% \section{Long-endurance Aircraft Trade Space}
+% 
+% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
+% 
+% \begin{equation}
+%     \label{e:structualfraction}
+%     W_{\text{structural}} \geq f_{\text{structural}}MTOW.
+% \end{equation}
+% 
+% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
 Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-
-% \begin{longtable}{llll}
-% \caption{Fixed Optimization Parameters} \\
-% \toprule
-% \toprule
-% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% \midrule
-% Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% $A$                 & 25     & $A$                 & 25              \\
-% $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% $BSFC$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-%                      &        & $h_{\text{battery}}$ [Whr/kg]  & 350 \\
-%                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-%                      &        & Day of the Year      & 355 (Dec 21st)\\
-% \bottomrule
-% \label{t:gassolarparams}
-%  \end{longtable}
-
-\subsection{Sizing for Different Latitude Bands}
-
-An analysis of the max take off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-This analysis is achieved by minimizing the max take off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-For this study, the gas-powered aircraft is constrained to fly for 8 days.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics{mtowvslatgassimple.pdf}}
-     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics{mtowvslatsolarsimple.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{ Trade study of max take off weight and latitude.
+% 
+% % \begin{longtable}{llll}
+% % \caption{Fixed Optimization Parameters} \\
+% % \toprule
+% % \toprule
+% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
+% % \midrule
+% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
+% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
+% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
+% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
+% % $A$                 & 25     & $A$                 & 25              \\
+% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
+% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
+% % $BSFC$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
+% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
+% %                      &        & $h_{\text{battery}}$ [Whr/kg]  & 350 \\
+% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
+% %                      &        & Day of the Year      & 355 (Dec 21st)\\
+% % \bottomrule
+% % \label{t:gassolarparams}
+% %  \end{longtable}
+% 
+% 
+% \subsection{Sizing for Different Latitude Bands}
+% 
+% An analysis of the max take off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
+% This analysis is achieved by minimizing the max take off weight in both the gas and solar-electric powered optimization models at different latitudes. 
+% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics{mtowvslatgassimple.pdf}}
+%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics{mtowvslatsolarsimple.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{ Trade study of max take off weight and latitude.
 Plots show required maximum take off weight required to operate between a band of latitudes.
@@
 }}
- \label{f:latvsmtowtrade}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-
-\subsection{Gas Powered Endurance}
-
-While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-The gas powered optimization model was set to minimize max take off weight for different endurance requirements to understand what endurance limits are possible.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth]{mtowvsendsimple.pdf}
-    \caption{ \textbf{ Trade study between endurance and max take off weight for a gas powered archicture. }}
-	\label{f:mtowvsendsimple}
-	\end{center}
-\end{figure}
-
-For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
+%  \label{f:latvsmtowtrade}
+% \end{figure}
+% 
+% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
+% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
+% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
+% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
+% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
+% 
+% \subsection{Gas Powered Endurance}
+% 
+% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
+% The gas powered optimization model was set to minimize max take off weight for different endurance requirements to understand what endurance limits are possible. 
+% 
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth]{mtowvsendsimple.pdf}
+%     \caption{ \textbf{ Trade study between endurance and max take off weight for a gas powered archicture. }}
+% 	\label{f:mtowvsendsimple}
+% 	\end{center}
+% \end{figure}
+% 
+% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
 

### 2017-01-25 gassolar/gas/flight_state.py
cl plot; sensitivity table

@@ -19,13 +19,14 @@ class FlightState(Model):
         density, vis = get_airvars(altitude)
 
         Vwind = Variable("V_{wind}", wind, "m/s", "wind velocity")
+        mfac = Variable("m_{fac}", 1.0, "-", "wind speed margin factor")
         V = Variable("V", "m/s", "true airspeed")
         rho = Variable("\\rho", density, "kg/m**3", "air density")
         mu = Variable("\\mu", vis, "N*s/m**2", "dynamic viscosity")
         h = Variable("h", altitude, "ft", "flight altitude")
         href = Variable("h_{ref}", 15000, "ft", "reference altitude")
 
-        constraints = [V >= Vwind,
+        constraints = [V/mfac >= Vwind,
                        rho == rho,
                        mu == mu,
                        h == h,

### 2017-01-25 gassolar/solar/plotting.py

@@ -8,7 +8,7 @@ path = os.path.abspath(__file__).replace(os.path.basename(__file__), "").replace
 path = "/Users/mjburton11/MIT/GPKIT/gpkit-projects/gas_solar_trade/gassolar/environment/"
 DF = pd.read_csv(path + "windaltfitdata.csv")
 
-def windalt_plot(latitude, sol):
+def windalt_plot(latitude, sol1, sol2):
     plt.rcParams.update({'font.size':15})
     alt = np.linspace(40000, 80000, 20)
     den = density(alt)
@@ -25,14 +25,11 @@ def windalt_plot(latitude, sol):
 
     vwind = (np.exp(softmax_affine(x, params)[0])*100).reshape(6, 20)[3]
     fig, ax = plt.subplots()
-    ax.plot(alt/1000.0, vwind*1.95384)
-    altsol = altitude(min([sol(sv).magnitude for sv in sol("\\rho")]))
-    vsol = max([sol(sv).to("knots").magnitude for sv in sol("V")])
-    ax.plot(altsol/1000, vsol, "o", markersize=10, label="operating point")
-    ax.annotate('operating point', xy=(altsol/1000.0, vsol),
-                xytext=(altsol/1000.0-15.0, vsol-15),
-                arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
-                                headwidth=10))
+    ax.plot(alt/1000.0, vwind*1.95384, linewidth=2)
+    for sol in [sol1, sol2]:
+        altsol = altitude(min([sol(sv).magnitude for sv in sol("\\rho")]))
+        vsol = max([sol(sv).to("knots").magnitude for sv in sol("V")])
+        ax.plot(altsol/1000, vsol, "o", markersize=10)
     ax.set_xlabel("Altitude [kft]")
     ax.set_ylabel("Wind Speed [knots]")
     ax.grid()

### 2017-01-25 gassolar/solar/print_sens.py

@@ -1,0 +1,23 @@
+"print sensitivities"
+
+def sens_table(sols, varnames):
+    with open("../../gassolarpaper/sens.generated.tex", "w") as f:
+        f.write("\\begin{longtable}{lccccccccccccc}\n")
+        f.write("\\caption{Sensitivities}\\\\\n")
+        f.write("\\toprule\n")
+        f.write("\\toprule\n")
+        f.write("\\label{t:sens}\n")
+        f.write("& Latitude $=25$ & Latitude $=25$ & Latitude $=25$ & Latitude $=25$ \\\\\n")
+        f.write("Variables & 85th Percentile Winds & 85th Percentile Winds & 90th Percentile Winds & 90th Percentile Winds \\\\\n")
+        f.write("\\midrule\n")
+        for vname in varnames:
+            sens = []
+            for s in sols:
+                sen = s["sensitivities"]["constants"][vname]
+                if hasattr(sen, "__len__"):
+                    sen = s["sensitivities"]["constants"][max(sen)]
+                sens.append(sen)
+            vals = "$" + vname + "$ &" + " & ".join(["%.3g" % x for x in sens])
+            f.write(vals + "\\\\\n")
+        f.write("\\bottomrule\n")
+        f.write("\\end{longtable}")

### 2017-01-25 gassolar/solar/solar.py

@@ -10,9 +10,10 @@ from gpkitmodels.aircraft.GP_submodels.tail_boom import TailBoomState
 from gpkitmodels.aircraft.GP_submodels.tail_boom_flex import TailBoomFlexibility
 from gpkitmodels.helpers import summing_vars
 
-path = os.path.abspath(__file__).replace(os.path.basename(__file__), "").replace(os.sep+"solar"+os.sep, os.sep+"environment"+os.sep)
-DF = pd.read_csv(path + "windaltfitdata.csv")
-DF2 = pd.read_csv("solarirrdata.csv")
+basepath = os.path.abspath(__file__).replace(os.path.basename(__file__), "")
+windpath = basepath.replace(os.sep+"solar"+os.sep, os.sep+"environment"+os.sep)
+DF = pd.read_csv(windpath + "windaltfitdata.csv")
+DF2 = pd.read_csv(basepath + "solarirrdata.csv")
 
 class Aircraft(Model):
     "vehicle"
@@ -129,7 +130,7 @@ class SolarCells(Model):
         g = Variable("g", 9.81, "m/s**2", "gravitational constant")
         S = Variable("S", "ft**2", "solar cell area")
         W = Variable("W", "lbf", "solar cell weight")
-        etasolar = Variable("\\eta", 0.2, "-", "Solar cell efficiency")
+        etasolar = Variable("\\eta", 0.22, "-", "Solar cell efficiency")
 
         constraints = [W >= rhosolar*S*g]
 

### 2017-01-25 gassolarpaper/gassolarpaper.tex
cl plot; sensitivity table
clean up results section

@@
 
-\subsection{Sensitivity Analysis}
-
-When a GP is solved, the sensitivity of the objective function with respect to each constraint is also returned.
-From this information, the sensitivity to each inputted variable with respect to the objective function can be extracted.\cite{hoburgthesis}
-For example, if the objective function were weight and the sensitivity to solar cell efficiency were -0.5, then a 1\% increase in the solar cell efficiecy would result in a -0.5\% decrease in weight.
-From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 
-To achieve a more accurate numbers for these variable they can be replaced with higher fidelity models.  
-For example, the structural weight fraction $f_{\text{structrual}}$, which has a high sensitivity, could be replaced by a structural model of the wing, fuselage and empennage. 
-Table~\ref{t:sens} shows the varialbes with the highest sensitivities for both the solar-electric powered aircraft optimized at the $\pm$40 degree latitude requirement minimizing weight.
-
-\begin{longtable}{ll}
-\caption{Sensitivities} \\
-\toprule
-\toprule
- Varialbe & Sensitivities \\
-\midrule
-$h_{\text{batt}}$         & -3.475 \\
-$\eta_{\text{solar}}$     & -1.268 \\
-$f_{\text{structural}}$   & 2.445  \\
-$\eta_{\text{discharge}}$ & -5.068 \\
-$\eta_{\text{prop}}$      & -4.743 \\
-$c_{d_0}$                 & 0.9935 \\
-\bottomrule
-\label{t:sens}
- \end{longtable}
-
 % \subsection{Solar-electric Powered Trade Space Exploration}
@@
 
-\section{Detailed Models}
-
-To improve the fidelity of the optimization, higher fidelity models can be implemented.  
-A spar and structural wing model are added to both the gas and solar-electric powered optimization models.  
-An empennage model is added to account for the additional weight and drag of horizontal and vertical tails and a tail boom.  
-Additional models were added to the gas powered aircraft optimization model to account for engine weight and performance. 
-These models, similar to those described in section III, are a set of constraints derived from governing first principles put in a GP-compatible form. 
-This results in a more comprehensive and detailed optimization model for aircraft sizing.   \\
-
 \subsection{Wing Spar Model}
@@
 
-With more detailed structural and performance models the accuracy of the optimization results improves. 
-Figure~\ref{f:latvsmtowtrade} was recalculated with the additional constraints included from Section V and is presented in Figure~\ref{f:latvsmtowtradet}. 
-The higher fidelity models reveal a more realistic limit to the size of the solar-electirc powered aircraft and actually improve the performance of the gas powered aircraft.
+To achieve sizing results for the solar-electric and gas powered long-endurance aircraft the optimization models were solve by minimizing the wing span across different bands of latitudes, shown in Figure~\ref{f:latvsspantrade}.
 
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Gas Powered\label{f:mtowvslatgas}]{\includegraphics{mtowvslatgas.pdf}}
-     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolar}]{\includegraphics{mtowvslatsolar.pdf}}
+     \subfigure[Gas Powered\label{f:spanvslatgas}]{\includegraphics{spanvslatgas.pdf}}
+     \subfigure[Solar-Electric Powered\label{f:spanvslatsolar}]{\includegraphics{spanvslatsolar.pdf}}
  \end{subfigmatrix}
- \caption{\textbf{ Recalulation of trade study of max take off weight and latitude with more detailed structural and performance models.   }}
- \label{f:latvsmtowtradet}
-\end{figure}
-
-One way in which adding higher fidelity models increases the accuracy of the optimization is that higher fidelity models allow new varibles to be solved by the optimizer instead of being inputs. 
-For example, the aspect ratio and non-wing drag coefficient, which were previously inputs, can now be optimized with more detailed models.  
-At 31 degrees latitude and 90th percentile winds the optimum aspect ratio is $A=32.6$ and the the non-wing drag coefficient is $C_{d_0} = 0.0043$ for the solar-electric powered aircraft.
-
-Additional models also improve the optimization accuracy because of new variable dependencies. 
-By adding in a wing structural model, air density has a larger effect on the operating altitude of the solar-electric powered aircraft.  
+ \caption{\textbf{Trade study of wingspan and latitude.
+Plots show required maximum take off weight required to operate between a band of latitudes.
+A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.}}
+ \label{f:latvsspantradet}
+\end{figure}
+
+One way to interpret Figure~\ref{f:latvsspantrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
+This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
+Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
+Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
+On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
+
+Because gas-powered aircraft are endurance limited by the amount of fuel that they can carry, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
+Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitude only depends on wind speed. 
+If a gas-powered aircraft can fly at the latitude with the worst wind speed (38 degrees latitude at 15,000 ft) it can theoretically fly at any latitude.  
+Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing wing span at $\pm$38 degrees latitude.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth]{spanvsendurance.pdf}
+    \caption{\textbf{Trade study of endurance and size for gas powered architecture.}}
+	\label{f:spanvsendurance}
+	\end{center}
+\end{figure}
+
+Changing the objective function can be informative to the design space. 
+High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
+By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires few solar cells.
+(Figure~\ref{f:solarobjcomp})
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[Wingspan\label{f:span}]{\includegraphics{span.pdf}}
+     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics{solarcell.pdf}}
+ \end{subfigmatrix}
+    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
+    \label{f:solarobjcomp}
+\end{figure}
+
+\subsection{Changing Assumptions}
+
+By altering the aircraft structural model it can be observed how air density trades for wing weight. 
 It might be assumed that because the wind speeds are lowest at 67,000 ft at 31 degrees latitude, that the aircraft will always fly at 67,000 ft.  
-Figure~\ref{f:altopersimple} confirms that without a structural model the aircraft will want to fly almost exactly at 67,000 ft.  
-However, by adding a structural model larger wings now have a weight penalty and according to Equation~\eqref{e:slfweight}, for the same $C_L$, with a smaller wing, higher air density will generate more lift.
+If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
+
+\begin{equation}
+    W_{\text{structural}} \geq \text{MTOW} f_{\text{structural}}
+\end{equation}
+
+where $f_{\text{structural}} = 0.35$ then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
+However, if the structural weight is modeled as explained in Section larger wings have a weight penalty and the optimization trades air density for wing weight.
 Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-Figure~\ref{f:altoper} shows that with a wing structural model the optimum operating point is not at the minimum wind speed.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Structural weight fraction\label{f:altopersimple}]{\includegraphics{windaltopersimple.pdf}}
-     \subfigure[Detailed structural model\label{f:wstruct}]{\includegraphics{windaltoper.pdf}}
- \end{subfigmatrix}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth]{windaltoper.pdf}
  \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.
 With only a structural fraction as the structural model the optimum point is at the minimum wind speed.
-With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.
-}}
+With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.}}
  \label{f:altoper}
-\end{figure}
-
-Higher fidelity results can also been seen using detailed structural and performance models in the gas powered aircraft. 
-Interesting, a direct comparison between the simple and detailed optimization models shows a cross over around an endurance of 6.5 days, Figure~\ref{f:simpdetcomp}.
-The comparison was run at a latitude of 38 degrees and 90th percentile winds.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Max take off weight comparison\label{f:simpdetcomp}]{\includegraphics{simpdetcomp.pdf}}
-     \subfigure[Non-wing drag coefficient comparison\label{f:nonwingdrag}]{\includegraphics{nonwingdrag.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Comparison of gas powered optimization with and without detailed structural and performance model.
-Non-wing drag variation due to Reynolds number effects cause of cross over in Figure~\ref{f:simpdetcomp}.
-}}
- \label{f:gasdetcomp}
-\end{figure}
-
-An analysis of the non-wing drag coefficienct, Figure~\ref{f:nonwingdrag}, reveals that cross over between the static non-wing drag coefficient $C_{d_0}=0.01$, and the more detailed component drag breakdown is one cause of the discrepancy between the detailed and non-detailed optimization models.
-The decrease in non-wing drag, even as the size of the aircraft increases, is due to a Reynolds number effect.  
-For a constant airspeed, air density and viscosity, as the aircraft increases in size, Reynolds number will increase which will decrease the non-wing drag coefficient.
-
-\subsection{Solar-electric Powered Trade Space Exploration}
-
-One way analyze the trade space of the solar-electirc powered aircraft it is to change the objective function.  
-High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-Altering the objective function to solar cell area can show difference configurations that meet the same requirements.
+	\end{center}
+\end{figure}
+
+Another interesting result is the operating lift to drag ratio for the gas powere aircraft.  
+The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
+However, while station keeping the aircraft will maintain a constant velocity during high wind speeds.  
+At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
+If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{solarobjcomp.pdf}
-    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$. }}
-	\label{f:solarobjcomp}
+	\includegraphics[width=0.6\textwidth]{polarmission.pdf}
+    \caption{\textbf{Comparision of operating lift to drag ratio with and without the wind speed constraint.
+Solid lines are drag polars at various Reynolds numbers.
+Without the wind speed constraint velocity is optimized to reach the maximum $C_L^{1.5}/C_D$ value.
+With the wind speed constraint, velocity and Reynolds number is constant operating off the maximum endurance lift to drag ratio.}}
+ \label{f:polarmission}
 	\end{center}
@@
 
-For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, solar cell efficiencies, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
+The solar aircraft size and design depends on the assumed solar cell efficiency and battery energy density. 
+By solving the model for different assumed solar cell efficiency and battery energy density values a broader picture of the design space is achieved.   
+Figure~\ref{f:solarcontours} shows a maxtrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
 
@@
  \end{subfigmatrix}
- \caption{ \textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
+ \caption{\textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft.
+Each point on the graph corresponds to a unique design.}}
  \label{f:solarcontours}
 \end{figure}
+
+\subsection{Sensitivities}
+
+When a GP is solved, the sensitivity of the objective function with respect to each constraint is also returned.
+From this information, the sensitivity to each inputted variable with respect to the objective function can be extracted.\cite{hoburgthesis}
+For example, if the objective function were weight and the sensitivity to solar cell efficiency were -0.5, then a 1\% increase in the solar cell efficiecy would result in a -0.5\% decrease in weight.
+From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 
+Table~\ref{t:sens} shows the varialbes with the highest sensitivities the solar-electric powered aircraft optimized at various latitudes and percentile wind speeds by minimizing weight.
+
+It is intersting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
+This ocurrs because the discharge efficiency is directly related to the operating power which goes by the cube of velocity. ($P_{\text{oper}}\sim V^3$)
+Thus, the discharge efficiency sensitivity shares similiar growth to the percentil wind speed, while the charge efficiency sensitivity does not.
+
+It is also interesting to note that the design is significantly more sensitivity to the battery energy density than to the solar cell efficiency.  
+Thus to design a smaller, lighter aircraft, the sensitivity analysis reveals that it is more benefical to invest in higher quality batteries than solar cells.
+
+The margin factor for the wing weight $m_{\text{fac-wing}}$, is a measure of the inaccuracy of the wind structural model.  
+If the wing weight is actually 5\% heavier than predicted then the aircraft will be approximately 4\% to 5\% heavier according Table~\ref{e:sens}.
+
+\input{sens.generated.tex}
 

### 2017-01-25 gassolarpaper/gassolarpaper.tex
revise intro

@@
 
-The design problem of a long-endurance, station-keeping aircraft is considered.
-Solar electric and gas powered architectures are proposed as solutions to this problem.
-Because sizing long-endurance aircraft is so multifaceted and the interaction between aerodynamics, structural weight, energy and other disciplines is so intertwined, a systematic and rapid approach is needed to explore this design space.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.
-
-The driving condition that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
+Long-endurance, station-keeping aircraft are potential solutions to providing internet, communication, and persistent survellience. 
+Recently, tech companies such as Google and Facebook have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
+Defense contractors like AeroVironment and small engineering firms like Vanailla Aircraft have developed gas powered long-endurance platforms as a means of providing continual survellience for days at a time.
+However, sizing a long-endurance aircraft is not obvious because of the multifacted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
+Additionally, because few long-endurance gas powered aircraft have been built and only Airbus's Zephyr solar aircraft has been successfully flown, the operational limits of these aircraft is relatively unknown.
+Even more unclear is understanding the trade offs between gas powered and solar-electric powered archictures.
+
+This paper presents a physics-based optimization model that uses geometric programming as systematic and rapid approach to explore this design space and highlight trade offs between the two proposed architecutres.
+Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
+Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
+The optimization is solved multiple times for different design parameters such as endurance, latitude, and percentile wind speed to observe how aircraft size and weight are affected by different requirements.
+
+The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
 To achieve a feasible design, solar-electric powered aircraft must carry enough solar cells and batteries required to fly during the day while storing enough energy to fly through the night.\cite{solartech}
@@
 Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-Figure~\ref{f:solarstart} shows that in order to operate at higher latitudes during the winter a larger aircraft is required.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth]{mtowvslatsolar.pdf}
-    \caption{ \textbf{ Trade study showing how solar-electric powered aircraft size increases with latitude operating capacity.  }}
-	\label{f:solarstart}
-	\end{center}
-\end{figure}
 
@@
 If a gas-powered aircraft can fly at the latitude with the worst wind speed in can theoretically fly anywhere else in the world.
- (Figure~\ref{f:mtowvsendurance})
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
-    \caption{\textbf{Trade study showing how gas powered aircraft weight increases with endurance.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-To capture the key trade offs like those presented in Figures~\ref{f:solarstart} and~\ref{f:mtowvsendurance} a geometric programm is built that models for both gas and solar-electric powered, long-endurance aircraft.
-Equations governing the underlying physics of aerodynamic performance, solar energy collection and storage, and fuel burn are written in a geometric programming form, which can then be solved as an optimization problem.
-Then by changing the design parameters, such as endurance, latitude, and percentile wind speeds, multiple solutions can be rapidly generated to observe how the aircraft size and weight is affected by different requirements.
 
@@
 A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-
-\subsubsection{Loading Case for Gas Powered Aircraft}
-
-It is assumed that the key spar sizing case for the gas powered aircraft is standard wing bending during steady level flight, fully fueled.
-The distributed load is a combination of the lifting distribution and weight distribution along the wing
+It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight bending case and gust load case.
+
+\subsubsection{Steady Level Flight Wing Loading Case}
+
+During steady level flight there exists a distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing
 
@@
 
-\subsubsection{Loading Case for Solar-electric Powered Aircraft}
-
-The key sizing case for the solar-electric aircraft spar is a gust load that varies across the span.  
-It is assumed that the solar aircraft is span loaded, meaning the majority of the weight lies along the wings,
-
-\begin{align}
-    W_{\text{wing}} &\geq W_{\text{cap}} + W_{\text{skin}} + W_{\text{battery}} + W_{\text{solar}}\\
-    W_{\text{cent}} &\geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{empennage}}.
-\end{align}
-
-For steady level flight, the same loading distribution as for the gas powered aircraft is assumed,
-
-\begin{equation}
-    \label{e:qbar}
-    q(y) = \frac{N_{\text{max}}W_{\text{cent}}}{b}\frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right).
-\end{equation}
-
-Because of the wing's large moment of inertia, caused by the weight of the batteries inside the wings, the solar-electric aircraft is considered a highly flexible aircraft. 
-To account for this flexiblity and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the loading distribution
+The center weight for the solar powered aircraft is
+
+\begin{equation}
+    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{empennage}}.
+\end{equation}
+
+\subsubsection{Gust Loading Case}
+
+Because long-endurance aircraft typically have high aspect ratio wing they are considered flexible aircraft. 
+To account for this flexiblity and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\ref{e:qbar}
 
@@
 where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
+The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the wing
+
+\begin{equation}
+    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
+\end{equation}
+
+For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
+
+\begin{equation}
+    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
+\end{equation}
 
@@
 	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
-    \caption{ \textbf{ Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.044) }}
+    \caption{\textbf{Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.044)}}
 	\label{f:arctanfit}

### 2017-01-25 gassolarpaper/gassolarpaper.tex
clean up nomenclature

@@
  \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Professor, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
+  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
+  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
   {\normalsize\itshape
@@
 \begin{abstract}
-    Using geometric programming, a comparison of solar-electric and gas powered, long-endurance aircraft is accomplished.
-    Long-endurance aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, velocity and coverage footprint.
+    Recently, several companies have invested in developing aircraft that can provide long-endurance (days to weeks) persistent aerial coverage. 
+    These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
+    In this paper, a comparison of solar-electric and gas powered, long-endurance aircraft is accomplished.
     Geometric programming, a form of convex optimization, can reliably solve problems with thousands of variables in seconds and is used as an optimization tool to evaluate the design trade offs between architectures and requirements.
-    Using this approach, a gas powered and a solar-electric powered aircraft were considered.  
     The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
@@
 $b$ \> wing span [ft] \\
-$BSFC$ \> brake specific fuel consumption [lb/hr/hp] \\
+BSFC \> brake specific fuel consumption [lb/hr/hp] \\
 $c$ \> wing chord [m] \\
@@
 $\Delta y$ \> wing section length [m] \\
-$DOY$ \> day of the year \\
+DOY \> day of the year \\
 $e$ \> Oswald efficiency factor \\
@@
 $\mathcal{M}$ \> wing bending moment [Nm] \\
-$MTOW$ \> max take-off weight [N] \\
+MTOW \> max take-off weight [N] \\
 $n$ \> number of wing segments \\
@@
 $\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> bending deflection angle 
+$\Theta$ \> wing bending deflection angle 
  \end{tabbing}
@@
 The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To achieve a feasible design, solar-electric powered aircraft must carry enough solar cells and batteries required to fly during the day while storing enough energy to fly through the night.\cite{solartech}
+To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries required to fly during the day while storing enough energy to fly through the night.\cite{solartech}
 If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then it can theoretically fly during any time of the year for long durations. \cite{solartech}
@@
     \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, DOY, p_{\text{wind}}).
+    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
     \end{equation}
@@
 
-    where $\beta = 2\pi (DOY-1)/365$.
+    where $\beta = 2\pi (\text{DOY}-1)/365$.
     The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation Equation~\eqref{e:solartheta}, \cite{solar}
@@
         \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{DOY-93}{365}\right)} \right],
+        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
     \end{align}
@@
     \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}BSFCg} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
+    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
 \end{equation}
@@
 
-Using the definition of $BSFC$
+Using the definition of $\text{BSFC}$
 
@@
     \label{e:brBSFC}
-    BSFC = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
+    \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
 \end{equation}
@@
     \label{e:brdiff2}
-    -dW = g P_{\text{shaft}} BSFC dt.
-\end{equation}
-
-It is assumed that $BSFC$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+    -dW = g P_{\text{shaft}} \text{BSFC} dt.
+\end{equation}
+
+It is assumed that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
 A constant power to weight ratio assumes constant velocity and constant lift coefficient.\cite{br2}
@@
     \label{e:brdiff2}
-    -\frac{dW}{W} = \frac{g P_{\text{shaft}}BSFC }{W} dt,
+    -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
 \end{equation}
@@
     \label{e:be1}
-    \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}BSFC}{W_{\text{ave}}} t,
+    \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
 \end{equation}
@@
     \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t BSFC g}{W}\\
+    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
     \label{e:brtaylor}
@@
     \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i BSFC g}{\sqrt{W_i W_{i+1}}}\\
+    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
     \label{e:brtaylord}
@@
     For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ is used.
-Additionally, it is assumed that the break specific fuel consumption is constant, $BSFC = 0.6$ [lb/hp/hr].\cite{bsfcperf}
+Additionally, it is assumed that the break specific fuel consumption is constant, $\text{BSFC} = 0.6$ [lb/hp/hr].\cite{bsfcperf}
 
@@
     \label{e:weightmtow}
-    MTOW &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} \\
+    \text{MTOW} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} \\
 \end{align}
@@
     \label{e:weightsmtow}
-    MTOW &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} \\
+    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} \\
     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
@@
 %      $f_{\text{structural}}$     & 0.35   \\
-%      $BSFC$ [lb/hp/hr]    & 0.6    \\
+%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
 %      \bottomrule
@@
 %     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}MTOW.
+%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
 % \end{equation}
@@
 % % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $BSFC$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
+% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
 % % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
@@
 Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is break specific fuel consumption, $BSFC$; a lower $BSFC$ will result in increased endurance.  
+The first is break specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
 The second is the lapse rate.  
@@
 
-A GP-compatible curve was generated from the DF70 data to relate $BSFC$ and shaft power,
+A GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
 
@@
     \label{e:rpmtobsfc}
-    \left( \frac{BSFC}{BSFC_{\text{min}}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-\end{equation}
-
-where $BSFC_{\text{min}} = 0.3162$ [lb/hp/hr].\cite{rcvengines}
-Using this approximations the required shaft power determines the $BSFC$.
+    \left( \frac{\text{BSFC}}{\text{BSFC}_{\text{min}}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
+\end{equation}
+
+where $\text{BSFC}_{\text{min}} = 0.3162$ [lb/hp/hr].\cite{rcvengines}
+Using this approximations the required shaft power determines the $\text{BSFC}$.
 A comparison of the GP-compatible approximations to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $BSFC$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of a similar class.
+A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of a similar class.
 

### 2017-01-26 gassolarpaper/gassolarpaper.tex
plot formatting and more edits

@@
 Long-endurance, station-keeping aircraft are potential solutions to providing internet, communication, and persistent survellience. 
-Recently, tech companies such as Google and Facebook have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Defense contractors like AeroVironment and small engineering firms like Vanailla Aircraft have developed gas powered long-endurance platforms as a means of providing continual survellience for days at a time.
+Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
+Engineering firms like Aurora Flight Sciences\cite{orion} and Vanailla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual survellience for days at a time.
 However, sizing a long-endurance aircraft is not obvious because of the multifacted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
-Additionally, because few long-endurance gas powered aircraft have been built and only Airbus's Zephyr solar aircraft has been successfully flown, the operational limits of these aircraft is relatively unknown.
+Additionally, because few long-endurance gas powered aircraft have been built and only Airbus's Zephyr\cite{zephyr} solar aircraft has been successfully flown, the operational limits of these aircraft is relatively unknown.
 Even more unclear is understanding the trade offs between gas powered and solar-electric powered archictures.
@@
 The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries required to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then it can theoretically fly during any time of the year for long durations. \cite{solartech}
-This becomes more difficult to achieve at higher latitudes as the solar flux during the winter solstice decreases.  
+To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
+If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
+\cite{solartech}
+This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
 Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
@@
 Therefore, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitude only depends on wind speed. 
+Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on wind speed. 
 If a gas-powered aircraft can fly at the latitude with the worst wind speed in can theoretically fly anywhere else in the world.
@@
 
-Design requirements to operate at specific altitudes, latitudes, and percentile wind speeds each of which will affect the performance and design of the aircraft.\cite{zephyr} 
+Design requirements to station keep at various altitudes, latitudes, and percentile wind speeds which will affect the performance and design of the aircraft.\cite{zephyr} 
 To station keep an aircraft must fly faster than the wind speed,
@@
 
-Long-endurance gas powered aircraft, the endurance requirement the amount of time the aircraft needs to remain in the air without refueling.  
-If solar-electric powered aircraft can fly during the shortest day of the year, December 21st, then it can theortically fly at any time the year.  
+For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can theortically fly at any time the year.  
 Therefore, the endurance requirement for solar-electric powere aircaft is assumed to be a requirement to fly during the winter solstice.  
 To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section , and with higher wind speeds as shown in Figure~\ref{f:windvsmonth}.  
+Because the endurance requirement for solar-electric powered aircraft is set by the winter solstice and because wind speeds are higher during winter months, only wind speed data from December will be used for the aircraft sizing and performance analysis presented in this paper.
 
@@
 
-Gas powered aircraft generally fly at a given altitude to meet a payload requirement.  
-A naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} mean that gas powered aircraft will generally be more efficiect at lower altitude and generally fly between 5,000 at 25,000 ft depending on the payload requirement.
-For this design study, it is assume d that the aircraft is flying at 15,000 ft.
-
-Solar-electric powered aircraft tend to fly around 60,000 ft to meet payload requirements, avoid cloud coverage and fly in lower wind speeds which have a minimum around 60,000 shown in Figure~\ref{f:altvswind}
+Gas powered aircraft generally fly between 5,000 to 25,000 ft due to payload requirements.  
+Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} gas powered aircraft will generally be more efficiect at lower altitudes and generally fly at a given latitude to meet a payload requirement.
+For this design study, it is assumed that the aircraft is flying at 15,000 ft.
+
+Solar-electric powered aircraft tend to fly around 60,000 ft to meet payload requirements, avoid cloud coverage and fly in lower wind speeds which have a minimum around 60,000 shown in Figure~\ref{f:altvswind}.
+For the solar-electric powered aircraft altitude is optimized as explained in Section.
 
@@
 
-The altitude also determines atmospheric properties such as air density and viscosity.  For the solar aircraft air density is calculated using equations specific to the tropopause,\cite{isaatm}
+The altitude also determines atmospheric properties such as air density and viscosity.
+For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+
+For the solar aircraft air density is calculated using equations specific to the tropopause,\cite{isaatm}
 
@@
 
-For the gas powered aircraft air density is calculated using atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computer prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K].
-
 \subsection{Latitude}
@@
 Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
-Wind data includes winds during December for the year 2005-2015.}}
+Wind data includes winds during December for the years 2005-2015.}}
 	\label{f:latvswind}
@@
 Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section.
+This is further discussed in Section .
 
@@
 For example, an aircraft optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude.
-This can be imposed in the optimization model as multiple constraints. For the 35 degree latitude case, 70 separate wind and solar flux constraints that must be satisfied, one for each latitude.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015. One lag is equivalent to 6 hours. }}
+ \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
  \label{f:bostonwinds}
@@
  \end{subfigmatrix}
- \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015. One lag is equivalent to 6 hours. }}
+ \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
  \label{f:romewinds}
@@
 An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to relocate or land.
+However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
 
@@
 
-where the functions $f_i$ must be \emph{monomial} functions and the functions $g_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
+where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
 
@@
 Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}  
-To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} \\
+To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} 
+A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.
+A commerical solver, mosek\cite{mosek}, is used to solve the GP.
+\\
 
@@
 To evaluate the size and performance of both gas and solar-electric powered aircraft underlying physics models are used.  
-Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar insolation and aircraft models governing the performance, aerodynamics and weight breakdown are used in the optimization model.
-Equations from these models are then expressed in a GP-compatible form to enable their use in the optimization. 
+Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar irradiance and aircraft models governing the performance, aerodynamics and structural weight are used in the optimization model.
+Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
 Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
@@
 The local minimum in wind speed around 63,000 ft and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed an altitude are imposed to optimize altitude. 
+To accomplish this, constraints relating wind speed to altitude are imposed. 
 This approach agrees from an operational perspective as a solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
@@
 Because wind speeds are low and solar flux is higher below 20 degrees latitude no constraints for latitudes below 20 degrees are actually included in the optimization. 
-Additionally, only wind speed data from the Northern Hemisphere are considered.
+Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
+Thus, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude. 
 Therefore, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
-The constraint of wind speed and air density at the 35 degree latitude is
+The constraint on wind speed and air density at the 35 degree latitude is
 
@@
 For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated will be less efficient at higher altitudes.  
+Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
 Additionally, wind speeds increase nearly linearly up to 30,000 ft.  
-Therefore, the altitude for gas-powered aircraft is determined as a requirement of the payload and the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h=15,000$ ft.
+Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
+For this sizing study, the altitude requirement for the gas powered aircraft is $h=15,000$ ft, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
 
@@
 Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_C$ are functions of $(P/S)_{\text{min}}$.
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions $(P/S)_{\text{min}}$, one for each latitude between 20 and 60 degrees latitude.
-Figure~\ref{f:energyapprox} sample monomial approximations at 30 degrees latitude.
+The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude.
 
@@
     where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation Equation~\eqref{e:solartheta}, \cite{solar}
+    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
 
@@
 
-    where noon is $t=0$. Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-
+    where noon is $t=0$.
+Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
 The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
@@
 
-The derivation begins with the differential form of Breguet Range,\cite{br2}
-
-\begin{equation}
-    \label{e:breguetdiff}
-    -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-\end{equation}
-
-Using the definition of $\text{BSFC}$
-
-\begin{equation}
-    \label{e:brBSFC}
-    \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-\end{equation}
-
-Equation~\eqref{e:breguetdiff} can be written as
-
-\begin{equation}
-    \label{e:brdiff2}
-    -dW = g P_{\text{shaft}} \text{BSFC} dt.
-\end{equation}
-
-It is assumed that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-A constant power to weight ratio assumes constant velocity and constant lift coefficient.\cite{br2}
-Dividing by $W$,
-
-\begin{equation}
-    \label{e:brdiff2}
-    -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-\end{equation}
-
-and integrating, the Breguet Range equation can be expressed as
-
-\begin{equation}
-    \label{e:be1}
-    \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-\end{equation}
-
-where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.  $W_{\text{ave}}$ is assumed to be the geometric mean, defined as
+% The derivation begins with the differential form of Breguet Range,\cite{br2}
+%
+% \begin{equation}
+%     \label{e:breguetdiff}
+%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
+% \end{equation}
+% 
+% Using the definition of $\text{BSFC}$
+% 
+% \begin{equation}
+%     \label{e:brBSFC}
+%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
+% \end{equation}
+% 
+% Equation~\eqref{e:breguetdiff} can be written as
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
+% \end{equation}
+
+This verison comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
+$W_{\text{ave}}$ is assumed to be the geometric mean, defined as
+
+% Dividing by $W$,
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
+% \end{equation}
+% 
+% and integrating, the Breguet Range equation can be expressed as
+% 
+% \begin{equation}
+%     \label{e:be1}
+%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
+% \end{equation}
+% 
+% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
 
@@
 
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynmial respectively and therefore GP compatible. \\
+    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible. \\
 
     For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.  
-Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized for more accurate results.
+    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this.
 
@@
     For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ is used.
-Additionally, it is assumed that the break specific fuel consumption is constant, $\text{BSFC} = 0.6$ [lb/hp/hr].\cite{bsfcperf}
 
@@
 
-Both the gas and solar powered aircraft are assumed to be in steady level flight.  The balance of the main forces of lift, drag, weight and thrust are modeled as\cite{hoburgthesis}
+Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
 
@@
     \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S . 
-\end{align}
-
-The shaft power produced by the engine or motor and can expressed as
+    W &= \frac{1}{2} \rho V^2 C_L S. 
+\end{align}
+
+The shaft power produced by the engine or motor can be expressed as
 
@@
 where the Oswald efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a constant.  
-For the gas powered aircraft, $C_{d_0} = 0.01$.  
-For the solar-electric powered aircraft, $C_{d_0} = 0.002$.  
-It is assumed that the solar powered aircraft will store its batteries in the wings. 
-The gas powered aircraft will use a fuselage to store its fuel, thus having a larger non-lifting surface and a higher drag coefficient.\cite{raymer}
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the \emph{JH01} airfoil. (See Appendix B) 
+The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
+
+    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
     Drag polars were produced using XFOIL at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg.\cite{fitting}
@@
 	\includegraphics[width=0.6\textwidth]{JH01polar.pdf}
-    \caption{ \textbf{ The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as a solid line.
+    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as a solid line.
 The open circles are the calculated drag polars from XFOIL.
-(Log space RMS error = 0.00489.)} }
+(Log space RMS error = 0.00489.)}}
 	\label{f:JH01polar}
@@
 
-\begin{align}
+\begin{equation}
     \label{e:weightmtow}
-    \text{MTOW} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} \\
-\end{align}
+    \text{MTOW} \geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}}
+\end{equation}
 
@@
     \label{e:weightsmtow}
-    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} \\
+    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} + W_{\text{motor}} \\
     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
@@
 
-where the solar cell density and the battery energy density are assumed to be $\rho_{\text{solar}} = 0.27$ [kg/m$^2$] and $h_{\text{batt}} = 350$ [Whr/kg].\cite{solartech}\cite{solarparam}
+where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
+Lithim-ion batteries are selected with a battery energy density $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumion}
 
@@
 
-A wing spar is modeled as one of the main structural elements of both aircraft. 
+A wing spar is one of the main structural elements of both aircraft. 
 A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
@@
 
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$.
-The local chord for a constant tapered wing\cite{bending} is defined as
-
-\begin{align}
+where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$. The local chord for a constant tapered wing\cite{bending} is
+
+\begin{equation}
     \label{e:localchord}
-    c(y) &= \frac{S}{b} \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right) \\
-    \bar{c}(y) \equiv \frac{c(y)}{S/b} &= \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-\end{align}
+    \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
     \label{e:qbar}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right) \\
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} &= \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-\end{align}
+    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
+\end{equation}
 
@@
 	\includegraphics[width=0.8\textwidth]{capspar2.jpg}
-    \caption{ \textbf{ Cross sectional view of a cap spar.  The two caps are made of carbon fiber, separated by a foam core. A thin shear web is wrapped around the caps and foam to prevent shearing.}}
+    \caption{\textbf{Cross sectional view of a cap spar.
+The two caps are made of carbon fiber, separated by a foam core.
+A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.}}
 	\label{f:capspar}
@@
 When a GP is solved, the sensitivity of the objective function with respect to each constraint is also returned.  
-From this information, the sensitivity to each inputted variable with respect to the objective function can be extracted.\cite{hoburgthesis}
-For example, if the objective function were weight and the sensitivity to solar cell efficiency were -0.5, then a 1\% increase in the solar cell efficiecy would result in a -0.5\% decrease in weight.
+From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
+For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiecy would result in a 0.5\% decrease in weight.  
 From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 

### 2017-01-26 gassolar/environment/wind_fit.py

@@ -65,9 +65,12 @@ def plot_fits(xdata, ydata, yfit, latitude, rm=None):
     fig, ax = plt.subplots()
     for p, y, yf, cl in zip(x2, ydata.reshape(len(x2), len(x1)),
                             yfit.reshape(len(x2), len(x1)), colors):
-        ax.plot(np.exp(x1), np.exp(y)*WIND_NORM, "o", markeredgecolor=cl, markerfacecolor="none")
-        ax.plot(np.exp(x1), np.exp(yf)*WIND_NORM, c=cl,
-                label="%d Percentile Winds" % np.rint(np.exp(p)*PERCT_NORM))
+        pp = np.exp(p)
+        if pp == 0.75 or pp == 0.85 or pp == 0.95:
+            ax.plot(np.exp(x1), np.exp(y)*WIND_NORM, "o", mec=cl, mfc="none",
+                    mew=1.5)
+            ax.plot(np.exp(x1), np.exp(yf)*WIND_NORM, c=cl, lw=2,
+                    label="%d Percentile Winds" % np.rint(np.exp(p)*PERCT_NORM))
     ax.legend(loc=2, fontsize=15)
     ax.set_xlabel("Air Density [kg/m$^3$]")
     ax.set_ylabel("Wind Speed [m/s]")

### 2017-01-26 gassolar/environment/wind_speeds.py

@@ -89,7 +89,7 @@ if __name__ == "__main__":
         Wind90 = [get_windspeed(l, 90, al, 355) for l in np.arange(70)]
         Wind95 = [get_windspeed(l, 95, al, 355) for l in np.arange(70)]
         Ax.fill_betweenx(np.arange(70), Wind85, Wind95, alpha=0.5, facecolor=c)
-        Ax.plot(Wind90, np.arange(70), c, label="Altitude=%d" % al)
+        Ax.plot(Wind90, np.arange(70), c, label="Altitude=%d" % al, lw=2)
     Ax.set_ylabel("Latitude [deg]")
     Ax.set_xlabel("Wind speed [m/s]")
     Ax.set_ylim([0, 70])
@@ -105,7 +105,7 @@ if __name__ == "__main__":
         Wind90 = get_windspeed(l, 90, Alt, 355)
         Wind95 = get_windspeed(l, 95, Alt, 355)
         Ax.fill_betweenx(Alt, Wind85, Wind95, alpha=0.5, facecolor=c)
-        Ax.plot(Wind90, Alt, c, label="Latitude=%d" % l)
+        Ax.plot(Wind90, Alt, c, label="Latitude=%d" % l, lw=2)
     Ax.set_ylabel("Altitude [ft]")
     Ax.set_xlabel("Wind speed [m/s]")
     Ax.set_ylim([0, 80000])
@@ -126,7 +126,7 @@ if __name__ == "__main__":
         Wind95 = [get_windspeed(45, 95, al, d) for d in Mid]
         Ax.fill_between(range(13), Wind85 + [Wind85[0]], Wind95 + [Wind95[0]],
                         alpha=0.5, facecolor=c)
-        Ax.plot(range(13), Wind90 + [Wind90[0]], c, label="Altitude=%d" % al)
+        Ax.plot(range(13), Wind90 + [Wind90[0]], c, label="Altitude=%d" % al, lw=2)
     Ax.set_xticks(np.arange(12))
     Ax.set_xticks(np.arange(12)+0.5, minor=True)
     Ax.set_xticklabels(Mos, minor=True)

### 2017-01-26 gassolarpaper/gassolarpaper.tex
more edits; plot formatting; variable tweaking

@@
  \usepackage[labelfont=bf]{caption}
+ \usepackage{multirow}
 
@@
   Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautical and Astronautical Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
+  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
   {\normalsize\itshape
@@
 $b$ \> wing span [ft] \\
-BSFC \> brake specific fuel consumption [lb/hr/hp] \\
+BSFC \> brake specific fuel consumption [kg/kW/hr] \\
 $c$ \> wing chord [m] \\
-$C_D$ \> drag coefficient \\
+$C_D$ \> aircraft drag coefficient \\
 $C_{d_0}$ \> non-wing drag coefficient \\
@@
 $E$ \> Young's Modulus [Pa] \\
-$E_{\text{battery}}$ \> energy stored in battery [J] \\
+$E_{\text{batt}}$ \> energy stored in battery [J] \\
 $E_{\text{sun}}$ \> total solar energy available [Whr] \\
+$(E/S)_{\text{C}}$ \> low irradiance solar energy per unit area [Whr/m$^2$] \\
+$(E/S)_{\text{day}}$ \> daily operational solar energy per unit area [Whr/m$^2$] \\
 $(E/S)_{\text{sun}}$ \> available solar energy per unit area [Whr/m$^2$] \\
@@
 $m'$ \> local wing mass [kg/m] \\
-$m_{\text{h}}$ \> modified horizontal tail lift coefficient \\
+$m_{\text{fac-wing}}$ \> margin factor\\
+$m_{\text{fac-wing}}$ \> wing weight margin factor \\
 $m_{\text{motor}}$ \> electric motor mass \\
-$m_{\text{w}}$ \> modified wing lift coefficient \\
 $\dot{m}_{\text{fuel}}$ \> fuel flow rate [kg/s] \\
@@
 $N$ \> number of flight segments \\
-$N_{\text{max}}$ \> load factor\\
+$N_{\text{max}}$ \> safety load factor\\
 $p$ \> air pressure [Pa]\\
@@
 $P_{\text{oper}}$ \> aircraft operating power [hp] \\
+$(P/S)_{\text{min}}$ \> minium operational solar power \\
+\>per unit area [Whr/m$^2$] \\
 $P_{\text{shaft}}$ \> engine/motor shaft power [hp] \\
@@
 $W_{\text{ave}}$ \> average weight over a flight segment [N] \\
-$W_{\text{battery}}$ \> battery weight [N] \\
+$W_{\text{batt}}$ \> battery weight [N] \\
 $w_{\text{cap}}$ \> spar cap width [m] \\
@@
 $\eta_{\text{discharge}}$ \> battery discharging efficiency \\
+$\eta_{\text{motor}}$ \> electric motor efficiency \\
 $\eta_{\text{prop}}$ \> propulsive efficiency \\
@@
 $\rho$ \> air density [kg/m$^3$] \\
-$\rho_{\text{carbon fiber}}$ \> density of carbon fiber [kg/m$^3$] \\
+$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber [kg/cm$^2$] \\
+$\rho_{\text{cfrp}}$ \> density of carbon fiber [kg/cm$^3$] \\
 $\rho_{\text{foam}}$ \> density of foam [lbf/ft$^3$] \\
 $\rho_{\text{solar}}$ \> solar cell density [kg/m$^2$] \\
-$\sigma_{\text{carbon fiber}}$ \> maximum carbon fiber stress [Pa] \\
+$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress [Pa] \\
 $\tau_{\text{h}}$ \> horiztonal tail thickness to chord ratio \\
@@
 
-Long-endurance, station-keeping aircraft are potential solutions to providing internet, communication, and persistent survellience. 
+Long-endurance, station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
 Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanailla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual survellience for days at a time.
-However, sizing a long-endurance aircraft is not obvious because of the multifacted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
-Additionally, because few long-endurance gas powered aircraft have been built and only Airbus's Zephyr\cite{zephyr} solar aircraft has been successfully flown, the operational limits of these aircraft is relatively unknown.
-Even more unclear is understanding the trade offs between gas powered and solar-electric powered archictures.
-
-This paper presents a physics-based optimization model that uses geometric programming as systematic and rapid approach to explore this design space and highlight trade offs between the two proposed architecutres.
+Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
+However, sizing a long-endurance aircraft is not obvious because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
+Additionally, because few long-endurance gas powered aircraft have been built and only Airbus has successfully built and tested a solar aircraft for multiple days\cite{zephyr}, the operational limits of these aircraft is relatively unknown.
+Even more unclear are the trade offs between gas powered and solar-electric powered architectures.
+
+This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to explore this design space and highlight trade offs between the two proposed architectures.
 Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
@@
 As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Therefore, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed in can theoretically fly anywhere else in the world.
+Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
+If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
 
@@
 For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can theortically fly at any time the year.  
-Therefore, the endurance requirement for solar-electric powere aircaft is assumed to be a requirement to fly during the winter solstice.  
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section , and with higher wind speeds as shown in Figure~\ref{f:windvsmonth}.  
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can thoretically fly at any time the year.  
+Therefore, the endurance requirement for solar-electric powered aircraft is assumed to be a requirement to fly during the winter solstice.  
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher wind speeds as shown in Figure~\ref{f:windvsmonth}.  
 Because the endurance requirement for solar-electric powered aircraft is set by the winter solstice and because wind speeds are higher during winter months, only wind speed data from December will be used for the aircraft sizing and performance analysis presented in this paper.
@@
 Gas powered aircraft generally fly between 5,000 to 25,000 ft due to payload requirements.  
-Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} gas powered aircraft will generally be more efficiect at lower altitudes and generally fly at a given latitude to meet a payload requirement.
-For this design study, it is assumed that the aircraft is flying at 15,000 ft.
+Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given latitude to meet a payload requirement.
 
 Solar-electric powered aircraft tend to fly around 60,000 ft to meet payload requirements, avoid cloud coverage and fly in lower wind speeds which have a minimum around 60,000 shown in Figure~\ref{f:altvswind}.
-For the solar-electric powered aircraft altitude is optimized as explained in Section.
+For the solar-electric powered aircraft altitude is optimized as explained in Section IV.A.
 
@@
 Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section .
+This is further discussed in Section IV.B.
 
@@
 
-For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data in December across all longitudes and spanning the years 2005-2015.  
-To understand what a certain percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
+For the purposes of this design study, a percentile wind speed corresponds to that percentile of the wind speed data in December, across all longitudes and spanning the years 2005-2015.  
+To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
+Wind speeds 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
 
@@
 A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.
-A commerical solver, mosek\cite{mosek}, is used to solve the GP.
+A commercial solver, mosek\cite{mosek}, is used to solve the GP.
 \\
@@
 Altitude can be post-calculated from the optimized air density. 
-Because wind speeds are low and solar flux is higher below 20 degrees latitude no constraints for latitudes below 20 degrees are actually included in the optimization. 
+Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
 Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
@@
 
-At a given time of year DOY, and latitude $\phi$, there exists an availble solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day
+At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day
 
@@
 
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$ and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
+To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
 
@@
     \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{battery}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
+    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
     \label{e:solarbatt}
-    E_{\text{battery}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_C \eta_{\text{solar}} S_{\text{solar}}
+    E_{\text{batt}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_C \eta_{\text{solar}} S_{\text{solar}}
 \end{align}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{lat30.pdf}
+	\includegraphics[width=0.7\textwidth]{lat30.pdf}
     \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.
-For a solar-electirc powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
+For a solar-electric powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
 	\label{f:lat30}
@@
 For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$ a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
+The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
 
@@
     Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night is calculated from the latitude and the day of the year prior to an optimization solve.
+Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
 
@@
 
-This verison comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
 One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
@@
 
-    To make Equation~\eqref{e:be1} GP compatible, a Taylor expansion of the exponentiated right hand side is used,\cite{hoburgthesis}
+    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion of the exponentiated right hand side is used,\cite{hoburgthesis}
 
@@
 
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible. \\
-
+    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
 For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
@@
 The open circles are the calculated drag polars from XFOIL.
-(Log space RMS error = 0.00489.)}}
+(Log space RMS error $= 0.00489$.)}}
 	\label{f:JH01polar}
@@
 \end{figure}
-
-As a preliminary sizing study the aspect ratio is assumed to be a constant, $A = 30$.
-A more detailed sizing study with a wing structural model, discussed later, will allow the aspect ratio to be a free variable.
-
-\subsection{Weight Breakdown}
-
-The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
-The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizer, and engine. 
-The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.
-
-\begin{equation}
-    \label{e:weightmtow}
-    \text{MTOW} \geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}}
-\end{equation}
-
-A similiar weight breakdown is done for the solar-electric powered aircraft,
-
-\begin{align}
-    \label{e:weightsmtow}
-    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{battery}} + W_{\text{motor}} \\
-    W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-    W_{\text{battery}} &\geq \frac{E_{\text{battery}}}{h_{\text{battery}}} g
-\end{align}
-
-where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-Lithim-ion batteries are selected with a battery energy density $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumion}
 
@@
 % % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{battery}}$ [Whr/kg]  & 350 \\
+% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
 % %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
@@
 A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight bending case and gust load case.
+It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight case and gust loads.
 
@@
 Because long-endurance aircraft typically have high aspect ratio wing they are considered flexible aircraft. 
-To account for this flexiblity and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\ref{e:qbar}
+To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\ref{e:qbar}
 
@@
 where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
+The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
 The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the wing
@@
 	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
-    \caption{\textbf{Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.044)}}
+    \caption{\textbf{Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.0246)}}
 	\label{f:arctanfit}
@@
 where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth]{gustvschord.pdf}
-%     \caption{ \textbf{ Graphical representation of assumption that the local chord is equal to the local lift generated from the elliptical distribution.
-Assumes the center to wing weight ratio, $\frac{W_{\text{cent}}}{W_{wing}} = 0.14$.
-}}
-% 	\label{f:gustvschord}
-% 	\end{center}
-% \end{figure}
 
@@
     \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2
+    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
 \end{equation}
@@
     \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2
+    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
 \end{equation}
@@
 The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed by no greater than 30\% of chord, $\tau_w = 0.3$.
+The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
 
@@
 
-The wing spar at the root must be strong enough to withstand the bending root moment and stiff enough to not exceed some deflection limit.  Both constraints are imposed in the optimization model as
+The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
+Both constraints are imposed in the optimization model as
 
@@
     \label{e:stresscont}
-    \sigma_{\text{carbon fiber}} &\geq \mathcal{M}_0 \frac{h_{\text{cap}}+t_{\text{cap}}}{I}\\
+    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
     \label{e:defcont}
@@
 
-The maximum stress for carbon fiber is, $\sigma_{\text{carbon fiber}} = 475$ [MPa].\cite{carbonfiber}
+The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 570$ [MPa].\cite{carbonfiber}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
     \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{carbon fiber}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
+    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
     \label{e:sparmasssum}
@@
 
-where $\rho_{\text{carbon fiber}} = 1.4$ [g/cm$^3$].\cite{carbonfiber}
+where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{carbonfiber}
 
@@
     \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{\text{carbon fiber}} S g 
-\end{equation}
-
-where $\rho_{\text{carbon fiber}} = 0.043$ [g/cm$^2$].\cite{carbonfiber}
+    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g 
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or the area density of one ply of carbon fiber.\cite{carbonfiber}
 
@@
 
-The tail boom has a diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E$, and density $\rho_{\text{carbon fiber}} = 1.4$ [g/cm$^3$], and length $l_{\text{h}}$. 
+The tail boom has a diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E$, and density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$], and length $l_{\text{h}}$. 
 The total mass and root bending inertia are imposed in the optimization model as
@@
 \begin{align}
-    m &\geq \pi \rho_{\text{carbon fiber}} t_0 d l_{\text{h}} \left( 1 + \frac{1}{2} k\right) \\
+    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 + \frac{1}{2} k\right) \\
     I_0 &\leq \pi t_0 d^3/8
@@
 
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis.
+where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.  For both the solar-electric and gas powered aircraft $k=0.8$.  
+When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
 
@@
     \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition, $V_{\text{h}} = 0.45$.\cite{aircraftrules}
+    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
+\end{align}
+
+The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition, $V_{\text{h}} = 0.45$,\cite{aircraftrules}
 
@@
 
-The vertical tail is sized to meet a conservative volume tail coefficient, $V_{\text{v}}= 0.02$,\cite{aircraftrules}
+The vertical tail is sized to meet a conservative tail volume coefficient, $V_{\text{v}}= 0.04$,\cite{aircraftrules}
 
@@
 
-where $l_{\text{v}}$ is the vertical tail moment arm and assume to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{\text{carbon fiber}} = 0.1$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
+where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
+
+Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
 The weight of the horizontal and vertical tails is
@@
     \label{e:htweight}
-    W_{\text{h}} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{\text{carbon fiber}} S_{\text{h}} \\
+    W_{\text{h}}/m_{fac} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{h}} \\
     \label{e:vtweight}
-    W_{\text{v}} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{\text{carbon fiber}} S_{\text{v}}
+    W_{\text{v}}/m_{fac} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{v}}
 \end{align}
@@
 where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-
-The drag of the empennage was considered as three separate parts with no interference drag.  The drag of the tail boom was calculated using a turbulent flat plate model,
+The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
+
+The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
 
@@
 
-The drag of the horizontal and vertical tails was computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
+The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
 
@@
 
-where $\tau_{\text{(v,h)}} = 0.08$ is the airfoil thickness.
-Assuming steady level flight where neither surface is generating lift, the XFOIL data was generated for a zero angle of attack.
+where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
+The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.  
 A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
@@
 	\includegraphics[width=0.6\textwidth]{taildragpolar.pdf}
-    \caption{ \textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
+    \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
 	\label{f:taildragpolar}
@@
 The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a smal fuselage whose effects will be ignored.
+The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
 
@@
 
-where the fuel is assumed to be Jet A fuel with density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The volume of the fuselage is constrained by
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The volume of the fuselage is constrained by
 
@@
 where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius the surface area can be calculated using Thomsen's approxmiation,\cite{ellipsoidSA}
+Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
 
@@
     \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + ((2R_{\text{fuse}})^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is based off of the surface area
+    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
+\end{equation}
+
+The weight of the fuselage is constrained by
 
@@
     \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{\text{carbon fiber}} g
-\end{equation}
-
-where $\rho_{\text{carbon fiber}} = 0.085$ [g/cm$^2$].\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
+    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
 
@@
 
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ration is $B_{PM} = 4140.8$ [W/kg]. \\
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
 
@@
 	\includegraphics[width=0.6\textwidth]{powervsweightfit.pdf}
-    \caption{ \textbf{ Comparison of power law equation and engine data.  Engine data taken from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
+    \caption{\textbf{Comparison of power law equation and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
 	\label{f:powervsweightfit}
@@
 Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is break specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
+The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
 The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aispirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
+Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
 To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
@@
 It is assumed that engines of a similar class will perform similarly to the DF70 engine.  
-Thus, this engine performance model is only valid for IC engines.
+This engine performance model is only valid for IC engines.
 
@@
 
-The lapse rate $L_{\text{eng}} \leq 1$, is assumed to affect the maximum power output,
+The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
 
@@
     \label{e:lapse}
-    L_{\text{eng}} = \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
+    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
 \end{equation}
@@
     \label{e:lapsefit}
-    L_{\text{eng}} = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
+    L_{\text{eng}}(h) = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
 \end{equation}
@@
 Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must be big enough to provide the necessary power to meet the minimum climb rate,
+The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
 
@@
 
-where $W$, is the weight of the aircraft during climb.
+where $W$ is the weight of the aircraft during climb.
+
+\subsection{Weight Breakdown}
+
+The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
+The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizer, and engine. 
+The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.
+
+\begin{align}
+    \label{e:weightmtow}
+    \text{MTOW} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
+    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
+\end{align}
+
+A similar weight breakdown is done for the solar-electric powered aircraft,
+
+\begin{align}
+    \label{e:weightsmtow}
+    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
+    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
+    W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
+    W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
+\end{align}
+
+where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
+Lithim-ion batteries are selected with a battery energy density $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumion}
 
@@
 
-To achieve sizing results for the solar-electric and gas powered long-endurance aircraft the optimization models were solve by minimizing the wing span across different bands of latitudes, shown in Figure~\ref{f:latvsspantrade}.
+To achieve sizing results for the solar-electric and gas powered long-endurance aircraft the optimization models were solve by minimizing the wing span across different bands of latitudes, shown in Figure~\ref{f:latvsmtowtrade}.
 
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Gas Powered\label{f:spanvslatgas}]{\includegraphics{spanvslatgas.pdf}}
-     \subfigure[Solar-Electric Powered\label{f:spanvslatsolar}]{\includegraphics{spanvslatsolar.pdf}}
+     \subfigure[Gas Powered\label{f:mtowvslatgas}]{\includegraphics{mtowvslatgas.pdf}}
+     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolar}]{\includegraphics{mtowvslatsolar.pdf}}
  \end{subfigmatrix}
- \caption{\textbf{Trade study of wingspan and latitude.
-Plots show required maximum take off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.}}
- \label{f:latvsspantradet}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsspantrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
+ \caption{\textbf{Trade study of max take off weight and operational latitude bands.
+A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
+The endurance requirement for the gas powered aircraft in this trade study is 7 days.}}
+ \label{f:latvsmtowtrade}
+\end{figure}
+
+One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$30 degrees North latitude in 90th percentile wind speeds.  
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
+Figure~\ref{f:mtowvslatgas} levels off around 38 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
+Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
+On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$35 degrees.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{spanvsendurance.pdf}
+	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
     \caption{\textbf{Trade study of endurance and size for gas powered architecture.}}
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Wingspan\label{f:span}]{\includegraphics{span.pdf}}
-     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics{solarcell.pdf}}
+     \subfigure[Wingspan\label{f:span}]{\includegraphics{solarobjcomp.pdf}}
+     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics{solarobjcomp2.pdf}}
  \end{subfigmatrix}
@@
 By altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 67,000 ft at 31 degrees latitude, that the aircraft will always fly at 67,000 ft.  
+It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
 If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
@@
 
-where $f_{\text{structural}} = 0.35$ then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is modeled as explained in Section larger wings have a weight penalty and the optimization trades air density for wing weight.
+where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
+However, if the structural weight is modeled as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
 Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
@@
 
-Another interesting result is the operating lift to drag ratio for the gas powere aircraft.  
+Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
 The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
@@
 Solid lines are drag polars at various Reynolds numbers.
-Without the wind speed constraint velocity is optimized to reach the maximum $C_L^{1.5}/C_D$ value.
-With the wind speed constraint, velocity and Reynolds number is constant operating off the maximum endurance lift to drag ratio.}}
+For this result, the max take off weight was minimized for an endurance requirement of 7 days.}}
  \label{f:polarmission}
@@
 By solving the model for different assumed solar cell efficiency and battery energy density values a broader picture of the design space is achieved.   
-Figure~\ref{f:solarcontours} shows a maxtrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
+Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft.
+ \caption{\textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft where the wing span is minimized.
 Each point on the graph corresponds to a unique design.}}
@@
 From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiecy would result in a 0.5\% decrease in weight.  
+For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% decrease in weight.
 From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 
-Table~\ref{t:sens} shows the varialbes with the highest sensitivities the solar-electric powered aircraft optimized at various latitudes and percentile wind speeds by minimizing weight.
-
-It is intersting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This ocurrs because the discharge efficiency is directly related to the operating power which goes by the cube of velocity. ($P_{\text{oper}}\sim V^3$)
-Thus, the discharge efficiency sensitivity shares similiar growth to the percentil wind speed, while the charge efficiency sensitivity does not.
+Table~\ref{t:sens} shows the variables with the highest sensitivities for the solar-electric powered aircraft optimized at various latitudes and percentile wind speeds by minimizing weight.
+
+It is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
+This occurs because the discharge efficiency is directly related to the operating power which goes by the cube of velocity. ($P_{\text{oper}}\sim V^3$)
+Thus, the discharge efficiency sensitivity shares similar growth to the percentile wind speed, while the charge efficiency sensitivity does not.
 
 It is also interesting to note that the design is significantly more sensitivity to the battery energy density than to the solar cell efficiency.  
-Thus to design a smaller, lighter aircraft, the sensitivity analysis reveals that it is more benefical to invest in higher quality batteries than solar cells.
+Thus to design a smaller, lighter aircraft, the sensitivity analysis reveals that it is more beneficial to invest in higher quality batteries than solar cells.
 
 The margin factor for the wing weight $m_{\text{fac-wing}}$, is a measure of the inaccuracy of the wind structural model.  
-If the wing weight is actually 5\% heavier than predicted then the aircraft will be approximately 4\% to 5\% heavier according Table~\ref{e:sens}.
+If the wing weight is actually 5\% heavier than predicted then the aircraft will be approximately 4\% to 5\% heavier according Table~\ref{t:sens}.
 
@@
 Using geometric programming as a form of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the important of wind speed.  
-If station keeping at all times during the flight is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of then time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
+One high level discovery is the importance of wind speed.  
+If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
+However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
 An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery energy density and solar cell efficiency.
@@
 The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between two options.
-While the results of these findings are dependent on the accuracy of the models used in the optimization, it is shown that higher fidelity models can be incorporated to increase accuracy.
 
@@
 Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
+Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
 
@@
 Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=3e5$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela who helped in the redesign.
+The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{ Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.  }}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.}}
  \label{f:jhcps}

### 2017-01-26 gassolar/gas/engine_law.py

@@ -19,10 +19,10 @@ weight = np.linspace(min(df["lbs"]), max(df["lbs"]), 100)
 yfit = cn.evaluate(np.log(weight/WEIGHT))
 
 fig, ax = plt.subplots()
-ax.plot(u*WEIGHT, w*POWER, "o", markerfacecolor="None")
-ax.plot(weight, np.exp(yfit)*POWER)
+ax.plot(u*WEIGHT, w*POWER, "o", mfc="None", mew=1.5)
+ax.plot(weight, np.exp(yfit)*POWER, lw=2)
 ax.set_xlabel("Engine Weight [lbs]")
 ax.set_ylabel("Maximum Shaft Power [hp]")
-ax.legend(["Sample Data", "Power Fit"], loc=2)
+ax.legend(["UND Engine Data", "Power Fit"], loc=2)
 ax.grid()
 fig.savefig("../../gassolarpaper/powervsweightfit.pdf", bbox_inches="tight")

### 2017-01-26 gassolar/gas/gas.py

@@ -31,6 +31,7 @@ class Aircraft(Model):
         Wwing = Variable("W_{wing}", "lbf", "wing weight for loading")
 
         self.empennage.substitutions["V_h"] = 0.55
+        self.empennage.substitutions["V_v"] = 0.04
         self.empennage.substitutions["m_h"] = 0.4
         constraints = [
             Wzfw >= sum(summing_vars(components, "W")) + Wpay + Wavn,
@@ -175,6 +176,12 @@ class Mission(Model):
             constraints.extend([
                 mission[i]["W_{end}"][-1] == fs["W_{start}"][0]
                 ])
+
+        for vk in loading.varkeys["N_{max}"]:
+            if "ChordSparL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 5})
+            if "GustL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 2})
 
         return JHO, mission, loading, constraints
 

### 2017-01-26 gassolar/solar/arctan_fit.py

@@ -15,8 +15,8 @@ print "RMS error: %.4f" % rm
 
 yfit = evaluate_fit(cn, x, "MA")
 fig, ax = plt.subplots()
-ax.plot(u, w)
-ax.plot(u, np.exp(yfit))
+ax.plot(u, w, lw=2)
+ax.plot(u, np.exp(yfit), "--", lw=2)
 ax.set_xlim([0, 0.7])
 ax.grid()
 ax.set_xlabel("$V_{\\mathrm{gust}}/V$")

### 2017-01-26 gassolar/solar/plotting.py

@@ -25,7 +25,7 @@ def windalt_plot(latitude, sol1, sol2):
 
     vwind = (np.exp(softmax_affine(x, params)[0])*100).reshape(6, 20)[3]
     fig, ax = plt.subplots()
-    ax.plot(alt/1000.0, vwind*1.95384, linewidth=2)
+    l = ax.plot(alt/1000.0, vwind*1.95384, linewidth=2)
     for sol in [sol1, sol2]:
         altsol = altitude(min([sol(sv).magnitude for sv in sol("\\rho")]))
         vsol = max([sol(sv).to("knots").magnitude for sv in sol("V")])

### 2017-01-26 gassolar/solar/solar.py

@@ -34,7 +34,7 @@ class Aircraft(Model):
         Wcent = Variable("W_{cent}", "lbf", "center weight")
 
         self.empennage.substitutions["V_h"] = 0.45
-        self.empennage.substitutions["V_v"] = 0.02
+        self.empennage.substitutions["V_v"] = 0.04
         self.empennage.substitutions["m_h"] = 5.514
 
         constraints = [
@@ -286,7 +286,10 @@ class Mission(Model):
         loading = self.solar.loading(self.solar["W_{cent}"], self.solar["W_{wing}"], mission[-1]["V"], mission[-1]["C_L"])
         # loading = self.solar.loading(self.solar["W_{cent}"], mission[-1]["\\rho"], mission[-1]["V"], self.solar.wing["S"])
         for vk in loading.varkeys["N_{max}"]:
-            loading.substitutions.update({vk: 1.5})
+            if "ChordSparL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 5})
+            if "GustL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 2})
 
         return self.solar, mission, loading
 
@@ -296,7 +299,7 @@ def test():
     M.solve()
 
 if __name__ == "__main__":
-    M = Mission(latitude=31)
+    M = Mission(latitude=25)
     M.cost = M["W_{total}"]
     sol = M.solve("mosek")
     mn = [max(M[sv].descr["modelnums"]) for sv in sol("(E/S)_{irr}") if

### 2017-01-27 gassolarpaper/gassolarpaper.tex
small edits

@@
 Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-However, sizing a long-endurance aircraft is not obvious because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
+However, sizing a long-endurance aircraft is not immediately obvious because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
 Additionally, because few long-endurance gas powered aircraft have been built and only Airbus has successfully built and tested a solar aircraft for multiple days\cite{zephyr}, the operational limits of these aircraft is relatively unknown.
@@
 For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can thoretically fly at any time the year.  
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can thoretically fly at any time of the year.  
 Therefore, the endurance requirement for solar-electric powered aircraft is assumed to be a requirement to fly during the winter solstice.  
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher wind speeds as shown in Figure~\ref{f:windvsmonth}.  
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
 Because the endurance requirement for solar-electric powered aircraft is set by the winter solstice and because wind speeds are higher during winter months, only wind speed data from December will be used for the aircraft sizing and performance analysis presented in this paper.
@@
 Gas powered aircraft generally fly between 5,000 to 25,000 ft due to payload requirements.  
-Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind} gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given latitude to meet a payload requirement.
+Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind}, gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given latitude to meet a payload requirement.
 
@@
     \label{e:tropoalt}
-    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}} 
+    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
 \end{align}
@@
 
-For the purposes of this design study, a percentile wind speed corresponds to that percentile of the wind speed data in December, across all longitudes and spanning the years 2005-2015.  
+For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
 To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-Wind speeds 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
+Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
 
@@
 Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
-Thus, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude. 
-Therefore, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
+Thus, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitudebetween 20 and 30 degrees. 
+In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
 The constraint on wind speed and air density at the 35 degree latitude is
@@
 
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day
+At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
 
@@
 where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_C$ is included in Equation~\ref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on battery power.
-Equations~\ref{e:solarreq} and~\ref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
+An additional energy term $(E/S)_C$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on battery power.
+Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude.}}
+ \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude. (Log-space RMS error = 0.007 and 0.0180)}}
  \label{f:energyapprox}
@@
 
-For the purposes of this design study, it is assumed that the solar cells are can be placed on the wings and the horizontal tail but are not placed on the vertical tail. 
+For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
 The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are placed on the horizontal tail,
@@
 
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion of the exponentiated right hand side is used,\cite{hoburgthesis}
+    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
 
@@
     where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.75$.
-Equations~\eqref{e:slfthrust},~\eqref{e:slfweight}, and~\eqref{e:slfpower} are all monomials and therefore GP-compatible.
 
@@
 The open circles are the calculated drag polars from XFOIL.
-(Log space RMS error $= 0.00489$.)}}
+(Log space RMS error~=~0.00489.)}}
 	\label{f:JH01polar}
@@
 A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight case and gust loads.
+It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight and gust wind loads.
 
@@
 Because long-endurance aircraft typically have high aspect ratio wing they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\ref{e:qbar}
+To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\eqref{e:qbar}
 
@@
 
-For this sizing study a cap spar is considered.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}
+For this sizing study a cap spar is considered.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
 
@@
     \label{e:htweight}
-    W_{\text{h}}/m_{fac} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{h}} \\
+    W_{\text{h}}/m_{\text{fac}} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{h}} \\
     \label{e:vtweight}
-    W_{\text{v}}/m_{fac} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{v}}
+    W_{\text{v}}/m_{\text{fac}} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{v}}
 \end{align}
@@
 To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
+RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
 It is assumed that engines of a similar class will perform similarly to the DF70 engine.  
@@
 
-where $\text{BSFC}_{\text{min}} = 0.3162$ [lb/hp/hr].\cite{rcvengines}
+where $\text{BSFC}_{\text{min}} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
 Using this approximations the required shaft power determines the $\text{BSFC}$.
@@
 Function is representative of engine performance for engines of similar size.
-The log-space RMS error is 0.007.
-}}
+(Log-space RMS error is 0.007) }}
 	\label{f:powervsweightfit}
@@
 The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
-The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizer, and engine. 
+The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
 The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.
@@
 where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is modeled as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
+However, if the structural weight is represented by a more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
 Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
@@
 From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% decrease in weight.
-From a design standpoint the sensitivity result is useful in that variables with high sensitivities are important to the vehicle design. 
+For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
+From a design standpoint the sensitivity result is useful in that variables with high absolute sensitivities are important to the vehicle design. 
 Table~\ref{t:sens} shows the variables with the highest sensitivities for the solar-electric powered aircraft optimized at various latitudes and percentile wind speeds by minimizing weight.

### 2017-01-27 gassolar/solar/solar_irradiance.py

@@ -45,12 +45,12 @@ def get_Eirr(latitude, day, N=50.0):
 if __name__ == "__main__":
     ES, td, tn, p = get_Eirr(30, 355, N=1000)
     fig, ax = plt.subplots()
-    ax.fill_between([-12, -td/2], 0, 80, alpha=0.5, facecolor="b",
+    ax.fill_between([-12, -td/2], 0, 80, alpha=0.3, facecolor="b",
                     linewidth=2, color="b")
-    ax.fill_between([td/2, 12], 0, 80, alpha=0.5, facecolor="b", linewidth=2,
+    ax.fill_between([td/2, 12], 0, 80, alpha=0.3, facecolor="b", linewidth=2,
                     color="b")
     ax.fill_between(p[1], p[0], 80, where=p[0] < 80, linewidth=2, color="b",
-                    facecolor="m", alpha=0.5)
+                    facecolor="m", alpha=0.3)
     ax.fill_between(p[1], 0, p[0], hatch="xx", color="r", linewidth=2,
                     facecolor="none")
     newp = []
@@ -60,7 +60,7 @@ if __name__ == "__main__":
         else:
             newp.append(80)
     ax.fill_between(p[1], 0, newp, linewidth=2, color="g", facecolor="g",
-                    alpha=0.5)
+                    alpha=0.3)
     ax.set_xlabel("Time [hr]")
     ax.set_ylabel("Available Solar Power [W/m$^2$]")
     ax.set_xlim([-12, 12])

### 2017-02-01 gassolarpaper/gassolarpaper.tex
final edits

@@
     However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by limited sunlight during the winter and wind speeds at higher latitudes.
+    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
     Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
@@
 BSFC \> brake specific fuel consumption [kg/kW/hr] \\
+$\text{BSFC}_{100\%}$ \> BSFC at full throttle [kg/kW/hr] \\
 $c$ \> wing chord [m] \\
@@
 $E_{\text{sun}}$ \> total solar energy available [Whr] \\
-$(E/S)_{\text{C}}$ \> low irradiance solar energy per unit area [Whr/m$^2$] \\
+$(E/S)_{\text{twilight}}$ \> low irradiance solar energy per unit area [Whr/m$^2$] \\
 $(E/S)_{\text{day}}$ \> daily operational solar energy per unit area [Whr/m$^2$] \\
@@
 $h$ \> flight altitude [ft] \\
-$h_{\text{batt}}$ \> battery energy density [Whr/kg] \\
+$h_{\text{batt}}$ \> battery specific energy [Whr/kg] \\
 $h_{\text{cap}}$ \> spar cap separation [m] \\
@@
 $L$ \> atmospheric lapse rate [K/m] \\
-$L'$ \> local lifting force [N/m] \\
 $L_\text{h}$ \> horizontal tail lift [N] \\
@@
 
-Long-endurance, station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
+Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
 Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
 Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-However, sizing a long-endurance aircraft is not immediately obvious because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines.
-Additionally, because few long-endurance gas powered aircraft have been built and only Airbus has successfully built and tested a solar aircraft for multiple days\cite{zephyr}, the operational limits of these aircraft is relatively unknown.
+Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
+Additionally, because few long-endurance gas and solar-electric powered aircraft have been built, the operational limits of these aircraft are relatively unknown. 
 Even more unclear are the trade offs between gas powered and solar-electric powered architectures.
@@
 This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Similar trade offs are calculated for other requirements and performance metrics to determine under what conditions which platform is best suited to meet a set of requirements.
+Similar trade offs are calculated for other requirements and performance metrics to determine under what conditions each platform is best suited to meet a set of requirements.
 
@@
 
-Design requirements to station keep at various altitudes, latitudes, and percentile wind speeds which will affect the performance and design of the aircraft.\cite{zephyr} 
-To station keep an aircraft must fly faster than the wind speed,
+To station keep an aircraft must fly at least as fast as the wind speed,
 
@@
 Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speeds can be approximated by using wind data across all longitudes.
+It is assumed that the wind speed distribution is independent of longitude.
 
@@
     \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Lower bound, upper bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.}}
+Lower, upper and middle lines represent 85th, 95th, and 90th percentile winds respectively.}}
 	\label{f:windvsmonth}
@@
 
-Gas powered aircraft generally fly between 5,000 to 25,000 ft due to payload requirements.  
-Because of a naturally aspirated engine and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind}, gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given latitude to meet a payload requirement.
+Gas powered aircraft generally fly between 5,000 and 25,000 ft due to payload requirements.  
+Because of the limitations of naturally aspirated engines and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind}, gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given altitude specified by the mission.
 
@@
     \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
-Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
-Wind data includes winds during December for the year 2005-2015.}}
+Left, right and middle lines represent 85th, 95th, and 90th percentile winds respectively.
+Wind data includes winds during December for the years 2005-2015.}}
 	\label{f:altvswind}
@@
 
-For the solar aircraft air density is calculated using equations specific to the tropopause,\cite{isaatm}
+For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
 
@@
     \label{e:tropoalt}
-    h &= 11,000 - \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
 \end{align}
@@
 
-Operationally, long-endurance aircraft will fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
-For example, an aircraft optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude.
+It is assumed that long-endurance aircraft will have a requirement to fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
+For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude.
 
@@
 
+Sizing long-edurance aircraft for the stated requirements was accomplished using geometric programming. 
 Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions. GPs have the form
@@
 
-To evaluate the size and performance of both gas and solar-electric powered aircraft underlying physics models are used.  
-Environmental models describing the effect of altitude, latitude and season on wind speeds, air density, and solar irradiance and aircraft models governing the performance, aerodynamics and structural weight are used in the optimization model.
+To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
+Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
 Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
@@
 To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach agrees from an operational perspective as a solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds also vary with latitude several constraints are needed to optimize altitude. 
-Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg\cite{fitting}.
+This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
+
+Because wind speeds also vary with latitude, several constraints are needed to optimize altitude. 
+Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg et. al\cite{fitting}.
 Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
 Altitude can be post-calculated from the optimized air density.
+
+For example, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
 Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
-Thus, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitudebetween 20 and 30 degrees. 
 In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
@@
 
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
+To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
 
@@
     \label{e:solarbatt}
-    E_{\text{batt}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_C \eta_{\text{solar}} S_{\text{solar}}
+    E_{\text{batt}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
 \end{align}
@@
 where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_C$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on battery power.
+An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
 Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
@@
 where $\eta_{\text{motor}} = 0.95$.
-Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_C$ are functions of $(P/S)_{\text{min}}$.
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.
+The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
 Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude.
@@
      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics{Benergy.pdf}}
-     \subfigure[Monomial Approximation for $(E/S)_C$\label{f:Cenergy}]{\includegraphics{Cenergy.pdf}}
+     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics{Cenergy.pdf}}
  \end{subfigmatrix}
- \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_C$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude. (Log-space RMS error = 0.007 and 0.0180)}}
+ \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude.
+(Log-space RMS error = 0.007 and 0.0180)}}
  \label{f:energyapprox}
@@
 
-\subsection{Breguet Range}
+\subsection{Breguet Endurance}
 
@@
 
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ is used.
+    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
 
@@
 
-The aircraft aerodynamics is modeled as a drag build up of wing profile drag, induced drag, and non-wing drag coefficients,
+The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
 
@@
     \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L}{\pi e A}.
+    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
     \end{equation}
@@
     To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg.\cite{fitting}
+    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
     The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
@@
 	\includegraphics[width=0.6\textwidth]{JH01polar.pdf}
-    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as a solid line.
+    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.
 The open circles are the calculated drag polars from XFOIL.
@@
 %  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery energy density and structural weight fraction vs wing span for solar-electric powered aircraft.
+%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
 Each point on the graph corresponds to a unique design.} }
@@
 
-A wing spar is one of the main structural elements of both aircraft. 
+A wing spar is one of the primary structural elements of both aircraft. 
 A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: steady level flight and gust wind loads.
-
-\subsubsection{Steady Level Flight Wing Loading Case}
-
-During steady level flight there exists a distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing
-
-\begin{equation}
-    \label{e:wloading}
-    q(y) = L'(y) - N_{\text{max}}gm'(y)
-\end{equation}
-
-where $q(y)$ is the spanwise distributed load. 
+It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
+
+\subsubsection{Standard Wing Bending Case}
+
+The distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing
 The wing loading distribution, $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
@@
 
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$. The local chord for a constant tapered wing\cite{bending} is
+where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$.
+$N_{\text{max}}=1$ corresponds to steady level flight.
+The local chord for a constant tapered wing\cite{bending} is
 
@@
 
-Because long-endurance aircraft typically have high aspect ratio wing they are considered flexible aircraft. 
+Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
 To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\eqref{e:qbar}
@@
 The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the wing
+The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
 
@@
 
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculating using techniques described by Hoburg\cite{fitting}.
+Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et. al\cite{fitting}.
 This approximation is graphically shown in Figure~\ref{f:arctanfit} for the expected range of $V_{\text{gust}}/V \in [0, 0.7]$.
@@
 	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
-    \caption{\textbf{Comparison of monomial approximation of $f(x) = \arctan{x}$ to real function, $x \in [0,0.7]$. (Log space RMS error = 0.0246)}}
+    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha}, $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
 	\label{f:arctanfit}
@@
 
-The gust velocity has a profile along the wing\cite{acgust},
+The gust velocity has an assumed profile along the wing\cite{acgust},
 
@@
 
-For this sizing study a cap spar is considered.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
+A cap spar is considered for the solar-electric and gas powered aircraft.
+Both aircraft optimization models share the same governing equations.
+A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
 
@@
     \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g 
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or the area density of one ply of carbon fiber.\cite{carbonfiber}
+    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{carbonfiber} The wing skin is asssumed not to contribute to the bending stiffness.
 
@@
 
-An empennage model with a single tail consisting of a tail boom, horizontal and vertical tail, is added to both the solar-electric and gas powered aircraft.  
-The empennage adds both weight and drag to the each aircraft.
+An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
+The empennage adds both weight and drag to each aircraft.
 
@@
 \begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 + \frac{1}{2} k\right) \\
+    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
     I_0 &\leq \pi t_0 d^3/8
@@
 
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.  For both the solar-electric and gas powered aircraft $k=0.8$.  
+where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
+For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
 When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
@@
 
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The volume of the fuselage is constrained by
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage is constrained by
 
@@
 where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].
-Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg\cite{fitting}.
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}
+Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg et.
+al\cite{fitting}.
+Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
 
@@
 	\includegraphics[width=0.6\textwidth]{powervsweightfit.pdf}
-    \caption{\textbf{Comparison of power law equation and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
+    \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
 	\label{f:powervsweightfit}
@@
 RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar class will perform similarly to the DF70 engine.  
-This engine performance model is only valid for IC engines.
-
-\subsubsection{Break Specific Fuel Consumption Constraint}
-
-A GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
+It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
+This engine performance model is only valid for internal combustion engines.
+
+\subsubsection{Brake Specific Fuel Consumption Constraint}
+
+To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
 
@@
     \label{e:rpmtobsfc}
-    \left( \frac{\text{BSFC}}{\text{BSFC}_{\text{min}}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-\end{equation}
-
-where $\text{BSFC}_{\text{min}} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-Using this approximations the required shaft power determines the $\text{BSFC}$.
-A comparison of the GP-compatible approximations to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
+    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
+\end{equation}
+
+where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
+Using this approximation the required shaft power determines the $\text{BSFC}$.
+A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
 A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of a similar class.
@@
 where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-Lithim-ion batteries are selected with a battery energy density $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumion}
+Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
+The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
 
@@
 \begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Gas Powered\label{f:mtowvslatgas}]{\includegraphics{mtowvslatgas.pdf}}
-     \subfigure[Solar-Electric Powered\label{f:mtowvslatsolar}]{\includegraphics{mtowvslatsolar.pdf}}
- \end{subfigmatrix}
+	\begin{center}
+	\includegraphics[width=0.6\textwidth]{mtowvslat.pdf}
     \caption{\textbf{Trade study of max take off weight and operational latitude bands.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-The endurance requirement for the gas powered aircraft in this trade study is 7 days.}}
+Lower, upper and middle curves were evaluated at the 80th, 95th, and 90th percentile winds respectively.
+The endurance requirement for the gas powered aircraft in this trade study is 7 days.
+The next integer latitude value for the solar-electric curves is infeasible.}}
     \label{f:latvsmtowtrade}
+	\end{center}
 \end{figure}
@@
 Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
-On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$35 degrees.
+On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
 
 Because gas-powered aircraft are endurance limited by the amount of fuel that they can carry, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitude only depends on wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed (38 degrees latitude at 15,000 ft) it can theoretically fly at any latitude.  
-Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing wing span at $\pm$38 degrees latitude.
+Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude.
 
@@
 High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires few solar cells.
+By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
 (Figure~\ref{f:solarobjcomp})
@@
 where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by a more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
+However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
 Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
@@
 The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping the aircraft will maintain a constant velocity during high wind speeds.  
+However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
 At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
@@
 	\includegraphics[width=0.6\textwidth]{polarmission.pdf}
-    \caption{\textbf{Comparision of operating lift to drag ratio with and without the wind speed constraint.
+    \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint.
 Solid lines are drag polars at various Reynolds numbers.
-For this result, the max take off weight was minimized for an endurance requirement of 7 days.}}
+The maximum $C_L^{1.5}/C_D$ are marked by a plus sign on the polar curve.}}
  \label{f:polarmission}
@@
 
-The solar aircraft size and design depends on the assumed solar cell efficiency and battery energy density. 
-By solving the model for different assumed solar cell efficiency and battery energy density values a broader picture of the design space is achieved.   
+The solar aircraft size and design depends on the assumed solar cell efficiency and battery specific energy. 
+By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
 Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
@@
  \end{subfigmatrix}
- \caption{\textbf{Matrix of contour plots of battery energy density and solar cell efficiency vs wing span for solar-electric powered aircraft where the wing span is minimized.
-Each point on the graph corresponds to a unique design.}}
+ \caption{\textbf{Matrix of contour plots of battery specific energy and solar cell efficiency vs wing span for solar-electric powered aircraft where the wing span is minimized.
+Each point on the graph corresponds to a unique design.
+Values of constants other than solar cell efficiency and battery specific energy are given throughout the text.
+The infeasible region and contour shapes would change for different assumed contant values.}}
  \label{f:solarcontours}
@@
 
-When a GP is solved, the sensitivity of the objective function with respect to each constraint is also returned.  
+When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
 From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-For example, if the objective function were $MTOW$ and the sensitivity to battery energy density were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-From a design standpoint the sensitivity result is useful in that variables with high absolute sensitivities are important to the vehicle design. 
-Table~\ref{t:sens} shows the variables with the highest sensitivities for the solar-electric powered aircraft optimized at various latitudes and percentile wind speeds by minimizing weight.
-
-It is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency is directly related to the operating power which goes by the cube of velocity. ($P_{\text{oper}}\sim V^3$)
-Thus, the discharge efficiency sensitivity shares similar growth to the percentile wind speed, while the charge efficiency sensitivity does not.
-
-It is also interesting to note that the design is significantly more sensitivity to the battery energy density than to the solar cell efficiency.  
-Thus to design a smaller, lighter aircraft, the sensitivity analysis reveals that it is more beneficial to invest in higher quality batteries than solar cells.
-
-The margin factor for the wing weight $m_{\text{fac-wing}}$, is a measure of the inaccuracy of the wind structural model.  
-If the wing weight is actually 5\% heavier than predicted then the aircraft will be approximately 4\% to 5\% heavier according Table~\ref{t:sens}.
+While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
+For example, if the objective function were MTOW and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.  
+Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take off weight.
+
+For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
+This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
+
+Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
+For example, both batteries and solar cells are expensive and important to the design.  
+The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
+The ratio of their magnitudes is 1.76.  
+Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
 
 \input{sens.generated.tex}
+
+\input{gassens.generated.tex}
 
@@
 However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery energy density and solar cell efficiency.
+An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
 Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between two options.
+The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
 

### 2017-02-01 gassolar/gas/engine_law.py

@@ -23,6 +23,6 @@ ax.plot(u*WEIGHT, w*POWER, "o", mfc="None", mew=1.5)
 ax.plot(weight, np.exp(yfit)*POWER, lw=2)
 ax.set_xlabel("Engine Weight [lbs]")
 ax.set_ylabel("Maximum Shaft Power [hp]")
-ax.legend(["UND Engine Data", "Power Fit"], loc=2)
+ax.legend(["UND Engine Data", "Power Law Fit"], loc=2)
 ax.grid()
 fig.savefig("../../gassolarpaper/powervsweightfit.pdf", bbox_inches="tight")

### 2017-02-01 gassolar/gas/flight_state.py

@@ -13,12 +13,12 @@ class FlightState(Model):
     percent: percentile wind speeds [%]
     day: day of the year [Jan 1st = 1]
     """
-    def setup(self, latitude=45, percent=90, altitude=15000, day=355):
+    def setup(self, Vwind, latitude=45, percent=90, altitude=15000, day=355):
 
-        wind = get_windspeed(latitude, percent, altitude, day)
+        # wind = get_windspeed(latitude, percent, altitude, day)
         density, vis = get_airvars(altitude)
 
-        Vwind = Variable("V_{wind}", wind, "m/s", "wind velocity")
+        # Vwind = Variable("V_{wind}", wind, "m/s", "wind velocity")
         mfac = Variable("m_{fac}", 1.0, "-", "wind speed margin factor")
         V = Variable("V", "m/s", "true airspeed")
         rho = Variable("\\rho", density, "kg/m**3", "air density")

### 2017-02-01 gassolar/gas/gas.py

@@ -29,6 +29,7 @@ class Aircraft(Model):
         Wpay = Variable("W_{pay}", 10, "lbf", "payload weight")
         Wavn = Variable("W_{avn}", 8, "lbf", "avionics weight")
         Wwing = Variable("W_{wing}", "lbf", "wing weight for loading")
+        etaprop = Variable("\\eta_{prop}", 0.75, "-", "propulsive efficiency")
 
         self.empennage.substitutions["V_h"] = 0.55
         self.empennage.substitutions["V_v"] = 0.04

### 2017-02-01 gassolar/gas/print_sens.py

@@ -1,0 +1,22 @@
+"print sensitivities"
+
+def sens_table(sols, varnames):
+    with open("../../gassolarpaper/gassens.generated.tex", "w") as f:
+        f.write("\\begin{longtable}{lccccccccccccc}\n")
+        f.write("\\caption{Gas Sensitivities}\\\\\n")
+        f.write("\\toprule\n")
+        f.write("\\toprule\n")
+        f.write("\\label{t:sens}\n")
+        f.write("Variable & 5 Day Endurance & 7 Day Endurance & 9 Day Endurance\\\\\n")
+        f.write("\\midrule\n")
+        for vname in varnames:
+            sens = []
+            for s in sols:
+                sen = s["sensitivities"]["constants"][vname]
+                if hasattr(sen, "__len__"):
+                    sen = s["sensitivities"]["constants"][max(sen)]
+                sens.append(sen)
+            vals = "$" + vname + "$ &" + " & ".join(["%.3g" % x for x in sens])
+            f.write(vals + "\\\\\n")
+        f.write("\\bottomrule\n")
+        f.write("\\end{longtable}")

### 2017-02-01 gassolar/solar/plotting.py

@@ -31,7 +31,7 @@ def windalt_plot(latitude, sol1, sol2):
         vsol = max([sol(sv).to("knots").magnitude for sv in sol("V")])
         ax.plot(altsol/1000, vsol, "o", markersize=10)
     ax.set_xlabel("Altitude [kft]")
-    ax.set_ylabel("Wind Speed [knots]")
+    ax.set_ylabel("90th Percentile Wind Speed [knots]")
     ax.grid()
     ax.set_ylim([0, 200])
     return fig, ax
@@ -113,7 +113,7 @@ def labelLine(line,x,label=None,align=True,**kwargs):
 
     ax.text(x,y,label,rotation=trans_angle,**kwargs)
 
-def labelLines(lines,align=True,xvals=None,**kwargs):
+def labelLines(lines,align=True,xvals=None,zorder=[],**kwargs):
 
     ax = lines[0].get_axes()
     labLines = []
@@ -130,5 +130,5 @@ def labelLines(lines,align=True,xvals=None,**kwargs):
         xmin,xmax = ax.get_xlim()
         xvals = np.linspace(xmin,xmax,len(labLines)+2)[1:-1]
 
-    for line,x,label in zip(labLines,xvals,labels):
-        labelLine(line,x,label,align,**kwargs)
+    for line,x,label,zo in zip(labLines,xvals,labels, zorder):
+        labelLine(line,x,label,align,zorder=zo,**kwargs)

### 2017-02-01 gassolar/solar/solar_irradiance.py

@@ -65,13 +65,11 @@ if __name__ == "__main__":
     ax.set_ylabel("Available Solar Power [W/m$^2$]")
     ax.set_xlim([-12, 12])
     ax.text(-1.2, 400, "$(E/S)_{\mathrm{sun}}$", fontsize=15)
-    ax.text(-10, 30, "$E_{\mathrm{batt}}/S_{\mathrm{solar}}$", fontsize=15)
-    ax.text(7, 30, "$E_{\mathrm{batt}}/S_{\mathrm{solar}}$", fontsize=15)
     ax.text(-1.2, 30, "$(E/S)_{\mathrm{day}}$", fontsize=15)
-    ax.annotate("$(E/S)_C$", xy=(-4.9, 70), xytext=(-9, 220),
+    ax.annotate("$(E/S)_{\mathrm{twilight}}$", xy=(-4.9, 70), xytext=(-10.1, 220),
                 arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
                                 headwidth=10, frac=0.1))
-    ax.annotate("", xy=(5.5, 70), xytext=(-6.8, 205),
+    ax.annotate("", xy=(5.5, 70), xytext=(-6.7, 205),
                 arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
                                 headwidth=10, frac=0.025))
     ax.grid()
@@ -106,7 +104,7 @@ if __name__ == "__main__":
         ax1.plot(P[1:-15], 2*C[:-14], "o", c=col, markerfacecolor="none")
         ax1.plot(P[1:-15], np.exp(yfit), c=col, label="%dth Latitude" % l)
         ax1.set_xlabel("Minimum Necessary Power $(P/S)_{\mathrm{min}}$ [W/m$^2$]")
-        ax1.set_ylabel("Extra Required Battery Energy $(E/S)_C}$ [Whr/m$^2$]")
+        ax1.set_ylabel("Extra Required Battery Energy $(E/S)_{\mathrm{twilight}}}$ [Whr/m$^2$]")
         ax1.grid()
         params.append(cn[0].right.c)
         params.append(cn[0].right.exp[list(cn[0].varkeys["u_fit_(0,)"])[0]])

### 2017-02-01 gassolar/solar/testcon.py

@@ -1,0 +1,107 @@
+from solar import Mission
+import matplotlib.pyplot as plt
+import numpy as np
+from plotting import windalt_plot, labelLines
+from gpkit.tools.autosweep import sweep_1d
+
+LATITUDE = False
+WIND = False
+CON = True
+COMP = False
+SENS = False
+
+""" contour """
+
+N = 100
+plt.rcParams.update({'font.size':19})
+hmin = 250.0
+hmax = 400.0
+for av in [80, 85, 90]:
+    for lat in [25, 30, 35]:
+        zo = 10
+        etamin = 0.15
+        etamax = 0.4
+        if lat == 25:
+            bs = range(30, 80, 5)
+        elif lat == 30:
+            bs = range(35, 80, 5)
+            if av == 90:
+                del bs[0]
+        else:
+            bs = range(45, 80, 5)
+            if av == 90:
+                del bs[0]
+        fig, ax = plt.subplots()
+        M = Mission(latitude=lat)
+        M.substitutions.update({"W_{pay}": 10})
+        for vk in M.varkeys["\\eta_{prop}"]:
+            M.substitutions.update({vk: 0.75})
+        for vk in M.varkeys["p_{wind}"]:
+            M.substitutions.update({vk: av/100.0})
+        del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
+        del M.substitutions["h_{batt}"]
+        M.cost = M["h_{batt}"]
+        lines = []
+        midx = []
+        for b in bs:
+            M.substitutions.update({"b_Mission, Aircraft, Wing": b})
+            M.substitutions.update({"\\eta_Mission, Aircraft, SolarCells": etamax})
+            sol = M.solve("mosek")
+            if sol("h_{batt}").magnitude < hmin:
+                M.substitutions.update({"h_{batt}": hmin})
+                del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
+                M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
+                sol = M.solve("mosek")
+                etamax = sol("\\eta_Mission, Aircraft, SolarCells")
+            else:
+                etammax = 0.5
+                del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
+
+            M.substitutions.update({"h_{batt}": hmax})
+            M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
+            sol = M.solve("mosek")
[…]

### 2017-02-01 gassolarpaper/gassolarpaper.tex
format lat sweep

@@
  \usepackage{multirow}
+ \usepackage{setspace}
+ \doublespacing
 
@@
 \begin{abstract}
-    Recently, several companies have invested in developing aircraft that can provide long-endurance (days to weeks) persistent aerial coverage. 
+    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
     These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
@@
 $E_{\text{sun}}$ \> total solar energy available [Whr] \\
-$(E/S)_{\text{twilight}}$ \> low irradiance solar energy per unit area [Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily operational solar energy per unit area [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy per unit area [Whr/m$^2$] \\
+$(E/S)_{\text{twilight}}$ \> twilight solar energy [Whr/m$^2$] \\
+$(E/S)_{\text{day}}$ \> daily required solar energy [Whr/m$^2$] \\
+$(E/S)_{\text{sun}}$ \> available solar energy [Whr/m$^2$] \\
 $f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
@@
 $P_{\text{oper}}$ \> aircraft operating power [hp] \\
-$(P/S)_{\text{min}}$ \> minium operational solar power \\
-\>per unit area [Whr/m$^2$] \\
+$(P/S)_{\text{min}}$ \> minium operational solar power [Whr/m$^2$] \\
 $P_{\text{shaft}}$ \> engine/motor shaft power [hp] \\
@@
     \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Lower, upper and middle lines represent 85th, 95th, and 90th percentile winds respectively.}}
+Lower, middle and upper lines represent 85th, 90th and 95th percentile winds respectively.}}
 	\label{f:windvsmonth}
@@
     \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
-Left, right and middle lines represent 85th, 95th, and 90th percentile winds respectively.
+Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively.
 Wind data includes winds during December for the years 2005-2015.}}
@@
     \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Left bound, right bound and middle lines represent 85th, 95th, and 90th percentile winds respectively.
+Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively.
 Wind data includes winds during December for the years 2005-2015.}}
@@
 	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
-    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha}, $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
+    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
 	\label{f:arctanfit}
@@
 
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage is constrained by
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
 
@@
 A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of a similar class.
+A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
 
@@
 
-To achieve sizing results for the solar-electric and gas powered long-endurance aircraft the optimization models were solve by minimizing the wing span across different bands of latitudes, shown in Figure~\ref{f:latvsmtowtrade}.
+To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes, shown in Figure~\ref{f:latvsmtowtrade}.
+The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.
 
@@
 	\includegraphics[width=0.6\textwidth]{mtowvslat.pdf}
-    \caption{\textbf{Trade study of max take off weight and operational latitude bands.
-Lower, upper and middle curves were evaluated at the 80th, 95th, and 90th percentile winds respectively.
+    \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements.
+Each point is a unique solution from the optimization.
+Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.
 The endurance requirement for the gas powered aircraft in this trade study is 7 days.
-The next integer latitude value for the solar-electric curves is infeasible.}}
+The next integer latitude value for each of the solar-electric curves is infeasible for the given assumptions.}}
     \label{f:latvsmtowtrade}
@@
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-Figure~\ref{f:mtowvslatgas} levels off around 38 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
+The gas powered aircraft curves level off around 38 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
 Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
-On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
+On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
 
@@
 	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
-    \caption{\textbf{Trade study of endurance and size for gas powered architecture.}}
+    \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.
+Max take off weight was minimized for this analysis.}}
 	\label{f:spanvsendurance}
@@
  \end{subfigmatrix}
-    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
+    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.
+Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.}}
     \label{f:solarobjcomp}
@@
   \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl25a85.pdf}}
-  \subfigure[25h Latitude, 90\% Wind Speed]{\includegraphics{bcontourl25a90.pdf}}
+  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl25a90.pdf}}
  \end{subfigmatrix}
@@
 Values of constants other than solar cell efficiency and battery specific energy are given throughout the text.
-The infeasible region and contour shapes would change for different assumed contant values.}}
+The infeasible region and contour shapes would change for different assumed constant values.}}
  \label{f:solarcontours}
@@
 
-Using geometric programming as a form of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
+Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
 One high level discovery is the importance of wind speed.  

### 2017-02-01 gassolar/solar/solar_irradiance.py

@@ -101,8 +101,8 @@ if __name__ == "__main__":
         cn, rm = fit(x, y, 1, "MA")
         print "RMS error: %.4f" % rm
         yfit = cn.evaluate(x)
-        ax1.plot(P[1:-15], 2*C[:-14], "o", c=col, markerfacecolor="none")
-        ax1.plot(P[1:-15], np.exp(yfit), c=col, label="%dth Latitude" % l)
+        ax1.plot(P[1:-15], 2*C[:-14], "o", c=col, markerfacecolor="none", mew=1.5)
+        ax1.plot(P[1:-15], np.exp(yfit), c=col, label="%dth Latitude" % l, lw=2)
         ax1.set_xlabel("Minimum Necessary Power $(P/S)_{\mathrm{min}}$ [W/m$^2$]")
         ax1.set_ylabel("Extra Required Battery Energy $(E/S)_{\mathrm{twilight}}}$ [Whr/m$^2$]")
         ax1.grid()
@@ -114,8 +114,8 @@ if __name__ == "__main__":
         cn, rm = fit(x, y, 1, "MA")
         print "RMS error: %.4f" % rm
         yfit = cn.evaluate(x)
-        ax2.plot(P[1:-15], 2*B[:-14], "o", c=col, markerfacecolor="none")
-        ax2.plot(P[1:-15], np.exp(yfit), c=col, label="%dth Latitude" % l)
+        ax2.plot(P[1:-15], 2*B[:-14], "o", c=col, markerfacecolor="none", mew=1.5)
+        ax2.plot(P[1:-15], np.exp(yfit), c=col, label="%dth Latitude" % l, lw=2)
         ax2.grid()
         ax2.set_xlabel("Minimum Necessary Power $(P/S)_{\mathrm{min}}$ [W/m$^2$]")
         ax2.set_ylabel("Daytime Energry for Solar Cells $(E/S)_{\mathrm{day}}$ [Whr/m$^2$]")

### 2017-02-01 gassolarpaper/gassolarpaper.tex
submission draft

@@
  \usepackage{setspace}
- \doublespacing
+ % \usepackage[dvipdfmx]{graphicx} 
+ % \usepackage{bmpsize}
+ % \doublespacing
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{windvsmonth.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
     \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{altvswind.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
     \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{latvswind.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
     \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics{Boston_dec_2015auto.pdf}}
-     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics{Boston_dec_2015.pdf}}
+     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
+     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
  \end{subfigmatrix}
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics{Rome_dec_2015auto.pdf}}
-     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics{Rome_dec_2015.pdf}}
+     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
+     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
  \end{subfigmatrix}
@@
 \text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p \nonumber 
+g_i (\bold{x}) &= 1, i = 1,\ldots,p 
 \end{align}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{windfitl35.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
     \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}.
@@
 	\begin{center}
-	\includegraphics[width=0.7\textwidth]{lat30.pdf}
+	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
     \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.
@@
     (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}. \\
+    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
 \end{align}
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics{Benergy.pdf}}
-     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics{Cenergy.pdf}}
+     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
+     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
  \end{subfigmatrix}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{JH01polar.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
     \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.
@@
 %  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics{bvehiclecomp.pdf}}
+%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
+%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
 %  \end{subfigmatrix}
@@
 %  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics{mtowvslatsolarsimple.pdf}}
+%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
+%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
 %  \end{subfigmatrix}
@@
 % 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth]{mtowvsendsimple.pdf}
+% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
 %     \caption{ \textbf{ Trade study between endurance and max take off weight for a gas powered archicture. }}
@@
 %  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl30a90.pdf}}
+%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
+%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
+%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
+%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
+%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
+%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
+%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
+%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
+%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
 %  \end{subfigmatrix}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
     \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
@@
 	\begin{center}
-	\includegraphics[width=0.8\textwidth]{capspar2.jpg}
+	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
     \caption{\textbf{Cross sectional view of a cap spar.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{taildragpolar.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
     \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{powervsweightfit.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
     \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{powertobsfcfit.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
     \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{mtowvslat.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
     \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
     \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Wingspan\label{f:span}]{\includegraphics{solarobjcomp.pdf}}
-     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics{solarobjcomp2.pdf}}
+     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
+     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
  \end{subfigmatrix}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{windaltoper.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
  \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth]{polarmission.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
     \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint.
@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl35a80.pdf}}
-  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl35a85.pdf}}
-  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl35a90.pdf}}
-  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl30a80.pdf}}
-  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl30a85.pdf}}
-  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl30a90.pdf}}
-  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl25a80.pdf}}
-  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl25a85.pdf}}
-  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl25a90.pdf}}
+  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.pdf}}
+  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.pdf}}
+  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.pdf}}
+  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.pdf}}
+  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.pdf}}
+  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.pdf}}
+  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.pdf}}
+  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.pdf}}
+  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.pdf}}
  \end{subfigmatrix}
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics{cpmin.pdf}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics{cpmax.pdf}}
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmin.pdf}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmax.pdf}}
  \end{subfigmatrix}

### 2017-02-02 gassolarpaper/gassolarpaper.tex
fix no units text

@@
 
-For example, an aircraft optimized for $\pm$30 will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
+For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees.
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 

### 2017-02-02 gassolarpaper/gassolarpaper.tex
shorten captions

@@
 $\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span [ft] \\
-BSFC \> brake specific fuel consumption [kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle [kg/kW/hr] \\
-$c$ \> wing chord [m] \\
+$b$ \> wing span \\ % [ft] \\
+BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
+$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
+$c$ \> wing chord \\ %[m] \\
 $C_D$ \> aircraft drag coefficient \\
@@
 $C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter [m] \\
-$D_{\text{boom}}$ \> tail boom drag [N] \\
-$\Delta W$ \> wing section weight [N] \\
-$\Delta y$ \> wing section length [m] \\
+$d$ \> tail boom diameter \\ %[m] \\
+$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
+$\Delta W$ \> wing section weight \\ %[N] \\
+$\Delta y$ \> wing section length \\ %[m] \\
 DOY \> day of the year \\
 $e$ \> Oswald efficiency factor \\
-$E$ \> Young's Modulus [Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery [J] \\
-$E_{\text{sun}}$ \> total solar energy available [Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy [Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy [Whr/m$^2$] \\
+$E$ \> Young's Modulus \\ %[Pa] \\
+$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
+$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
+$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
+$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
+$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
 $f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
 $f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant [m/s$^2$] \\
-$h$ \> flight altitude [ft] \\
-$h_{\text{batt}}$ \> battery specific energy [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation [m] \\
-$I$ \> cap spar moment of inertia [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia [m$^4$] \\
+$g$ \> gravitational constant \\ % [m/s$^2$] \\
+$h$ \> flight altitude \\ % [ft] \\
+$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
+$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
+$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
+$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
 $k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant [N/m$^2$] \\
-$L$ \> atmospheric lapse rate [K/m] \\
-$L_\text{h}$ \> horizontal tail lift [N] \\
-$l_\text{fuse}$ \> fuselage length [m] \\
-$l_\text{h}$ \> horizontal tail moment arm [m] \\
-$l_\text{v}$ \> vertical tail moment arm [m] \\
-$m$ \> tail boom mass [kg] \\
-$m'$ \> local wing mass [kg/m] \\
+$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
+$L$ \> atmospheric lapse rate \\ % [K/m] \\
+$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
+$l_\text{fuse}$ \> fuselage length \\ % [m] \\
+$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
+$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
+$m$ \> tail boom mass \\ % [kg] \\
+$m'$ \> local wing mass \\ % [kg/m] \\
 $m_{\text{fac-wing}}$ \> margin factor\\
@@
 $m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate [kg/s] \\
-$\mathcal{M}$ \> wing bending moment [Nm] \\
-MTOW \> max take-off weight [N] \\
+$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
+$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
+MTOW \> max take-off weight \\ % [N] \\
 $n$ \> number of wing segments \\
@@
 $N_{\text{max}}$ \> safety load factor\\
-$p$ \> air pressure [Pa]\\
-$p_{11}$ \> air pressure at 11,000 meters [Pa]\\
+$p$ \> air pressure \\ % [Pa]\\
+$p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
 $p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power [hp] \\
-$(P/S)_{\text{min}}$ \> minium operational solar power [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface [W/m$^2$] \\
+$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
+$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
+$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
+$(P/S)_{\text{min}}$ \> minium operational solar power \\ % [Whr/m$^2$] \\
+$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
+$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
+$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
 $r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range [nmi] \\
+$R$ \> aircraft range \\ % [nmi] \\
 $R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius [m] \\
-$R_{\text{spec}}$ \> specific gas constant [J/k/K] \\
+$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
+$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
 $R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading [N/m] \\
+$q$ \> distributed wing loading \\ % [N/m] \\
 $\bar{q}$ \> normalized distributed wing loading \\
@@
 $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area [m$^2$]\\
-$\mathcal{S}$ \> shear forces [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area [m$^2$]\\
-$t$ \> flight time [days] \\
-$t_0$ \> tail boom root wall thickness [m] \\
-$t_{\text{cap}}$ \> spar cap thickness [m] \\
-$t_{\text{day}}$ \> hours of daylight [hr] \\
-$t_{\text{night}}$ \> hours of darkness [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise [hr] \\
-$t_{\text{sun set}}$ \> time of sun set [hr] \\
-$T$ \> aircraft thrust [N] \\
-$T_{11}$ \> air temperature at 11,000 meters [K] \\
-$T_{\text{atm}}$ \> air temperature at flight altitude [K] \\
-$T_{\text{sl}}$ \> air temperature at sea level [K] \\
-$V$ \> true airspeed [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity [m/s] \\
+$S$ \> wing planform area \\ % [m$^2$]\\
+$\mathcal{S}$ \> shear forces \\ % [N] \\
+$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
+$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
+$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
+$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
+$t$ \> flight time \\ % [days] \\
+$t_0$ \> tail boom root wall thickness \\ % [m] \\
+$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
+$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
+$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
+$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
+$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
+$T$ \> aircraft thrust \\ % [N] \\
+$T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
+$T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
+$T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
+$V$ \> true airspeed \\ % [m/s] \\
+$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
+$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
 $V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed [m/s] \\
-$V_{\text{wind}}$ \>  wind speed [m/s] \\
+$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
+$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
 $V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection [m] \\
-$W$ \> aircraft weight [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment [N] \\
-$W_{\text{batt}}$ \> battery weight [N] \\
-$w_{\text{cap}}$ \> spar cap width [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight [N] \\
-$W_{\text{empennage}}$ \> empennage weight [N] \\
-$W_{\text{engine}}$ \> engine weight [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight [N] \\
-$W_{\text{fuel}}$ \> fuel weight [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight [N] \\
-$W_{\text{h}}$ \> horizontal tail weight [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight [N] \\
-$w_{\text{max}}$ \> maximum deflection limit [m] \\
-$W_{\text{payload}}$ \> payload weight [N] \\
-$W_{\text{skin}}$ \> wing skin weight [N] \\
-$W_{\text{solar}}$ \> solar cell weight [N] \\
-$W_{\text{spar}}$ \> wing spar weight [N] \\
-$W_{\text{structural}}$ \> structural weight [N] \\
-$W_{\text{v}}$ \> vertical tail weight [N] \\
-$W_{\text{wing}}$ \> wing weight [N] \\
-$y$ \> distance from wing root along span [m] \\
+$w$ \> wing deflection \\ % [m] \\
+$W$ \> aircraft weight \\ % [N] \\
+$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
+$W_{\text{batt}}$ \> battery weight \\ % [N] \\
+$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
+$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
+$W_{\text{empennage}}$ \> empennage weight \\ % [N] \\
+$W_{\text{engine}}$ \> engine weight \\ % [N] \\
+$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
+$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
+$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
+$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
+$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
+$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
+$W_{\text{payload}}$ \> payload weight \\ % [N] \\
+$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
+$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
+$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
+$W_{\text{structural}}$ \> structural weight \\ % [N] \\
+$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
+$W_{\text{wing}}$ \> wing weight \\ % [N] \\
+$y$ \> distance from wing root along span \\ % [m] \\
 $z_{\text{bre}}$ \> Breguet Range helper variable \\
 $\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities [rad] \\
+$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
 $\Delta$ \> solar declination angle \\
@@
 $\eta_{\text{solar}}$ \> solar cell efficiency \\
-$\mu$ \> air viscosity [kg/m/s] \\
-$\mu_0$ \> reference air viscosity [kg/m/s] \\
+$\mu$ \> air viscosity \\ % [kg/m/s] \\
+$\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
 $\phi$ \> latitude \\
-$\rho$ \> air density [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress [Pa] \\
+$\rho$ \> air density \\ % [kg/m$^3$] \\
+$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
+$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
+$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
+$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
+$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
 $\tau_{\text{h}}$ \> horiztonal tail thickness to chord ratio \\
@@
 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-    \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Lower, middle and upper lines represent 85th, 90th and 95th percentile winds respectively.}}
+    \caption{\textbf{Wind speeds by month at 30 degrees latitude.  Bands represent 80th, 90th and 95th percentile winds.}}
 	\label{f:windvsmonth}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
-    \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.
-Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively.
-Wind data includes winds during December for the years 2005-2015.}}
+    \caption{\textbf{Wind speeds by altitude. Bands represent 80th, 90th and 95th percentile winds.}}
 	\label{f:altvswind}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
-    \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.
-Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively.
-Wind data includes winds during December for the years 2005-2015.}}
+    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds. }}
 	\label{f:latvswind}
@@
 In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
-The constraint on wind speed and air density at the 35 degree latitude is
+The constraint on December wind speeds and air density at the 35 degree latitude is
 
@@
 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
-    \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}.
-(Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.}}
+    \caption{\textbf{Posynomial approximation to wind speed data comparison at 35$^{\circ}$ latitude. (Log space RMS error = 0.043.)}}
 	\label{f:windfitl35}
@@
 	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$.
-For a solar-electric powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
+    \caption{\textbf{Solar power on Dec 21st at $30^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
 	\label{f:lat30}
@@
  \end{subfigmatrix}
- \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude.
-(Log-space RMS error = 0.007 and 0.0180)}}
+ \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
  \label{f:energyapprox}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
-    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.
-The open circles are the calculated drag polars from XFOIL.
-(Log space RMS error~=~0.00489.)}}
+    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles) comparison.  (Log-space RMS error~=~0.00489.)}}
 	\label{f:JH01polar}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
-    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
+    \caption{\textbf{Monomial approximation to Equation~\ref{e:gustalpha} comparison for $V/V_{\text{gust}} \in [0,0.7]$. (Log-space RMS error = 0.0246)}}
 	\label{f:arctanfit}
@@
 A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
+A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
 
@@
 	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
-    \caption{\textbf{Cross sectional view of a cap spar.
-The two caps are made of carbon fiber, separated by a foam core.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.}}
+    \caption{\textbf{Cross sectional view of a cap spar.}}
 	\label{f:capspar}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-    \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
+    \caption{\textbf{Posynomial equation (solid lines) to XFOIL data (circles) comparison. (Log space RMS error = 0.0173)}}
 	\label{f:taildragpolar}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
-    \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
+    \caption{\textbf{Power law fit to University of North Dakota data\cite{gasengine} comparison. (Log space RMS error = 0.34)}}
 	\label{f:powervsweightfit}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
-    \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.
-Function is representative of engine performance for engines of similar size.
-(Log-space RMS error is 0.007) }}
+    \caption{\textbf{Representative engine performance function based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
 	\label{f:powervsweightfit}
@@
 
-To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes, shown in Figure~\ref{f:latvsmtowtrade}.
+To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes.
+Figure~\ref{latvsmtowtrade} shows this trade study evaluated at the 80th, 90th and 95th percentile wind speeds.
+Each point on the graph is a unique solution to the optimization model for the stated assumptions and contants.
 The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.
@@
 	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
-    \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements.
-Each point is a unique solution from the optimization.
-Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.
-The endurance requirement for the gas powered aircraft in this trade study is 7 days.
-The next integer latitude value for each of the solar-electric curves is infeasible for the given assumptions.}}
+    \caption{\textbf{Gas optimization model feasible for all latitudes. Next integer latitude value for each solar-electric curve is infeasible.}}
     \label{f:latvsmtowtrade}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
-    \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.
-Max take off weight was minimized for this analysis.}}
+    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
 	\label{f:spanvsendurance}
@@
  \end{subfigmatrix}
-    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.
-Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.}}
+    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
     \label{f:solarobjcomp}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
- \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.
-With only a structural fraction as the structural model the optimum point is at the minimum wind speed.
-With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.}}
+ \caption{\textbf{Comparison of structural fraction and detailed structural models highlights trade between wing weight and air density.}}
  \label{f:altoper}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
-    \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint.
-Solid lines are drag polars at various Reynolds numbers.
-The maximum $C_L^{1.5}/C_D$ are marked by a plus sign on the polar curve.}}
+    \caption{\textbf{Wind speed constraint moves lift to drag ratio from optimum endurance ratio $\text{max}(C_L^{1.5}/C_D)$ (plus sign). Solid lines are drag polars.}}
  \label{f:polarmission}
@@
 Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
+Each point in Figure~\ref{f:solarcontours} is a unique design and was solved by minizing wing span. 
+Values of constants other than solar cell efficiency are given throughout the text.
+The infeasible regions and contour shapes would charge for different assumed constant values.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Matrix of contour plots of battery specific energy and solar cell efficiency vs wing span for solar-electric powered aircraft where the wing span is minimized.
-Each point on the graph corresponds to a unique design.
-Values of constants other than solar cell efficiency and battery specific energy are given throughout the text.
-The infeasible region and contour shapes would change for different assumed constant values.}}
+ \caption{\textbf{Operating at higher latitudes requires both high performing batteries and solar cells.}}
  \label{f:solarcontours}
@@
 \label{t:windvals} \\
+    \caption{Wind Speed Function Coefficients and Exponents} \\
     \toprule
     \toprule
-    \caption{Wind Speed Function Coefficients and Exponents} \\
     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
@@
 The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=3e5$.
+Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
 The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
@@
  \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.}}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=\num{3d5}$.}}
  \label{f:jhcps}

### 2017-02-08 gassolar/environment/solar_irradiance.py
initial start on presentation; presentation abstract

@@ -78,6 +78,30 @@ if __name__ == "__main__":
     ax2.set_yticklabels(["", "$(P/S)_{\mathrm{min}}$", ""])
     fig.savefig("../../gassolarpaper/lat30.pdf", bbox_inches="tight")
 
+    # solar irradiance by year
+    Fig, Ax = plt.subplots()
+    Mos = ["jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep",
+           "oct", "nov", "dec", "jan"]
+    Dayinmo = [0, 31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31]
+    Moday = [sum(Dayinmo[:i+1]) for i in range(len(Dayinmo))]
+    Mid = [(Moday[i]+Moday[i+1])/2 for i in range(len(Moday)-1)]
+    es = []
+    for d in range(365):
+        E, _, _, _ = get_Eirr(30, d)
+        es.append(E)
+    ind = es.index(min(es))
+    Ax.annotate("winter solstice", xy=(ind, min(es)), xytext=(200, 5500),
+               arrowprops=dict(arrowstyle="->"))
+    Ax.plot(range(365), es)
+    Ax.set_xticks(Moday)
+    Ax.set_xticks(Mid, minor=True)
+    Ax.set_xticklabels(Mos, minor=True)
+    Ax.set_xticklabels([])
+    Ax.set_ylabel("Solar Energy [Whr/m$^2$]")
+    Ax.grid()
+    Ax.set_xlim([0, 365])
+    Fig.savefig("../../gassolarpaper/eirrvsmonth.pdf", bbox_inches="tight")
+
     data = {}
     fig1, ax1 = plt.subplots()
     fig2, ax2 = plt.subplots()

### 2017-02-13 gassolar/solar/battsolarcon.py
finalize presentation

@@ -1,0 +1,59 @@
+" contour plots "
+import matplotlib.pyplot as plt
+import numpy as np
+from solar import Mission
+from plotting import windalt_plot, labelLines
+from gpkit.tools.autosweep import autosweep_1d
+
+N = 100
+plt.rcParams.update({'font.size':15})
+fig, ax = plt.subplots()
+
+etamax = []
+lines = []
+midx = []
+passing = True
+for lat in range(20, 35, 1):
+    if passing:
+        hmax = 500
+        M = Mission(latitude=lat)
+        M.substitutions.update({"W_{pay}": 10})
+        for vk in M.varkeys["\\eta_{prop}"]:
+            M.substitutions.update({vk: 0.75})
+        for vk in M.varkeys["p_{wind}"]:
+            M.substitutions.update({vk: 90/100.0})
+        del M.substitutions["h_{batt}"]
+        notpassing = True
+        M.substitutions.update({"\\eta_Mission, Aircraft, SolarCells": 0.4})
+        M.cost = M["h_{batt}"]
+        try:
+            sol = M.solve("mosek")
+            hmin = sol["cost"].magnitude + 1e-3
+            del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
+            M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
+            xmin_ = np.linspace(hmin, hmax, 100)
+            tol = 0.01
+            bst = autosweep_1d(M, tol, M["h_{batt}"], [hmin, hmax],
+                               solver="mosek")
+            if lat % 4 == 0:
+                l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "k",
+                            label="%d" % lat)
+                lines.append(l[0])
+                etamax.append(max(bst.sample_at(xmin_)["cost"].magnitude))
+                midx.append(np.median(xmin_))
+            elif lat % 2 == 0:
+                l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "--", c="0.5",
+                            label="%d" % lat)
+        except RuntimeWarning:
+            passing = False
+
+ax.fill_between(xmin_, bst.sample_at(xmin_)["cost"], max(etamax),
+                edgecolor="r", lw=2, hatch="/", facecolor="None", zorder=100)
+labelLines(lines, align=False, xvals=midx, zorder=[10]*len(lines))
[…]

### 2017-02-21 gassolar/environment/solar_irradiance.py
bag of constraints; additional motivational slides

@@ -100,6 +100,7 @@ if __name__ == "__main__":
     Ax.set_ylabel("Solar Energy [Whr/m$^2$]")
     Ax.grid()
     Ax.set_xlim([0, 365])
+    Ax.set_ylim([0, 12000])
     Fig.savefig("../../gassolarpaper/eirrvsmonth.pdf", bbox_inches="tight")
 
     data = {}

### 2017-02-21 gassolar/environment/wind_speeds.py

@@ -83,7 +83,8 @@ def interpolate(xs, ys, x):
 if __name__ == "__main__":
 
     Fig, Ax = plt.subplots()
-    Colors = ["b", "g", "r"]
+    # Colors = ["#253B6E", "#1F5F8B", "#1891AC"]
+    Colors = ["#1C226B", "#3E31AE", "#4AA9AF"]
     for al, c in zip([15000, 50000, 60000], Colors):
         Wind85 = [get_windspeed(l, 80, al, 355) for l in np.arange(70)]
         Wind90 = [get_windspeed(l, 90, al, 355) for l in np.arange(70)]
@@ -100,6 +101,7 @@ if __name__ == "__main__":
 
     Fig, Ax = plt.subplots()
     Alt = range(1000, 80000, 1000)
+    Colors = ["#470031", "#971549", "#CF455C"]
     for l, c in zip([30, 35, 45], Colors):
         Wind85 = get_windspeed(l, 80, Alt, 355)
         Wind90 = get_windspeed(l, 90, Alt, 355)
@@ -115,6 +117,7 @@ if __name__ == "__main__":
     Fig.savefig("../../gassolarpaper/altvswind.pdf", bbox_inches="tight")
 
     Fig, Ax = plt.subplots()
+    Colors = ["#1C226B", "#3E31AE", "#4AA9AF"]
     Mos = ["jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep",
            "oct", "nov", "dec", "jan"]
     Dayinmo = [0, 31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31]

### 2017-02-22 gassolar/environment/wind_speeds.py
presentation refinement

@@ -85,12 +85,23 @@ if __name__ == "__main__":
     Fig, Ax = plt.subplots()
     # Colors = ["#253B6E", "#1F5F8B", "#1891AC"]
     Colors = ["#1C226B", "#3E31AE", "#4AA9AF"]
+    lat = np.arange(70)
     for al, c in zip([15000, 50000, 60000], Colors):
-        Wind85 = [get_windspeed(l, 80, al, 355) for l in np.arange(70)]
-        Wind90 = [get_windspeed(l, 90, al, 355) for l in np.arange(70)]
-        Wind95 = [get_windspeed(l, 95, al, 355) for l in np.arange(70)]
+        Wind85 = [get_windspeed(l, 80, al, 355) for l in lat]
+        Wind90 = [get_windspeed(l, 90, al, 355) for l in lat]
+        Wind95 = [get_windspeed(l, 95, al, 355) for l in lat]
         Ax.fill_betweenx(np.arange(70), Wind85, Wind95, alpha=0.5, facecolor=c)
         Ax.plot(Wind90, np.arange(70), c, label="Altitude=%d" % al, lw=2)
+        if al == 50000:
+            Ax.annotate("80%", xy=(Wind85[np.where(lat==31)[0][0]],31),
+                        xytext=(3,0.1), textcoords="offset points",
+                        arrowprops=dict(arrowstyle="-"), fontsize=12)
+            Ax.annotate("90%", xy=(Wind90[np.where(lat==31)[0][0]],31),
+                        xytext=(3,0.1), textcoords="offset points",
+                        arrowprops=dict(arrowstyle="-"), fontsize=12)
+            Ax.annotate("95%", xy=(Wind95[np.where(lat==31)[0][0]],31),
+                        xytext=(3,0.1), textcoords="offset points",
+                        arrowprops=dict(arrowstyle="-"), fontsize=12)
     Ax.set_ylabel("Latitude [deg]")
     Ax.set_xlabel("Wind speed [m/s]")
     Ax.set_ylim([0, 70])
@@ -108,6 +119,16 @@ if __name__ == "__main__":
         Wind95 = get_windspeed(l, 95, Alt, 355)
         Ax.fill_betweenx(Alt, Wind85, Wind95, alpha=0.5, facecolor=c)
         Ax.plot(Wind90, Alt, c, label="Latitude=%d" % l, lw=2)
+        if l == 45:
+            Ax.annotate("80%", xy=(Wind85[Alt.index(63000)], 63000),
[…]

### 2017-02-24 gassolar/environment/wind_speeds.py
change problem set up; min h altitude

@@ -129,9 +129,14 @@ if __name__ == "__main__":
             Ax.annotate("95%", xy=(Wind95[Alt.index(63000)], 63000),
                         xytext=(8,-0.1), textcoords="offset points",
                         arrowprops=dict(arrowstyle="-"), fontsize=12)
+
+    Ax.plot([0, 80], [15000, 15000], "k", lw=2)
+    Ax.fill_between(np.linspace(0, 80, 10), 12000, 15000, hatch="//", facecolor="None", edgecolor="k", linewidth=0.0)
+    Ax.text(64, 18000, "$h_{\\mathrm{min}}$", fontsize=15)
     Ax.set_ylabel("Altitude [ft]")
     Ax.set_xlabel("Wind speed [m/s]")
     Ax.set_ylim([0, 80000])
+    Ax.set_xlim([0, 80])
     Ax.grid()
     Ax.legend(loc=1, fontsize=15)
     # Ax.set_title("85%-95% Wind Speeds in Dec")

### 2017-03-10 gassolar/environment/solar_irradiance.py
update imports to reflect gpkit-modles structure

@@ -101,7 +101,7 @@ if __name__ == "__main__":
     Ax.set_xlim([0, 365])
     Ax.set_ylim([0, 12000])
     Fig.savefig("../../gassolarpaper/eirrvsmonth.pdf", bbox_inches="tight")
-    
+
     data = {}
     fig1, ax1 = plt.subplots()
     fig2, ax2 = plt.subplots()

### 2017-03-10 gassolar/gas/disp_sens.py

@@ -42,8 +42,8 @@ if __name__ == "__main__":
         sol = M.solve("mosek")
         sols.append(sol)
 
-    varnames = ["V_{wind}_Mission, Loiter, FlightSegment", "W_{pay}", "\\eta_{prop}", "BSFC_{min}", "t_Mission, Loiter", "N_{max}_Mission, AircraftLoading, WingLoading, ChordSparL"]
+    varnames = ["V_{wind}_Mission, Loiter, FlightSegment", "W_{pay}", "\\eta_{prop}", "BSFC_{min}", "t_Mission, Loiter", "N_{max}_Mission, AircraftLoading, WingLoading, GustL"]
     latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
     sens_table(sols, varnames, filename="sens.generated.tex")
-    fig, ax = plot_sens(sols[2], varnames, latns=latns)
+    fig, ax = plot_sens(M, sols[2], varnames, latns=latns)
     fig.savefig("../../gassolarpaper/gassensbar.pdf", bbox_inches="tight")

### 2017-03-10 gassolar/gas/flight_segment.py

@@ -1,6 +1,6 @@
 " flight segment model "
 from gpkit import Model, Variable, Vectorize
-from gpkitmodels.aircraft.GP_submodels.breguet_endurance import BreguetEndurance
+from gpkitmodels.GP.aircraft.mission.breguet_endurance import BreguetEndurance
 from steady_level_flight import SteadyLevelFlight
 from flight_state import FlightState
 from gassolar.environment.wind_speeds import get_windspeed

### 2017-03-10 gassolar/gassolarlat.py

@@ -4,6 +4,10 @@ from gassolar.gas.gas import Mission as Mgas
 from gassolar.environment.wind_speeds import get_windspeed
 import matplotlib.pyplot as plt
 import numpy as np
+import os
+
+FIGPATH = (os.path.abspath(__file__).replace(os.path.basename(__file__), "")
+           + "figs" + os.sep)
 
 plt.rcParams.update({'font.size':15})
 fig, ax = plt.subplots()
@@ -24,7 +28,6 @@ for a in [80, 90, 95]:
         if runagains:
             Ms = Msolar(latitude=l)
             Ms.substitutions.update({"W_{pay}": 10})
-            print Ms.substitutions["W_{pay}"]
             for vk in Ms.varkeys["p_{wind}"]:
                 Ms.substitutions.update({vk: a/100.0})
             Ms.cost = Ms["W_{total}"]

### 2017-03-10 gassolar/solar/print_sens.py

@@ -26,7 +26,7 @@ def sens_table(sols, varnames,
         f.write("\\bottomrule\n")
         f.write("\\end{longtable}")
 
-def plot_sens(sol, varnames, latns=None):
+def plot_sens(model, sol, varnames, latns=None):
     fig, ax = plt.subplots()
     pss = []
     ngs = []
@@ -34,13 +34,16 @@ def plot_sens(sol, varnames, latns=None):
     for vname, latn in zip(varnames, latns):
         sen = sol["sensitivities"]["constants"][vname]
         if hasattr(sen, "__len__"):
-            sen = sol["sensitivities"]["constants"][max(sen)]
-        sens[latn] = sen
+            vk = max(sen)
+            sen = sol["sensitivities"]["constants"][vk]
+        else:
+            vk = model[vname]
+        sens[vk] = sen
 
     labels = []
     for s in sorted(np.absolute(sens.values()), reverse=True):
         vn = [se for se in sens if abs(sens[se]) == s][0]
-        labels.append(vn)
+        labels.append(model[vn].descr["label"])
         if sens[vn] > 0:
             pss.append(sens[vn])
             ngs.append(0)
@@ -48,13 +51,12 @@ def plot_sens(sol, varnames, latns=None):
             ngs.append(abs(sens[vn]))
             pss.append(0)
 
-    ind = np.arange(0.75, len(varnames)*1.5, 1.5)
-    ax.bar(ind, pss, 1.0, color="#4D606E")
-    ax.bar(ind, ngs, 1.0, color="#3FBAC2")
-    ax.set_xlim([0.25, ind[-1]+1.5])
-    if latns:
-        ax.set_xticks(ind+0.75)
-        ax.set_xticklabels(labels, rotation=-45)
+    ind = np.arange(0.5, len(varnames) + 0.5, 1)
+    ax.bar(ind - 0.25, pss, 0.5, color="#4D606E")
+    ax.bar(ind - 0.25, ngs, 0.5, color="#3FBAC2")
+    ax.set_xlim([0.0, ind[-1]+0.5])
+    ax.set_xticks(ind)
+    ax.set_xticklabels(labels, rotation=-45, ha="left")
     ax.legend(["Positive", "Negative"])
     ax.set_ylabel("sensitivities")
     return fig, ax
@@ -82,6 +84,6 @@ if __name__ == "__main__":
[…]

### 2017-03-13 gassolar/gas/disp_sens.py
move latex to docs folder

@@ -46,4 +46,10 @@ if __name__ == "__main__":
     latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
     sens_table(sols, varnames, filename="sens.generated.tex")
     fig, ax = plot_sens(M, sols[2], varnames, latns=latns)
-    fig.savefig("../../gassolarpaper/gassensbar.pdf", bbox_inches="tight")
+
+    if len(sys.argv) > 1:
+        path = sys.argv[1]
+        fig.savefig(path + "gassensbar.pdf", bbox_inches="tight")
+    else:
+        fig.savefig("gassensbar.pdf", bbox_inches="tight")
+

### 2017-03-13 gassolar/gas/ldratios.py

@@ -4,76 +4,105 @@ import numpy as np
 from gassolar.gas.gas import Mission
 from gassolar.environment.wind_speeds import get_windspeed
 from gassolar.solar.plotting import labelLines
+import sys
 
 def return_cd(cl, re):
     cd = (0.0247*cl**2.49*re**-1.11 + 2.03e-7*cl**12.7*re**-0.338 +
           6.35e10*cl**-0.243*re**-3.43 + 6.49e-6*cl**-1.9*re**-0.681)**(1/3.72)
     return cd
 
-M = Mission()
-M.cost = 1/M["t_Mission, Loiter"]
-for e in M.varkeys["\\eta_{prop}"]:
-    M.substitutions.update({e: 0.75})
-fig, ax = plt.subplots()
-wind = get_windspeed(38, 90, 15000, 355)
-cwind = get_windspeed(38, 90, np.linspace(0, 15000, 11)[1:], 355)
-for vk in M.varkeys["V_{wind}"]:
-    if "Climb" in vk.models:
-        M.substitutions.update({vk: cwind[vk.idx[0]]})
-    else:
-        M.substitutions.update({vk: wind})
-M.substitutions.update({"MTOW": 200})
-sol = M.solve("mosek")
-re = sol("Re_Mission, Loiter, FlightSegment, AircraftPerf, WingAero")[-1]
-clm = sol("C_L_Mission, Loiter, FlightSegment, AircraftPerf, WingAero")
-cdm = sol("c_{dp}_Mission, Loiter, FlightSegment, AircraftPerf, WingAero")
-cl = np.linspace(0.2, 1.5, 100)
-cd = return_cd(cl, re)
-cl15 = cl**1.5/cd
-clmax = cl[cl15 == max(cl15)]
-cdmax = cd[cl15 == max(cl15)]
-l = ax.plot(clm, cdm, "o", mfc="None", ms=7, mew=1.5,
-            label="With Wind Constraint")
-ax.annotate("mission start", xy=(clm[0], cdm[0]), xytext=(0.87, 0.0062),
-            arrowprops=dict(arrowstyle="->"), fontsize=13)
-ax.annotate("mission end", xy=(clm[-1], cdm[-1]), xytext=(0.42, 0.0072),
-            arrowprops=dict(arrowstyle="->"), fontsize=13)
+def ld_plot(model, num):
+    model.cost = 1/model["t_Mission, Loiter"]
+    for e in model.varkeys["\\eta_{prop}"]:
+        model.substitutions.update({e: 0.75})
 
[…]

### 2017-03-13 gassolar/solar/battsolarcon.py

@@ -37,13 +37,13 @@ for lat in range(20, 44, 1):
                                solver="mosek")
             if lat % 4 == 0:
                 l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "k",
-                            label="%d" % lat)
+                            label="%d$^{\\circ}$ Lat" % lat)
                 lines.append(l[0])
                 etamax.append(max(bst.sample_at(xmin_)["cost"].magnitude))
                 midx.append(np.median(xmin_))
             elif lat % 2 == 0:
                 l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "--", c="0.5",
-                            label="%d" % lat)
+                            label="%d$^{\\circ}$ Lat" % lat)
         except RuntimeWarning:
             passing = False
 
@@ -56,4 +56,4 @@ ax.set_xlabel("Battery Specific Energy [Whr/kg]")
 ax.set_xlim([250, 500])
 ax.set_ylim([0.1, max(etamax)])
 ax.grid()
-fig.savefig("../../gassolarpaper/battsolarcontour.pdf", bbox_inches="tight")
+fig.savefig("../../docs/battsolarcontour.pdf", bbox_inches="tight")

### 2017-03-14 docs/paperin.tex
initial try and comprehensive makefile

--- 2017-02-02 gassolarpaper/gassolarpaper.tex
@@
-% `advanced_example.tex', an advanced example employing the AIAA class
-% plus other third-party LaTeX packages.
-%
-% For a bare-bones usage, see `template.tex'.
-%
-% Typical processing for PostScript (PS) output:
-%
-%  latex advanced_example
-%  bibtex advanced_example  (bibliography)
-%  makeindex -s nomencl.ist -o advanced_example.gls advanced_example.glo
-%                            (nomenclature)
-%  latex advanced_example   (repeat as needed to resolve references)
-%
-%  xdvi advanced_example    (onscreen draft display)
-%  dvips advanced_example   (postscript)
-%  gv advanced_example.ps   (onscreen display)
-%  lpr advanced_example.ps  (hardcopy)
-%
-% With the above, only Encapsulated PostScript (EPS) images can be used.
-%
-%
-%  pdflatex advanced_example
-%  bibtex advanced_example    (bibliography)
-%  makeindex -s nomencl.ist -o advanced_example.gls advanced_example.glo
-%                              (nomenclature)
-%  pdflatex advanced_example  (repeat as needed to resolve references)
-%
-%  acroread advanced_example.pdf  (onscreen display)
-%
-% If you have EPS figures, you will need to use the epstopdf script
-% to convert them to PDF because PDF is a limmited subset of EPS.
-% pdflatex accepts a variety of other image formats such as JPG, TIFF,
-% PNG, and so forth -- check the documentation for your version.
-%
-% If you do *not* specify suffixes when using the graphicx package's
-% \includegraphics command, latex and pdflatex will automatically select
-% the appropriate figure format from those available.  This allows you
-% to produce PS and PDF output from the same LaTeX source file.
-%
-% To generate a large format (e.g., 11"x17") PostScript copy for editing
-% purposes, use
-%
-%  dvips -x 1467 -O -0.65in,0.85in -t tabloid advanced_example
-%
-% For further details and support, read the Users Manual, aiaa.pdf.
-
-\documentclass[]{aiaa-tc}% insert '[draft]' option to show overfull boxes
-
- \usepackage{varioref}%  smart page, figure, table, and equation referencing
- \usepackage{wrapfig}%   wrap figures/tables in text (i.e., Di Vinci style)
- \usepackage{threeparttable}% tables with footnotes
- \usepackage{dcolumn}%   decimal-aligned tabular math columns
-  \newcolumntype{d}{D{.}{.}{-1}}
- \usepackage{nomencl}%   nomenclature generation via makeindex
-  \makeglossary
- \usepackage{amssymb,amsmath}
- \usepackage{subfigure}% subcaptions for subfigures
- \usepackage{subfigmat}% matrices of similar subfigures, aka small mulitples
- \usepackage{fancyvrb}%  extended verbatim environments
- \fvset{fontsize=\footnotesize,xleftmargin=2em}
- \usepackage{lettrine}%  dropped capital letter at beginning of paragraph
-%  \usepackage[dvips]{dropping}% alternative dropped capital package
- \usepackage[colorlinks]{hyperref}%  hyperlinks [must be loaded after dropping]
- \usepackage{float}
- \usepackage{longtable,booktabs,tabularx}
- \restylefloat{table}
- \usepackage{graphicx}
- \usepackage{caption}
- \usepackage{siunitx}
- \usepackage{multicol}
- \usepackage{indentfirst}
- \usepackage[labelfont=bf]{caption}
- \usepackage{multirow}
- \usepackage{setspace}
- % \usepackage[dvipdfmx]{graphicx} 
- % \usepackage{bmpsize}
- % \doublespacing
-
  \title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-    \caption{\textbf{Wind speeds by month at 30 degrees latitude.  Bands represent 80th, 90th and 95th percentile winds.}}
+    \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
 	\label{f:windvsmonth}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
-    \caption{\textbf{Wind speeds by altitude. Bands represent 80th, 90th and 95th percentile winds.}}
+    \caption{\textbf{Winds tend to increase up to 35$^{\circ}$ latitude. Bands represent 80th, 90th and 95th percentiles.}}
 	\label{f:altvswind}
@@
 
-\section{Geometric Programming\cite{gp}}
+\section{Geometric Programming}
 
 Sizing long-edurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions. GPs have the form
+Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
 
@@
 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
-    \caption{\textbf{Posynomial approximation to wind speed data comparison at 35$^{\circ}$ latitude. (Log space RMS error = 0.043.)}}
+    \caption{\textbf{Posynomial approximation to wind speed data at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
 	\label{f:windfitl35}
@@
 	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at $30^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
+    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
 	\label{f:lat30}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles) comparison.  (Log-space RMS error~=~0.00489.)}}
+    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
 	\label{f:JH01polar}
@@
 
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et. al\cite{fitting}.
+Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting}.
 This approximation is graphically shown in Figure~\ref{f:arctanfit} for the expected range of $V_{\text{gust}}/V \in [0, 0.7]$.
@@
 	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
-    \caption{\textbf{Monomial approximation to Equation~\ref{e:gustalpha} comparison for $V/V_{\text{gust}} \in [0,0.7]$. (Log-space RMS error = 0.0246)}}
+    \caption{\textbf{Monomial approximation to Equation~\ref{e:gustalpha}. (Log-space RMS error = 0.0246)}}
 	\label{f:arctanfit}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-    \caption{\textbf{Posynomial equation (solid lines) to XFOIL data (circles) comparison. (Log space RMS error = 0.0173)}}
+    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
 	\label{f:taildragpolar}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
-    \caption{\textbf{Power law fit to University of North Dakota data\cite{gasengine} comparison. (Log space RMS error = 0.34)}}
+    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
 	\label{f:powervsweightfit}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
-    \caption{\textbf{Representative engine performance function based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
+    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
 	\label{f:powervsweightfit}
@@
 To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes.
-Figure~\ref{latvsmtowtrade} shows this trade study evaluated at the 80th, 90th and 95th percentile wind speeds.
-Each point on the graph is a unique solution to the optimization model for the stated assumptions and contants.
+Figure~\ref{f:latvsmtowtrade} shows this trade study evaluated at the 80th, 90th and 95th percentile wind speeds.
+Each point on the graph is a unique solution to the optimization model for the stated assumptions and constants.
 The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.
@@
 	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
-    \caption{\textbf{Gas optimization model feasible for all latitudes. Next integer latitude value for each solar-electric curve is infeasible.}}
+    \caption{\textbf{Gas archicture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
     \label{f:latvsmtowtrade}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
- \caption{\textbf{Comparison of structural fraction and detailed structural models highlights trade between wing weight and air density.}}
+ \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
  \label{f:altoper}
@@
 	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio from optimum endurance ratio $\text{max}(C_L^{1.5}/C_D)$ (plus sign). Solid lines are drag polars.}}
+    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
  \label{f:polarmission}
@@
 Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design and was solved by minizing wing span. 
-Values of constants other than solar cell efficiency are given throughout the text.
-The infeasible regions and contour shapes would charge for different assumed constant values.
+Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
+The infeasible regions and contour shapes would change for different assumed constant values.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Operating at higher latitudes requires both high performing batteries and solar cells.}}
+ \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
  \label{f:solarcontours}
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmin.pdf}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmax.pdf}}
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmin.pdf}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmax.pdf}}
  \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=\num{3d5}$.}}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
  \label{f:jhcps}

### 2017-03-14 gassolar/environment/print_fits.py

@@ -1,24 +1,29 @@
 import pandas as pd
+import sys
 
-df = pd.read_csv("windaltfitdata.csv")
-PRINT = False
+def fits_table(df, filename):
+    with open(filename, "w") as f:
+        f.write("\\begin{longtable}{lccccccccccccc}\n")
+        f.write("\\caption{Wind Fit Data}\\\\\n")
+        f.write("\\toprule\n")
+        f.write("\\toprule\n")
+        f.write("\\label{t:windvals}\n")
+        f.write("Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$& $c_3$ & $e_{3,1}$ & $e_{3,2}$& $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $\\alpha$\\\\\n")
+        f.write("\\midrule\n")
+        for i in range(1, len(df)+1):
+            d = [df[i-1:i][n].iloc[0] for n in df[i-1:i]]
+            vals = " & ".join(["%.3g" % v for v in d[1:]])
+            f.write(vals + "\\\\\n")
+        f.write("\\bottomrule\n")
+        f.write("\\end{longtable}")
 
-if PRINT:
-    filename = "../../gassolarpaper/windfitdata.generated.tex"
-else:
-    filename = "windfitdata.generated.tex"
+if __name__ == "__main__":
+    df = pd.read_csv("windaltfitdata.csv")
 
-with open(filename, "w") as f:
-    f.write("\\begin{longtable}{lccccccccccccc}\n")
-    f.write("\\caption{Wind Fit Data}\\\\\n")
-    f.write("\\toprule\n")
-    f.write("\\toprule\n")
-    f.write("\\label{t:windvals}\n")
-    f.write("Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$& $c_3$ & $e_{3,1}$ & $e_{3,2}$& $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $\\alpha$\\\\\n")
-    f.write("\\midrule\n")
-    for i in range(1, len(df)+1):
-        d = [df[i-1:i][n].iloc[0] for n in df[i-1:i]]
-        vals = " & ".join(["%.3g" % v for v in d[1:]])
-        f.write(vals + "\\\\\n")
-    f.write("\\bottomrule\n")
-    f.write("\\end{longtable}")
[…]

### 2017-03-14 gassolar/environment/solar_irradiance.py

@@ -2,8 +2,11 @@
 import numpy as np
 from numpy import sin, tan, cos, arccos, deg2rad
 import matplotlib.pyplot as plt
+from gpfit.fit import fit
+import sys
 import pandas as pd
 plt.rcParams.update({'font.size':15})
+GENERATE = False
 
 def get_Eirr(latitude, day, N=50.0):
     """
@@ -42,6 +45,11 @@ def get_Eirr(latitude, day, N=50.0):
     return E, tday, tnight, plot
 
 if __name__ == "__main__":
+    if len(sys.argv) > 1:
+        path = sys.argv[0]
+    else:
+        path = ""
+
     ES, td, tn, p = get_Eirr(30, 355, N=1000)
     fig, ax = plt.subplots()
     ax.fill_between([-12, -td/2], 0, 80, alpha=0.3, facecolor="b",
@@ -75,7 +83,7 @@ if __name__ == "__main__":
     ax2 = ax.twinx()
     ax2.set_yticks([0, 80, 900])
     ax2.set_yticklabels(["", "$(P/S)_{\mathrm{min}}$", ""])
-    fig.savefig("../../gassolarpaper/lat30.pdf", bbox_inches="tight")
+    fig.savefig(path + "lat30.pdf", bbox_inches="tight")
 
     # solar irradiance by year
     Fig, Ax = plt.subplots()
@@ -100,13 +108,17 @@ if __name__ == "__main__":
     Ax.grid()
     Ax.set_xlim([0, 365])
     Ax.set_ylim([0, 12000])
-    Fig.savefig("../../gassolarpaper/eirrvsmonth.pdf", bbox_inches="tight")
+    Fig.savefig(path + "eirrvsmonth.pdf", bbox_inches="tight")
 
     data = {}
     fig1, ax1 = plt.subplots()
     fig2, ax2 = plt.subplots()
-    # for l in range(20, 61):
-    for l, col in zip([30], ["g"]):
+
+    if GENERATE:
+        lat = range(20, 61)
+    else:
+        lat = [30]
+    for l in lat:
 
         ES, td, tn, p = get_Eirr(l, 355)
         params = [l]
@@ -125,8 +137,8 @@ if __name__ == "__main__":
         cn, rm = fit(x, y, 1, "MA")
         print "RMS error: %.4f" % rm
         yfit = cn.evaluate(x)
[…]

### 2017-03-14 gassolar/environment/wind_fit.py

@@ -2,6 +2,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas as pd
+import sys
 from gassolar.environment.wind_speeds import get_windspeed, interpolate
 from gpfit.fit import fit
 plt.rc("text", usetex=True)
@@ -91,6 +92,11 @@ def plot_fits(xdata, ydata, yfit, latitude):
 
 if __name__ == "__main__":
 
+    if len(sys.argv) > 1:
+        path = sys.argv[0]
+    else:
+        path = ""
+
     if GENERATE:
         latitude = range(20, 61, 1)
     else:
@@ -151,8 +157,7 @@ if __name__ == "__main__":
                 print "RMS Error: %.3f, Alt iter=%d" % (rm, j)
         fig, ax = plot_fits(X, Y, yfit, l)
         if not GENERATE:
-            fig.savefig("../../gassolarpaper/windfitl%d.pdf" % l,
-                        bbox_inches="tight")
+            fig.savefig(path + "windfitl%d.pdf" % l, bbox_inches="tight")
         plt.close()
 
     if GENERATE:

### 2017-03-14 gassolar/environment/wind_speeds.py

@@ -2,6 +2,7 @@
 import pandas as pd
 import numpy as np
 import os
+import sys
 import matplotlib.pyplot as plt
 plt.rcParams.update({'font.size':15})
 
@@ -81,6 +82,10 @@ def interpolate(xs, ys, x):
     return y
 
 if __name__ == "__main__":
+    if len(sys.argv) > 1:
+        path = sys.argv[0]
+    else:
+        path = ""
 
     Fig, Ax = plt.subplots()
     # Colors = ["#253B6E", "#1F5F8B", "#1891AC"]
@@ -108,7 +113,7 @@ if __name__ == "__main__":
     Ax.grid()
     Ax.legend(loc=2, fontsize=15)
     # Ax.set_title("85%-95% Wind Speeds in Dec")
-    Fig.savefig("../../gassolarpaper/latvswind.pdf", bbox_inches="tight")
+    Fig.savefig(path + "latvswind.pdf", bbox_inches="tight")
 
     Fig, Ax = plt.subplots()
     Alt = range(1000, 80000, 1000)
@@ -140,7 +145,7 @@ if __name__ == "__main__":
     Ax.grid()
     Ax.legend(loc=1, fontsize=15)
     # Ax.set_title("85%-95% Wind Speeds in Dec")
-    Fig.savefig("../../gassolarpaper/altvswind.pdf", bbox_inches="tight")
+    Fig.savefig(path + "altvswind.pdf", bbox_inches="tight")
 
     Fig, Ax = plt.subplots()
     Colors = ["#1C226B", "#3E31AE", "#4AA9AF"]
@@ -175,4 +180,4 @@ if __name__ == "__main__":
     Ax.grid()
     Ax.legend(loc=1, fontsize=15)
     # Ax.set_title("85%-95% Wind Speeds at 45 deg Lat")
-    Fig.savefig("../../gassolarpaper/windvsmonth.pdf", bbox_inches="tight")
+    Fig.savefig(path + "windvsmonth.pdf", bbox_inches="tight")

### 2017-03-14 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -2,6 +2,7 @@
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
+import sys
 plt.rcParams.update({'font.size':15})
 
 def text_to_df(filename):
@@ -77,4 +78,8 @@ if __name__ == "__main__":
     Re = range(200, 750, 50)
     X, Y = fit_setup(Re) # call fit(X, Y, 4, "SMA") to get fit
     F, A = plot_fits([300, 350, 400, 450, 500])
-    F.savefig("../../../gassolarpaper/jho1polarfit1.pdf", bbox_inches="tight")
+    if len(sys.argv) > 1:
+        path = sys.argv[0]
+        F.savefig(path + "jho1polarfit1.pdf", bbox_inches="tight")
+    else:
+        F.savefig(path + "jho1polarfit1.pdf", bbox_inches="tight")

### 2017-03-14 gassolar/fits/tailpolars/naca_cl0fits.py

@@ -89,5 +89,9 @@ if __name__ == "__main__":
     NACA = ["0005", "0008", "0009", "0010", "0015", "0020"]
     X, Y = fit_setup(NACA, Re) # call fit(X, Y, 4, "SMA") to get fit
     F, A = plot_fits(NACA, Re)
-    F.savefig("../../../gassolarpaper/taildragpolar.pdf",
-              bbox_inches="tight")
+    if len(sys.argv) > 1:
+        path = sys.argv[0]
+        F.savefig(path + "taildragpolar.pdf", bbox_inches="tight")
+    else:
+        F.savefig("taildragpolar.pdf", bbox_inches="tight")
+

### 2017-03-14 gassolar/gas/disp_sens.py

@@ -1,5 +1,6 @@
 "print sensitivities"
 import numpy as np
+import sys
 from gassolar.gas.gas import Mission
 from gassolar.environment.wind_speeds import get_windspeed
 from gassolar.solar.print_sens import plot_sens

### 2017-03-14 gassolar/gas/fits/DF70/fitDF70.py

@@ -1,5 +1,6 @@
 import pandas as pd
 import numpy as np
+import sys
 from numpy import logspace, log, log10
 import matplotlib.pyplot as plt
 from gpfit.fit import fit
@@ -8,111 +9,111 @@ plt.rcParams.update({'font.size':19})
 
 np.random.seed(0)
 
-# Fitting BSFC vs. Power
-df = pd.read_csv('Dataset_Power_Kw.csv')
-p = df['P']
-rpm = df["RPM"]
-f = interpolate.interp1d(rpm, p)
-df = pd.read_csv('Dataset_BSFC_kgKwh.csv')
-df = df[(df["RPM"] > min(rpm)) & (df["RPM"] < max(rpm))]
-rpmnew = df["RPM"]
-u = f(rpmnew)/max(p)
-w = df['BSFC']/min(df["BSFC"])
-x = np.array(log(u))
-y = np.array(log(w))
-Type = 'SMA'
-K = 2
+def plot_BSFCtoPower():
+    # Fitting BSFC vs. Power
+    df = pd.read_csv('Dataset_Power_Kw.csv')
+    p = df['P']
+    rpm = df["RPM"]
+    f = interpolate.interp1d(rpm, p)
+    df = pd.read_csv('Dataset_BSFC_kgKwh.csv')
+    df = df[(df["RPM"] > min(rpm)) & (df["RPM"] < max(rpm))]
+    rpmnew = df["RPM"]
+    u = f(rpmnew)/max(p)
+    w = df['BSFC']/min(df["BSFC"])
+    x = np.array(log(u))
+    y = np.array(log(w))
+    Type = 'SMA'
+    K = 2
 
-cstrt, rmserror = fit(x, y, K, Type)
-print "RMS error = %.4f" % rmserror
-yfit = cstrt.evaluate(x)
+    cstrt, rmserror = fit(x, y, K, Type)
+    print "RMS error = %.4f" % rmserror
+    yfit = cstrt.evaluate(x)
 
-fig, ax = plt.subplots()
-ax.plot(u*100, w*min(df["BSFC"]), "o", mfc="None", ms=7, mew=1.5)
-ax.plot(u*100, np.exp(yfit)*min(df["BSFC"]), linewidth=2)
-ax.set_xlabel("Percent Power")
-ax.set_ylabel("$BSFC$ [kg/kW/hr]")
[…]

### 2017-03-14 gassolar/gas/fits/powerlaw/engine_law.py

@@ -7,22 +7,31 @@ plt.rcParams.update({'font.size':15})
 WEIGHT = 10.0
 POWER = 10.0
 
-df = pd.read_csv("powervsweight.csv")
-u = np.array(df["lbs"]/WEIGHT)
-w = np.array(df["hp"]/POWER)
+def plot_powerlaw()
+    df = pd.read_csv("powervsweight.csv")
+    u = np.array(df["lbs"]/WEIGHT)
+    w = np.array(df["hp"]/POWER)
 
-x = np.log(u)
-y = np.log(w)
+    x = np.log(u)
+    y = np.log(w)
 
-cn, rm = fit(x, y, 1, "MA")
-weight = np.linspace(min(df["lbs"]), max(df["lbs"]), 100)
-yfit = cn.evaluate(np.log(weight/WEIGHT))
+    cn, rm = fit(x, y, 1, "MA")
+    weight = np.linspace(min(df["lbs"]), max(df["lbs"]), 100)
+    yfit = cn.evaluate(np.log(weight/WEIGHT))
 
-fig, ax = plt.subplots()
-ax.plot(u*WEIGHT, w*POWER, "o", mfc="None", mew=1.5)
-ax.plot(weight, np.exp(yfit)*POWER, lw=2)
-ax.set_xlabel("Engine Weight [lbs]")
-ax.set_ylabel("Maximum Shaft Power [hp]")
-ax.legend(["UND Engine Data", "Power Law Fit"], loc=2)
-ax.grid()
-fig.savefig("../../gassolarpaper/powervsweightfit.pdf", bbox_inches="tight")
+    fig, ax = plt.subplots()
+    ax.plot(u*WEIGHT, w*POWER, "o", mfc="None", mew=1.5)
+    ax.plot(weight, np.exp(yfit)*POWER, lw=2)
+    ax.set_xlabel("Engine Weight [lbs]")
+    ax.set_ylabel("Maximum Shaft Power [hp]")
+    ax.legend(["UND Engine Data", "Power Law Fit"], loc=2)
+    ax.grid()
+
+if __name__ == "__main__":
+    fig, ax = plot_powerlaw()
+    if len(sys.argv) > 1:
+        path = sys.argv[1]
+        fig.savefig(path + "powervsweightfit.pdf", bbox_inches="tight")
+    else:
+        fig.savefig("powervsweightfit.pdf", bbox_inches="tight")
+

### 2017-03-14 gassolar/gas/flight_segment.py

@@ -4,6 +4,7 @@ from gpkitmodels.GP.aircraft.mission.breguet_endurance import BreguetEndurance
 from steady_level_flight import SteadyLevelFlight
 from flight_state import FlightState
 from gassolar.environment.wind_speeds import get_windspeed
+from gpkit.constraints.tight import Tight as TCS
 
 class FlightSegment(Model):
     "flight segment"

### 2017-03-14 gassolar/gas/ldratios.py

@@ -92,14 +92,14 @@ def ld_plot(model, num):
     return fig, ax
 
 if __name__ == "__main__":
+    M = Mission()
 
     nums = [0, 1, 2]
     if len(sys.argv) > 1:
         path = sys.argv[1]
         for num in nums:
-            M = Mission()
             fig, ax = ld_plot(M, num)
-            fig.savefig(path + "polarmission%d.pdf" % num)
+            fig.savefig(path + "polarmission%d.pdf" % num, bbox_inches="tight")
     else:
         for num in nums:
             M = Mission()

### 2017-03-14 gassolar/solar/battsolarcon.py

@@ -1,59 +1,70 @@
 " contour plots "
 import matplotlib.pyplot as plt
 import numpy as np
+import sys
 from solar import Mission
 from plotting import windalt_plot, labelLines
 from gpkit.tools.autosweep import autosweep_1d
 
 N = 100
 plt.rcParams.update({'font.size':15})
-fig, ax = plt.subplots()
 
-etamax = []
-lines = []
-midx = []
-passing = True
-for lat in range(20, 44, 1):
-    if passing:
-        hmax = 500
-        M = Mission(latitude=lat)
-        M.substitutions.update({"W_{pay}": 10})
-        for vk in M.varkeys["\\eta_{prop}"]:
-            M.substitutions.update({vk: 0.75})
-        for vk in M.varkeys["p_{wind}"]:
-            M.substitutions.update({vk: 90/100.0})
-        del M.substitutions["h_{batt}"]
-        M.substitutions.update({"\\eta_Mission, Aircraft, SolarCells": 0.4})
-        M.cost = M["h_{batt}"]
-        sol = M.solve("mosek")
-        hmin = sol["cost"].magnitude + 1e-3
-        del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
-        M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
-        xmin_ = np.linspace(hmin, hmax, 100)
-        tol = 0.01
-        notpassing = True
-        try:
-            bst = autosweep_1d(M, tol, M["h_{batt}"], [hmin, hmax],
-                               solver="mosek")
-            if lat % 4 == 0:
-                l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "k",
-                            label="%d$^{\\circ}$ Lat" % lat)
-                lines.append(l[0])
-                etamax.append(max(bst.sample_at(xmin_)["cost"].magnitude))
-                midx.append(np.median(xmin_))
-            elif lat % 2 == 0:
-                l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "--", c="0.5",
-                            label="%d$^{\\circ}$ Lat" % lat)
-        except RuntimeWarning:
-            passing = False
+def plot_battsolarcon():
+    fig, ax = plt.subplots()
 
-# ax.fill_between(xmin_, bst.sample_at(xmin_)["cost"], max(etamax),
-#                 edgecolor="r", lw=2, hatch="/", facecolor="None", zorder=100)
[…]

### 2017-03-14 gassolar/solar/fits/arctan/arctan_fit.py

@@ -4,23 +4,31 @@ import matplotlib.pyplot as plt
 import numpy as np
 plt.rcParams.update({'font.size':15})
 
-u = np.linspace(0.01, 0.7, 100)
-w = np.arctan(u)
+def arctanfit():
+    u = np.linspace(0.01, 0.7, 100)
+    w = np.arctan(u)
 
-x = np.log(u)
-y = np.log(w)
+    x = np.log(u)
+    y = np.log(w)
 
-cn, rm = fit(x, y, 1, "MA")
-print "RMS error: %.4f" % rm
+    cn, rm = fit(x, y, 1, "MA")
+    print "RMS error: %.4f" % rm
 
-yfit = evaluate_fit(cn, x, "MA")
-fig, ax = plt.subplots()
-ax.plot(u, w, lw=2)
-ax.plot(u, np.exp(yfit), "--", lw=2)
-ax.set_xlim([0, 0.7])
-ax.grid()
-ax.set_xlabel("$V_{\\mathrm{gust}}/V$")
-ax.set_ylabel("$\\alpha_{\\mathrm{gust}}$")
-ax.legend(["$\\arctan{(V_{\\mathrm{gust}}/V)}$",
-           "$0.905 (V_{\\mathrm{gust}}/V)^{0.961}$"], loc=2, fontsize=15)
-fig.savefig("../../../../gassolarpaper/arctanfit.pdf", bbox_inches="tight")
+    yfit = evaluate_fit(cn, x, "MA")
+    fig, ax = plt.subplots()
+    ax.plot(u, w, lw=2)
+    ax.plot(u, np.exp(yfit), "--", lw=2)
+    ax.set_xlim([0, 0.7])
+    ax.grid()
+    ax.set_xlabel("$V_{\\mathrm{gust}}/V$")
+    ax.set_ylabel("$\\alpha_{\\mathrm{gust}}$")
+    ax.legend(["$\\arctan{(V_{\\mathrm{gust}}/V)}$",
+               "$0.905 (V_{\\mathrm{gust}}/V)^{0.961}$"], loc=2, fontsize=15)
+    return fig, ax
+
+if __name__ == "__main__":
+    fig, ax = arctanfit()
+    if len(sys.argv) > 1:
+        fig.savefig(path + "arctanfit.pdf", bbox_inches="tight")
+    else:
+        fig.savefig("arctanfit.pdf", bbox_inches="tight")

### 2017-03-14 gassolar/solar/objectivecomp.py

@@ -1,66 +1,72 @@
 " change objective plot "
 import matplotlib.pyplot as plt
 import numpy as np
+import sys
 from gassolar.solar.solar import Mission
 
-LATITUDE = False
-WIND = False
-CON = True
-COMP = False
-SENS = False
-
-plt.rcParams.update({'font.size':15})
-fig, ax = plt.subplots()
-fig2, ax2 = plt.subplots()
-lat = np.arange(20, 40, 1)
-l1 = []
-l2 = []
-for obj in ["b_Mission, Aircraft, Wing", "S_Mission, Aircraft, SolarCells"]:
-    W = []
-    SS = []
-    runagain = True
-    for l in lat:
-        if runagain:
-            M = Mission(latitude=l)
-            M.substitutions.update({"W_{pay}": 10})
-            for vk in M.varkeys["p_{wind}"]:
-                M.substitutions.update({vk: 90/100.0})
-            M.cost = M[obj]
-            try:
-                sol = M.solve("mosek")
-                W.append(sol("b_Mission, Aircraft, Wing").magnitude)
-                SS.append(sol("S_Mission, Aircraft, SolarCells").magnitude)
-            except RuntimeWarning:
+def plot_objectivecomp(path=None):
+    plt.rcParams.update({'font.size':15})
+    fig, ax = plt.subplots()
+    fig2, ax2 = plt.subplots()
+    lat = np.arange(20, 40, 1)
+    l1 = []
+    l2 = []
+    for obj in ["b_Mission, Aircraft, Wing", "S_Mission, Aircraft, SolarCells"]:
+        W = []
+        SS = []
+        runagain = True
+        for l in lat:
+            if runagain:
+                M = Mission(latitude=l)
+                M.substitutions.update({"W_{pay}": 10})
+                for vk in M.varkeys["p_{wind}"]:
+                    M.substitutions.update({vk: 90/100.0})
+                M.cost = M[obj]
+                try:
+                    sol = M.solve("mosek")
+                    W.append(sol("b_Mission, Aircraft, Wing").magnitude)
+                    SS.append(sol("S_Mission, Aircraft, SolarCells").magnitude)
+                except RuntimeWarning:
[…]

### 2017-03-14 gassolar/solar/print_sens.py

@@ -2,6 +2,7 @@
 from gassolar.solar.solar import Mission
 import matplotlib.pyplot as plt
 import numpy as np
+import sys
 
 def sens_table(sols, varnames,
                filename="../../gassolarpaper/solarsens.generated.tex"):
@@ -85,5 +86,9 @@ if __name__ == "__main__":
               "$\\eta_{\\mathrm{prop}}$"]
     sens_table(sols, varns, filename="test.tex")
     fig, ax = plot_sens(M, sols[3], varns, latns=latns)
-    fig.savefig("../../gassolarpaper/solarsensbar.pdf", bbox_inches="tight")
+    if len(sys.argv) > 1:
+        path = sys.argv[1]
+        fig.savefig(path + "solarsensbar.pdf", bbox_inches="tight")
+    else:
+        fig.savefig("solarsensbar.pdf", bbox_inches="tight")
 

### 2017-03-14 gassolar/solar/windstruct.py

@@ -1,24 +1,46 @@
 "wind structural model comparison"
-from gassolar.solar.solar_simple.solarsimple import Mission
 from gassolar.solar.plotting import windalt_plot
 import matplotlib.pyplot as plt
+import sys
 
-lat = 29
-M = Mission(latitude=lat)
-for vk in M.varkeys["CDA_0"]:
-    M.substitutions.update({vk: 0.002})
-M.cost = M["W"]
-sol1 = M.solve("mosek")
-from gassolar.solar.solar import Mission
-M = Mission(latitude=lat)
-M.cost = M["W_{total}"]
-sol2 = M.solve("mosek")
-fig, ax = windalt_plot(lat, sol1, sol2)
-# fig, ax = windalt_plot(lat, sol1=sol2)
-ax.annotate("structural weight fraction", xy=(66.2, 30), xytext=(46, 10),
-            arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
-                            headwidth=10))
-ax.annotate("detailed structural model", xy=(58, 50), xytext=(61, 65),
-            arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
-                            headwidth=10))
-fig.savefig("../../gassolarpaper/windaltoper.pdf", bbox_inches="tight")
+def plot_structm(num, lat=29):
+    if num == 0:
+        fig, ax = windalt_plot(lat)
+    elif num == 1:
+        from gassolar.solar.solar import Mission
+        M = Mission(latitude=lat)
+        M.cost = M["W_{total}"]
+        sol2 = M.solve("mosek")
+        fig, ax = windalt_plot(lat, sol1=sol2)
+        ax.annotate("detailed structural model", xy=(58, 50), xytext=(61, 65),
+                    arrowprops=dict(facecolor='black', shrink=0.05, width=1.5,
+                                    headwidth=10))
+    elif num == 2:
+        from gassolar.solar.solar_simple.solarsimple import Mission
+        M = Mission(latitude=lat)
+        for vk in M.varkeys["CDA_0"]:
+            M.substitutions.update({vk: 0.002})
+        M.cost = M["W"]
+        sol1 = M.solve("mosek")
+        from gassolar.solar.solar import Mission
+        M = Mission(latitude=lat)
+        M.cost = M["W_{total}"]
+        sol2 = M.solve("mosek")
+        fig, ax = windalt_plot(lat, sol1, sol2)
+        ax.annotate("structural weight fraction", xy=(66.2, 30), xytext=(46, 10),
[…]

### 2017-04-17 docs/paperin_review.tex
add new wind fitted files; add north and south wind data; generate new wind fit file; add weight breakdown and configuration drawings to paper; reorder paper

--- 2017-03-14 docs/paperin.tex
@@
+\usepackage{tikz}
 \title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
@@
 
+\graphicspath{{./figs/}} 
 \maketitle
@@
 Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-Additionally, because few long-endurance gas and solar-electric powered aircraft have been built, the operational limits of these aircraft are relatively unknown. 
-Even more unclear are the trade offs between gas powered and solar-electric powered architectures.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to explore this design space and highlight trade offs between the two proposed architectures.
+The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures even more unclear.
+
+This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two archictures for a given mission.
 Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
@@
 
-\section{Requirements Analysis}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
+\section{Mission Requirements}
+
+A comparision of the gas and solar-elecrtric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements, shown in Table~\ref{t:mreqs}.
+These requirements can and will be changed in the results section to observe which archicture is best suited to meet various sets of requirements.
+A discussion of each requirement is included in this section.
+
+\begin{longtable}{lccccccccccccc}
+\caption{Mission Requirements}\\
+\toprule
+\toprule
+\label{t:mreqs}
+Payload & 10 lbs\\
+Endurance & 6 days\\
+Season & all seasons\\
+Altitude & 4,600 m\\
+Latitude & $\pm30^{\circ}$\\
+\bottomrule
+\end{longtable}
+
+\subsection{Payload}
+
+A 10 lbs payload is a baseline requirement for both the gas and solar-electric powered aircraft.
+This requirement can be changed to observe how each architeure is able to handle larger or smaller payloads.
+
+Inherent in a payload requirement is the ability to station keep. To station keep an aircraft must fly at least as fast as the wind speed,
 
@@
 For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can thoretically fly at any time of the year.  
-Therefore, the endurance requirement for solar-electric powered aircraft is assumed to be a requirement to fly during the winter solstice.  
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, thus more than satisfying the 6 day endurance requirement.
+Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
 To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-Because the endurance requirement for solar-electric powered aircraft is set by the winter solstice and because wind speeds are higher during winter months, only wind speed data from December will be used for the aircraft sizing and performance analysis presented in this paper.
+Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June will be used for the aircraft sizing and performance analysis presented in this paper.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Winds peak at summer and winter solstice in Southern and Northern Hemispheres respectively.\cite{wind}}}
+ \label{f:windvsmonth}
+\end{figure}
+
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
+%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
+% 	\label{f:windvsmonth}
+% 	\end{center}
+% \end{figure}
+
+\subsection{Altitude}
+
+The altitude requirement is generally set by a minimum payload required to meet a certain coverage
+
+\begin{equation}
+ h \geq h_{\text{min}}.
+\end{equation}
+
+Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswind} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
+Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m depending on the engine.
+Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-    \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-	\label{f:windvsmonth}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
+    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
+	\label{f:altvswindhmin}
 	\end{center}
@@
 
-\subsection{Altitude}
-
-Gas powered aircraft generally fly between 5,000 and 25,000 ft due to payload requirements.  
-Because of the limitations of naturally aspirated engines and linearly increasing wind speeds from 0 to 35,000 ft shown in Figure~\ref{f:altvswind}, gas powered aircraft will generally be more efficient at lower altitudes and generally fly at a given altitude specified by the mission.
-
-Solar-electric powered aircraft tend to fly around 60,000 ft to meet payload requirements, avoid cloud coverage and fly in lower wind speeds which have a minimum around 60,000 shown in Figure~\ref{f:altvswind}.
-For the solar-electric powered aircraft altitude is optimized as explained in Section IV.A.
+\subsection{Latitude}
+
+It is assumed that long-endurance aircraft will have a requirement to fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
+For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude. 
+Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
+Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
-    \caption{\textbf{Winds tend to increase up to 35$^{\circ}$ latitude. Bands represent 80th, 90th and 95th percentiles.}}
-	\label{f:altvswind}
-	\end{center}
-\end{figure}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\subsection{Latitude}
-
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds. }}
+	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
+    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{winds} }}
 	\label{f:latvswind}
@@
 
-It is assumed that long-endurance aircraft will have a requirement to fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude.
-
-\subsection{Percentile Wind Speed}
-
-For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
- \label{f:bostonwinds}
-\end{figure}
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
- \label{f:romewinds}
-\end{figure}
-
-This analysis shows high wind speeds can last for days at a time.  
-An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
+% \subsection{Percentile Wind Speed}
+% 
+% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
+% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
+% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
+%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
+%  \label{f:bostonwinds}
+% \end{figure}
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
+%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
+%  \label{f:romewinds}
+% \end{figure}
+% 
+% This analysis shows high wind speeds can last for days at a time.  
+% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
+% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
 
@@
 Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}  
-To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} 
+To evaulate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
+Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
+Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
+The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.8\textwidth,natwidth=480,natheight=200]{flowchart.pdf}
+    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
+	\label{f:flowchart}
+	\end{center}
+\end{figure}
+
 A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
 A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-\\
-
-\section{Aircraft Sizing Models}
+Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gas_solar_trade} \\
+
+\section{Configuration}
+
+Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
+Both aircraft have a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
+The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
+The solar-electric powered aircraft holds the batteries in the wings.  
+It is assummed that the added weight and complexity of wet wings outweights the drag and weight penalty of a fuselage for the gas powered archicture.  
+The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
+A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{simpleaircraft.pdf}
+    \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
+	\label{f:simpleaircraft}
+	\end{center}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.7\textwidth,natwidth=358,natheight=334]{wbreak.pdf}
+   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
+	\label{f:wbreak}
+	\end{center}
+\end{figure}
+
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      % \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=1.0\textwidth,natwidth=144,natheight=320]{solarwbreak.pdf}}
+%      % \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=1.0\textwidth,natwidth=178,natheight=320]{gaswbreak.pdf}}
+%      \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=0.30\textwidth]{solarwbreak.pdf}}
+%      \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=0.40\textwidth]{gaswbreak.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
+%  \label{f:wbreak}
+% \end{figure}
+
+These assumptions can be changed to evaulate different configuration.  Changes in configuration assumptions may require modification of models or equations.
+
+\section{Aircraft Physical Models}
 
@@
 
-\subsection{Wind Speeds}
+\subsection{Solar-Electric Aircraft Physical Models}
+
+\subsubsection{Wind Speeds}
 
 There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 63,000 ft and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
+The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
 To accomplish this, constraints relating wind speed to altitude are imposed. 
@@
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
-Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
-In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
-The constraint on December wind speeds and air density at the 35 degree latitude is
-
-\begin{align}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.213} &\geq 2918\rho^{7.587}p_{\text{wind}}^{9.37} + \num{1.58d-10}\rho^{-6.074}p_{\text{wind}}^{60.421} \nonumber \\
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
+I n total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
+The constraint on December wind speeds and air density at the 30 degree latitude is
+
+\begin{align}
+    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
     \label{e:windfit35}
-    &+ \num{4.05d-16}\rho^{-9.879}p_{\text{wind}}^{13.204} + 1048\rho^{6.33}p_{\text{wind}}^{46.232}
+    &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
 \end{align}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
     \caption{\textbf{Posynomial approximation to wind speed data at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
@@
 Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase nearly linearly up to 30,000 ft.  
+Additionally, wind speeds increase nearly linearly up to 9,100 m.  
 Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h=15,000$ ft, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsection{Solar-Electric Power}
+For this sizing study, the altitude requirement for the gas powered aircraft is $h=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
+
+\subsubsection{Solar-Electric Power}
 
@@
 The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
- \label{f:energyapprox}
-\end{figure}
+% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
+%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
+%  \label{f:energyapprox}
+% \end{figure}
 
@@
 
-\subsection{Breguet Endurance}
+\subsection{Gas Powered Aircraft Physical Models}
+
+\subsubsection{Breguet Endurance}
 
@@
 
+\subsubsection{Elliptcal Fuselage for Gas Powered Aircraft}
+
+For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
+The fuselage will increase the overall weight and drag of the aircraft.
+The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
+
+The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
+
+\begin{equation}
+    \label{e:fusevol}
+    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
+\end{equation}
+
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
+
+\begin{equation}
+    \label{e:fusevol2}
+    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
+\end{equation}
+
+where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
+Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
+
+\begin{equation}
+    \label{e:fusesa}
+    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
+\end{equation}
+
+The weight of the fuselage is constrained by
+
+\begin{equation}
+    \label{e:fuseweight}
+    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
+
+\begin{align}
+    \label{e:fusedrag}
+    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
+    C_f &\geq \frac{0.455}{Re^{0.3}}
+\end{align}
+
+where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
+
+\begin{equation}
+    \label{e:fuseform}
+    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
+\end{equation}
+
+\subsubsection{Engine Weight}
+
+The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
+
+\begin{equation}
+    \label{e:electricengine}
+    P_{\text{max}} = B_{PM} m_{\text{motor}}
+\end{equation}
+
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
+
+The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
+
+\begin{equation}
+    \label{e:powerlaw}
+    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
+\end{equation}
+
+where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].
+Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg et.
+al\cite{fitting}.
+Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
+    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
+	\label{f:powervsweightfit}
+	\end{center}
+\end{figure}
+
+\subsubsection{Gas Powered Engine Performance}
+
+Two characteristics of gas engines affect the performance of long-endurance aircraft.  
+The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
+The second is the lapse rate.  
+Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
+To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
+RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
+It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
+This engine performance model is only valid for internal combustion engines.
+
+To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
+
+\begin{equation}
+    \label{e:rpmtobsfc}
+    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
+\end{equation}
+
+where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
+Using this approximation the required shaft power determines the $\text{BSFC}$.
+A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
+A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
+    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
+	\label{f:powervsweightfit}
+	\end{center}
+\end{figure}
+
+The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
+
+\begin{equation}
+    \label{e:lapse}
+    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
+\end{equation}
+
+The lapse is calculated from the required flight altitude $h=15,000$ [ft], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 1000 ft,\cite{enginelapse}
+
+\begin{equation}
+    \label{e:lapsefit}
+    L_{\text{eng}}(h) = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
+\end{equation}
+
+\subsubsection{Climb Constraints}
+
+Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
+The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
+
+\begin{equation}
+    \label{e:climbrate}
+    \dot{h} \geq 100 \text{ [ft/min]}.
+\end{equation}
+
+The climb rate affects the required thrust, and therefore the required power during climb,
+
+\begin{equation}
+    \label{e:climb}
+    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}}{V}
+\end{equation}
+
+where $W$ is the weight of the aircraft during climb.
+
+\subsection{Shared Physical Models}
+
+The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
+
 \subsection{Steady Level Flight}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.pdf}
     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
@@
 
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting}.
-This approximation is graphically shown in Figure~\ref{f:arctanfit} for the expected range of $V_{\text{gust}}/V \in [0, 0.7]$.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
-    \caption{\textbf{Monomial approximation to Equation~\ref{e:gustalpha}. (Log-space RMS error = 0.0246)}}
-	\label{f:arctanfit}
-	\end{center}
-\end{figure}
+Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
+
+\begin{equation}
+    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
+\end{equation}
+
+and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
 
@@
 	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
+	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capsparcross.pdf}
     \caption{\textbf{Cross sectional view of a cap spar.}}
@@
 
-The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 570$ [MPa].\cite{carbonfiber}
+The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{carbonfiber}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
 Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the horizontal and vertical tails is
-
-\begin{align}
-    \label{e:htweight}
-    W_{\text{h}}/m_{\text{fac}} &= \rho_{\text{foam}} \frac{S_{\text{h}}^2}{b_{\text{h}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{h}} \\
-    \label{e:vtweight}
-    W_{\text{v}}/m_{\text{fac}} &= \rho_{\text{foam}} \frac{S_{\text{v}}^2}{b_{\text{v}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{v}}
-\end{align}
+The weight of the tails is
+
+\begin{equation}
+    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
+\end{equation}
 
@@
 \begin{align}
-    D_{\text{h}} &\geq \frac{1}{2} c_{d_{\text{h}}} \rho V^2 S_{\text{h}} \\
-    D_{\text{v}} &\geq \frac{1}{2} c_{d_{\text{v}}} \rho V^2 S_{\text{v}} \\
+    D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
     \label{e:taildrag}
@@
                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-    Re_{\text{v}} &= \frac{V\rho S_{\text{v}}/b_{\text{v}}}{\mu} \\
-    Re_{\text{h}} &= \frac{V\rho S_{\text{h}}/b_{\text{h}}}{\mu} 
 \end{align}
@@
 The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.  
-A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-	\label{f:taildragpolar}
-	\end{center}
-\end{figure}
-
-\subsection{Elliptcal Fuselage for Gas Powered Aircraft}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsection{Engine Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{PM} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].
-Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg et.
-al\cite{fitting}.
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-\subsubsection{Brake Specific Fuel Consumption Constraint}
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
-
-\begin{equation}
-    \label{e:rpmtobsfc}
-    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-\end{equation}
-
-where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Engine Lapse Rate Calculation}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=15,000$ [ft], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 1000 ft,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
-\end{equation}
-
-\subsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h} \geq 100 \text{ [ft/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
+% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
+
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
+%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
+% 	\label{f:taildragpolar}
+% 	\end{center}
+% \end{figure}
 
@@
 
-To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes.
-Figure~\ref{f:latvsmtowtrade} shows this trade study evaluated at the 80th, 90th and 95th percentile wind speeds.
-Each point on the graph is a unique solution to the optimization model for the stated assumptions and constants.
+The gas and solar-electric powered aircraft optimization models were solved by minimizing take of weight for the requirements listed in Table~\ref{t:mreqs}. 
+Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
+The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take off weight of 84.0 lbs.  
+The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take off weight of 320.5 lbs.
+According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
+This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
+
+\begin{longtable}{llll}
+\caption{Fixed Optimization Parameters} \\
+\toprule
+\toprule
+\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
+\midrule
+$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
+$\tau$                       & 0.5  & $\tau$                       & 0.115 \\
+$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
+$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
+$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
+$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
+$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
+Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
+$\dot{h}$ [ft/min]           & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
+                             &     & $\eta_{\text{charge}}$          & 0.98 \\
+                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
+\bottomrule
+\label{t:gassolarparams}
+ \end{longtable}
+
+\subsection{Changing Requirements}
+
+Important trade offs between the gas powered and solar-electric powered archictures are highlighted and quantified by changing the latitude and endurance endurance requirements.   
+Both optimization models were solved by minimizing the max take off weight across different bands of latitudes. 
+Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
+The gas powered model was solved x times and the solar-electric powered model was solved x times to produce Figure~\ref{f:latvsmtowtrade}.
 The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.
@@
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-The gas powered aircraft curves level off around 38 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
+The gas powered aircraft curves level off around 38 degrees latitude because wind speeds are highest at that latitude at 4,500 m. 
 Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
@@
 
-Because gas-powered aircraft are endurance limited by the amount of fuel that they can carry, a trade study was done for the gas powered aircraft to determine how weight increases for longer endurance.
+While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
+Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
 Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude.
@@
 
-Changing the objective function can be informative to the design space. 
-High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
+% Changing the objective function can be informative to the design space. 
+% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
+% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
 (Figure~\ref{f:solarobjcomp})
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
- \end{subfigmatrix}
-    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-    \label{f:solarobjcomp}
-\end{figure}
-
-\subsection{Changing Assumptions}
-
-By altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq \text{MTOW} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
+%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
+%  \end{subfigmatrix}
+%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
+%     \label{f:solarobjcomp}
+% \end{figure}
+
+\subsection{Changing Input Values}
+
+The previous results are dependent on the assumed input values and parameters.  
+Changing the input values can help show where the designs becomes infeasible. 
+Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
+By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
+Figure~\ref{f:battsolarcontour} shows contours of latitude for a gien solar cell efficiency and battery energy density.
+Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.pdf}
+    \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
+	\label{f:spanvsendurance}
 	\end{center}
@@
 
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-The solar aircraft size and design depends on the assumed solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
 Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
@@
  \label{f:solarcontours}
+\end{figure}
+
+\subsection{Changing Physical Modeling Assumptions}
+
+By altering the aircraft structural model it can be observed how air density trades for wing weight. 
+It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
+If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
+
+\begin{equation}
+    W_{\text{structural}} \geq \text{MTOW} f_{\text{structural}}
+\end{equation}
+
+where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
+However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
+Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.pdf}
+ \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
+ \label{f:altoper}
+	\end{center}
+\end{figure}
+
+Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
+The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
+However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
+At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
+If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.pdf}
+    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
+ \label{f:polarmission}
+	\end{center}
 \end{figure}
@@
 
+\section*{Appendix C}
+
+\subsection{Altitude and relevant atmosphereic quantities}
+
+The altitude also determines atmospheric properties such as air density and viscosity.
+For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+
+For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
+
+\begin{align}
+    \label{e:tropopress}
+    p &= \rho R_{\text{spec}}T_{11} \\
+    \label{e:tropoalt}
+    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+\end{align}
+
+According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+\\
+
 \end{document}

### 2017-04-17 gassolar/environment/wind_fit.py

@@ -8,7 +8,7 @@ from gpfit.fit import fit
 plt.rc("text", usetex=True)
 plt.rcParams.update({'font.size':15})
 
-GENERATE = True
+GENERATE = False
 PERCT_NORM = 100.0
 WIND_NORM = 100.0
 RHO_NORM = 1.0
@@ -99,7 +99,7 @@ if __name__ == "__main__":
     if GENERATE:
         latitude = range(20, 61, 1)
     else:
-        latitude = [20]
+        latitude = [30]
 
     data = []
     np.random.seed(0)

### 2017-04-17 gassolar/environment/wind_speeds.py

@@ -8,10 +8,9 @@ plt.rcParams.update({'font.size':15})
 
 # PATH = (os.path.abspath(__file__).replace(os.path.basename(__file__), "")
 #         + "windspeeds" + os.sep)
-PATH = (os.path.abspath(__file__).replace(os.path.basename(__file__), "")
-        + "windsNS" + os.sep)
+PATH = os.path.abspath(__file__).replace(os.path.basename(__file__), "")
 
-def get_windspeed(latitude, perc, altitude, day, path=PATH):
+def get_windspeed(latitude, perc, altitude, day, NS="NS", path=PATH):
     """
     Method to return windspeeds for different latitudes
     altitudes/percentiles
@@ -27,6 +26,7 @@ def get_windspeed(latitude, perc, altitude, day, path=PATH):
     -------
     wind: wind speed [m/s] (array if altitude is array)
     """
+    path = path + "winds%s" % NS + os.sep
 
     mos = ["jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep",
            "oct", "nov", "dec"]
@@ -89,102 +89,103 @@ if __name__ == "__main__":
     else:
         path = ""
 
-    Fig, Ax = plt.subplots()
-    # Colors = ["#253B6E", "#1F5F8B", "#1891AC"]
-    Colors = ["#1C226B", "#3E31AE", "#4AA9AF"]
-    lat = np.arange(0, 70)
-    for al, c in zip([15000, 50000, 60000], Colors):
-        Wind85 = [get_windspeed(l, 80, al, 355) for l in lat]
-        Wind90 = [get_windspeed(l, 90, al, 355) for l in lat]
-        Wind95 = [get_windspeed(l, 95, al, 355) for l in lat]
-        Ax.fill_betweenx(lat, Wind85, Wind95, alpha=0.5, facecolor=c)
-        Ax.plot(Wind90, lat, c, label="Altitude=%d" % al, lw=2)
-        if al == 50000:
-            Ax.annotate("80%", xy=(Wind85[np.where(lat==31)[0][0]],31),
-                        xytext=(3,0.1), textcoords="offset points",
-                        arrowprops=dict(arrowstyle="-"), fontsize=12)
-            Ax.annotate("90%", xy=(Wind90[np.where(lat==31)[0][0]],31),
-                        xytext=(3,0.1), textcoords="offset points",
[…]

### 2017-04-17 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -38,19 +38,33 @@ def fit_setup(Re_range):
     CL = []
     CD = []
     RE = []
+    fig, ax = plt.subplots()
     for r in Re_range:
         dataf = text_to_df("jho1.ncrit09.Re%dk.pol" % r)
-        CL.append(dataf["CL"])
-        CD.append(dataf["CD"])
-        RE.append([r*1000.0]*len(dataf["CL"]))
+        if r < 150:
+            cl = dataf["CL"].values.astype(np.float)
+            cd = dataf["CD"].values.astype(np.float)
+            CL.append(cl[cl >= 1.0])
+            CD.append(cd[cl >= 1.0])
+        elif r < 200:
+            cl = dataf["CL"].values.astype(np.float)
+            cd = dataf["CD"].values.astype(np.float)
+            CL.append(cl[cl >= 0.8])
+            CD.append(cd[cl >= 0.8])
+        else:
+            CL.append(dataf["CL"].values.astype(np.float))
+            CD.append(dataf["CD"].values.astype(np.float))
+        ax.plot(dataf["CL"].values.astype(np.float), dataf["CD"].values.astype(np.float))
+        ax.legend(["%d" % re for re in Re_range])
+        RE.append([r*1000.0]*len(CL[-1]))
+
+    fig.savefig("polarstest.pdf")
 
     u1 = np.hstack(CL)
     u2 = np.hstack(RE)
     w = np.hstack(CD)
-    u1 = u1.astype(np.float)
-    u2 = u2.astype(np.float)
-    w = w.astype(np.float)
     u = [u1, u2]
+    xx = np.log(u2)
     x = np.log(u)
     y = np.log(w)
     return x, y
@@ -86,7 +100,7 @@ def plot_fits(re, cnstr, x, y):
     return fig, ax
 
 if __name__ == "__main__":
-    Re = np.arange(150, 750, 50)
+    Re = np.arange(100, 750, 50)
     X, Y = fit_setup(Re) # call fit(X, Y, 4, "SMA") to get fit
     np.random.seed(0)
     cn, err = fit(X, Y, 4, "SMA")
@@ -98,7 +112,7 @@ if __name__ == "__main__":
     else:
         df.to_csv("jho_fitdata.csv")
 
-    replot = np.array([150, 200, 300, 350, 400])
+    replot = np.array([100, 150, 200, 300, 350])
     # replot = np.array([300, 350, 400, 450, 500])
     F, A = plot_fits(replot, cn, X, Y)
     if len(sys.argv) > 1:

### 2017-04-17 gassolar/fits/jho1polars/jho1polarsweep.sh

@@ -1,5 +1,5 @@
 AIRFOIL=jho1
-Re="100 150 200 250 300 350 400 450 500 550 600"
+Re="50 100 150 200 250 300 350 400 450 500 550 600"
 for r in $Re
 do
     ./genpolar.sh $AIRFOIL $r

### 2017-04-17 gassolar/gas/gas.py

@@ -15,11 +15,11 @@ from loiter import Loiter
 
 class Aircraft(Model):
     "the JHO vehicle"
-    def setup(self, Wfueltot, DF70=False):
+    def setup(self, Wfueltot):
 
         self.fuselage = Fuselage(Wfueltot)
         self.wing = Wing()
-        self.engine = Engine(DF70)
+        self.engine = Engine()
         self.empennage = Empennage()
 
         components = [self.fuselage, self.wing, self.engine, self.empennage]
@@ -29,9 +29,9 @@ class Aircraft(Model):
         Wpay = Variable("W_{pay}", 10, "lbf", "payload weight")
         Wavn = Variable("W_{avn}", 8, "lbf", "avionics weight")
         Wwing = Variable("W_{wing}", "lbf", "wing weight for loading")
-        etaprop = Variable("\\eta_{prop}", 0.75, "-", "propulsive efficiency")
+        etaprop = Variable("\\eta_{prop}", 0.8, "-", "propulsive efficiency")
 
-        self.empennage.substitutions["V_h"] = 0.55
+        self.empennage.substitutions["V_h"] = 0.45
         self.empennage.substitutions["V_v"] = 0.04
         self.empennage.substitutions["m_h"] = 0.4
         constraints = [
@@ -148,20 +148,22 @@ class Climb(Model):
 
 class Mission(Model):
     "creates flight profile"
-    def setup(self, DF70=False):
+    def setup(self, latitude=38, percent=90):
 
         mtow = Variable("MTOW", "lbf", "max-take off weight")
         Wcent = Variable("W_{cent}", "lbf", "center aircraft weight")
         Wfueltot = Variable("W_{fuel-tot}", "lbf", "total aircraft fuel weight")
         LS = Variable("(W/S)", "lbf/ft**2", "wing loading")
 
-        JHO = Aircraft(Wfueltot, DF70)
+        JHO = Aircraft(Wfueltot)
 
-        climb1 = Climb(JHO, 10, altitude=np.linspace(0, 15000, 11)[1:])
-        cruise1 = Cruise(JHO, 1, R=200)
-        loiter1 = Loiter(JHO, 5)
-        cruise2 = Cruise(JHO, 1)
-        mission = [climb1, cruise1, loiter1, cruise2]
+        climb1 = Climb(JHO, 10, latitude=latitude, percent=percent,
+                       altitude=np.linspace(0, 15000, 11)[1:])
[…]

### 2017-04-17 gassolar/solar/battsolarcon.py

@@ -9,9 +9,20 @@ from gpkit.tools.autosweep import autosweep_1d
 N = 100
 plt.rcParams.update({'font.size':15})
 
+def find_sols(bsts):
+    for s in bsts:
+        if s.splits is None:
+            continue
+        for sn in s.splits:
+            if sn in bsts:
+                continue
+            bsts.append(sn)
+        return bsts
+
 def plot_battsolarcon():
     fig, ax = plt.subplots()
-
+    time = 0
+    numsolves = 0
     etamax = []
     lines = []
     midx = []
@@ -29,7 +40,11 @@ def plot_battsolarcon():
             M.substitutions.update({"\\eta_Mission, Aircraft, SolarCells": 0.4})
             M.cost = M["h_{batt}"]
             sol = M.solve("mosek")
+            time += sol["soltime"]
+            numsolves += 1
             hmin = sol["cost"].magnitude + 1e-3
+            if hmin >= hmax:
+                break
             del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
             M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
             xmin_ = np.linspace(hmin, hmax, 100)
@@ -38,6 +53,10 @@ def plot_battsolarcon():
             try:
                 bst = autosweep_1d(M, tol, M["h_{batt}"], [hmin, hmax],
                                    solver="mosek")
+                bsts = find_sols([bst])
+                sols = np.hstack([b.sols for b in bsts])
+                stime = sum(np.unique([s["soltime"] for s in sols]))
+                numsolves += bst.nsols
                 if lat % 4 == 0:
                     l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "k",
                                 label="%d$^{\\circ}$ Lat" % lat)
@@ -52,6 +71,7 @@ def plot_battsolarcon():
 
     # ax.fill_between(xmin_, bst.sample_at(xmin_)["cost"], max(etamax),
     #                 edgecolor="r", lw=2, hatch="/", facecolor="None", zorder=100)
+    print "%d solves in %.4f seconds" % (numsolves, time)
     labelLines(lines, align=False, xvals=midx, zorder=[10]*len(lines))
     # ax.text(425, 0.36, "Infeasible")
     ax.set_ylabel("Solar Cell Efficiency")

### 2017-04-18 docs/reviews/review_response.tex
initial reponse at reviewer comments; more edits to paper

\documentclass[10pt, a4paper]{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\newcommand{\dbar}[1]{\Bar{\Bar{#1}}}

\topmargin = 0.1in \textwidth=5.7in \textheight=8.6in

\oddsidemargin = 0.1in \evensidemargin = 0.1in

% headers
\usepackage{fancyhdr}
\pagestyle{fancy}
\chead{} 
\rhead{\thepage} 
% footer
\lfoot{\small\scshape } 
\cfoot{} 
%%%% insert your name here %%%%
% \rfoot{\footnotesize Author} 
\renewcommand{\headrulewidth}{.3pt} 
\renewcommand{\footrulewidth}{.3pt}
\setlength\voffset{-0.25in}
\setlength\textheight{648pt}

\begin{document}

\title{Response to referees \\
C034405: "Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming``}
\author{Warren Hoburg and Michael Burton}
\maketitle

We thank the anonymous referees for their thoughtful and insightful comments on the draft manuscript.
We feel that the paper has improved significantly thanks to the feedback. 
Specific comments from the reviews (in italics) are addressed herein.

\section*{Referee 1}

\emph{
This article presents a description of a methodology to design a hale type aircraft using geometric programming.
As such, the approach is probably of more relevance than the specifics.
Review of Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming
This article presents a description of a methodology to design a hale type aircraft using geometric programming.
As such, the approach is probably of more relevance than the specifics.
It lacks focus. Frankly this article is far more appropriate as a conference paper than a journal article. It reads like a recipe with a significant number of assumptions that are not supported.
While the approach is definitely of value, the presentation is less so. If published, this article should be shortened to an engineering note. 
A flow chart/figure could be included that shows the parameters (equations) that feed into the formulation. This way the reader gets the “bigger” picture. 
The most significant figures to include are the results only. 
Perhaps show a sketch of what a typical vehicle would look like. 
Include only the most relevant equations.
}

\section*{Referee 2}

\emph{Optimization: I would like to see more information on what were the design variables for both aircraft, how long a single analysis took and how many analyses were required by the GP approach.}

We would like to thank the referee for this suggestion.  We claim that speed is o of great benefits of using GP and we should have absolutely included more information about our design variables, solve times, and number of analysis.  We refer the referee to the results section of the revised paper where we have included a table of key design parameters and solve times and number of analysis throughout.

\emph{Approximation: In order to accommodate the use of GP posynomial approximations were used. It is important then to check the final designs and see if they perform as approximated. Often the optimization process is steered to regions where the quality of the approximations is poor.}

We thank the referee for this insightful point.  We have since included bounds on each posynmial approxmiation to ensure that the solution does not enter a region where we do not have representative data.  We noticed that the solar-electric solution tended to operate a lower Reynolds numbers than was approximated by our posynomial approxmiation.  We have since generated more drag data from XFOIL at low Reynolds numbers and re-generated the posynomial fits for the airfoil drag data.

\emph{Assumptions:
The paper appears to insist that a single model of the UAV need to be used for all latitudes. This seems to be very restrictive. I would like to know how much better the performance is for UAVs that are tailored for a narrow band of latitudes.
In a similar vein, why not have both kind of UAVs depending on the latitude and the season.
}

We thank the referee for this comment.  These are both interesting trade studies that we could certainly explore.  We chose not to include them in this paper because we felt they were not central take aways of our findings.  This would have additionally increased the length of our paper.  If the referee feels that these two trade studies should be explored and included we would be happy to discuss this futher.

\emph{Length: The paper is tedious to read with the long listing of models and constraints. It may make more sense to put some of the material as additional on-line material, which I assume is available with the Journal of Aircraft.}

We recognize that this paper is long.  We have cut a few non-essential equations and graphs to try and shorten the paper.  However, because GP solves a list of constraints we can explicity state all of our assumptions.  We feel that this is important to justify our results.  We are willing to discuss cutting additionally sections/figures/equations if the referee has specific suggestions.

\emph{Minor issues}
\begin{itemize}
    \item \emph{The fact that the design is limited to the northern hemisphere should be articulated clearly in the Abstract.}
        We thank the referee for this comment.  We recongnize that this may have seemed like a limitation in our assumptions.  We have since revised our study to include wind speed data from the Southern Hemisphere.  We would refer the referee to Figure 1 of the revised paper.  We initially chose not to include wind speed data from the Southern Hemisphere because the wind speeds are consistently lower in the Southern Hemisphere versus the Northern Hemisphere.  Thus our assumption was conservative.  However, given that our results do not change drastically by including the Southern Hemisphere wind speed data we have chosen to included it to avoid misconceptions about our results.

    \item \emph{Sources for data presented in figures should be given in the captions.}
        We thank the reader for this comment.  Citations have since been added to all capations where data is used.

    \item \emph{It was not clear to me why the solar UAV not fly substantially higher as there is less drag even if wind speeds are higher.}
    There are two separate, but related issues that affect the operating altitude.  The first is wind speed.  It is true that drag is lower at higher altitudes.  But, if we want to be faster than the wind speeds then we pay a higher penalty for wind speeds than we do for air density,

    \begin{equation}
        D = \frac{1}{2} \rho V^2 S C_D
    \end{equation}

    as drag goes by $V^2$ but only varies linearly with $\rho$.  Additionally, wind speeds cause an even greater constraint on power because power goes by velcoity cubed, $P \sim V^3$.  This affects the number of solar cells and number of batteries, which in turn affects the wing size and therefore aircraft weight. This would suggest that we should fly at the minimum wind speed always.

    There is a second factor that is also going on, which is the trade off between wing weight and air density.  Because lift also scales equally with air density and wing area, like drag,

    \begin{equation}
        L = \frac{1}{2} \rho V^2 S C_L
    \end{equation}

    for lower air densities a bigger wing is required.  Because of the cubed square law, a larger wing is very costly.  Therefore, it is actually more advantageous to fly at a lower altitude to accomodate a slightly larger wing.

    If we assume that a large wing doesn't affect the weight at all, then we should see the aircraft fly at the minimum wind speed.  This is precisely the point we are trying to make in Figure 17.

    \item \emph{I did not understand what the autocorrelation indicated in terms of performance. That is, I could not make sense of the meaning of Figs 4a and 5a.}
        We thank the referee for this comment.  We realized that these autocorrelation plots, while interesting, are not actually pertinent to the optimization model and we chose to cut them from the paper.  We initially included them because as we were doing the research we wanted to understand what a percentile wind speed meant operationally.  We were asking ourselves does the 90th percentile wind speed mean that everyday we will be blown off station 10\% of the time, or that there will be 5 days in the month during which it is impossible to station keep.  The autocorrelation plots showed that the later was true.  In other words, when there are high wind speeds they can occur for days a time. `

    \item \emph{When data is given for a certain latitude, it is not clear whether the implication is that there is no difference with longitude, or that this is the worst for any longitude.}
        While wind speeds certianly vary with longitude, our claim is that the variation with longitude is minimial compared to the variation to latitude and can be ignored.

    \end{itemize}

\section*{Referee 3}
\emph{Overall the paper is well written and follows a logical progression. The analysis is clearly explained in sufficient detail. }

    We thank the referee for this supportive comment.

\emph{When I started reading the paper I was under the impression that it would be a direct comparison between an optimized solar electric aircraft and an internal combustion driven aircraft for a specific mission or mission type. However since the mission requirements on the 2 aircraft types are not the same the paper is mainly an example of how the geometric programming method can be applied to 2 different types of aircraft (combustion engine and solar electric). This is fine but it should be made clear at the beginning of the paper.}

    We thank the referee for this comment.  We realize that we may have been unclear in our comparison.  What we are really comparing is the capability of gas and solar-electric powered aircraft to meet the same set of requirements.  We have changed the second section in the paper to convey that statement.  We believe this to be a fair comparison and will hopefully address other comments made by this referee.

\emph{Overall I only had a few other comments:}

\begin{itemize}
    \item \emph{On page 5: again this speaks to not having the same mission constraints for both aircraft. Using the winter solstice to set the endurance level for the solar aircraft does not truly examine the design space for that aircraft. From a mission perspective having an aircraft that can fly 6 months of the year can be desirable when you are comparing it to an aircraft with a 7 day endurance. Even though that 7 day endurance can happen at any time during the year.}

        We believe this comment is best addressed by a clarification of the requirements.  One requirement states that the mission must be greater than 6 days.  Obviously, the solar-electric powered aircraft will likely exceed 6 days with ease, but it is same requiement for both aircraft.  Another requirement is that the aircraft be able to fly in all seasons.  This requires the solar-electric aircraft to fly during the winter-solstice.  As we explored changing the latitude requirement, we could also explore changing the season requirement.  We chose not to explore this trade study in the paper because of the additional length to the paper.  We would be happy to include a season trade study if the referee feels it is an important result.

    \item \emph{On page 8: it should be stated that the temperature is constant between the altitude range of 11 km up to 20 km not above 11 km. Also figure 3 shows the altitude in ft. This is not consistent with the text which states altitude in meters. A single unit system should be used, otherwise it gets confusing. I would suggest listing everything in meters and then if needed put the equivalent value in ft in parenthesis next to it.}

        We thank the referee for noticing this error.  It has since been corrected.  Additionally, all altitude references in the paper are now in meters.

    \item \emph{On page 12: in figure 6 it should be indicated that the markers "o" are data points for the wind speed and their source should be referenced on the figure.}

        We thank the referee for noticing this.  The data points have since been labeled and their source referenced.

    \item \emph{On page 26: Using the fuselage to store all of the fuel for the combustion engine aircraft and then using the wings to store the batteries for the solar electric aircraft is not consistent between the designs. Again, this is fine for just demonstrating the optimization method but is an issue if you are trying to compare the results between the two aircraft types and draw a conclusion.}

        We thank the referee for the insightful comment.  Because we are attempting to do a fair comparison between the two archictures we have put a lot of thought into whether or not to putting fuel in the wings is the best possible configuration for the gas powered archicture.  We could certainly add it as a constraint to the optimization but to be fair we would have to add additional constraints to account for the added weight and complexity of wet wings.  For the smaller gas powered UAVs that are being considered in this problem wet wings would add a significant amount of weight and complexity whereas a fuselage of this size is easy to build and light weight.  For the solar-electric aircraft putting batteries in the wings is much simpler than putting fuel in wings.  Batteries do not require a special fuel tank and have low volume density.  Therefore, we believe that putting batteries in the wings for the solar-electric aircraft and fuel in a fuselage for the gas aircraft is a fair comparison.

    \item \emph{On page 29: This is just another point showing the significant differences between the proposed flight requirements between the two aircraft types and why the should not be compared. From equation 105 it can be seen that above ~44,000 ft the engine would produce no output due to the air density. Therefore to fly at altitudes comparable to the solar electric aircraft a compressor system would be needed which would greatly increase the mass and impact the performance of the combustion aircraft.}

        We thank the referee for this comment.  We hope that previous comments about the requirements have address this concern.  Because the requirement is to be at an altitude > 4,500 meters we don't necessarily care where each aircraft flies so long as it meets the requirement.  In this respect, we believe it to be a fair comparison.  An obvious take away from the lapse rate of the gas engine is that any payloads requiring an altitude greater than 44,000 ft is that the solar-electric is the preferred option.

    \item \emph{On Page 31 line 28: it should state ±30 degrees Latitude not North latitude}

        We thank the referee for noticing this error.  It has been corrected.

    \item \emph{On page 38, I could not get to the file given in reference 19 from the link provided. The actual report should be referenced not just the web site.}
    \end{itemize}

\end{document}

### 2017-04-18 docs/paperin.tex
initial reponse at reviewer comments; more edits to paper

@@
 $\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-MTOW \> max take-off weight \\ % [N] \\
+$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
 $n$ \> number of wing segments \\
@@
 Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-Additionally, because few long-endurance gas and solar-electric powered aircraft have been built, the operational limits of these aircraft are relatively unknown. 
-Even more unclear are the trade offs between gas powered and solar-electric powered architectures.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to explore this design space and highlight trade offs between the two proposed architectures.
+The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures even more unclear.
+
+This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two archictures for a given mission.
 Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
 Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters such as endurance, latitude, and percentile wind speed to observe how aircraft size and weight are affected by different requirements.
+The optimization is solved multiple times to observe how aircraft size and weight are affected by different design parameters and requirements.
 
@@
 
-\section{Requirements Analysis}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
+\section{Mission Requirements}
+
+A comparision of the gas and solar-elecrtric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
+These requirements can and will be changed in the results section to observe which archicture is best suited to meet various sets of requirements.
+A discussion of each requirement is included in this section.
+
+\begin{longtable}{lccccccccccccc}
+\caption{Mission Requirements}\\
+\toprule
+\toprule
+\label{t:mreqs}
+Payload & 10 lbs\\
+Endurance & 6 days\\
+Season & all seasons\\
+Altitude & > 4,600 m\\
+Latitude & $\pm30^{\circ}$\\
+\bottomrule
+\end{longtable}
+
+\subsection{Payload}
+
+A 10 lbs payload is a baseline requirement for both the gas and solar-electric powered aircraft.
+This requirement can be changed to observe how each architeure is able to handle larger or smaller payloads.
+
+Inherent in a payload requirement is the ability to station keep. To station keep an aircraft must fly at least as fast as the wind speed,
 
@@
 For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st, then it can thoretically fly at any time of the year.  
-Therefore, the endurance requirement for solar-electric powered aircraft is assumed to be a requirement to fly during the winter solstice.  
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement.
+Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
 To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-Because the endurance requirement for solar-electric powered aircraft is set by the winter solstice and because wind speeds are higher during winter months, only wind speed data from December will be used for the aircraft sizing and performance analysis presented in this paper.
+Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June will be used for the aircraft sizing and performance analysis presented in this paper.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Winds peak at summer and winter solstice in Southern and Northern Hemispheres respectively.\cite{wind}}}
+ \label{f:windvsmonth}
+\end{figure}
+
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
+%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
+% 	\label{f:windvsmonth}
+% 	\end{center}
+% \end{figure}
+
+\subsection{Altitude}
+
+The altitude requirement is generally set by a minimum payload required to meet a certain coverage
+
+\begin{equation}
+ h \geq h_{\text{min}}.
+\end{equation}
+
+Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
+Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m depending on the engine.
+Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
-    \caption{\textbf{Winds tend to increase up to 35$^{\circ}$ latitude. Bands represent 80th, 90th and 95th percentiles.}}
-	\label{f:altvswind}
-	\end{center}
-\end{figure}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\subsection{Latitude}
-
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds. }}
+	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
+    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
 	\label{f:latvswind}
@@
 Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-To solve an aircraft design problem, all physical models are expressed as constraints on the GP.\cite{hoburgthesis} 
+
+To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
+Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
+Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
+The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.8\textwidth,natwidth=480,natheight=200]{flowchart.pdf}
+    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
+	\label{f:flowchart}
+	\end{center}
+\end{figure}
+
 A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
 A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-\\
-
-\section{Aircraft Sizing Models}
+Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gas_solar_trade} \\
+
+\section{Configuration}
+
+Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
+Common characteristics between the two models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
+The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
+The solar-electric powered aircraft holds the batteries in the wings.  
+The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
+A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.8\textwidth,natwidth=517,natheight=408]{simpleaircraft.pdf}
+    \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
+	\label{f:simpleaircraft}
+	\end{center}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.7\textwidth,natwidth=358,natheight=334]{wbreak.pdf}
+   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
+	\label{f:wbreak}
+	\end{center}
+\end{figure}
+
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      % \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=1.0\textwidth,natwidth=144,natheight=320]{solarwbreak.pdf}}
+%      % \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=1.0\textwidth,natwidth=178,natheight=320]{gaswbreak.pdf}}
+%      \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=0.30\textwidth]{solarwbreak.pdf}}
+%      \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=0.40\textwidth]{gaswbreak.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
+%  \label{f:wbreak}
+% \end{figure}
+
+These assumptions can be changed to evaulate different configuration.  Changes in configuration assumptions may require modification of models or equations.
+
+\section{Aircraft Physical Models}
 
@@
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
-Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
-In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
-The constraint on December wind speeds and air density at the 35 degree latitude is
-
-\begin{align}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.213} &\geq 2918\rho^{7.587}p_{\text{wind}}^{9.37} + \num{1.58d-10}\rho^{-6.074}p_{\text{wind}}^{60.421} \nonumber \\
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees. (See Appendix A) 
+The constraint on December wind speeds and air density at the 30 degree latitude is
+
+\begin{align}
+    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
     \label{e:windfit35}
-    &+ \num{4.05d-16}\rho^{-9.879}p_{\text{wind}}^{13.204} + 1048\rho^{6.33}p_{\text{wind}}^{46.232}
+    &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
 \end{align}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
-    \caption{\textbf{Posynomial approximation to wind speed data at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
 	\label{f:windfitl35}
@@
 
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes, the shear forces and moments can be computed from the distributed loads $q(y)$, with boundary conditions of zero shear forces and moments at the wing tips.\cite{bending}
-
-\begin{align}
-    \label{e:shear}
-    \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-    \label{e:moment}
-    M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-    \label{e:shearboundary}
-    \mathcal{S}_n &= 0 \\
-    \label{e:momentboundary}
-    M_n &= 0
-\end{align}
-
-Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-
-\begin{align}
-    \label{e:angle}
-    \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-    \label{e:angleboundary}
-    \Theta_0 &= 0 \\
-    \label{e:defboundary}
-    w_0 &= 0 
-\end{align}
-
-Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
+Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input.
+Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending}
+
+% \begin{align}
+%     \label{e:shear}
+%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
+%     \label{e:moment}
+%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
+%     \label{e:shearboundary}
+%     \mathcal{S}_n &= 0 \\
+%     \label{e:momentboundary}
+%     M_n &= 0
+% \end{align}
+% 
+% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
+% 
+% \begin{align}
+%     \label{e:angle}
+%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
+%     \label{e:deflection}
+%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
+%     \label{e:angleboundary}
+%     \Theta_0 &= 0 \\
+%     \label{e:defboundary}
+%     w_0 &= 0 
+% \end{align}
+%  
+% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
 
@@
 The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.  
-A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-	\label{f:taildragpolar}
-	\end{center}
-\end{figure}
-
-\subsection{Elliptcal Fuselage for Gas Powered Aircraft}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsection{Engine Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{PM} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].
-Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg et.
-al\cite{fitting}.
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-\subsubsection{Brake Specific Fuel Consumption Constraint}
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
-
-\begin{equation}
-    \label{e:rpmtobsfc}
-    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-\end{equation}
-
-where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Engine Lapse Rate Calculation}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=15,000$ [ft], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 1000 ft,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
-\end{equation}
-
-\subsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h} \geq 100 \text{ [ft/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\subsection{Weight Breakdown}
-
-The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
-The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.
-
-\begin{align}
-    \label{e:weightmtow}
-    \text{MTOW} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-\end{align}
-
-A similar weight breakdown is done for the solar-electric powered aircraft,
-
-\begin{align}
-    \label{e:weightsmtow}
-    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-    W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-    W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-\end{align}
-
-where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
+% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
+
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
+%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
+% 	\label{f:taildragpolar}
+% 	\end{center}
+% \end{figure}
+
+% \subsection{Weight Breakdown}
+% 
+% The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
+% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
+% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
+% 
+% \begin{align}
+%     \label{e:weightmtow}
+%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
+%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
+% \end{align}
+% 
+% A similar weight breakdown is done for the solar-electric powered aircraft, 
+% 
+% \begin{align}
+%     \label{e:weightsmtow}
+%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
+%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
+%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
+%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
+% \end{align}
+% 
+% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
+% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
+% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
 
@@
 
-Changing the objective function can be informative to the design space. 
-High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
+% Changing the objective function can be informative to the design space. 
+% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
+% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
 (Figure~\ref{f:solarobjcomp})
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
- \end{subfigmatrix}
-    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-    \label{f:solarobjcomp}
-\end{figure}
-
-\subsection{Changing Assumptions}
-
-By altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq \text{MTOW} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
+%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
+%  \end{subfigmatrix}
+%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
+%     \label{f:solarobjcomp}
+% \end{figure}
+
+\subsection{Changing Input Values}
+
+The previous results are dependent on the assumed input values and parameters.  
+Changing the input values can help show where the designs becomes infeasible. 
+Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
+By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
+Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.
+Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude.
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.pdf}
+    \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
+	\label{f:battsolarcontour}
 	\end{center}
@@
 
+\subsection{Changing Physical Modeling Assumptions}
+
+By altering the aircraft structural model it can be observed how air density trades for wing weight. 
+It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
+If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
+
+\begin{equation}
+    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
+\end{equation}
+
+where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
+However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
+Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.pdf}
+ \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
+ \label{f:altoper}
+	\end{center}
+\end{figure}
+
+Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
+The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
+However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
+At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
+If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.pdf}
+    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
+ \label{f:polarmission}
+	\end{center}
+\end{figure}
+
 \subsection{Sensitivities}
@@
 While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were MTOW and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.  
+For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
 Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take off weight.
@@
 
+\section*{Appendix C}
+
+\subsection{Altitude and relevant atmosphereic quantities}
+
+The altitude also determines atmospheric properties such as air density and viscosity.
+For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+
+For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
+
+\begin{align}
+    \label{e:tropopress}
+    p &= \rho R_{\text{spec}}T_{11} \\
+    \label{e:tropoalt}
+    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+\end{align}
+
+According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+\\
+
 \end{document}

### 2017-04-18 docs/paperin_review.tex
initial reponse at reviewer comments; more edits to paper

@@
 Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters such as endurance, latitude, and percentile wind speed to observe how aircraft size and weight are affected by different requirements.
+The optimization is solved multiple times to observe how aircraft size and weight are affected by different design parameters and requirements.
 
@@
 
-A comparision of the gas and solar-elecrtric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements, shown in Table~\ref{t:mreqs}.
+A comparision of the gas and solar-elecrtric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
 These requirements can and will be changed in the results section to observe which archicture is best suited to meet various sets of requirements.
@@
 Season & all seasons\\
-Altitude & 4,600 m\\
+Altitude & > 4,600 m\\
 Latitude & $\pm30^{\circ}$\\
@@
 For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, thus more than satisfying the 6 day endurance requirement.
+For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement.
 Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
@@
 
-Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswind} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
+Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
 Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m depending on the engine.
@@
 	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{winds} }}
+    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
 	\label{f:latvswind}
@@
 Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-To evaulate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
+
+To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
 Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
@@
 Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
-Both aircraft have a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
+Common characteristics between the two models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
 The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
 The solar-electric powered aircraft holds the batteries in the wings.  
-It is assummed that the added weight and complexity of wet wings outweights the drag and weight penalty of a fuselage for the gas powered archicture.  
 The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{simpleaircraft.pdf}
+	\includegraphics[width=0.8\textwidth,natwidth=517,natheight=408]{simpleaircraft.pdf}
     \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
@@
 The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
-I n total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees. (See Appendix A) 
 The constraint on December wind speeds and air density at the 30 degree latitude is
@@
 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-    \caption{\textbf{Posynomial approximation to wind speed data at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
 	\label{f:windfitl35}
@@
 
-\subsection{Weight Breakdown}
-
-The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
-The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight.
-
-\begin{align}
-    \label{e:weightmtow}
-    \text{MTOW} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-\end{align}
-
-A similar weight breakdown is done for the solar-electric powered aircraft,
-
-\begin{align}
-    \label{e:weightsmtow}
-    \text{MTOW} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-    W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-    W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-    W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-\end{align}
-
-where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
+% \subsection{Weight Breakdown}
+% 
+% The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
+% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
+% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
+% 
+% \begin{align}
+%     \label{e:weightmtow}
+%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
+%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
+% \end{align}
+% 
+% A similar weight breakdown is done for the solar-electric powered aircraft, 
+% 
+% \begin{align}
+%     \label{e:weightsmtow}
+%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
+%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
+%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
+%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
+% \end{align}
+% 
+% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
+% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
+% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
 
@@
 
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$30 degrees North latitude in 90th percentile wind speeds.  
+One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$30 degrees latitude in 90th percentile wind speeds.  
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
@@
 By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a gien solar cell efficiency and battery energy density.
+Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.
 Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude.
@@
     \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
-	\label{f:spanvsendurance}
+	\label{f:battsolarcontour}
 	\end{center}
@@
 \begin{equation}
-    W_{\text{structural}} \geq \text{MTOW} f_{\text{structural}}
+    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
 \end{equation}
@@
 According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant above 11,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
 \\

### 2017-04-18 gassolar/environment/wind_fit.py

@@ -86,6 +86,7 @@ def plot_fits(xdata, ydata, yfit, latitude):
             ax.text(np.exp(x1)[0]+0.005, yfits[i][0]-1.0, p)
     ax.set_xlabel("Air Density [kg/m$^3$]")
     ax.set_ylabel("Wind Speed [m/s]")
+    ax.legend(["ECF Wind Data", "GP approximation"], loc=2)
     ax.grid()
     return fig, ax
 

### 2017-04-19 docs/reviews/review_response.tex
create latex diff file; finish review paper response; add few citations

@@
 \begin{itemize}
-    \item \emph{The fact that the design is limited to the northern hemisphere should be articulated clearly in the Abstract.}
+    \item \emph{The fact that the design is limited to the northern hemisphere should be articulated clearly in the Abstract.} \\
         We thank the referee for this comment.
@@
 
-    \item \emph{Sources for data presented in figures should be given in the captions.}
+    \item \emph{Sources for data presented in figures should be given in the captions.} \\
         We thank the reader for this comment.  Citations have since been added to all capations where data is used.
 
-    \item \emph{It was not clear to me why the solar UAV not fly substantially higher as there is less drag even if wind speeds are higher.}
+    \item \emph{It was not clear to me why the solar UAV not fly substantially higher as there is less drag even if wind speeds are higher.} \\
     There are two separate, but related issues that affect the operating altitude.
@@
 
-    \item \emph{I did not understand what the autocorrelation indicated in terms of performance. That is, I could not make sense of the meaning of Figs 4a and 5a.}
+    \item \emph{I did not understand what the autocorrelation indicated in terms of performance. That is, I could not make sense of the meaning of Figs 4a and 5a.} \\
         We thank the referee for this comment.
@@
 
-    \item \emph{When data is given for a certain latitude, it is not clear whether the implication is that there is no difference with longitude, or that this is the worst for any longitude.}
+    \item \emph{When data is given for a certain latitude, it is not clear whether the implication is that there is no difference with longitude, or that this is the worst for any longitude.} \\
         While wind speeds certianly vary with longitude, our claim is that the variation with longitude is minimial compared to the variation to latitude and can be ignored.
@@
 From a mission perspective having an aircraft that can fly 6 months of the year can be desirable when you are comparing it to an aircraft with a 7 day endurance.
-Even though that 7 day endurance can happen at any time during the year.}
+Even though that 7 day endurance can happen at any time during the year.} \\
 
@@
 A single unit system should be used, otherwise it gets confusing.
-I would suggest listing everything in meters and then if needed put the equivalent value in ft in parenthesis next to it.}
+I would suggest listing everything in meters and then if needed put the equivalent value in ft in parenthesis next to it.} \\
 
@@
 
-    \item \emph{On page 12: in figure 6 it should be indicated that the markers "o" are data points for the wind speed and their source should be referenced on the figure.}
+    \item \emph{On page 12: in figure 6 it should be indicated that the markers "o" are data points for the wind speed and their source should be referenced on the figure.} \\
 
@@
     \item \emph{On page 26: Using the fuselage to store all of the fuel for the combustion engine aircraft and then using the wings to store the batteries for the solar electric aircraft is not consistent between the designs.
-Again, this is fine for just demonstrating the optimization method but is an issue if you are trying to compare the results between the two aircraft types and draw a conclusion.}
+Again, this is fine for just demonstrating the optimization method but is an issue if you are trying to compare the results between the two aircraft types and draw a conclusion.} \\
 
@@
 From equation 105 it can be seen that above ~44,000 ft the engine would produce no output due to the air density.
-Therefore to fly at altitudes comparable to the solar electric aircraft a compressor system would be needed which would greatly increase the mass and impact the performance of the combustion aircraft.}
+Therefore to fly at altitudes comparable to the solar electric aircraft a compressor system would be needed which would greatly increase the mass and impact the performance of the combustion aircraft.} \\
 
@@
 
-    \item \emph{On Page 31 line 28: it should state ±30 degrees Latitude not North latitude}
+    \item \emph{On Page 31 line 28: it should state ±30 degrees Latitude not North latitude} \\
 
@@
 
-    \item \emph{On page 38, I could not get to the file given in reference 19 from the link provided. The actual report should be referenced not just the web site.}
+    \item \emph{On page 38, I could not get to the file given in reference 19 from the link provided. The actual report should be referenced not just the web site.} \\
+
+        We thank the referee for noticing this error.
+We have corrected the bad reference and have replaced it two more reliable sources for composite properties.
+See references 16 and 26 in the revised paper.
     \end{itemize}

### 2017-04-19 docs/paperin_review.tex
create latex diff file; finish review paper response; add few citations

@@
 A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gas_solar_trade} \\
+Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
 
@@
 
-\subsubsection{Elliptcal Fuselage for Gas Powered Aircraft}
+\subsubsection{Elliptcal Fuselage}
 
@@
 
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{carbonfiber}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
+where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
 
@@
 
-The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{carbonfiber}
+The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
 
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{carbonfiber}
+where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
 
@@
 
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{carbonfiber} The wing skin is asssumed not to contribute to the bending stiffness.
+where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
 

### 2017-04-19 docs/paperin_review.tex
change ft to m for lapse rate and climb rate

@@
 
-The lapse is calculated from the required flight altitude $h=15,000$ [ft], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 1000 ft,\cite{enginelapse}
+The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
 
@@
     \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{1000 \text{ [ft]}} h.
+    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
 \end{equation}
@@
     \label{e:climbrate}
-    \dot{h} \geq 100 \text{ [ft/min]}.
+    \dot{h} \geq 30 \text{ [m/min]}.
 \end{equation}

### 2017-04-21 docs/reviews/review_response.tex
change fitting style; more edits; better flowchart

@@
 
+Note to the editor:
+
+The peer review of this manuscript was insightful and helpful.
+At a high level, the reviewers had comments about the clarity of our assumptions, objective, and approach.
+We have added material that explicitly discusses the requirements against which we are comparing the two aircraft, charts explaining our approach and process, and configuration drawings and text to explain some of our assumptions.
+Reviewers also stated that the manuscript was long.
+We have cut unnecessary figures and equations to help shorten it.
+Overall feedback was positive with minor additional questions that we have responded to herein.
+Only one referee asked that this manuscript be submitted as an engineering note.
+Given the overall positive feedback and our responses to the reviewers comments we respectfully submit that this manuscript be reconsidered as full paper instead of an engineering note.
+
 We thank the anonymous referees for their thoughtful and insightful comments on the draft manuscript.
@@
 As such, the approach is probably of more relevance than the specifics.
-Review of Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming
-This article presents a description of a methodology to design a hale type aircraft using geometric programming.
-As such, the approach is probably of more relevance than the specifics.
 It lacks focus. Frankly this article is far more appropriate as a conference paper than a journal article. It reads like a recipe with a significant number of assumptions that are not supported.
-While the approach is definitely of value, the presentation is less so. If published, this article should be shortened to an engineering note. 
-A flow chart/figure could be included that shows the parameters (equations) that feed into the formulation. This way the reader gets the “bigger” picture. 
-The most significant figures to include are the results only. 
-Perhaps show a sketch of what a typical vehicle would look like. 
+While the approach is definitely of value, the presentation is less so. If published, this article should be shortened to an engineering note. }
+
+We thank the referee for this comment.
+We have made organizational changes to the paper, including a more detailed description of the requirements, GP approach, and results.
+As with any aircraft design, we made many assumptions and explicitly stated them.
+We argue that a great benefit to the GP approach is the ability to change those assumptions and to resolve the problem quickly.
+We do this in the results section of the paper and believe this shows important insights about long-endurance UAVs.
+We would be happy to respond to any sepecific assumptions that the referee feels are not supported.
+
+\emph{
+A flow chart/figure could be included that shows the parameters (equations) that feed into the formulation. This way the reader gets the “bigger” picture. }
+
+We thank the referee for this helpful comment.  We have since added a flow chart.
+
+\emph{
+The most significant figures to include are the results only. }
+
+We thank the referee for this comment.  We did cut multiple figures.
+
+\emph{
+Perhaps show a sketch of what a typical vehicle would look like. }
+
+We thank the referee for this comment.  We included a vehicle sketch.
+
+\emph{
 Include only the most relevant equations.
 }
+
+We thank the referee for this comment.  We did cut a few unnecessary equations to help shorten the paper.
 

### 2017-04-21 docs/paperin_review.tex
change fitting style; more edits; better flowchart

@@
 Payload & 10 lbs\\
-Endurance & 6 days\\
+Station Keeping & 90\% \\
+Endurance & $>6$ days\\
 Season & all seasons\\
-Altitude & > 4,600 m\\
+Altitude & $>$ 4,600 m\\
 Latitude & $\pm30^{\circ}$\\
@@
 
-\subsection{Payload}
-
-A 10 lbs payload is a baseline requirement for both the gas and solar-electric powered aircraft.
-This requirement can be changed to observe how each architeure is able to handle larger or smaller payloads.
-
-Inherent in a payload requirement is the ability to station keep. To station keep an aircraft must fly at least as fast as the wind speed,
+\subsection{Station Keeping}
+
+To station keep an aircraft must fly at least as fast as the wind speed,
 
@@
 
+Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
+Therefore, the station keeping requirement is defined as a percentage which is the percentile wind speed. 
 The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,
@@
 The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees. (See Appendix A) 
-The constraint on December wind speeds and air density at the 30 degree latitude is
-
-\begin{align}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-    \label{e:windfit35}
-    &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-\end{align}
-
-where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
+% The constraint on December wind speeds and air density at the 30 degree latitude is 
+% 
+% \begin{align}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
+%     \label{e:windfit35}
+%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
+% \end{align}
+% 
+% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
 The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-	\label{f:windfitl35}
-	\end{center}
-\end{figure}
+% 
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+% 	\label{f:windfitl35}
+% 	\end{center}
+% \end{figure}
 
@@
 The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix B to generate the fitted equations. 
 % Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
@@
 
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
 \subsection{Gas Powered Aircraft Physical Models}
@@
 
-\subsection{Steady Level Flight}
+\subsubsection{Steady Level Flight}
 
@@
 
-\subsection{Aerodynamics}
+\subsubsection{Aerodynamics}
 
@@
 
-\subsection{Wing Spar Model}
+\subsubsection{Wing Spar Model}
 
@@
 
-\subsubsection{Standard Wing Bending Case}
+\begin{itemize}
+
+\item[Standard Wing Bending Case]
 
@@
 
-\subsubsection{Gust Loading Case}
+\item[Gust Loading Case]
 
@@
 
-\subsubsection{Discretized Beam Model}
+\item[Discretized Beam Model]
 
@@
 
-\subsubsection{Cap Spar for Bending Loads}
+\item[Cap Spar for Bending Loads]
 
@@
 
-\subsection{Wing Skin Weight}
+\end{itemize}
+
+\subsubsection{Wing Skin Weight}
 
@@
 
-\subsection{Empennage}
+\subsubsection{Empennage}
 
@@
 The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-
-\begin{align}
-    D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-    \label{e:taildrag}
-    c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-                 & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-                 &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-\end{align}
-
 where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
 The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
+
+% \begin{align}
+%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
+%     \label{e:taildrag}
+%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
+%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
+%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
+% \end{align}
+
 % A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
@@
 
+\section*{Appendix D}
+
+\subsection{Solar energy calculations}
+
+The total available solar energy per day is an integral of the available solar power during the course of the day,
+
+    \begin{equation}
+        \label{e:solares}
+        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
+    \end{equation}
+
+For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
+The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
+
+\begin{equation}
+    \label{e:solarp}
+    P_{\text{sun}} = P_0 \cos{\theta}
+\end{equation}
+
+The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
+
+    \begin{equation}
+        \label{e:solartheta}
+        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
+    \end{equation}
+
+ The declination angle $\Delta$, can be found using the relation\cite{solar}
+
+    \begin{align}
+        \label{e:solardelta}
+        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
+        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
+    \end{align}
+
+    where $\beta = 2\pi (\text{DOY}-1)/365$.
+    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
+
+    \begin{align}
+        \label{e:solartday}
+        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
+        \label{e:solarsunrise}
+        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
+        \label{e:solartday2}
+        t_{\text{day}} &= 2t_{\text{sun rise}} \\
+        \label{e:solartnight}
+        t_{\text{night}} &= 24 - t_{\text{day}}
+    \end{align}
+
+    where noon is $t=0$.
+Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
+The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
+
+    \begin{align}
+        \label{e:solarp0}
+        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
+        \label{e:solareo}
+        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
+    \end{align}
+
+    where
+
+    \[ \begin{array}{lcl}
+        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
+        R_{\text{sun}} & : & \text{Radius of the sun} \\
+        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
+        r_0 & : & \text{Average distance from earth to sun} \\
+        0.017 & : & \text{Eccentricity of earth-sun orbit}.
+    \end{array} \]
+
+    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
+Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
+
 \end{document}

### 2017-04-21 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -44,13 +44,13 @@ def fit_setup(Re_range):
         if r < 150:
             cl = dataf["CL"].values.astype(np.float)
             cd = dataf["CD"].values.astype(np.float)
-            CL.append(cl[cl >= 1.0])
-            CD.append(cd[cl >= 1.0])
+            CL.append(np.hstack([cl[cl >= 1.0]]*1))
+            CD.append(np.hstack([cd[cl >= 1.0]]*1))
         elif r < 200:
             cl = dataf["CL"].values.astype(np.float)
             cd = dataf["CD"].values.astype(np.float)
-            CL.append(cl[cl >= 0.8])
-            CD.append(cd[cl >= 0.8])
+            CL.append(cl[cl >= 0.9])
+            CD.append(cd[cl >= 0.9])
         else:
             CL.append(dataf["CL"].values.astype(np.float))
             CD.append(dataf["CD"].values.astype(np.float))
@@ -100,7 +100,7 @@ def plot_fits(re, cnstr, x, y):
     return fig, ax
 
 if __name__ == "__main__":
-    Re = np.arange(100, 750, 50)
+    Re = np.arange(150, 750, 50)
     X, Y = fit_setup(Re) # call fit(X, Y, 4, "SMA") to get fit
     np.random.seed(0)
     cn, err = fit(X, Y, 4, "SMA")
@@ -112,7 +112,7 @@ if __name__ == "__main__":
     else:
         df.to_csv("jho_fitdata.csv")
 
-    replot = np.array([100, 150, 200, 300, 350])
+    replot = np.array([150, 200, 300, 350, 400])
     # replot = np.array([300, 350, 400, 450, 500])
     F, A = plot_fits(replot, cn, X, Y)
     if len(sys.argv) > 1:

### 2017-04-25 docs/paperin_review.tex
tweak plotting functions to accomodate modeling and fitting changes

@@
 \usepackage{tikz}
+\usepackage{bm}
 \title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
@@
 $C_{d_0}$ \> non-wing drag coefficient \\
-$c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
+% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
 $c_{d_p}$ \> wing profile drag coefficient \\
-$c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
+% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
 $C_f$ \> skin friction coefficient \\
@@
 DOY \> day of the year \\
-$e$ \> Oswald efficiency factor \\
+$e$ \> span efficiency factor \\
 $E$ \> Young's Modulus \\ %[Pa] \\
@@
 $m'$ \> local wing mass \\ % [kg/m] \\
-$m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac-wing}}$ \> wing weight margin factor \\
+% $m_{\text{fac-wing}}$ \> margin factor\\
+$m_{\text{fac}}$ \> weight margin factor \\
 $m_{\text{motor}}$ \> electric motor mass \\
@@
 $S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear forces \\ % [N] \\
+$\mathcal{S}$ \> shear force \\ % [N] \\
 $S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
@@
 $W_{\text{engine}}$ \> engine weight \\ % [N] \\
+$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
 $W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
@@
 Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
-The optimization is solved multiple times to observe how aircraft size and weight are affected by different design parameters and requirements.
+The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
 
@@
 This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Similar trade offs are calculated for other requirements and performance metrics to determine under what conditions each platform is best suited to meet a set of requirements.
+Varying mission requirements reveals how the best architecture changes depending on the requirements.
 
@@
 Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is defined as a percentage which is the percentile wind speed. 
-The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,
+Therefore, the station keeping requirement is defined as the percentage of time the aircraft can remain on station, which is the percentile wind speed, $p_{\text{wind}}$. 
+The wind speed at a station is also a function of latitude, altitude, season or day of the year,
 
@@
 Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June will be used for the aircraft sizing and performance analysis presented in this paper.
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.A, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
+Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June respectively, will be used for the aircraft sizing and performance analysis presented in this paper.
 
@@
  \end{subfigmatrix}
- \caption{\textbf{Winds peak at summer and winter solstice in Southern and Northern Hemispheres respectively.\cite{wind}}}
+ \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
  \label{f:windvsmonth}
@@
 Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m depending on the engine.
+Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
 Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.
@@
 Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section IV.B.
+This is further discussed in Section IV.A.
 
@@
 A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
+The configuration is assumed to be static for this study.
 
@@
 
-These assumptions can be changed to evaulate different configuration.  Changes in configuration assumptions may require modification of models or equations.
-
 \section{Aircraft Physical Models}
@@
 Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
-Altitude can be post-calculated from the optimized air density.
+Altitude can be post-calculated from the optimized air density. (See Appendix A)
 
@@
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
 % The constraint on December wind speeds and air density at the 30 degree latitude is 
@@
     \label{e:solarbatt}
-    E_{\text{batt}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
+    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
 \end{align}
@@
 where $\eta_{\text{motor}} = 0.95$. 
-Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.
+Thus, for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
 The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix B to generate the fitted equations. 
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
 % Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
@@
 
+\subsubsection{Motor Weight}
+
+The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
+
+\begin{equation}
+    \label{e:electricengine}
+    P_{\text{max}} = B_{PM} m_{\text{motor}}
+\end{equation}
+
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
+
 \subsection{Gas Powered Aircraft Physical Models}
@@
 One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-$W_{\text{ave}}$ is assumed to be the geometric mean, defined as
+The average weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
 
@@
 \subsubsection{Engine Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{PM} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
 
@@
 
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
+    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix D for a discussion of the airfoil choice). 
     Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
@@
 
-    \begin{align}
-        \label{e:aerodragprof}
-        c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-                       &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-        Re &= \frac{\rho V S/b}{\mu}
-    \end{align}
+%     \begin{align}
+%         \label{e:aerodragprof}
+%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
+%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
+%         Re &= \frac{\rho V S/b}{\mu}
+%     \end{align}
 
@@
 
-\item[Standard Wing Bending Case]
+\item{Standard Wing Bending Case}
 
@@
 
-\item[Gust Loading Case]
+\item{Gust Loading Case}
 
@@
 
-Because the gust velocity is assumed to be vertical to the flight path, the local angle of attack is approximated by
+The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
 
@@
 
-\item[Discretized Beam Model]
+\item{Discretized Beam Model}
 
@@
 
-\item[Cap Spar for Bending Loads]
+\item{Cap Spar for Bending Loads}
 
 A cap spar is considered for the solar-electric and gas powered aircraft.
-Both aircraft optimization models share the same governing equations.
 A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
@@
 
-\subsubsection{Wing Skin Weight}
+\subsubsection{Additional Wing Weight}
 
@@
 where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
+
+Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight in the wing,
+
+\begin{equation}
+    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
+\end{equation}
+
+where $m_{\text{fac}} = 1.2$.
 
@@
 
-The gas and solar-electric powered aircraft optimization models were solved by minimizing take of weight for the requirements listed in Table~\ref{t:mreqs}. 
+The gas and solar-electric powered aircraft optimization models were solved by minimizing the take of weight for the requirements listed in Table~\ref{t:mreqs}. 
 Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
@@
 $\dot{h}$ [ft/min]           & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-                             &     & $\eta_{\text{charge}}$          & 0.98 \\
+$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
                              &     & $\eta_{\text{discharge}}$       & 0.98 \\
@@
 Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved x times and the solar-electric powered model was solved x times to produce Figure~\ref{f:latvsmtowtrade}.
-The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.
+The gas powered model was solved 63 times in 5.6178 seconds and the solar-electric powered model was solved 31 times in 2.5529 seconds to produce Figure~\ref{f:latvsmtowtrade}.
 
@@
 Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
-Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude.
+Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude. 
+Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
 
@@
     \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:spanvsendurance}
+	\label{f:mtowvsendurance}
 	\end{center}
@@
 Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
+Figure~\ref{f:battsolarcontour} was produced using 157 separate optimiation solutions that took a total of 14.379 seconds to solve.
 
@@
 By altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
+It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
 If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
@@
 
-\section*{Appendix A}
-
-\subsection{Wind Speed Constraints}
-
-The each wind speed constraint follows the following form
-
-\begin{equation}
-    \label{e:windspeedgen}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-\end{equation}
-
-Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
+\section*{Appendix C}
+
+\subsection{Altitude and relevant atmosphereic quantities}
+
+The altitude also determines atmospheric properties such as air density and viscosity.
+For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+
+For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
+
+\begin{align}
+    \label{e:tropopress}
+    p &= \rho R_{\text{spec}}T_{11} \\
+    \label{e:tropoalt}
+    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+\end{align}
+
+According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+\\
+
+\section*{Appendix B}
+
+\subsection{Convex Fitted Functions}
+
+Many constraints in this optimization study are convex equations that approximate data or other functions.  
+These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
+Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
+All fitted equations are a combination of \emph{max-affine} and \emph{softmax-affine} functions,
+
+\begin{align}
+    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
+    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
+\end{align}
+
+The log-space RMS error,
+
+\begin{equation}
+\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
+\end{equation}
+
+ is given throughout the paper for each equation, where $m$ is the number of data points.
+
+Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
+In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
+
+% The each wind speed constraint follows the following form
+% 
+% \begin{equation}
+%     \label{e:windspeedgen}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
+% \end{equation}
+% 
+% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
 Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-
-\tiny
-\begin{longtable}{lccccccccccccc} \\
-\label{t:windvals} \\
-    \caption{Wind Speed Function Coefficients and Exponents} \\
-    \toprule
-    \toprule
-    Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-    \midrule
-20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-\bottomrule
-\end{longtable}
-\normalsize
-
-\section*{Appendix B}
+% 
+% \tiny
+% \begin{longtable}{lccccccccccccc} \\
+% \label{t:windvals} \\
+%     \caption{Wind Speed Function Coefficients and Exponents} \\
+%     \toprule
+%     \toprule
+%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
+%     \midrule
+% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
+% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
+% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
+% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
+% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
+% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
+% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
+% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
+% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
+% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
+% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
+% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
+% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
+% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
+% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
+% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
+% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
+% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
+% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
+% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
+% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
+% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
+% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
+% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
+% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
+% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
+% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
+% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
+% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
+% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
+% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
+% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
+% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
+% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
+% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
+% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
+% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
+% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
+% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
+% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
+% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
+% \bottomrule
+% \end{longtable}
+% \normalsize
+
+% \section*{Appendix C}
+% 
+% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
+% 
+% \begin{align}
+%     \label{e:mcenter}
+%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
+%     \label{e:eq6}
+%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
+% \end{align}
+% 
+% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
+% 
+% Using moment and lift approximations
+% 
+% \begin{align}
+%     C_{m_{\text{w}}} &= constant \\
+%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
+%     \label{e:eq10}
+%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
+%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
+%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
+% \end{align}
+% 
+% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
+% 
+% \begin{equation}
+%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
+% \end{equation}
+% 
+% so that \eqref{e:eq10} can be written as
+% 
+% \begin{equation}
+%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
+% \end{equation}
+% 
+% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
+% 
+% \begin{align}
+%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
+%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
+% \end{align}
+% 
+% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
+% 
+% \begin{align}
+%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
+%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
+% \end{align}
+% 
+% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
+% 
+% \begin{equation}
+%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
+% \end{equation}
+% 
+% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
+% 
+% \begin{equation}
+%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
+% \end{equation}
+% 
+% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
+% 
+% \begin{equation}
+%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
+% \end{equation}
+% 
+% The pitch authority requirement is that at the forward-most c.g.
+position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
+Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
+% 
+% \begin{align}
+%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
+% \end{align}
+% 
+% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
+
+\section*{Appendix C}
+
+\subsection{Solar energy calculations}
+
+The total available solar energy per day is an integral of the available solar power during the course of the day,
+
+    \begin{equation}
+        \label{e:solares}
+        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
+    \end{equation}
+
+For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
+The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
+
+\begin{equation}
+    \label{e:solarp}
+    P_{\text{sun}} = P_0 \cos{\theta}
+\end{equation}
+
+The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
+
+    \begin{equation}
+        \label{e:solartheta}
+        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
+    \end{equation}
+
+ The declination angle $\Delta$, can be found using the relation\cite{solar}
+
+    \begin{align}
+        \label{e:solardelta}
+        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
+        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
+    \end{align}
+
+    where $\beta = 2\pi (\text{DOY}-1)/365$.
+    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
+
+    \begin{align}
+        \label{e:solartday}
+        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
+        \label{e:solarsunrise}
+        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
+        \label{e:solartday2}
+        t_{\text{day}} &= 2t_{\text{sun rise}} \\
+        \label{e:solartnight}
+        t_{\text{night}} &= 24 - t_{\text{day}}
+    \end{align}
+
+    where noon is $t=0$.
+Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
+The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
+
+    \begin{align}
+        \label{e:solarp0}
+        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
+        \label{e:solareo}
+        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
+    \end{align}
+
+    where
+
+    \[ \begin{array}{lcl}
+        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
+        R_{\text{sun}} & : & \text{Radius of the sun} \\
+        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
+        r_0 & : & \text{Average distance from earth to sun} \\
+        0.017 & : & \text{Eccentricity of earth-sun orbit}.
+    \end{array} \]
+
+    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
+Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
+
+\section*{Appendix D}
 
@@
 
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\section*{Appendix C}
-
-\subsection{Altitude and relevant atmosphereic quantities}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section*{Appendix D}
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
 \end{document}

### 2017-04-25 gassolar/environment/solar_irradiance.py

@@ -6,7 +6,7 @@ from gpfit.fit import fit
 import sys
 import pandas as pd
 plt.rcParams.update({'font.size':15})
-GENERATE = True
+GENERATE = False
 
 def get_Eirr(latitude, day, N=50.0):
     """

### 2017-04-25 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -112,8 +112,8 @@ if __name__ == "__main__":
     else:
         df.to_csv("jho_fitdata.csv")
 
-    replot = np.array([150, 200, 300, 350, 400])
-    # replot = np.array([300, 350, 400, 450, 500])
+    # replot = np.array([150, 200, 300, 350, 400])
+    replot = np.array([300, 350, 400, 450, 500])
     F, A = plot_fits(replot, cn, X, Y)
     if len(sys.argv) > 1:
         path = sys.argv[1]

### 2017-04-25 gassolar/fits/tailpolars/naca_cl0fits.py

@@ -8,7 +8,7 @@ from gpfit.fit import fit
 import os
 plt.rcParams.update({'font.size':15})
 
-GENERATE = True
+GENERATE = False
 REFAC = 1e6
 NACAFAC = 100.0
 

### 2017-04-25 gassolar/gas/disp_sens.py

@@ -46,7 +46,7 @@ if __name__ == "__main__":
     varnames = ["V_{wind}_Mission, Loiter, FlightSegment", "W_{pay}", "\\eta_{prop}", "BSFC_{min}", "t_Mission, Loiter", "N_{max}_Mission, AircraftLoading, WingLoading, ChordSparL"]
     latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
     sens_table(sols, varnames, filename="sens.generated.tex")
-    fig, ax = plot_sens(M, sols[2], varnames, latns=latns)
+    fig, ax = plot_sens(M, sols[2], varnames)
 
     if len(sys.argv) > 1:
         path = sys.argv[1]

### 2017-04-25 gassolar/gas/fits/DF70/fitDF70.py

@@ -12,7 +12,7 @@ plt.rcParams.update({'font.size':19})
 np.random.seed(0)
 PATH = (os.path.abspath(__file__).replace(os.path.basename(__file__), "")
         + os.sep)
-GENERATE = True
+GENERATE = False
 
 def plot_BSFCtoPower(power_csv, bsfc_csv):
     # Fitting BSFC vs. Power

### 2017-04-25 gassolar/gas/fits/powerlaw/engine_law.py

@@ -11,7 +11,7 @@ WEIGHT = 10.0
 POWER = 10.0
 PATH = (os.path.abspath(__file__).replace(os.path.basename(__file__), "")
         + os.sep)
-GENERATE = True
+GENERATE = False
 
 def plot_powerlaw(csv):
     df = pd.read_csv(csv)

### 2017-04-25 gassolar/gas/mtowendtrade.py

@@ -3,6 +3,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 from gassolar.gas.gas import Mission
 from gassolar.environment.wind_speeds import get_windspeed
+from gassolar.solar.battsolarcon import find_sols
 from gpkit.tools.autosweep import autosweep_1d
 import sys
 
@@ -11,8 +12,10 @@ plt.rcParams.update({'font.size':15})
 def mtow_plot(model):
     model.cost = model["MTOW"]
     fig, ax = plt.subplots()
-    x = np.linspace(1, 12, 500)
+    x = np.linspace(1, 15, 500)
     tol = 0.05
+    time = 0.0
+    nsolves = 0
     ws = []
     xs = []
     for p in [85, 90, 95]:
@@ -27,6 +30,8 @@ def mtow_plot(model):
         model.substitutions.update({"MTOW": 1000})
         model.cost = 1/model["t_Mission, Loiter"]
         sol = model.solve("mosek")
+        time += sol["soltime"]
+        nsolves += 1
         upper = sol("t_Mission, Loiter").magnitude
         xmin_ = x[x < upper + 0.03]
         xs.append(xmin_)
@@ -35,8 +40,13 @@ def mtow_plot(model):
         model.cost = model["MTOW"]
         bst = autosweep_1d(model, tol, model["t_Mission, Loiter"],
                            [1, xmin_[-1]], solver="mosek")
+        bsts = find_sols([bst])
+        sols = np.hstack([b.sols for b in bsts])
+        time += sum(np.unique([s["soltime"] for s in sols]))
+        nsolves += bst.nsols
         ws.append(bst.sample_at(xmin_)["cost"])
 
+    print "%d solves in %.4f seconds" % (nsolves, time)
     ax.fill_between(xs[0], ws[0],
                     np.append(ws[2], [1000]*(len(xs[0])-len(xs[2]))),
                     facecolor="r", edgecolor="None", alpha=0.3)

### 2017-04-25 gassolar/gassolarlat.py

@@ -2,6 +2,7 @@
 from gassolar.solar.solar import Mission as Msolar
 from gassolar.gas.gas import Mission as Mgas
 from gassolar.environment.wind_speeds import get_windspeed
+from gassolar.solar.battsolarcon import find_sols
 import matplotlib.pyplot as plt
 import numpy as np
 import os
@@ -22,6 +23,10 @@ def plot_lats():
     psolar = []
     pgas = []
     faillat = []
+    gtime = 0.0
+    gnsols = 0
+    stime = 0.0
+    snsols = 0.0
     for a in [80, 90, 95]:
         wg = []
         ws = []
@@ -36,6 +41,8 @@ def plot_lats():
                 Ms.cost = Ms["W_{total}"]
                 try:
                     sol = Ms.solve("mosek")
+                    stime += sol["soltime"]
+                    snsols += 1
                     ws.append(sol("W_{total}").magnitude)
                 except RuntimeWarning:
                     ws.append(np.nan)
@@ -55,6 +62,8 @@ def plot_lats():
                         Mg.substitutions.update({vk: wind})
             try:
                 sol = Mg.solve("mosek")
+                gtime += sol["soltime"]
+                gnsols += 1
                 wg.append(sol("MTOW").magnitude)
             except RuntimeWarning:
                 wg.append(np.nan)
@@ -62,6 +71,10 @@ def plot_lats():
 
         pgas.append(wg)
         psolar.append(ws)
+
+
+    print "Solar: %d solves in %.4f seconds" % (snsols, stime)
+    print "Gas: %d solves in %.4f seconds" % (gnsols, gtime)
 
     indl = psolar[0].index(max(psolar[0]))
     indh = psolar[2].index(max(psolar[2]))
@@ -81,11 +94,11 @@ def plot_lats():
     for i, p in enumerate(["80%", "90%", "95%"]):
         ax.annotate(p, xy=(36,pgas[i][np.where(lat==36)[0][0]]), xytext=(0.1,-20), textcoords="offset points", arrowprops=dict(arrowstyle="-"), fontsize=12)
 
-    ax.annotate("80%", xy=(25,psolar[0][np.where(lat==25)[0][0]]), xytext=(0.1,-20), textcoords="offset points", arrowprops=dict(arrowstyle="-"), fontsize=12)
[…]

### 2017-04-25 gassolar/solar/battsolarcon.py

@@ -10,6 +10,7 @@ N = 100
 plt.rcParams.update({'font.size':15})
 
 def find_sols(bsts):
+    runagain = False
     for s in bsts:
         if s.splits is None:
             continue
@@ -17,6 +18,10 @@ def find_sols(bsts):
             if sn in bsts:
                 continue
             bsts.append(sn)
+            runagain = True
+    if runagain:
+        return find_sols(bsts)
+    else:
         return bsts
 
 def plot_battsolarcon():
@@ -55,7 +60,7 @@ def plot_battsolarcon():
                                    solver="mosek")
                 bsts = find_sols([bst])
                 sols = np.hstack([b.sols for b in bsts])
-                stime = sum(np.unique([s["soltime"] for s in sols]))
+                time += sum(np.unique([s["soltime"] for s in sols]))
                 numsolves += bst.nsols
                 if lat % 4 == 0:
                     l = ax.plot(xmin_, bst.sample_at(xmin_)["cost"], "k",

### 2017-04-25 gassolar/solar/contourplots.py

@@ -5,17 +5,18 @@ import sys
 from solar import Mission
 from plotting import windalt_plot, labelLines
 from gpkit.tools.autosweep import autosweep_1d
+from battsolarcon import find_sols
 
 def plot_contours(path=None):
     N = 100
+    stime = 0.0
+    numsolves = 0
     plt.rcParams.update({'font.size':19})
     hmin = 250.0
     hmax = 400.0
     for av in [80, 85, 90]:
         for lat in [25, 30, 35]:
             zo = 10
-            etamin = 0.15
-            etamax = 0.4
             if lat == 25:
                 bs = range(30, 80, 5)
             elif lat == 30:
@@ -39,31 +40,44 @@ def plot_contours(path=None):
             lines = []
             midx = []
             for b in bs:
+                etamax = 0.4
+                etamin = 0.15
                 M.substitutions.update({"b_Mission, Aircraft, Wing": b})
                 M.substitutions.update({"\\eta_Mission, Aircraft, SolarCells":
                                         etamax})
                 sol = M.solve("mosek")
+                stime += sol["soltime"]
+                numsolves += 1
+                del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
                 if sol("h_{batt}").magnitude < hmin:
                     M.substitutions.update({"h_{batt}": hmin})
-                    del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
                     M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
                     sol = M.solve("mosek")
+                    stime += sol["soltime"]
+                    numsolves += sol["soltime"]
                     etamax = sol("\\eta_Mission, Aircraft, SolarCells")
-                else:
-                    etammax = 0.5
-                    del M.substitutions["\\eta_Mission, Aircraft, SolarCells"]
 
                 M.substitutions.update({"h_{batt}": hmax})
                 M.cost = M["\\eta_Mission, Aircraft, SolarCells"]
                 sol = M.solve("mosek")
+                stime += sol["soltime"]
+                numsolves += 1
                 etamin = sol("\\eta_Mission, Aircraft, SolarCells")
+
+                if etamin > etamax:
+                    continue
 
                 del M.substitutions["h_{batt}"]
                 M.cost = M["h_{batt}"]
                 xmin_ = np.linspace(etamin, etamax, 100)
                 tol = 0.01
[…]

### 2017-04-25 gassolar/solar/fits/arctan/arctan_fit.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import sys
 plt.rcParams.update({'font.size':15})
-GENERATE = True
+GENERATE = False
 
 def arctanfit():
     u = np.linspace(1e-15, 0.7, 100)

### 2017-04-25 gassolar/solar/plotting.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from gpfit.softmax_affine import softmax_affine
+from gpfit.max_affine import max_affine
 import os
 import pandas as pd
 
@@ -15,14 +16,18 @@ def windalt_plot(latitude, sol1=None, sol2=None):
                 np.hstack([[p/100.0]*len(den)
                            for p in range(75, 100, 5) + [99]])]).T
 
-    df = DF[DF["latitude"] == latitude]
-    params = np.append(np.hstack([[
-        np.log((df["c%d" % i]**(1/df["alpha"])).iloc[0]),
-        (df["e%d1" % i]/df["alpha"]).iloc[0],
-        (df["e%d2" % i]/df["alpha"]).iloc[0]] for i in range(1, 5)]),
-                       1/df["alpha"].iloc[0])
+    # df = DF[DF["latitude"] == latitude]
+    df = pd.read_csv(path + "windfits/windaltfit_lat%d.csv" % latitude)
+    ftype = df["ftype"].iloc[0]
+    K = df["K"].iloc[0]
+    if ftype == "SMA":
+        params = np.append(np.hstack([[
+            np.log((df["c%d" % i]**(1/df["a1"])).iloc[0]),
+            (df["e%d1" % i]/df["a1"]).iloc[0],
+            (df["e%d2" % i]/df["a1"]).iloc[0]] for i in range(1, K)]),
+                           1/df["a1"].iloc[0])
 
-    vwind = (np.exp(softmax_affine(x, params)[0])*100).reshape(6, 20)[3]
+        vwind = (np.exp(softmax_affine(x, params)[0])*100).reshape(6, 20)[3]
     fig, ax = plt.subplots()
     l = ax.plot(alt/1000.0, vwind*1.95384, linewidth=2)
     if sol1:

### 2017-04-25 docs/reviews/review_response.tex
edit review response

@@
 Given the overall positive feedback and our responses to the reviewers comments we respectfully submit that this manuscript be reconsidered as full paper instead of an engineering note.
+\\
 
@@
 We have made organizational changes to the paper, including a more detailed description of the requirements, GP approach, and results. 
-As with any aircraft design, we made many assumptions and explicitly stated them.
-We argue that a great benefit to the GP approach is the ability to change those assumptions and to resolve the problem quickly.
-We do this in the results section of the paper and believe this shows important insights about long-endurance UAVs.
+As with any aircraft design, assumptions had to be made.  
+We did our best to explicity state and justify major assumptions about the requirements.   
 We would be happy to respond to any sepecific assumptions that the referee feels are not supported. 
+We would like to point out that justifying all of our assumptions would significantly lengthen the paper.
+We also aim to show that because of the GP approach we can quickly change our assumptions and resolve. 
+We do this in the results section of the paper and believe this shows important insights about long-endurance UAVs.  \\
 
@@
 
-We thank the referee for this helpful comment.  We have since added a flow chart.
+We thank the referee for this helpful comment.  We have since added a flow chart. \\
 
@@
 
-We thank the referee for this comment.  We did cut multiple figures.
+We thank the referee for this comment.  We did cut multiple figures. \\
 
@@
 
-We thank the referee for this comment.  We included a vehicle sketch.
+We thank the referee for this comment.  We included a vehicle sketch.  \\
 
@@
 
-We thank the referee for this comment.  We did cut a few unnecessary equations to help shorten the paper.
+We thank the referee for this comment.  We did cut a few unnecessary equations to help shorten the paper. \\
 
@@
 We would like to thank the referee for this suggestion.  
-We claim that speed is o of great benefits of using GP and we should have absolutely included more information about our design variables, solve times, and number of analysis.
-We refer the referee to the results section of the revised paper where we have included a table of key design parameters and solve times and number of analysis throughout.
+We claim that speed is o of great benefits of using GP and we should have included more information about our design variables, solve times, and number of analysis.  
+We refer the referee to the results section of the revised paper where we have included a table of key design parameters and solve times and number of analysis throughout.  \\
 
@@
 We have since included bounds on each posynmial approxmiation to ensure that the solution does not enter a region where we do not have representative data.  
-We noticed that the solar-electric solution tended to operate a lower Reynolds numbers than was approximated by our posynomial approxmiation.
+We noticed when setting these bounds that the solar-electric solution tended to operate a lower Reynolds numbers than was approximated by our posynomial approxmiation.  
 We have since generated more drag data from XFOIL at low Reynolds numbers and re-generated the posynomial fits for the airfoil drag data. 
+We also implemented a post-solution ``check''  for any posynomial fit that relies on XFOIL data.  
+Once a solution has been found the optimized lift coefficient and Reynolds number are entered into XFOIL and the drag coefficient is compared.  
+We guarentee that all drag coefficients are within 5\% of the predicted XFOIL drag number.\\
 
@@
 If the referee feels that these two trade studies should be explored and included we would be happy to discuss this futher.
+\\
 
@@
 We have cut a few non-essential equations and graphs to try and shorten the paper.  
-However, because GP solves a list of constraints we can explicity state all of our assumptions.
-We feel that this is important to justify our results.
-We are willing to discuss cutting additionally sections/figures/equations if the referee has specific suggestions.
+However, because GP solves a list of constraints we felt it important to explicity state all of our equations to justify our results.  
+We are willing to discuss cutting additionally sections/figures/equations if the referee has specific suggestions. \\
 
@@
 Thus our assumption was conservative.  
-However, given that our results do not change drastically by including the Southern Hemisphere wind speed data we have chosen to included it to avoid misconceptions about our results.
+However, given that our results do not change drastically by including the Southern Hemisphere wind speed data we have chosen to included it to avoid misconceptions about our results. \\
 
     \item \emph{Sources for data presented in figures should be given in the captions.} \\
-        We thank the reader for this comment.  Citations have since been added to all capations where data is used.
+        We thank the reader for this comment.  Citations have since been added to all capations where data is used.  \\
 
@@
 
-    There is a second factor that is also going on, which is the trade off between wing weight and air density.  Because lift also scales equally with air density and wing area, like drag,
+    There is a second factor that is also going on, which is the trade off between wing weight and air density.  Because lift also scales equally with air density and wing area,
 
@@
 
-    If we assume that a large wing doesn't affect the weight at all, then we should see the aircraft fly at the minimum wind speed.  This is precisely the point we are trying to make in Figure 17.
+    If we assume that a large wing doesn't affect the weight at all, then we should see the aircraft fly at the minimum wind speed.  This is precisely the point we are trying to make in Figure 17.  \\
 
     \item \emph{I did not understand what the autocorrelation indicated in terms of performance. That is, I could not make sense of the meaning of Figs 4a and 5a.}  \\
-        We thank the referee for this comment.
-We realized that these autocorrelation plots, while interesting, are not actually pertinent to the optimization model and we chose to cut them from the paper.
+We thank the referee for this comment.  We realized that these autocorrelation plots, while interesting, are not actually included to the optimization model and we chose to cut them from the paper.  
 We initially included them because as we were doing the research we wanted to understand what a percentile wind speed meant operationally.  
 We were asking ourselves does the 90th percentile wind speed mean that everyday we will be blown off station 10\% of the time, or that there will be 5 days in the month during which it is impossible to station keep.
-The autocorrelation plots showed that the later was true.
-In other words, when there are high wind speeds they can occur for days a time.
-`
+The autocorrelation plots showed that the later was true.  In other words, when there are high wind speeds they can occur for days a time. \\
 
     \item \emph{When data is given for a certain latitude, it is not clear whether the implication is that there is no difference with longitude, or that this is the worst for any longitude.} \\
-        While wind speeds certianly vary with longitude, our claim is that the variation with longitude is minimial compared to the variation to latitude and can be ignored.
+While wind speeds certianly vary with longitude, our claim is that the variation with longitude is minimial compared to the variation to latitude and can be ignored.  \\
 
@@
 
-    We thank the referee for this supportive comment.
+    We thank the referee for this supportive comment. \\
 
@@
 We believe this to be a fair comparison and will hopefully address other comments made by this referee.
+\\
 
@@
 We chose not to explore this trade study in the paper because of the additional length to the paper.
-We would be happy to include a season trade study if the referee feels it is an important result.
+We would be happy to discuss includind a season trade study if the referee feels it is an important result.
+\\
 
@@
 
-        We thank the referee for noticing this error.  It has since been corrected.  Additionally, all altitude references in the paper are now in meters.
+        We thank the referee for noticing this error.  It has since been corrected.  Additionally, all altitude references in the paper are now in meters.\\
 
@@
 
-        We thank the referee for noticing this.  The data points have since been labeled and their source referenced.
+        We thank the referee for noticing this.  The data points have since been labeled and their source referenced.  \\
 
@@
         For the solar-electric aircraft putting batteries in the wings is much simpler than putting fuel in wings.  
-Batteries do not require a special fuel tank and have low volume density.
-Therefore, we believe that putting batteries in the wings for the solar-electric aircraft and fuel in a fuselage for the gas aircraft is a fair comparison.
+        Batteries do not require a special fuel tank to hold liquid and have low volume density.  
+        Therefore, we believe that putting batteries in the wings for the solar-electric aircraft and fuel in a fuselage for the gas aircraft is a fair comparison.  \\
 
@@
 
-        We thank the referee for this comment.
-We hope that previous comments about the requirements have address this concern.
+        We thank the referee for this comment.  We hope that previous comments about the requirements have address this concern.  
         Because the requirement is to be at an altitude > 4,500 meters we don't necessarily care where each aircraft flies so long as it meets the requirement.  
         In this respect, we believe it to be a fair comparison.  
-An obvious take away from the lapse rate of the gas engine is that any payloads requiring an altitude greater than 44,000 ft is that the solar-electric is the preferred option.
+        An obvious take away from the lapse rate of the gas engine is that any payloads requiring an altitude greater than 44,000 ft is that the solar-electric is the preferred option. \\
 
@@
 
-        We thank the referee for noticing this error.  It has been corrected.
+        We thank the referee for noticing this error.  It has been corrected. \\
 
@@
 See references 16 and 26 in the revised paper.
+\\
     \end{itemize}

### 2017-04-26 docs/paperin.tex
print to eps not pdf where possible

@@
 
-The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,
+Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
+Therefore, the station keeping requirement is defined as the percentage of time the aircraft can remain on station, which is the percentile wind speed, $p_{\text{wind}}$. 
+The wind speed at a station is also a function of latitude, altitude, season or day of the year,
 
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.eps}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
  \end{subfigmatrix}
- \caption{\textbf{Winds peak at summer and winter solstice in Southern and Northern Hemispheres respectively.\cite{wind}}}
+ \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
  \label{f:windvsmonth}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.eps}
     \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.eps}
     \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
@@
 	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=480,natheight=200]{flowchart.pdf}
+	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
     \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
@@
 	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=517,natheight=408]{simpleaircraft.pdf}
+	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
     \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
@@
 	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=358,natheight=334]{wbreak.pdf}
+	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
    \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
@@
 Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees. (See Appendix A) 
-The constraint on December wind speeds and air density at the 30 degree latitude is
-
-\begin{align}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-    \label{e:windfit35}
-    &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-\end{align}
-
-where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
+% The constraint on December wind speeds and air density at the 30 degree latitude is 
+% 
+% \begin{align}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
+%     \label{e:windfit35}
+%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
+% \end{align}
+% 
+% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
 The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-	\label{f:windfitl35}
-	\end{center}
-\end{figure}
+% 
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+% 	\label{f:windfitl35}
+% 	\end{center}
+% \end{figure}
 
@@
 	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
+	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.eps}
     \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
@@
 where $\eta_{\text{motor}} = 0.95$. 
-Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.
+Thus, for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
 The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
 % Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
@@
 
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
+\subsubsection{Motor Weight}
+
+The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
+
+\begin{equation}
+    \label{e:electricengine}
+    P_{\text{max}} = B_{PM} m_{\text{motor}}
+\end{equation}
+
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
     \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
     \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
@@
 
-\subsubsection{Standard Wing Bending Case}
+\begin{itemize}
+
+\item{Standard Wing Bending Case}
 
@@
 
-\subsubsection{Gust Loading Case}
+\item{Gust Loading Case}
 
@@
 
-\subsubsection{Discretized Beam Model}
+\item{Discretized Beam Model}
 
@@
 
-\subsubsection{Cap Spar for Bending Loads}
+\item{Cap Spar for Bending Loads}
 
 A cap spar is considered for the solar-electric and gas powered aircraft.
-Both aircraft optimization models share the same governing equations.
 A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
@@
 	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capsparcross.pdf}
+	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
     \caption{\textbf{Cross sectional view of a cap spar.}}
@@
 
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{carbonfiber}
-
-\subsection{Wing Skin Weight}
+where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
+
+\end{itemize}
+
+\subsubsection{Additional Wing Weight}
 
@@
 
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{carbonfiber} The wing skin is asssumed not to contribute to the bending stiffness.
-
-\subsection{Empennage}
+where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
+
+Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight in the wing,
+
+\begin{equation}
+    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
+\end{equation}
+
+where $m_{\text{fac}} = 1.2$.
+
+\subsubsection{Empennage}
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
     \caption{\textbf{Gas archicture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
     \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:spanvsendurance}
+	\label{f:mtowvsendurance}
 	\end{center}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
     \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.pdf}}
-  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.pdf}}
-  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.pdf}}
-  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.pdf}}
-  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.pdf}}
-  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.pdf}}
-  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.pdf}}
-  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.pdf}}
-  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.pdf}}
+  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
+  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
+  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
+  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
+  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
+  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
+  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
+  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
+  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
  \end{subfigmatrix}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
  \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
     \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
@@
 
-\section*{Appendix A}
-
-\subsection{Wind Speed Constraints}
-
-The each wind speed constraint follows the following form
-
-\begin{equation}
-    \label{e:windspeedgen}
-    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-\end{equation}
-
-Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
+\section*{Appendix C}
+
+\subsection{Altitude and relevant atmosphereic quantities}
+
+The altitude also determines atmospheric properties such as air density and viscosity.
+For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
+
+\begin{align}
+    \label{e:Talt}
+    T_{\text{atm}} = T_{\text{sl}} - Lh \\
+    \label{e:rhot}
+    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+\end{align}
+
+Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+
+\begin{equation}
+    \label{e:sutherland}
+    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+\end{equation}
+
+where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+
+For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
+
+\begin{align}
+    \label{e:tropopress}
+    p &= \rho R_{\text{spec}}T_{11} \\
+    \label{e:tropoalt}
+    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+\end{align}
+
+According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
+\\
+
+\section*{Appendix B}
+
+\subsection{Convex Fitted Functions}
+
+Many constraints in this optimization study are convex equations that approximate data or other functions.  
+These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
+Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
+All fitted equations are a combination of \emph{max-affine} and \emph{softmax-affine} functions,
+
+\begin{align}
+    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
+    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
+\end{align}
+
+The log-space RMS error,
+
+\begin{equation}
+\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
+\end{equation}
+
+ is given throughout the paper for each equation, where $m$ is the number of data points.
+
+Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
+In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
+
+% The each wind speed constraint follows the following form
+% 
+% \begin{equation}
+%     \label{e:windspeedgen}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
+% \end{equation}
+% 
+% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
 Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-
-\tiny
-\begin{longtable}{lccccccccccccc} \\
-\label{t:windvals} \\
-    \caption{Wind Speed Function Coefficients and Exponents} \\
-    \toprule
-    \toprule
-    Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-    \midrule
-20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-\bottomrule
-\end{longtable}
-\normalsize
-
-\section*{Appendix B}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmin.pdf}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmax.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
+% 
+% \tiny
+% \begin{longtable}{lccccccccccccc} \\
+% \label{t:windvals} \\
+%     \caption{Wind Speed Function Coefficients and Exponents} \\
+%     \toprule
+%     \toprule
+%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
+%     \midrule
+% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
+% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
+% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
+% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
+% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
+% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
+% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
+% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
+% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
+% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
+% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
+% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
+% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
+% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
+% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
+% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
+% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
+% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
+% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
+% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
+% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
+% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
+% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
+% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
+% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
+% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
+% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
+% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
+% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
+% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
+% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
+% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
+% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
+% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
+% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
+% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
+% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
+% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
+% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
+% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
+% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
+% \bottomrule
+% \end{longtable}
+% \normalsize
 
@@
 
-\subsection{Altitude and relevant atmosphereic quantities}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
+\subsection{Solar energy calculations}
+
+The total available solar energy per day is an integral of the available solar power during the course of the day,
+
+    \begin{equation}
+        \label{e:solares}
+        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
+    \end{equation}
+
+For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
+The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
+
+\begin{equation}
+    \label{e:solarp}
+    P_{\text{sun}} = P_0 \cos{\theta}
+\end{equation}
+
+The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
+
+    \begin{equation}
+        \label{e:solartheta}
+        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
+    \end{equation}
+
+ The declination angle $\Delta$, can be found using the relation\cite{solar}
+
+    \begin{align}
+        \label{e:solardelta}
+        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
+        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
+    \end{align}
+
+    where $\beta = 2\pi (\text{DOY}-1)/365$.
+    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
+
+    \begin{align}
+        \label{e:solartday}
+        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
+        \label{e:solarsunrise}
+        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
+        \label{e:solartday2}
+        t_{\text{day}} &= 2t_{\text{sun rise}} \\
+        \label{e:solartnight}
+        t_{\text{night}} &= 24 - t_{\text{day}}
+    \end{align}
+
+    where noon is $t=0$.
+Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
+The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
+
+    \begin{align}
+        \label{e:solarp0}
+        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
+        \label{e:solareo}
+        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
+    \end{align}
+
+    where
+
+    \[ \begin{array}{lcl}
+        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
+        R_{\text{sun}} & : & \text{Radius of the sun} \\
+        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
+        r_0 & : & \text{Average distance from earth to sun} \\
+        0.017 & : & \text{Eccentricity of earth-sun orbit}.
+    \end{array} \]
+
+    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
+Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
+
+\section*{Appendix D}
+
+\subsection{Discussion on the use of the \emph{JH01} airfoil}
+
+The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
+Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
+The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     % \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmin.eps}}
+     % \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmax.eps}}
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
+ \end{subfigmatrix}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
+ \label{f:jhcps}
+\end{figure}
 

### 2017-04-26 docs/paperin_review.tex
print to eps not pdf where possible

@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.eps}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
  \end{subfigmatrix}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
     \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
@@
     Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because only Equations~\eqref{e:solarreq}-\eqref{e:solarssolar} are GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
+Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
 
@@
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmin.pdf}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmax.pdf}}
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
  \end{subfigmatrix}

### 2017-04-26 gassolar/environment/wind_fit.py

@@ -145,7 +145,7 @@ if __name__ == "__main__":
         if not GENERATE:
             if not yfit is None:
                 fig, ax = plot_fits(X, Y, yfit, l)
-                fig.savefig(path + "windfitl%d.pdf" % l, bbox_inches="tight")
+                fig.savefig(path + "windfitl%d.eps" % l, bbox_inches="tight")
                 plt.close()
 
 #     if GENERATE:

### 2017-04-26 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -96,6 +96,7 @@ def plot_fits(re, cnstr, x, y):
             i += 1
     ax.set_xlabel("$C_L$")
     ax.set_ylabel("$c_{d_p}$")
+    ax.legend(loc=2)
     ax.grid()
     return fig, ax
 
@@ -117,6 +118,6 @@ if __name__ == "__main__":
     F, A = plot_fits(replot, cn, X, Y)
     if len(sys.argv) > 1:
         path = sys.argv[1]
-        F.savefig(path + "jho1polarfit1.pdf", bbox_inches="tight")
+        F.savefig(path + "jho1polarfit1.eps", bbox_inches="tight")
     else:
-        F.savefig("jho1polarfit1.pdf", bbox_inches="tight")
+        F.savefig("jho1polarfit1.eps", bbox_inches="tight")

### 2017-04-26 gassolar/fits/tailpolars/naca_cl0fits.py

@@ -112,7 +112,7 @@ if __name__ == "__main__":
     F, A = plot_fits(NACA, cn, X, Y)
     if len(sys.argv) > 1:
         path = sys.argv[1]
-        F.savefig(path + "taildragpolar.pdf", bbox_inches="tight")
+        F.savefig(path + "taildragpolar.eps", bbox_inches="tight")
     else:
-        F.savefig("taildragpolar.pdf", bbox_inches="tight")
+        F.savefig("taildragpolar.eps", bbox_inches="tight")
 

### 2017-04-26 gassolar/gas/ldratios.py

@@ -99,7 +99,7 @@ if __name__ == "__main__":
         path = sys.argv[1]
         for num in nums:
             fig, ax = ld_plot(M, num)
-            fig.savefig(path + "polarmission%d.pdf" % num, bbox_inches="tight")
+            fig.savefig(path + "polarmission%d.eps" % num, bbox_inches="tight")
     else:
         for num in nums:
             M = Mission()

### 2017-04-26 gassolar/solar/battsolarcon.py

@@ -90,6 +90,6 @@ if __name__ == "__main__":
     fig, ax = plot_battsolarcon()
     if len(sys.argv) > 1:
         path = sys.argv[1]
-        fig.savefig(path + "battsolarcontour.pdf", bbox_inches="tight")
+        fig.savefig(path + "battsolarcontour.eps", bbox_inches="tight")
     else:
         fig.savefig("battsolarcontour.pdf", bbox_inches="tight")

### 2017-04-26 gassolar/solar/contourplots.py

@@ -138,7 +138,7 @@ def plot_contours(path=None):
             ax.set_xlim([0.15, 0.4])
             ax.set_ylim([250, 400])
             if path:
-                fig.savefig(path + "bcontourl%da%d.pdf" % (lat, av),
+                fig.savefig(path + "bcontourl%da%d.eps" % (lat, av),
                             bbox_inches="tight")
             else:
                 fig.savefig("bcontourl%da%d.pdf" % (lat, av))

### 2017-04-26 gassolar/solar/windstruct.py

@@ -41,6 +41,6 @@ if __name__ == "__main__":
         fig, ax = plot_structm(num)
         if len(sys.argv) > 1:
             path = sys.argv[1]
-            fig.savefig(path + "windaltoper%d.pdf" % num, bbox_inches="tight")
+            fig.savefig(path + "windaltoper%d.eps" % num, bbox_inches="tight")
         else:
             fig.savefig("windaltoper.pdf", bbox_inches="tight")

### 2017-04-26 docs/paperin_review.tex
minor changes for compiling reasons

@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
+  \subfigure[35th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
+  \subfigure[35th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
+  \subfigure[35th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
+  \subfigure[30th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
+  \subfigure[30th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
+  \subfigure[30th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
+  \subfigure[25th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
+  \subfigure[25th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
+  \subfigure[25th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
  \end{subfigmatrix}

### 2017-05-02 docs/paperin_review.tex
incorporate whoburg review edits

@@
 $h$ \> flight altitude \\ % [ft] \\
+$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
 $h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
@@
 $P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minium operational solar power \\ % [Whr/m$^2$] \\
+$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
 $P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
@@
 $\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horiztonal tail thickness to chord ratio \\
+$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
 $\tau_t$ \> wing thickness to chord ratio \\
@@
 Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures even more unclear.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two archictures for a given mission.
+The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
+
+This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
 Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
+Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
 The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
@@
 
-Using this methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
+Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
 The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
 Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
+The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
 Varying mission requirements reveals how the best architecture changes depending on the requirements.
@@
 
-A comparision of the gas and solar-elecrtric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which archicture is best suited to meet various sets of requirements.
+A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
+These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
 A discussion of each requirement is included in this section.
@@
 
-Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is defined as the percentage of time the aircraft can remain on station, which is the percentile wind speed, $p_{\text{wind}}$. 
+Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
+Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
 The wind speed at a station is also a function of latitude, altitude, season or day of the year,
@@
 
-\subsection{Endurance}
-
-For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement.
-Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
+\subsection{Endurance and Season}
+
+To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
+If it can do so, it automatically exceeds the endurance requirement. 
 To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.A, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June respectively, will be used for the aircraft sizing and performance analysis presented in this paper.
 
@@
  \begin{subfigmatrix}{2}% number of columns
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
      \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.eps}}
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
  \end{subfigmatrix}
@@
 \end{figure}
+
+For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
+This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
+
+Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
 
@@
 
-The altitude requirement is generally set by a minimum payload required to meet a certain coverage
+One altitude requirement is a minimum height required to meet a certain coverage footprint,
 
@@
 
-Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.
+Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement, 4,600 m, or at high altitudes between 16,700-20,000 m.
+Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
+Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
 
@@
 
-It is assumed that long-endurance aircraft will have a requirement to fly within a band of latitudes (i.e. $\pm$45 degrees latitude).  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees latitude. 
+It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
+For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
 Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
@@
 
-Sizing long-edurance aircraft for the stated requirements was accomplished using geometric programming. 
+Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
 Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
@@
 
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
+To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
 Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
@@
 
-Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
-Common characteristics between the two models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
+For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
+Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
 The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
@@
 The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
+A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
 The configuration is assumed to be static for this study.
@@
 	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
+    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
 	\label{f:simpleaircraft}
@@
 Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, higher fidelity models can be implemented.
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds also vary with latitude, several constraints are needed to optimize altitude. 
-Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg et. al\cite{fitting}.
-Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
-Altitude can be post-calculated from the optimized air density. (See Appendix A)
-
-For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees.
-Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase nearly linearly up to 9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.eps}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{PM} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The average weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this.
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24} 
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptcal Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plys of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].
-Equation~\eqref{e:powerlaw} was found using techniques described in Hoburg et.
-al\cite{fitting}.
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two affects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power,
-
-\begin{equation}
-    \label{e:rpmtobsfc}
-    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-\end{equation}
-
-where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the GP-compatible approximation to the manufacturing data from RCV engines is presented in Figure~\ref{f:powervsweightfit}.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powervsweightfit} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
+To obtain higher fidelity results from the optimization, more detailed models can be implemented.
 
@@
 
-where the Oswald efficiency factor is assumed to be constant, $e=0.9$. 
+where the span efficiency factor is assumed to be constant, $e=0.9$. 
 The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
 
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix D for a discussion of the airfoil choice). 
+    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
     Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
@@
 % By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known pramaters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
+% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
 % A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
@@
 % 
-% An analysis of the max take off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take off weight in both the gas and solar-electric powered optimization models at different latitudes. 
+% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
+% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
 % For this study, the gas-powered aircraft is constrained to fly for 8 days. 
@@
 %  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take off weight and latitude.
-Plots show required maximum take off weight required to operate between a band of latitudes.
+%  \caption{\textbf{ Trade study of max take-off weight and latitude.
+Plots show required maximum take-off weight required to operate between a band of latitudes.
 A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
@@
 % While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take off weight for different endurance requirements to understand what endurance limits are possible. 
+% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
 % 
@@
 % 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take off weight for a gas powered archicture. }}
+%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
 % 	\label{f:mtowvsendsimple}
@@
 
-\begin{itemize}
-
-\item{Standard Wing Bending Case}
-
-The distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing
+%\begin{itemize}
+
+\paragraph{\textbf{Standard Wing Bending Case}} 
+The distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing.
 The wing loading distribution, $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
@@
 
-\item{Gust Loading Case}
+\paragraph{\textbf{Gust Loading Case}}
 
@@
 
-\item{Discretized Beam Model}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input.
-Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending}
+\paragraph{\textbf{Discretized Beam Model}}
+
+Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
 
@@
     \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y 
-\end{align}
-
-where the Young's Modulous of carbon fiber is $E = 20$ [MPa].
-
-\item{Cap Spar for Bending Loads}
+    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
+\end{align}
+
+where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
+
+\textbf{Cap Spar for Bending Loads} \\
 
@@
 
-The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
+The ultimate tensile strength for unidirectional carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
 
-\end{itemize}
+% \end{itemize}
 
@@
 
-Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight in the wing,
+Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
 
@@
 
-The tail boom has a diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E$, and density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$], and length $l_{\text{h}}$. 
+The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, and density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
 The total mass and root bending inertia are imposed in the optimization model as
@@
 % 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take off weight.  
+% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
 % The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
@@
 
+\subsection{Solar-Electric Aircraft Physical Models}
+
+\subsubsection{Wind Speeds}
+
+There exists an optimum flight altitude for solar-electric powered aircraft.  
+The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
+To accomplish this, constraints relating wind speed to altitude are imposed. 
+This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
+
+Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
+To handle this, a multipoint set of latitude constraints is imposed. 
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
+% The constraint on December wind speeds and air density at the 30 degree latitude is 
+% 
+% \begin{align}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
+%     \label{e:windfit35}
+%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
+% \end{align}
+% 
+% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
+The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
+% 
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+% 	\label{f:windfitl35}
+% 	\end{center}
+% \end{figure}
+
+For the gas powered aircraft, the altitude is not optimized.  
+Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
+Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
+Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
+For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
+
+\subsubsection{Solar-Electric Power}
+
+At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
+
+\begin{equation}
+    \label{e:solarfunc}
+    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
+\end{equation}
+
+To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
+
+\begin{align}
+    \label{e:solarreq}
+    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
+    \label{e:solarbatt}
+    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
+\end{align}
+
+where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
+An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
+Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.eps}
+    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
+	\label{f:lat30}
+	\end{center}
+\end{figure}
+
+In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
+
+\begin{align}
+    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
+    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
+\end{align}
+
+where $\eta_{\text{motor}} = 0.95$. 
+Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
+The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
+% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
+%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
+%  \label{f:energyapprox}
+% \end{figure}
+
+For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
+The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
+
+\begin{equation}
+    \label{e:solarssolar}
+    S_{\text{solar}} \leq f_{\text{solar}}S.
+\end{equation}
+
+\subsubsection{Motor Weight}
+
+The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
+
+\begin{equation}
+    \label{e:electricengine}
+    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
+\end{equation}
+
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
+
+\subsection{Gas Powered Aircraft Physical Models}
+
+\subsubsection{Breguet Endurance}
+
+A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
+For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
+
+\begin{equation}
+    \label{e:breguetendurance}
+    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
+\end{equation}
+
+% The derivation begins with the differential form of Breguet Range,\cite{br2}
+%
+% \begin{equation}
+%     \label{e:breguetdiff}
+%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
+% \end{equation}
+% 
+% Using the definition of $\text{BSFC}$
+% 
+% \begin{equation}
+%     \label{e:brBSFC}
+%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
+% \end{equation}
+% 
+% Equation~\eqref{e:breguetdiff} can be written as
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
+% \end{equation}
+
+This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
+The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
+
+% Dividing by $W$,
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
+% \end{equation}
+% 
+% and integrating, the Breguet Range equation can be expressed as
+% 
+% \begin{equation}
+%     \label{e:be1}
+%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
+% \end{equation}
+% 
+% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
+
+\begin{equation}
+    \label{e:gpmean}
+    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
+\end{equation}
+
+    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
+
+\begin{align}
+    \label{e:brzbre}
+    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
+    \label{e:brtaylor}
+    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
+\end{align}
+
+    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
+For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
+    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this.
+
+\begin{align}
+    \label{e:slfweightd}
+    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
+    \label{e:brzbred}
+    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
+    \label{e:brtaylord}
+    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24} 
+    \end{align}
+
+    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
+
+\subsubsection{Elliptical Fuselage}
+
+For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
+The fuselage will increase the overall weight and drag of the aircraft.
+The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
+
+The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
+
+\begin{equation}
+    \label{e:fusevol}
+    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
+\end{equation}
+
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
+
+\begin{equation}
+    \label{e:fusevol2}
+    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
+\end{equation}
+
+where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
+Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
+
+\begin{equation}
+    \label{e:fusesa}
+    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
+\end{equation}
+
+The weight of the fuselage is constrained by
+
+\begin{equation}
+    \label{e:fuseweight}
+    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
+
+\begin{align}
+    \label{e:fusedrag}
+    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
+    C_f &\geq \frac{0.455}{Re^{0.3}}
+\end{align}
+
+where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
+
+\begin{equation}
+    \label{e:fuseform}
+    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
+\end{equation}
+
+\subsubsection{Engine Weight}
+
+The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
+
+\begin{equation}
+    \label{e:powerlaw}
+    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
+\end{equation}
+
+where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
+Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
+    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
+	\label{f:powervsweightfit}
+	\end{center}
+\end{figure}
+
+\subsubsection{Gas Powered Engine Performance}
+
+Two characteristics of gas engines affect the performance of long-endurance aircraft.  
+The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
+The second is the lapse rate.  
+Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
+To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
+RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
+It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
+This engine performance model is only valid for internal combustion engines.
+
+To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
+Using this approximation the required shaft power determines the $\text{BSFC}$.
+A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
+
+% \begin{equation}
+%     \label{e:rpmtobsfc}
+%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
+% \end{equation}
+
+% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
+    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
+	\label{f:powertobsfc}
+	\end{center}
+\end{figure}
+
+The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
+
+\begin{equation}
+    \label{e:lapse}
+    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
+\end{equation}
+
+The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
+
+\begin{equation}
+    \label{e:lapsefit}
+    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
+\end{equation}
+
+\subsubsection{Climb Constraints}
+
+Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
+The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
+
+\begin{equation}
+    \label{e:climbrate}
+    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
+\end{equation}
+
+The climb rate affects the required thrust, and therefore the required power during climb,
+
+\begin{equation}
+    \label{e:climb}
+    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
+\end{equation}
+
+where $W$ is the weight of the aircraft during climb.
+
 \section{Results}
@@
 The gas and solar-electric powered aircraft optimization models were solved by minimizing the take of weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take off weight of 320.5 lbs.
+Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
+The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
+The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 320.5 lbs.
 According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
@@
 $\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.5  & $\tau$                       & 0.115 \\
+$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
 $\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
@@
 Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}$ [ft/min]           & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
+$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
 $\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
@@
 
-Important trade offs between the gas powered and solar-electric powered archictures are highlighted and quantified by changing the latitude and endurance endurance requirements.   
-Both optimization models were solved by minimizing the max take off weight across different bands of latitudes. 
+Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
+Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
 Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds and the solar-electric powered model was solved 31 times in 2.5529 seconds to produce Figure~\ref{f:latvsmtowtrade}.
+The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
 
@@
 	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas archicture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
+    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
     \label{f:latvsmtowtrade}
@@
 
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$30 degrees latitude in 90th percentile wind speeds.  
+One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-The gas powered aircraft curves level off around 38 degrees latitude because wind speeds are highest at that latitude at 4,500 m. 
-Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
 On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
@@
 While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude. 
+Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
+Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
 Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
@@
 
-\subsection{Changing Input Values}
+\subsection{Changing Parameter Values}
 
 The previous results are dependent on the assumed input values and parameters.  
-Changing the input values can help show where the designs becomes infeasible. 
-Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
+Changing parameter values can help show where the designs becomes infeasible. 
+As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
 By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
@@
 Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimiation solutions that took a total of 14.379 seconds to solve.
+Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
 
@@
 	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
+    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
 	\label{f:battsolarcontour}
@@
 
-By altering the aircraft structural model it can be observed how air density trades for wing weight. 
+Insight into the design space can also be gained by changing the physical modeling assumptions.
+For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
 It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
@@
 For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take off weight.
+Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
 
@@
 
-Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-For example, both batteries and solar cells are expensive and important to the design.  
-The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-The ratio of their magnitudes is 1.76.  
-Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
+% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
+% For example, both batteries and solar cells are expensive and important to the design.  
+% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
+% The ratio of their magnitudes is 1.76.  
+% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
 
@@
 
-\section*{Appendix C}
-
-\subsection{Altitude and relevant atmosphereic quantities}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
+% \section*{Appendix A}
+% 
+% \subsection{Altitude and relevant atmosphereic quantities}
+% 
+% The altitude also determines atmospheric properties such as air density and viscosity.
 For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+% 
+% \begin{align}
+%     \label{e:Talt}
+%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
+%     \label{e:rhot}
+%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+% \end{align}
+% 
+% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+% 
+% \begin{equation}
+%     \label{e:sutherland}
+%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+% \end{equation}
+% 
+% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+% 
+% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
+% 
+% \begin{align}
+%     \label{e:tropopress}
+%     p &= \rho R_{\text{spec}}T_{11} \\
+%     \label{e:tropoalt}
+%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+% \end{align}
+% 
+% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
 Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
@@
 
-\section*{Appendix B}
-
-\subsection{Convex Fitted Functions}
-
-Many constraints in this optimization study are convex equations that approximate data or other functions.  
+\section*{Appendix A}
+
+\textbf{Convex Fitted Functions} \\
+
+Some constraints in this optimization study are convex equations that approximate data or other functions.  
 These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
 Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are a combination of \emph{max-affine} and \emph{softmax-affine} functions,
+All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
 
@@
 
+\section*{Appendix B} \\
+
+\textbf{Discussion on the use of the \emph{JH01} airfoil} \\
+
+The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
+Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
+The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
+ \end{subfigmatrix}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
+ \label{f:jhcps}
+\end{figure}
+
 \section*{Appendix C}
 
-\subsection{Solar energy calculations}
+\textbf{Solar energy calculations} \\
 
@@
 
-\section*{Appendix D}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
 \end{document}

### 2017-05-02 gassolar/fits/jho1polars/jho1_polarfits.py

@@ -90,9 +90,10 @@ def plot_fits(re, cnstr, x, y):
     fig, ax = plt.subplots()
     i = 0
     for r, cl, cd, fi in zip(xre, xcl, cds, cdf):
-        if int(np.round(r)/1000) in re:
+        roundre = int(np.round(r)/1000)
+        if roundre in re:
             ax.plot(cl, cd, "o", mec=colors[i], mfc="none", mew=1.5)
-            ax.plot(cl, fi, c=colors[i], label="Re = %dk" % r, lw=2)
+            ax.plot(cl, fi, c=colors[i], label="Re = %dk" % roundre, lw=2)
             i += 1
     ax.set_xlabel("$C_L$")
     ax.set_ylabel("$c_{d_p}$")

### 2017-05-02 gassolar/gas/ldratios.py

@@ -25,6 +25,7 @@ def ld_plot(model, num):
         else:
             model.substitutions.update({vk: wind})
     model.substitutions.update({"MTOW": 200})
+    model.substitutions.update({"N_{max}_Mission, AircraftLoading, WingLoading, GustL":  0.1})
     sol = model.solve("mosek")
     re = sol("Re_Mission, Loiter, FlightSegment, AircraftPerf, WingAero")[-1]
 

### 2017-05-08 docs/paperin.tex
final edits

@@
 $K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-$L$ \> atmospheric lapse rate \\ % [K/m] \\
+% $L$ \> atmospheric lapse rate \\ % [K/m] \\
 $L_\text{h}$ \> horizontal tail lift \\ % [N] \\
@@
 $\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
 $n$ \> number of wing segments \\
@@
 $N_{\text{max}}$ \> safety load factor\\
-$p$ \> air pressure \\ % [Pa]\\
-$p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
+% $p$ \> air pressure \\ % [Pa]\\
+% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
 $p_{\text{wind}}$ \> percentile wind speed \\
@@
 $Re_{\text{boom}}$ \> tail boom Reynolds number \\
-$Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-$Re_{\text{v}}$ \> vertical tail Reynolds number \\
+% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
+% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
 $S$ \> wing planform area \\ % [m$^2$]\\
@@
 $T$ \> aircraft thrust \\ % [N] \\
-$T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-$T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-$T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
+% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
+% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
+% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
 $V$ \> true airspeed \\ % [m/s] \\
@@
 $W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{empennage}}$ \> empennage weight \\ % [N] \\
+$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
 $W_{\text{engine}}$ \> engine weight \\ % [N] \\
@@
 $w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
+$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
 $W_{\text{payload}}$ \> payload weight \\ % [N] \\
@@
 $\eta_{\text{solar}}$ \> solar cell efficiency \\
-$\mu$ \> air viscosity \\ % [kg/m/s] \\
-$\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
+% $\mu$ \> air viscosity \\ % [kg/m/s] \\
+% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
 $\phi$ \> latitude \\
@@
 Payload & 10 lbs\\
-Station Keeping & 90\% \\
+Station Keeping & 90\% winds \\
 Endurance & $>6$ days\\
@@
 
-\subsection{Endurance}
-
-For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
-For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement.
-Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.A, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June respectively, will be used for the aircraft sizing and performance analysis presented in this paper.
-
-\begin{figure}[H]
+To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
+If it can do so, it automatically exceeds the endurance requirement. 
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
+
+\begin{figure}[h!]
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.eps}}
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
  \end{subfigmatrix}
@@
 
-Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[H]
+Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
+Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
+Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
+
+\begin{figure}[h!]
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
     \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
     \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
@@
 Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section IV.A.
+This is further discussed in Section V.B.
 
@@
 
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assummed values are expressed as constraints on the GP.\cite{hoburgthesis} 
-Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
+To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
+Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
 Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
@@
 The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
+A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
 The configuration is assumed to be static for this study.
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
 	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas archictures respectively.  Black are shared characteristics.}}
+    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
 	\label{f:simpleaircraft}
@@
 
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.75$.
+    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
 
@@
 
-where the Oswald efficiency factor is assumed to be constant, $e=0.9$. 
+where the span efficiency factor is assumed to be constant, $e=0.9$. 
 The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
 
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix D for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
+    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
+    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
     The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
@@
 
-\begin{itemize}
-
-\item{Standard Wing Bending Case}
-
-The distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing
-The wing loading distribution, $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
+%\begin{itemize}
+
+\paragraph{\textbf{Standard Wing Bending Case}} 
+The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
+The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
 
@@
 
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$.
-$N_{\text{max}}=1$ corresponds to steady level flight.
-The local chord for a constant tapered wing\cite{bending} is
-
-\begin{equation}
-    \label{e:localchord}
-    \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-\end{equation}
-
-where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
+where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
+Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
+
+% \begin{equation}
+%     \label{e:localchord}
+%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
+% \end{equation}
+
+% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
 
@@
     \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
+    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
 \end{equation}
@@
 \begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{empennage}}.
+    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
 \end{equation}
@@
 \begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{empennage}}.
-\end{equation}
-
-\item{Gust Loading Case}
+    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
+\end{equation}
+
+\paragraph{\textbf{Gust Loading Case}}
 
 Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\eqref{e:qbar}
+To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
 
@@
 and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-
 The gust velocity has an assumed profile along the wing\cite{acgust},
@@
     \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y 
-\end{align}
-
-where the Young's Modulous of carbon fiber is $E = 20$ [MPa].
-
-\item{Cap Spar for Bending Loads}
-
+    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
+\end{align}
+
+where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
+
+\textbf{Cap Spar for Bending Loads} 
 A cap spar is considered for the solar-electric and gas powered aircraft.
@@
     \label{e:defcont}
-    w_n &\leq w_{\text{max}}.
-\end{align}
-
-The ultimate compressive strength for carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
+    w_n &\leq w_{\text{max}}
+\end{align}
+
+where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
 
-Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight in the wing,
+Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
 
@@
 
-The tail boom has a diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E$, and density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$], and length $l_{\text{h}}$. 
+The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
 The total mass and root bending inertia are imposed in the optimization model as
@@
 
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition, $V_{\text{h}} = 0.45$,\cite{aircraftrules}
+The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
 
@@
 
-The vertical tail is sized to meet a conservative tail volume coefficient, $V_{\text{v}}= 0.04$,\cite{aircraftrules}
+The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
 
@@
 
+\subsection{Solar-Electric Aircraft Physical Models}
+
+\subsubsection{Wind Speeds}
+
+There exists an optimum flight altitude for solar-electric powered aircraft.  
+The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
+To accomplish this, constraints relating wind speed to altitude are imposed. 
+This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
+
+Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
+To handle this, a multipoint set of latitude constraints is imposed. 
+The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
+% The constraint on December wind speeds and air density at the 30 degree latitude is 
+% 
+% \begin{align}
+%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
+%     \label{e:windfit35}
+%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
+% \end{align}
+% 
+% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
+The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
+% 
+% \begin{figure}[H]
+% 	\begin{center}
+% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
+% 	\label{f:windfitl35}
+% 	\end{center}
+% \end{figure}
+
+For the gas powered aircraft, the altitude is not optimized.  
+Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
+Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
+Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
+For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
+
+\subsubsection{Solar-Electric Power}
+
+At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
+
+\begin{equation}
+    \label{e:solarfunc}
+    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
+\end{equation}
+
+To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
+
+\begin{align}
+    \label{e:solarreq}
+    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
+    \label{e:solarbatt}
+    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
+\end{align}
+
+where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
+An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
+Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
+    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
+	\label{f:lat30}
+	\end{center}
+\end{figure}
+
+In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
+
+\begin{align}
+    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
+    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
+\end{align}
+
+where $\eta_{\text{motor}} = 0.95$. 
+Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
+The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
+% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
+% 
+% \begin{figure}[H]
+%  \begin{subfigmatrix}{2}% number of columns
+%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
+%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
+%  \end{subfigmatrix}
+%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
+%  \label{f:energyapprox}
+% \end{figure}
+
+For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
+The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
+
+\begin{equation}
+    \label{e:solarssolar}
+    S_{\text{solar}} \leq f_{\text{solar}}S.
+\end{equation}
+
+\subsubsection{Motor Weight}
+
+The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
+
+\begin{equation}
+    \label{e:electricengine}
+    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
+\end{equation}
+
+where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
+
+\subsection{Gas Powered Aircraft Physical Models}
+
+\subsubsection{Breguet Endurance}
+
+A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
+For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
+
+\begin{equation}
+    \label{e:breguetendurance}
+    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
+\end{equation}
+
+% The derivation begins with the differential form of Breguet Range,\cite{br2}
+%
+% \begin{equation}
+%     \label{e:breguetdiff}
+%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
+% \end{equation}
+% 
+% Using the definition of $\text{BSFC}$
+% 
+% \begin{equation}
+%     \label{e:brBSFC}
+%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
+% \end{equation}
+% 
+% Equation~\eqref{e:breguetdiff} can be written as
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
+% \end{equation}
+
+This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
+The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
+
+% Dividing by $W$,
+% 
+% \begin{equation}
+%     \label{e:brdiff2}
+%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
+% \end{equation}
+% 
+% and integrating, the Breguet Range equation can be expressed as
+% 
+% \begin{equation}
+%     \label{e:be1}
+%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
+% \end{equation}
+% 
+% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
+
+\begin{equation}
+    \label{e:gpmean}
+    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
+\end{equation}
+
+    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
+
+\begin{align}
+    \label{e:brzbre}
+    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
+    \label{e:brtaylor}
+    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
+\end{align}
+
+    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
+For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
+    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
+
+\begin{align}
+    \label{e:slfweightd}
+    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
+    \label{e:brzbred}
+    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
+    \label{e:brtaylord}
+    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
+    \end{align}
+
+    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
+
+\subsubsection{Elliptical Fuselage}
+
+For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
+The fuselage will increase the overall weight and drag of the aircraft.
+The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
+
+The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
+
+\begin{equation}
+    \label{e:fusevol}
+    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
+\end{equation}
+
+where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
+
+\begin{equation}
+    \label{e:fusevol2}
+    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
+\end{equation}
+
+where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
+Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
+
+\begin{equation}
+    \label{e:fusesa}
+    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
+\end{equation}
+
+The weight of the fuselage is constrained by
+
+\begin{equation}
+    \label{e:fuseweight}
+    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
+\end{equation}
+
+where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
+
+\begin{align}
+    \label{e:fusedrag}
+    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
+    C_f &\geq \frac{0.455}{Re^{0.3}}
+\end{align}
+
+where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
+
+\begin{equation}
+    \label{e:fuseform}
+    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
+\end{equation}
+
+\subsubsection{Engine Weight}
+
+The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
+
+\begin{equation}
+    \label{e:powerlaw}
+    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
+\end{equation}
+
+where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
+Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
+    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
+	\label{f:powervsweightfit}
+	\end{center}
+\end{figure}
+
+\subsubsection{Gas Powered Engine Performance}
+
+Two characteristics of gas engines affect the performance of long-endurance aircraft.  
+The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
+The second is the lapse rate.  
+Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
+To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
+RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
+It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
+This engine performance model is only valid for internal combustion engines.
+
+To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
+Using this approximation the required shaft power determines the $\text{BSFC}$.
+A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
+
+% \begin{equation}
+%     \label{e:rpmtobsfc}
+%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
+% \end{equation}
+
+% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
+    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
+	\label{f:powertobsfc}
+	\end{center}
+\end{figure}
+
+The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
+
+\begin{equation}
+    \label{e:lapse}
+    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
+\end{equation}
+
+The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
+
+\begin{equation}
+    \label{e:lapsefit}
+    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
+\end{equation}
+
+\subsubsection{Climb Constraints}
+
+Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
+The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
+
+\begin{equation}
+    \label{e:climbrate}
+    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
+\end{equation}
+
+The climb rate affects the required thrust, and therefore the required power during climb,
+
+\begin{equation}
+    \label{e:climb}
+    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
+\end{equation}
+
+where $W$ is the weight of the aircraft during climb.
+
 \section{Results}
 
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take of weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take off weight of 320.5 lbs.
+The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
+Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
+The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
+The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 320.5 lbs.
 According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
@@
 
+\newpage
 \begin{longtable}{llll}
@@
 
-Important trade offs between the gas powered and solar-electric powered archictures are highlighted and quantified by changing the latitude and endurance endurance requirements.   
-Both optimization models were solved by minimizing the max take off weight across different bands of latitudes. 
+Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
+Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
 Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds and the solar-electric powered model was solved 31 times in 2.5529 seconds to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[H]
+The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
+
+\begin{figure}[h!]
 	\begin{center}
 	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas archicture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
+    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
     \label{f:latvsmtowtrade}
@@
 
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$30 degrees latitude in 90th percentile wind speeds.  
+One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-The gas powered aircraft curves level off around 38 degrees latitude because wind speeds are highest at that latitude at 4,500 m. 
-Thus, if the gas powered aircraft can fly at 38 degrees latitude it can fly in any band of latitudes.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
+On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
 
 While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude. 
+Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
+Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
 Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
@@
 
-\subsection{Changing Input Values}
+\newpage
+\subsection{Changing Parameter Values}
 
 The previous results are dependent on the assumed input values and parameters.  
-Changing the input values can help show where the designs becomes infeasible. 
-Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
+Changing parameter values can help show where the designs becomes infeasible. 
+As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
 By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
@@
 Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimiation solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[H]
+Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
+
+\begin{figure}[h!]
 	\begin{center}
 	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes required better solar cells and batteries.}}
+    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
 	\label{f:battsolarcontour}
@@
 
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
+Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
 Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
  \end{subfigmatrix}
@@
 where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section IV, larger wings have a weight penalty and the optimization trades air density for wing weight.
+However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
 Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
@@
 
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
@@
 
+Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
+The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
+However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
+At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
+If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
+
 \subsection{Sensitivities}
@@
 
-Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-For example, both batteries and solar cells are expensive and important to the design.  
-The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-The ratio of their magnitudes is 1.76.  
-Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
+% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
+% For example, both batteries and solar cells are expensive and important to the design.  
+% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
+% The ratio of their magnitudes is 1.76.  
+% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
 
 \input{sens.generated.tex}
+
+\newpage
 
@@
 % produces the bibliography section when processed by BibTeX
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\section*{Appendix C}
-
-\subsection{Altitude and relevant atmosphereic quantities}
-
-The altitude also determines atmospheric properties such as air density and viscosity.
+
+% \section*{Appendix A}
+% 
+% \subsection{Altitude and relevant atmosphereic quantities}
+% 
+% The altitude also determines atmospheric properties such as air density and viscosity.
 For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-
-\begin{align}
-    \label{e:Talt}
-    T_{\text{atm}} = T_{\text{sl}} - Lh \\
-    \label{e:rhot}
-    \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-\end{align}
-
-Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-
-\begin{equation}
-    \label{e:sutherland}
-    \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-\end{equation}
-
-where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-
-For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm}
-
-\begin{align}
-    \label{e:tropopress}
-    p &= \rho R_{\text{spec}}T_{11} \\
-    \label{e:tropoalt}
-    h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-\end{align}
-
-According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
+% 
+% \begin{align}
+%     \label{e:Talt}
+%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
+%     \label{e:rhot}
+%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
+% \end{align}
+% 
+% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
+% 
+% \begin{equation}
+%     \label{e:sutherland}
+%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
+% \end{equation}
+% 
+% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
+% 
+% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
+% 
+% \begin{align}
+%     \label{e:tropopress}
+%     p &= \rho R_{\text{spec}}T_{11} \\
+%     \label{e:tropoalt}
+%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
+% \end{align}
+% 
+% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
 Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
@@
 
-\section*{Appendix B}
-
-\subsection{Convex Fitted Functions}
-
-Many constraints in this optimization study are convex equations that approximate data or other functions.  
+\section{Appendix}
+
+\subsection{Discussion on the use of the \emph{JH01} airfoil}
+
+The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
+Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
+The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
+
+\begin{figure}[h!]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
+ \end{subfigmatrix}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
+ \label{f:jhcps}
+\end{figure}
+
+\subsection{Convex fitted functions}
+
+Some constraints in this optimization study are convex equations that approximate data or other functions.  
 These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
 Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are a combination of \emph{max-affine} and \emph{softmax-affine} functions,
+All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
 
@@
 
-\section*{Appendix C}
-
 \subsection{Solar energy calculations}
@@
 
-\section*{Appendix D}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     % \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmin.eps}}
-     % \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792]{cpmax.eps}}
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
+\bibliography{biblibrary}
+\bibliographystyle{aiaa}
 

### 2017-05-08 docs/paperin_review.tex
final edits

@@
-\usepackage{tikz}
 \usepackage{bm}
@@
 
-\subsection{Endurance and Season}
-
 To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
 If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.A, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[H]
+To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
+
+\begin{figure}[h!]
  \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.eps}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.eps}}
+     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
+     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
  \end{subfigmatrix}
@@
 
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement, 4,600 m, or at high altitudes between 16,700-20,000 m.
+Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
 Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
     \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
@@
 To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programming optimization models are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
+Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
 Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
@@
 
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
+    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
+    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
     The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
@@
 \paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing that is a combination of the lifting distribution and weight distribution along the wing.
-The wing loading distribution, $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
+The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
+The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
 
@@
 \begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{empennage}}.
+    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
 \end{equation}
@@
 Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution in Equation~\eqref{e:qbar}
+To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
 
@@
 
-\textbf{Cap Spar for Bending Loads} \\
-
+\textbf{Cap Spar for Bending Loads} 
 A cap spar is considered for the solar-electric and gas powered aircraft.
@@
     \label{e:defcont}
-    w_n &\leq w_{\text{max}}.
-\end{align}
-
-The ultimate tensile strength for unidirectional carbon fiber is, $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
+    w_n &\leq w_{\text{max}}
+\end{align}
+
+where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
 The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
@@
 
-Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
+Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
 
@@
 
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, and density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
+The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
 The total mass and root bending inertia are imposed in the optimization model as
@@
 The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
 % The constraint on December wind speeds and air density at the 30 degree latitude is 
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.eps}
+	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
     \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
@@
 For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this.
+    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
 
@@
     \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24} 
+    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
     \end{align}
@@
 
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take of weight for the requirements listed in Table~\ref{t:mreqs}. 
+The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
 Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
@@
 This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
+On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
 
@@
 
+\newpage
 \subsection{Changing Parameter Values}
@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, 80$p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, 85$p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, 90$p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
  \end{subfigmatrix}
@@
 \input{sens.generated.tex}
+
+\newpage
 
@@
 % produces the bibliography section when processed by BibTeX
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
 
@@
 
-\section*{Appendix A}
-
-\textbf{Convex Fitted Functions} \\
+\section{Appendix}
+
+\subsection{Discussion on the use of the \emph{JH01} airfoil}
+
+The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
+Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
+The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
+
+\begin{figure}[h!]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
+     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
+ \end{subfigmatrix}
+ \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
+ \label{f:jhcps}
+\end{figure}
+
+\subsection{Convex fitted functions}
 
@@
 
-\section*{Appendix B} \\
-
-\textbf{Discussion on the use of the \emph{JH01} airfoil} \\
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[H]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\section*{Appendix C}
-
-\textbf{Solar energy calculations} \\
+\subsection{Solar energy calculations}
 
@@
 
+\bibliography{biblibrary}
+\bibliographystyle{aiaa}
+
 \end{document}

### 2017-05-08 gassolar/gassolarlat.py

@@ -106,7 +106,7 @@ def plot_lats():
     labels = ["$\\pm$" + item.get_text() for item in ax.get_xticklabels()]
     labels = ["$\\pm$%d" % l for l in np.linspace(20, 40, len(labels))]
     ax.set_xticklabels(labels)
-    ax.legend(["Solar-electric Powered", "Gas Powered (7-day endurance)"], fontsize=15, loc=2)
+    ax.legend(["Solar-electric Powered", "Gas Powered"], fontsize=15, loc=2)
     return fig, ax
 
 if __name__ == "__main__":

### 2017-05-11 docs/paperin.tex
change from eps to pdf; calculate loading during each flight seg

@@
     These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
-    In this paper, a comparison of solar-electric and gas powered, long-endurance aircraft is accomplished.
-    Geometric programming, a form of convex optimization, can reliably solve problems with thousands of variables in seconds and is used as an optimization tool to evaluate the design trade offs between architectures and requirements.
+    In this paper, the feasible limits of solar-elecrtic and gas powered, long-endurance aircraft are analyzed for various requirements and design variables to be used for initial vehicle sizing and design.
+    Additionally, this paper demonstrates the capability of geometric programming, a form of convex optimization, as a tool to evaluate the design trade offs between architectures and requirements.
     The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
@@
 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      % \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=1.0\textwidth,natwidth=144,natheight=320]{solarwbreak.pdf}}
-%      % \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=1.0\textwidth,natwidth=178,natheight=320]{gaswbreak.pdf}}
-%      \subfigure[Solar Weight Breakdown\label{f:solarwbreak}]{\includegraphics[width=0.30\textwidth]{solarwbreak.pdf}}
-%      \subfigure[Gas Weight Breakdown\label{f:gaswbreak}]{\includegraphics[width=0.40\textwidth]{gaswbreak.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-%  \label{f:wbreak}
-% \end{figure}
-
 \section{Aircraft Physical Models}
@@
 The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 320.5 lbs.
+The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
 According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
@@
 Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
+Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
 
@@
 
+\input{svals.generated.tex}
+
 \input{gassens.generated.tex}
+
+\input{gvals.generated.tex}
 

### 2017-05-11 docs/paperin_review.tex
change from eps to pdf; calculate loading during each flight seg

@@
     Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
-    In this paper, a comparison of solar-electric and gas powered, long-endurance aircraft is accomplished.
-    Geometric programming, a form of convex optimization, can reliably solve problems with thousands of variables in seconds and is used as an optimization tool to evaluate the design trade offs between architectures and requirements.
+    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
+    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
     The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  

### 2017-05-11 gassolar/gas/disp_sens.py

@@ -3,7 +3,7 @@ import numpy as np
 import sys
 from gassolar.gas.gas import Mission
 from gassolar.environment.wind_speeds import get_windspeed
-from gassolar.solar.print_sens import plot_sens, sens_table
+from gassolar.solar.print_sens import plot_sens, sens_table, sol_table
 
 if __name__ == "__main__":
     sols = []
@@ -22,14 +22,19 @@ if __name__ == "__main__":
         sols.append(sol)
 
     varnames = ["V_{wind}_Mission, Loiter, FlightSegment", "W_{pay}", "\\eta_{prop}", "BSFC_{min}", "t_Mission, Loiter", "N_{max}_Mission, AircraftLoading, WingLoading, ChordSparL"]
-    latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
+    latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$\\mathrm{BSFC}_{100\%}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
     fig, ax = plot_sens(M, sols[2], varnames)
+
+    dvarns = ["MTOW", "b_Mission, Aircraft, Wing", "AR_Mission, Aircraft, Wing", "W_{fuel-tot}", "W_Mission, Aircraft, Wing", "W_Mission, Aircraft, Engine", "BSFC", "C_L", "C_D"]
+    dlatns = ["MTOW", "$b$", "$A$", "$W_{\\mathrm{fuel}}$", "$W_{\\mathrm{wing}}$", "$W_{\\mathrm{engine}}$", "BSFC", "$C_L$", "$C_D$"]
 
     if len(sys.argv) > 1:
         path = sys.argv[1]
-        sens_table(sols, [M], varnames, solar=False, filename=path.replace("figs/", "") + "gassens.generated.tex", latns=latns, title="Gas Powered Aircraft Sensitivities (90th Percentile Winds)")
+        sens_table(sols, [M], varnames, solar=False, filename=path.replace("figs/", "") + "gassens.generated.tex", latns=latns, title="Gas Powered Aircraft Sensitivities (90th Percentile Winds)", label="gassens")
[…]

### 2017-05-11 gassolar/gas/gas.py

@@ -171,7 +171,7 @@ class Mission(Model):
             mtow >= climb1["W_{start}"][0],
             Wfueltot >= sum(fs["W_{fuel-fs}"] for fs in mission),
             mission[-1]["W_{end}"][-1] >= JHO["W_{zfw}"],
-            Wcent >= Wfueltot + sum(summing_vars(JHO.smeared_loads, "W")),
+            Wcent >= Wfueltot + JHO["W_{pay}"] + JHO["W_{avn}"] + sum(summing_vars(JHO.smeared_loads, "W")),
             LS == mtow/JHO.wing["S"]
             ]
 

### 2017-05-11 gassolar/gas/ldratios.py

@@ -100,7 +100,7 @@ if __name__ == "__main__":
         path = sys.argv[1]
         for num in nums:
             fig, ax = ld_plot(M, num)
-            fig.savefig(path + "polarmission%d.eps" % num, bbox_inches="tight")
+            fig.savefig(path + "polarmission%d.pdf" % num, bbox_inches="tight")
     else:
         for num in nums:
             M = Mission()

### 2017-05-11 gassolar/solar/battsolarcon.py

@@ -90,6 +90,6 @@ if __name__ == "__main__":
     fig, ax = plot_battsolarcon()
     if len(sys.argv) > 1:
         path = sys.argv[1]
-        fig.savefig(path + "battsolarcontour.eps", bbox_inches="tight")
+        fig.savefig(path + "battsolarcontour.pdf", bbox_inches="tight")
     else:
         fig.savefig("battsolarcontour.pdf", bbox_inches="tight")

### 2017-05-11 gassolar/solar/contourplots.py

@@ -138,7 +138,7 @@ def plot_contours(path=None):
             ax.set_xlim([0.15, 0.4])
             ax.set_ylim([250, 400])
             if path:
-                fig.savefig(path + "bcontourl%da%d.eps" % (lat, av),
+                fig.savefig(path + "bcontourl%da%d.pdf" % (lat, av),
                             bbox_inches="tight")
             else:
                 fig.savefig("bcontourl%da%d.pdf" % (lat, av))

### 2017-05-11 gassolar/solar/print_sens.py

@@ -3,6 +3,53 @@ from gassolar.solar.solar import Mission
 import matplotlib.pyplot as plt
 import numpy as np
 import sys
+from gpkit.small_scripts import unitstr
+
+def sol_table(sols, models, varnames, filename, solar, latns=[],
+              title="Design Variables", label="vals"):
+
+    with open(filename, "w") as f:
+        f.write("\\begin{longtable}{lccccccccccccc}\n")
+        f.write("\\caption{%s}\\\\\n" % title)
+        f.write("\\toprule\n")
+        f.write("\\toprule\n")
+        f.write("\\label{t:%s}\n" % label)
+        if solar:
+            f.write("\\multirow{2}{*}{Variable} & 25th Latitude & 30th Latitude & 25th Latitude & 30th Latitude \\\\\n")
+            f.write("& 85th Percentile Winds & 85th Percentile Winds & 90th Percentile Winds & 90th Percentile Winds \\\\\n")
+        else:
+            f.write("Variable & 5 Day Endurance & 7 Day Endurance & 9 Day Endurance\\\\\n")
+        f.write("\\midrule\n")
+
+        for vnm, ltnm in zip(varnames, latns):
+            vals = []
+            for s, m in zip(sols, models):
+                if isinstance(s(vnm), float):
+                    val = s(vnm)
+                    units = ""
+                elif not hasattr(s(vnm), "magnitude"):
+                    if solar:
+                        mn = [max(m[sv].descr["modelnums"]) for sv in
+                              s("(E/S)_{irr}") if
+                              abs(s["sensitivities"]["constants"][sv])>0.01][0]
+                        val = [s(vk) for vk in m.varkeys[vnm] if
+                               max(vk.modelnums) == mn][0]
+                    else:
+                        val = [s(sv) for sv in s(vnm) if "Loiter"
+                               in sv.models][0][1]
+                    if hasattr(val, "magnitude"):
+                        units = " [" + unitstr(val) + "]"
+                        val = val.magnitude
+                    else:
+                        units = ""
+                else:
+                    val = s(vnm).magnitude
+                    units = " [" + unitstr(s(vnm)) + "]"
+                vals.append(val)
[…]

### 2017-05-11 gassolar/solar/solar.py

@@ -278,6 +278,20 @@ class SteadyLevelFlight(Model):
 
         return constraints
 
+class Flight(Model):
+    "define mission for aircraft"
+    def setup(self, aircraft, latitude, day):
+
+        flight = FlightSegment(aircraft, latitude=latitude, day=day)
+        loading = aircraft.loading(aircraft["W_{cent}"], aircraft["W_{wing}"], flight["V"], flight["C_L"])
+        for vk in loading.varkeys["N_{max}"]:
+            if "ChordSparL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 5})
+            if "GustL" in vk.descr["models"]:
+                loading.substitutions.update({vk: 2})
+
+        return flight, loading
+
 class Mission(Model):
     "define mission for aircraft"
     def setup(self, latitude=35, day=355):
@@ -285,15 +299,9 @@ class Mission(Model):
         self.solar = Aircraft()
         mission = []
         for l in range(20, latitude+1, 1):
-            mission.append(FlightSegment(self.solar, latitude=l, day=day))
-        loading = self.solar.loading(self.solar["W_{cent}"], self.solar["W_{wing}"], mission[-1]["V"], mission[-1]["C_L"])
-        for vk in loading.varkeys["N_{max}"]:
-            if "ChordSparL" in vk.descr["models"]:
-                loading.substitutions.update({vk: 5})
-            if "GustL" in vk.descr["models"]:
-                loading.substitutions.update({vk: 2})
+            mission.append(Flight(self.solar, l, day))
 
-        return self.solar, mission, loading
+        return self.solar, mission
 
 def test():
     M = Mission(latitude=25)

### 2017-05-11 gassolar/solar/windstruct.py

@@ -41,6 +41,6 @@ if __name__ == "__main__":
         fig, ax = plot_structm(num)
         if len(sys.argv) > 1:
             path = sys.argv[1]
-            fig.savefig(path + "windaltoper%d.eps" % num, bbox_inches="tight")
+            fig.savefig(path + "windaltoper%d.pdf" % num, bbox_inches="tight")
         else:
             fig.savefig("windaltoper.pdf", bbox_inches="tight")

### 2017-05-11 docs/paperin.tex
move design var table; tweak abstract, intro, and conclusion

@@
     Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction of various requirements, such as endurance, wind speeds, operational capability and coverage footprint.
-    In this paper, the feasible limits of solar-elecrtic and gas powered, long-endurance aircraft are analyzed for various requirements and design variables to be used for initial vehicle sizing and design.
-    Additionally, this paper demonstrates the capability of geometric programming, a form of convex optimization, as a tool to evaluate the design trade offs between architectures and requirements.
+    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
+    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
     The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
@@
 Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
+The optimization can be solved in a fraction of a second and is used to the feasible limits of the solar-electric and gas powered, long-endurance aircraft.
 
@@
 
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
+Using this geometric programming methodology, trade offs between different aircraft configurations, power sources, and requirements can be calculated in seconds.  
 The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
@@
 
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
+The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs} and the assumptions and key design parameters described herein.
 The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
 The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
+Key design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements. 
+The rest of this section explores how these result may vary for different requirements, input parameter values, and physics modeling assumptions.
 
 \newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
+\input{svals.generated.tex}
+\input{gvals.generated.tex}
+% \begin{longtable}{llll}
+% \caption{Fixed Optimization Parameters} \\
+% \toprule
+% \toprule
+% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
+% \midrule
+% $\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
+% $\tau$                       & 0.115 & $\tau$                       & 0.115 \\
+% $\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
+% $\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
+% $N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
+% $N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
+% $\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
+% Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
+% $\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
+% $\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
+%                              &     & $\eta_{\text{discharge}}$       & 0.98 \\
+% \bottomrule
+% \label{t:gassolarparams}
+%  \end{longtable}
 
@@
 
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
+    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
+	\label{f:battsolarcontour}
+	\end{center}
+\end{figure}
+
 The previous results are dependent on the assumed input values and parameters.
@@
 Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
 
@@
 Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
 
@@
 \input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
 \input{gassens.generated.tex}
 
-\input{gvals.generated.tex}
-
 \section{Conclusion}
 
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
+Using geometric programming the feasibility limits and design space of solar-electric and gas powered aircraft is analyzed is second to a level of detail and speed not previously achieved in initial aircraft sizing and design.
+Based on the assumptions included herein, the gas powered aircraft is more capable of meeting high latitude and station keeping requirements. 
+The solar-electric powered aircraft sizing does not converge at latitude requirements above 31 degrees latitude. 
+However, solar-electric powered aircraft can meet high endurance and altitude requirements.  
+The effect of key design parameters on the size of each architecture was also achieved.  
+Using higher energy density batteries can result in significant weight and performance savings for the solar-electric powered aircraft.  
+The rapid solve time of geometric programming allows detailed sizing and design analysis of long-endurance aircraft and the capability of solar-electric and gas powered aircraft to meet various requirements.
 

### 2017-05-19 gassolar/solar/solar.py
enable sp wing capability

@@ -8,7 +8,7 @@ import numpy as np
 from gassolar.environment.solar_irradiance import get_Eirr, twi_fits
 from gpkit import Model, Variable
 from gpkitmodels.GP.aircraft.wing.wing import Wing as WingGP
-# from gpkitmodels.SP.aircraft.wing.wing import Wing as WingSP
+from gpkitmodels.SP.aircraft.wing.wing import Wing as WingSP
 from gpkitmodels.GP.aircraft.tail.empennage import Empennage
 from gpkitmodels.GP.aircraft.tail.tail_boom import TailBoomState
 from gpkitmodels.SP.aircraft.tail.tail_boom_flex import TailBoomFlexibility
@@ -27,10 +27,10 @@ class Aircraft(Model):
 
         self.sp = sp
         self.solarcells = SolarCells()
-        # if sp:
-        #     self.wing = WingSP(hollow=True)
-        # else:
-        self.wing = WingGP(hollow=True)
+        if sp:
+            self.wing = WingSP(hollow=True)
+        else:
+            self.wing = WingGP(hollow=True)
         self.battery = Battery()
         self.empennage = Empennage()
         self.motor = Motor()

### 2017-05-19 docs/thesis/abstract.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
+% $Log: abstract.tex,v $
+% Revision 1.1  93/05/14  14:56:25  starflt
+% Initial revision
+% 
+% Revision 1.1  90/05/04  10:41:01  lwvanels
+% Initial revision
+% 
+%
+%% The text of your abstract and nothing else (other than comments) goes here.
+%% It will be single-spaced and the rest of the text that is supposed to go on
+%% the abstract page will be generated by the abstractpage environment.  This
+%% file should be \input (not \include 'd) from cover.tex.
     Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
@@
     Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/app_fits.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
+\chapter{Convex fitted functions}
 
@@
 In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/app_jho.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
+\chapter{Discussion on the use of the \emph{JH01} airfoil}
 
@@
 \end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_concl.tex
add thesis folder

--- 2017-05-11 docs/paperin.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization can be solved in a fraction of a second and is used to the feasible limits of the solar-electric and gas powered, long-endurance aircraft.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs between different aircraft configurations, power sources, and requirements can be calculated in seconds.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs} and the assumptions and key design parameters described herein.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-Key design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements. 
-The rest of this section explores how these result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\input{svals.generated.tex}
-\input{gvals.generated.tex}
-% \begin{longtable}{llll}
-% \caption{Fixed Optimization Parameters} \\
-% \toprule
-% \toprule
-% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% \midrule
-% $\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-% $\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-% $\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-% $\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-% $N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-% $N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-% $\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-% Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-% $\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-% $\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-%                              &     & $\eta_{\text{discharge}}$       & 0.98 \\
-% \bottomrule
-% \label{t:gassolarparams}
-%  \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-The previous results are dependent on the assumed input values and parameters.
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-\input{gassens.generated.tex}
-
-\section{Conclusion}
-
+\chapter{Conclusion}
 Using geometric programming the feasibility limits and design space of solar-electric and gas powered aircraft is analyzed is second to a level of detail and speed not previously achieved in initial aircraft sizing and design.
@@
 The rapid solve time of geometric programming allows detailed sizing and design analysis of long-endurance aircraft and the capability of solar-electric and gas powered aircraft to meet various requirements.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_config.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
+\chapter{Configuration}
 
@@
 \end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_gp.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
+%% This is an example first chapter.  You should put chapter/appendix that you
+%% write into a separate file, and add a line \include{yourfilename} to
+%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
+%% You can process specific files by typing their names in at the 
+%% \files=
+%% prompt when you run the file main.tex through LaTeX.
+\chapter{Geometric Programming}
 
 Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming.
+
+\section{Properties of Geometric Programs}
+
 Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
@@
 \label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
+\text{minimize } f_0(\bm{x}) & \nonumber \\
+\text{subject to  } f_i(\bm{x}) &\leq 1, i=1,\ldots,m \\
+g_i (\bm{x}) &= 1, i = 1,\ldots,p 
 \end{align}
@@
  \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
+g(\bm{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
 \label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
+f(\bm{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
 \end{align}
@@
 Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
+
+\section{Geometric Program Utilization}
 
@@
 Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_intro.tex
add thesis folder

--- 2017-05-11 docs/paperin.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
+This thesis presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
 Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
@@
 Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs} and the assumptions and key design parameters described herein.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-Key design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements. 
-The rest of this section explores how these result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\input{svals.generated.tex}
-\input{gvals.generated.tex}
-% \begin{longtable}{llll}
-% \caption{Fixed Optimization Parameters} \\
-% \toprule
-% \toprule
-% \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% \midrule
-% $\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-% $\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-% $\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-% $\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-% $N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-% $N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-% $\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-% Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-% $\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-% $\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-%                              &     & $\eta_{\text{discharge}}$       & 0.98 \\
-% \bottomrule
-% \label{t:gassolarparams}
-%  \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-The previous results are dependent on the assumed input values and parameters.
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-\input{gassens.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming the feasibility limits and design space of solar-electric and gas powered aircraft is analyzed is second to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-Based on the assumptions included herein, the gas powered aircraft is more capable of meeting high latitude and station keeping requirements. 
-The solar-electric powered aircraft sizing does not converge at latitude requirements above 31 degrees latitude. 
-However, solar-electric powered aircraft can meet high endurance and altitude requirements.  
-The effect of key design parameters on the size of each architecture was also achieved.  
-Using higher energy density batteries can result in significant weight and performance savings for the solar-electric powered aircraft.  
-The rapid solve time of geometric programming allows detailed sizing and design analysis of long-endurance aircraft and the capability of solar-electric and gas powered aircraft to meet various requirements.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_mis.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
+\chapter{Mission Requirements}
 
@@
 
-\subsection{Station Keeping}
+\section{Station Keeping}
 
@@
 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
+\section{Altitude}
 
@@
 
-\subsection{Latitude}
+\section{Latitude}
 
@@
 
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
+\section{Percentile Wind Speed}
 
-\section{Geometric Programming}
+For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
+To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
+Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}.
 
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
+     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
+ \label{f:bostonwinds}
 \end{figure}
 
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
+     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
+ \label{f:romewinds}
 \end{figure}
 
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -
+This analysis shows high wind speeds can last for days at a time.  
+An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
+However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.

### 2017-05-19 docs/thesis/chap_phy.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
+\chapter{Aircraft Physical Models}
 
@@
 
-\subsection{Shared Physical Models}
+\section{Shared Physical Models}
 
@@
 
-\subsubsection{Steady Level Flight}
+\subsection{Steady Level Flight}
 
@@
 
-\subsubsection{Aerodynamics}
+\subsection{Aerodynamics}
 
@@
 
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
+To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil, Equation~\ref{aerodragprof} (See Appendix \ref{app:jho} for a discussion of the airfoil choice).
+Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix \ref{app:fits})
     The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
 
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
+    \begin{align}
+        \label{e:aerodragprof}
+        c_{d_p}^{3.09} &\geq \num{1.02d-7}C_L^{18.86}Re^{-0.18} + 0.0017C_L^{2.93}Re^{-0.66}  \nonumber \\
+                       &+ \num{2.28d-6}C_L^{-1.65}Re^{-3.43}+ \num{1.75e8}C_L^{-0.18}Re^{-2.71} \\
+        Re &= \frac{\rho V S/b}{\mu}
+    \end{align}
 
@@
 
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
+\subsection{Wing Spar Model}
 
@@
 
-\paragraph{\textbf{Standard Wing Bending Case}} 
+\subsubsection{Standard Wing Bending Case}
 The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
@@
 where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
+Using the equation for the local chord of a constant tapered wing\cite{bending}
+
+\begin{equation}
+    \label{e:localchord}
+    \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
+\end{equation}
+
+where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
 
@@
 
-\paragraph{\textbf{Gust Loading Case}}
+\subsubsection{Gust Loading Case}
 
@@
 
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
+Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting}.
+The fitted equation is compared to the actual equation in Figure~\ref{f:arctan}
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{arctanfit.pdf}
+    \caption{\textbf{Visual comparison of fitted equation to the arctan function.  (Log space RMS error $=0.039$ for $V_{\text{gust}}/V \in [0, 0.7]$)}}
+	\label{f:capspar}
+	\end{center}
+\end{figure}
+
 The gust velocity has an assumed profile along the wing\cite{acgust},
@@
 
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
+\subsubsection{Discretized Beam Model}
+
+Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, and moments can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input with zero forces and moments at the wing tip,
+
+\begin{align}
+    \label{e:shear}
+    \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
+    \label{e:moment}
+    M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
+    \label{e:shearboundary}
+    \mathcal{S}_n &= 0 \\
+    \label{e:momentboundary}
+    M_n &= 0
+\end{align}
+
+Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
+
+\begin{align}
+    \label{e:angle}
+    \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
+    \label{e:deflection}
+    w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
+    \label{e:angleboundary}
+    \Theta_0 &= 0 \\
+    \label{e:defboundary}
+    w_0 &= 0 
+\end{align}
+
+Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
 
@@
 
-\textbf{Cap Spar for Bending Loads} 
+\subsubsection{Cap Spar for Bending Loads} 
 A cap spar is considered for the solar-electric and gas powered aircraft.
@@
 
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
+\subsection{Additional Wing Weight}
 
@@
 
-\subsubsection{Empennage}
+\subsection{Empennage}
 
@@
 
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
+The drag of the horizontal and vertical tails is computed using a multiple monomial equations fitted from XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
+
+\begin{align}
+    D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
+    \label{e:taildrag}
+    c_{d_{\text{(v,h)}}} &\geq 0.34 Re_{\text{(v,h)}}^{-0.18} \tau_{\text{(v,h)}}^{0.77} \\
+    c_{d_{\text{(v,h)}}} &\geq 5.45 Re_{\text{(v,h)}}^{-0.48} \tau_{\text{(v,h)}}^{0.25} \\
+    c_{d_{\text{(v,h)}}} &\geq 16.3 Re_{\text{(v,h)}}^{-0.54} \tau_{\text{(v,h)}}^{0.47} \\
+    c_{d_{\text{(v,h)}}} &\geq 9.51 Re_{\text{(v,h)}}^{-0.48} \tau_{\text{(v,h)}}^{0.47} \\
+    c_{d_{\text{(v,h)}}} &\geq 218.7 Re_{\text{(v,h)}}^{-0.60} \tau_{\text{(v,h)}}^{1.31} \\
+    Re_{\text{(v,h)}} &= \frac{\rho V S_{\text{(v,h)}}/b_{\text{(v,h)}}}{\mu}
+\end{align}
+
 where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
 The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.  
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
+A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
+    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.040)}}
+	\label{f:taildragpolar}
+	\end{center}
+\end{figure}
+
+\section{Solar-Electric Aircraft Physical Models}
+
+\subsection{Wind Speeds}
 
@@
 The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
+In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%.
+(See Appendix \ref{app:fits}) The wind constraint at 30 degrees lattitude is shown in Figure~\ref{f:windfitl30}
+
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
+    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 30$^{\circ}$ latitude. (Log-space RMS error = 0.047.)}}
+	\label{f:windfitl30}
+	\end{center}
+\end{figure}
 
@@
 
-\subsubsection{Solar-Electric Power}
+\subsection{Solar-Electric Power}
 
@@
 The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
+$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix \ref{app:solar} to generate the fitted equations. 
+Figure~\ref{f:energyapprox} shows sample monomial approximations of the twilight and day energy equations at 30 degrees latitude.
+
+\begin{figure}[H]
+ \begin{subfigmatrix}{2}% number of columns
+     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
+     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
+ \end{subfigmatrix}
+ \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
+ \label{f:energyapprox}
+\end{figure}
 
@@
 
-\subsubsection{Motor Weight}
+\subsection{Motor Weight}
 
@@
 
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
+\section{Gas Powered Aircraft Physical Models}
+
+\subsection{Breguet Endurance}
 
@@
 
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
+The derivation of this form begins with the differential form of Breguet Range,\cite{br2}
+
+\begin{equation}
+    \label{e:breguetdiff}
+    -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
+\end{equation}
+
+Using the definition of $\text{BSFC}$
+
+\begin{equation}
+    \label{e:brBSFC}
+    \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
+\end{equation}
+
+Equation~\eqref{e:breguetdiff} can be written as
+
+\begin{equation}
+    \label{e:brdiff2}
+    -dW = g P_{\text{shaft}} \text{BSFC} dt.
+\end{equation}
+
+Dividing by $W$,
+
+\begin{equation}
+    \label{e:brdiff2}
+    -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
+\end{equation}
+
+and integrating, the Breguet Range equation can be expressed as
+
+\begin{equation}
+    \label{e:be1}
+    \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
+\end{equation}
+
+where the segment weight $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
 
@@
 \end{equation}
+
+This version of the Breguet Range equation comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
+One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
 
@@
 
-\subsubsection{Elliptical Fuselage}
+\subsection{Elliptical Fuselage}
 
@@
 
-\subsubsection{Engine Weight}
+\subsection{Engine Weight}
 
@@
 
-\subsubsection{Gas Powered Engine Performance}
+\subsection{Gas Powered Engine Performance}
 
@@
 
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
+\begin{align}
+    \label{e:rpmtobsfc}
+    \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} &\geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} \nonumber \\
+                &+ 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} 
+\end{align}
+
+where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
 
@@
 
-\subsubsection{Climb Constraints}
+\subsection{Climb Constraints}
 
@@
 where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
-
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
-The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
-
-\newpage
-\begin{longtable}{llll}
-\caption{Fixed Optimization Parameters} \\
-\toprule
-\toprule
-\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-\midrule
-$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
-$\tau$                       & 0.115 & $\tau$                       & 0.115 \\
-$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
-$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
-$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
-$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
-$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
-Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
-$\dot{h}_{\text{min}}$ [ft/min] & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
-$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
-                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
-\bottomrule
-\label{t:gassolarparams}
- \end{longtable}
-
-\subsection{Changing Requirements}
-
-Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.   
-Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
-Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
-The gas powered model was solved 63 times in 5.6178 seconds total and the solar-electric powered model was solved 31 times in 2.5529 seconds total to produce Figure~\ref{f:latvsmtowtrade}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
-    \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
-    \label{f:latvsmtowtrade}
-	\end{center}
-\end{figure}
-
-One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 190 lbs is able to operate between $\pm$28 degrees latitude in 90th percentile wind speeds.  
-This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-On the other hand, solar-electric powered aircraft design becomes infeasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 18,300 m, the combination of lower solar flux and higher wind speeds makes it difficult to reach latitude bands greater than $\pm$30 degrees.
-
-While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
-Solving the gas powered optimization model for different endurance requirements shows where the gas powered architecture becomes less feasible. 
-Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take-off weight for an aircraft capable of flying at any latitude. 
-Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
-    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
-	\label{f:mtowvsendurance}
-	\end{center}
-\end{figure}
-
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
-
-\newpage
-\subsection{Changing Parameter Values}
-
-The previous results are dependent on the assumed input values and parameters.  
-Changing parameter values can help show where the designs becomes infeasible. 
-As one example, two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
-By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
-Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
-Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
-Figure~\ref{f:battsolarcontour} was produced using 157 separate optimization solutions that took a total of 14.379 seconds to solve.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
-    \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
-	\label{f:battsolarcontour}
-	\end{center}
-\end{figure}
-
-Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft wing spans for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
-Each point in Figure~\ref{f:solarcontours} is a unique design for minimum wing span. 
-The infeasible regions and contour shapes would change for different assumed constant values.
-
- \begin{figure}[H]
- \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Maxtrix of minimum wing span solar-electric aircraft designs. Values of assumed constants are given throughout the text.}}
- \label{f:solarcontours}
-\end{figure}
-
-\subsection{Changing Physical Modeling Assumptions}
-
-Insight into the design space can also be gained by changing the physical modeling assumptions.
-For example, by altering the aircraft structural model it can be observed how air density trades for wing weight. 
-It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
-If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight
-
-\begin{equation}
-    W_{\text{structural}} \geq W_{\text{MTO}} f_{\text{structural}}
-\end{equation}
-
-where $f_{\text{structural}} = 0.35$, then the optimized flight altitude is almost exactly 67,000 ft as shown in Figure~\ref{f:altoper}.  
-However, if the structural weight is represented by the more detailed model as explained in Section V, larger wings have a weight penalty and the optimization trades air density for wing weight.
-Therefore, by adding a structural model the optimization seeks a smaller wing to save weight and operates at a lower altitude to increase density.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
- \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
- \label{f:altoper}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
-    \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
- \label{f:polarmission}
-	\end{center}
-\end{figure}
-
-Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
-The optimum lift to drag ratio to maximize endurance for gas powered aircraft is at the maximum $C_L^{1.5}/C_D$.\cite{br2}  
-However, while station keeping, the aircraft will maintain a constant velocity during high wind speeds.  
-At a constant velocity or constant Reynolds number the lift to drag ratio will not be at the maximum $C_L^{1.5}/C_D$.  
-If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
-
-\subsection{Sensitivities}
-
-When a GP is solved, the sensitivity of the optimal objective value with respect to each constraint is also returned.  
-From this information, the sensitivity of the optimal objective value to each fixed variable can be extracted.\cite{hoburgthesis} 
-While sensitivities are local and therefore only exact for small changes, they provide useful information about the relative importance of various design variables. 
-For example, if the objective function were $W_{\text{MTO}}$ and the sensitivity to battery specific energy were 0.5, then a 1\% increase in the solar cell efficiency would result in a 0.5\% increase in weight.
-Tables~\ref{t:sens} and~\ref{t:gassens} show the variables with the highest sensitivities for the solar-electric and gas powered architectures respectively, where the objective was max take-off weight.
-Tables~\ref{t:svals} and ~\ref{t:gvals} list a few deisgn variables from both architectures for reference.
-
-For the solar-electric aircraft, it is interesting to note that the battery discharge efficiency sensitivity is higher than the battery charge efficiency sensitivity.
-This occurs because the discharge efficiency directly affects the required battery size, whereas the charge efficiency only does so indirectly.
-
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
-
-\input{sens.generated.tex}
-
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -

### 2017-05-19 docs/thesis/chap_results.tex
add thesis folder

--- 2017-05-11 docs/paperin_review.tex
@@
-\usepackage{bm}
-\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}
-
- \author{
-  Michael Burton \thanks{Master's Candidate, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Student.}
-  \ and Warren Hoburg\thanks{Assistant Professor, Aeronautics and Astronautics Engineering, 77 Mass Ave, Cambridge MA, 02139, AIAA Member.}\\
-  {\normalsize\itshape
-   Massachusetts Institute of Technology, Cambridge, 02139, USA}\\
- }
-
- % Data used by 'handcarry' option
- \AIAApapernumber{YEAR-NUMBER}
- \AIAAconference{Conference Name, Date, and Location}
- \AIAAcopyright{\AIAAcopyrightD{YEAR}}
-
- % Define commands to assure consistent treatment throughout document
- \newcommand{\eqnref}[1]{(\ref{#1})}
- \newcommand{\class}[1]{\texttt{#1}}
- \newcommand{\package}[1]{\texttt{#1}}
- \newcommand{\file}[1]{\texttt{#1}}
- \newcommand{\BibTeX}{\textsc{Bib}\TeX}
- \usepackage{hyperref}
- \hypersetup{citecolor = blue}
-
-\begin{document}
-
-\graphicspath{{./figs/}} 
-\maketitle
-
-\begin{abstract}
-    Fueled by telecommunication needs and opportunities, there has been a recent push to develop aircraft that can provide long-endurance (days to weeks) persistent aerial coverage.
-    These aircraft present a complicated systems engineering problem because of the multifaceted interaction between aerodynamics, structural loads, environmental effects, and engine, battery, and other component performance.
-    Using geometric programming, models capturing the interaction between disciplines are used to analyze the feasible limits of solar-electric and gas powered, long-endurance aircraft in seconds to a level of detail and speed not previously achieved in initial aircraft sizing and design.
-    The results show that long-endurance, gas powered aircraft are generally more robust to higher wind speeds than solar-powered aircraft.  
-    However, gas powered aircraft are limited in their endurance by the amount of fuel that they can carry. 
-    While solar-electric powered aircraft can theoretically fly for months, they are operationally limited by reduced solar flux during the winter and wind speeds at higher latitudes.
-    Using geometric programming, a detailed trade study between gas-powered and solar-powered aircraft is performed to discover which architecture is best suited to meet a given set of requirements, and what is the optimum size and endurance of that platform.
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{multicols}{2}
-\small
-
-\begin{tabbing}
-  XXXXXXX \= \kill% this line sets tab stop
-$A$ \> wing aspect ratio \\
-$\bar{A}$ \> non-dimensional cross sectional area \\
-$b$ \> wing span \\ % [ft] \\
-BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
-$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
-$c$ \> wing chord \\ %[m] \\
-$C_D$ \> aircraft drag coefficient \\
-$C_{d_0}$ \> non-wing drag coefficient \\
-% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\
-$c_{d_p}$ \> wing profile drag coefficient \\
-% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\
-$C_f$ \> skin friction coefficient \\
-$C_L$ \> lift coefficient \\
-$c_{l_{\alpha}}$ \> lift slope coefficient \\
-$C_{L_{\text{max}}}$ \> maximum lift coefficient \\
-$d$ \> tail boom diameter \\ %[m] \\
-$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
-$\Delta W$ \> wing section weight \\ %[N] \\
-$\Delta y$ \> wing section length \\ %[m] \\
-DOY \> day of the year \\
-$e$ \> span efficiency factor \\
-$E$ \> Young's Modulus \\ %[Pa] \\
-$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
-$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
-$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
-$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
-$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\
-$f_{\text{solar}}$ \> planform area fraction covered in solar cells \\
-$f_{\text{structural}}$ \> fractional structural weight\\
-$g$ \> gravitational constant \\ % [m/s$^2$] \\
-$h$ \> flight altitude \\ % [ft] \\
-$\dot{h}_{\text{min}}$ \> minimum climb rate \\ % [ft] \\
-$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
-$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
-$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
-$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\
-$k$ \> tail boom taper index \\
-$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
-% $L$ \> atmospheric lapse rate \\ % [K/m] \\
-$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
-$l_\text{fuse}$ \> fuselage length \\ % [m] \\
-$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
-$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
-$m$ \> tail boom mass \\ % [kg] \\
-$m'$ \> local wing mass \\ % [kg/m] \\
-% $m_{\text{fac-wing}}$ \> margin factor\\
-$m_{\text{fac}}$ \> weight margin factor \\
-$m_{\text{motor}}$ \> electric motor mass \\
-$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
-$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
-$n$ \> number of wing segments \\
-$N$ \> number of flight segments \\
-$N_{\text{max}}$ \> safety load factor\\
-% $p$ \> air pressure \\ % [Pa]\\
-% $p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\
-$p_{\text{wind}}$ \> percentile wind speed \\
-$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
-$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
-$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
-$(P/S)_{\text{min}}$ \> minimum operational solar power \\ % [Whr/m$^2$] \\
-$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
-$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
-$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\
-$r_0$ \> average distance from earth to sun \\
-$R$ \> aircraft range \\ % [nmi] \\
-$R_{\text{earth orbit}}$ \> distance from earth to sun \\
-$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
-$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\
-$R_{\text{sun}}$ \> radius of the sun \\
-$q$ \> distributed wing loading \\ % [N/m] \\
-$\bar{q}$ \> normalized distributed wing loading \\
-$Re$ \> aircraft Reynolds number \\
-$Re_{\text{boom}}$ \> tail boom Reynolds number \\
-% $Re_{\text{h}}$ \> horizontal tail Reynolds number \\
-% $Re_{\text{v}}$ \> vertical tail Reynolds number \\
-$S$ \> wing planform area \\ % [m$^2$]\\
-$\mathcal{S}$ \> shear force \\ % [N] \\
-$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
-$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
-$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
-$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
-$t$ \> flight time \\ % [days] \\
-$t_0$ \> tail boom root wall thickness \\ % [m] \\
-$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
-$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
-$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
-$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
-$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
-$T$ \> aircraft thrust \\ % [N] \\
-% $T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
-% $T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
-% $T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
-$V$ \> true airspeed \\ % [m/s] \\
-$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
-$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\
-$V_{\text{h}}$ \> horizontal tail lift coefficient \\
-$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
-$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\
-$V_{\text{v}}$ \> vertical tail lift coefficient \\
-$w$ \> wing deflection \\ % [m] \\
-$W$ \> aircraft weight \\ % [N] \\
-$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
-$W_{\text{batt}}$ \> battery weight \\ % [N] \\
-$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
-$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
-$W_{\text{emp}}$ \> empennage weight \\ % [N] \\
-$W_{\text{engine}}$ \> engine weight \\ % [N] \\
-$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\
-$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
-$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
-$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
-$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
-$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
-$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
-$W_{\text{MTO}}$\> max take-off weight \\ % [N] \\
-$W_{\text{payload}}$ \> payload weight \\ % [N] \\
-$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
-$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
-$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
-$W_{\text{structural}}$ \> structural weight \\ % [N] \\
-$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
-$W_{\text{wing}}$ \> wing weight \\ % [N] \\
-$y$ \> distance from wing root along span \\ % [m] \\
-$z_{\text{bre}}$ \> Breguet Range helper variable \\
-$\alpha$ \> angle of attack \\
-$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\
-$\Delta$ \> solar declination angle \\
-$\eta_{\text{charge}}$ \> battery charging efficiency \\
-$\eta_{\text{discharge}}$ \> battery discharging efficiency \\
-$\eta_{\text{motor}}$ \> electric motor efficiency \\
-$\eta_{\text{prop}}$ \> propulsive efficiency \\
-$\eta_{\text{solar}}$ \> solar cell efficiency \\
-% $\mu$ \> air viscosity \\ % [kg/m/s] \\
-% $\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\
-$\phi$ \> latitude \\
-$\rho$ \> air density \\ % [kg/m$^3$] \\
-$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
-$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
-$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
-$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
-$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\
-$\tau_{\text{h}}$ \> horizontal tail thickness to chord ratio \\
-$\tau_t$ \> wing thickness to chord ratio \\
-$\tau_{\text{v}}$ \> vertical tail thickness to chord ratio \\
-$\tau_w$ \> cap spar width to chord ratio \\
-$\theta$ \> angle normal to surface \\
-$\theta_{\text{boom}}$ \> tail boom deflection angle \\
-$\Theta$ \> wing bending deflection angle 
- \end{tabbing}
-
-\end{multicols}
-% \printglossary% creates nomenclature section produced by MakeIndex
-
-\section{Introduction}
-
-Long-endurance station-keeping aircraft are potential solutions to providing internet, communication, and persistent surveillance. 
-Recently, tech companies such as Google\cite{googletitan} and Facebook\cite{aquila} have explored the possibility of using solar-electric powered aircraft to provide internet to parts of the world where 4G or 3G network is not available.
-Engineering firms like Aurora Flight Sciences\cite{orion} and Vanilla Aircraft\cite{vanilla} have developed gas powered long-endurance platforms as a means of providing continual surveillance for days at a time.
-Sizing of such long-endurance aircraft is complicated because of the multifaceted interaction between aerodynamics, structural weight, solar energy, wind speed, and other disciplines and environmental models.
-The complexity of sizing these kinds of aircraft makes the trade offs between gas powered and solar-powered architectures potentially non-intuitive.
-
-This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two architectures for a given mission.
-Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
-Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form and are then combined to form an optimization model. 
-The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.
-
-The driving requirement that sizes solar-electric powered aircraft is the ability to operate at multiple locations and during all seasons.  
-To fly multiple days, solar-electric powered aircraft must carry enough solar cells and batteries to fly during the day while storing enough energy to fly through the night.\cite{solartech}
-If this condition can be achieved during the winter solstice, when the solar flux is at a minimum, then a solar-electric powered aircraft can theoretically fly during any time of the year for long durations.
-\cite{solartech}
-This becomes more difficult to achieve at higher latitudes as solar flux during the winter solstice decreases.  
-Additionally, to station keep the aircraft must fly faster than the local wind speeds.  
-Wind speeds are a function of latitude, altitude, and season and tend to increase at higher latitudes and during winter months. 
-Therefore, a key sizing study of solar-electric powered aircraft is the effect of latitude on aircraft size.
-
-The key driving requirement for a gas powered aircraft is endurance.  
-As gas-powered aircraft are endurance limited by the amount of fuel that they can carry, longer endurance requires more fuel and hence a larger aircraft.  
-Because gas-powered aircraft are not affected by the solar flux, their station-keeping ability at different latitudes only depends on the local wind speed. 
-If a gas-powered aircraft can fly at the latitude with the worst wind speed, in can theoretically fly anywhere else in the world.
-
-Using this geometric programming methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
-The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
-Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
-The presented optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
-Varying mission requirements reveals how the best architecture changes depending on the requirements.
-
-\section{Mission Requirements}
-
-A comparison of the gas and solar-electric powered aircraft is achieved by comparing their respective capabilities at satisfying a common set of requirements shown in Table~\ref{t:mreqs}.
-These requirements can and will be changed in the results section to observe which architecture is best suited to meet various sets of requirements.
-A discussion of each requirement is included in this section.
-
-\begin{longtable}{lccccccccccccc}
-\caption{Mission Requirements}\\
-\toprule
-\toprule
-\label{t:mreqs}
-Payload & 10 lbs\\
-Station Keeping & 90\% winds \\
-Endurance & $>6$ days\\
-Season & all seasons\\
-Altitude & $>$ 4,600 m\\
-Latitude & $\pm30^{\circ}$\\
-\bottomrule
-\end{longtable}
-
-\subsection{Station Keeping}
-
-To station keep an aircraft must fly at least as fast as the wind speed,
-
-\begin{equation}
-    \label{e:availreq}
-    V \geq V_{\text{wind}}.
-\end{equation}
-
-Distributions of wind speed data\cite{wind} indicate that it is impractical for long-endurance UAVs to station keep in 100 percentile wind speeds.  
-Therefore, the station keeping requirement is parameterized by a percentile wind speed, $p_{\text{wind}}$, above which the aircraft is allowed to drift off station. 
-The wind speed at a station is also a function of latitude, altitude, season or day of the year,
-
-\begin{equation}
-    \label{e:windspeed}
-    V_{\text{wind}} = f(\phi, h, \text{DOY}, p_{\text{wind}}).
-    \end{equation}
-
-Wind speed data was collected from the ERA Interim atmospheric datasets for the years 2005-2015.\cite{wind} 
-It is assumed that the wind speed distribution is independent of longitude.
-
-To meet the season requirement, the solar-electric aircraft must be able to fly a full day/night cycle on the winter solstice.  
-If it can do so, it automatically exceeds the endurance requirement. 
-To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section V.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
-     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
- \end{subfigmatrix}
- \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
- \label{f:windvsmonth}
-\end{figure}
-
-For long-endurance gas powered aircraft, the endurance requirement is the amount of time the aircraft needs to remain in the air without refueling.  
-This is expected to be mostly independent of season, except insofar as the wind speeds depends on season.
-
-Because wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres, only wind speeds from December and June, respectively, will be used for the both the gas and solar-electric powered aircraft sizing and performance analysis.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
-%     \caption{\textbf{Winds at 30$^{\circ}$ latitude.  Bands represent 80th, 90th and 95th percentiles. Winds peak in winter.}}
-% 	\label{f:windvsmonth}
-% 	\end{center}
-% \end{figure}
-
-\subsection{Altitude}
-
-One altitude requirement is a minimum height required to meet a certain coverage footprint,
-
-\begin{equation}
- h \geq h_{\text{min}}.
-\end{equation}
-
-Long-endurance aircraft tend to fly at low speeds, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum height requirement of 4,600 m, or at high altitudes between 16,700-20,000 m.
-Gas powered aircraft generally fly in the lower altitude regime of 4,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.
-Solar-electric powered aircraft, which do not have naturally aspirated engines, will fly around 18,000 m to avoid cloud coverage.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswindhmin.pdf}
-    \caption{\textbf{Aircraft must fly above $h_{\text{min}}$ to meet altitude requirement. Bands represent 80th, 90th and 95th percentiles.\cite{wind}}}
-	\label{f:altvswindhmin}
-	\end{center}
-\end{figure}
-
-\subsection{Latitude}
-
-It is assumed that long-endurance aircraft will have a requirement to be capable of operating anywhere within a band of latitudes.  
-For example, an aircraft designed or optimized for the 35th latitude would be able to operate at any latitude between $\pm35$ degrees. 
-Latitude affects both the solar-electric and the gas powered aircraft because wind speed varies with latitude. 
-Figure~\ref{f:latvswind} shows how the wind speed varies with latitude in the Northern Hemisphere in December.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswindN.pdf}
-    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds.\cite{wind} }}
-	\label{f:latvswind}
-	\end{center}
-\end{figure}
-
-Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months.
-This is further discussed in Section V.B.
-
-% \subsection{Percentile Wind Speed}
-% 
-% For the purposes of this design study, a percentile wind speed corresponds to that percentile of wind speed data across all longitudes in December months spanning the years 2005-2015.  
-% To understand what a given percentile wind speed implies operationally an autocorrelation analysis was done for multiple cites at varying latitudes at 60,000 during December. 
-% Wind speeds at 60,000 ft over Boston and Rome during December of 2015 are shown in Figures~\ref{f:bostonwinds} and~\ref{f:romewinds}. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
-%      \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
-%  \label{f:bostonwinds}
-% \end{figure}
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
-%      \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
-%  \label{f:romewinds}
-% \end{figure}
-% 
-% This analysis shows high wind speeds can last for days at a time.  
-% An aircraft designed for 90th percentile winds will likely be able to remain on station for 90\% of the time.  
-% However, during times that wind speeds exceed the 90th percentile they could last for days at a time, possibly requiring the aircraft to drift off station, relocate or land.
-
-\section{Geometric Programming}
-
-Sizing long-endurance aircraft for the stated requirements was accomplished using geometric programming. 
-Geometric programs (GPs) are a mathematical optimization problem characterized by the convexity of the objective and constraint functions.\cite{gp} GPs have the form
-
-\begin{align} 
-\label{e:gpform}
-\text{minimize } f_0(\bold{x}) & \nonumber \\
-\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
-g_i (\bold{x}) &= 1, i = 1,\ldots,p 
-\end{align}
-
-where the functions $g_i$ must be \emph{monomial} functions and the functions $f_i$ must be \emph{posynomial} functions. \emph{Monomials} and \emph{posynomials} have the forms
-
-\begin{align}
- \label{e:mon}
-g(\bold{x}) &= c x_1^{a_1} x_2^{a_2} \dotsm x_n^{a_n} , \\
-\label{e:pos}
-f(\bold{x}) &= \displaystyle\sum_{k=1}^K c_k x_1^{a_{1_k}} x_2^{a_{2_k}} \dotsm x_n^{a_{n_k}}.
-\end{align}
-
-The properties of a geometric program allow solution algorithms to guarantee convergence to a global optimum, provided that there exists a feasible solution to the set of constraint functions.  
-Additionally, geometric programs can be solved rapidly with existing algorithms.  
-Using state of the art, standard interior-point algorithms, GPs with 1,000 variables and 10,000 constraints converge to a solution in tens of seconds.\cite{gp}
-
-To evaluate the gas and solar-electric powered design space, the requirements, architecture type, physical models, and assumed parameter values are expressed as constraints on the GP.\cite{hoburgthesis}
-Essentially, the gas and solar-electric powered geometric programs are long lists of constraints in a GP-form that can be rapidly solved using interior point methods.  
-Figure~\ref{f:flowchart} shows how the optimization problem is constructed and solved. 
-The following sections explain in detail the configuration, physical modeling, and input value assumptions and their resulting constraint equations that were used in both the gas and solar-electric powered optimization models.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=568,natheight=154]{flowchart.eps}
-    \caption{\textbf{Process of constructing and solving a GP optimization problem. }}
-	\label{f:flowchart}
-	\end{center}
-\end{figure}
-
-A python package for defining and manipulating geometric programs, GPkit\cite{gpkitdocs}, is used to create the models described in this paper.  
-A commercial solver, mosek\cite{mosek}, is used to solve the GP. 
-Both the gas and solar-electric powered GP optimization models are available for download and use. \cite{gassolartrade} \\
-
-\section{Configuration}
-
-For this study a simple fixed-wing tractor configuration was chosen as the baseline for both architectures. 
-Common characteristics between the gas and solar models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
-The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
-The solar-electric powered aircraft holds the batteries in the wings.  
-The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
-A simple diagram of each vehicle is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
-The configuration is assumed to be static for this study.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=447,natheight=265]{simpleaircraft.eps}
-    \caption{\textbf{Blue and red are specific to the solar-electric and gas architectures respectively.  Black are shared characteristics.}}
-	\label{f:simpleaircraft}
-	\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.8\textwidth,natwidth=378,natheight=335]{wbreak.eps}
-   \caption{\textbf{Representative weight breakdown of both aircraft optimization models.}}
-	\label{f:wbreak}
-	\end{center}
-\end{figure}
-
-\section{Aircraft Physical Models}
-
-To evaluate the size and performance of both gas and solar-electric powered aircraft, underlying physics models are used.  
-Environmental, aircraft performance, and structural models are used in the optimization model to capture the effects of the requirements on the size and design of the aircraft.
-Equations from these models are expressed in a GP-compatible form to enable their use in the optimization. 
-Understanding the models that are used in a GP is important because the solution to a GP is only as accurate as the models that are used to construct the program.  
-To obtain higher fidelity results from the optimization, more detailed models can be implemented.
-
-\subsection{Shared Physical Models}
-
-The models described in this section were applied to both the gas powered and solar-electric powered aircraft models.
-
-\subsubsection{Steady Level Flight}
-
-Both the gas and solar powered aircraft are assumed to be in steady level flight,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:slfthrust}
-    T &\geq \frac{1}{2} \rho V^2 C_D S\\
-    \label{e:slfweight}
-    W &= \frac{1}{2} \rho V^2 C_L S. 
-\end{align}
-
-The shaft power produced by the engine or motor can be expressed as
-
-\begin{equation}
-    \label{e:slfpower}
-    P_{\text{shaft}} = \frac{TV}{\eta_{\text{prop}}}
-    \end{equation}
-
-    where the propulsive efficiency is assumed to be a constant, $\eta_{\text{prop}} = 0.8$.
-
-\subsubsection{Aerodynamics}
-
-The aircraft aerodynamics is modeled as a drag build up of the non-wing drag, wing profile drag, and induced drag coefficients,
-
-\begin{equation}
-    \label{e:aerodragb}
-    C_D \geq C_{d_0} + c_{d_p} + \frac{C_L^2}{\pi e A}.
-    \end{equation}
-
-where the span efficiency factor is assumed to be constant, $e=0.9$. 
-The non-wing drag coefficient $C_{d_0}$, is estimated as a drag build up from the fuselage, tail boom, and horizontal and vertical surfaces discussed later.
-
-    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix A for a discussion of the airfoil choice). 
-    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation. (See Appendix B)
-    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.
-
-%     \begin{align}
-%         \label{e:aerodragprof}
-%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
-%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
-%         Re &= \frac{\rho V S/b}{\mu}
-%     \end{align}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{jho1polarfit1.eps}
-    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles).  (Log-space RMS error~=~0.00489.)}}
-	\label{f:JH01polar}
-	\end{center}
-\end{figure}
-
-% \subsection{Comparison to Existing Platforms}
-% 
-% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
-% Known parameters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
-% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}.
-The model is then solved by maximizing endurance.
-% 
-% \begin{longtable}{lcccc}
-%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
-%      \toprule
-%      \toprule
-%      \label{t:existingparams}
-%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
-%      \midrule
-%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
-%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
-%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
-%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
-%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
-%      \bottomrule
-%  \end{longtable}
-% 
-%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
-% 
-% \begin{longtable}{ll}
-%     \caption{Fixed Optimization Parameters} \\
-%      \toprule
-%      \toprule
-%      Variable &  Value \\
-%      \midrule
-%      $\eta_{prop}$        & 0.65   \\
-%      $A$                 & 25     \\
-%      $f_{\text{structural}}$     & 0.35   \\
-%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
-%      \bottomrule
-%      \label{t:compparams}
-%  \end{longtable}
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
-%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Comparison of optimization results to existing platforms.
-Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
-%  \label{f:vehiclecomp}
-% \end{figure}
-%  
-% \begin{longtable}{lcccc}
-% \caption{Optimization Results Comparison} \\
-% \label{t:existingplatforms} \\
-%      \toprule
-%      \toprule
-%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
-%      \midrule
-%      Orion      & 5 & 4.4   & 132 & 146.8\\
-%      Predator   & 1  & 1.7  & 55 & 39.0 \\
-%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
-%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
-%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
-%      \bottomrule
-%  \end{longtable}
-% 
-% 
-% Because of limited information about these aircraft the optimization predictions are not exact.
-However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.
-
-% \section{Long-endurance Aircraft Trade Space}
-% 
-% One way to analyze the trade space is to determine the structural weight of the aircraft as some percentage of the total aircraft weight, 
-% 
-% \begin{equation}
-%     \label{e:structualfraction}
-%     W_{\text{structural}} \geq f_{\text{structural}}\text{MTOW}.
-% \end{equation}
-% 
-% where the structural fraction is assumed to be, $f_{\text{structural}} = 0.35$.
-Using this assumption and models described in section III, a comparison of the long-endurance trade space is achieved by comparing solar-electric and gas powered aircraft for different parameters.
-% 
-% % \begin{longtable}{llll}
-% % \caption{Fixed Optimization Parameters} \\
-% % \toprule
-% % \toprule
-% % \multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
-% % \midrule
-% % Latitude [deg]       & 45     & Latitude [deg]       & 30 \\
-% % Altitude [ft]        & 16,000 & Altitude [ft]        & 50,000        \\
-% % Payload Weight [lbf] & 10     & Payload Weight [lbf] & 10            \\
-% % $\eta_{prop}$        & 0.75   & $\eta_{prop}$        & 0.75          \\
-% % $A$                 & 25     & $A$                 & 25              \\
-% % $f_{\text{structural}}$ & 0.35   & $f_{\text{structural}}$ & 0.35    \\
-% % $c_{d_0}$            & 0.005  & $c_{d_0}$            & 0.002         \\ 
-% % $\text{BSFC}$               & 0.6    & $\eta_{\text{charge}}$      & 0.98   \\
-% % Endurance [days]     & 8      & $\eta_{\text{discharge}}$   & 0.98   \\
-% %                      &        & $h_{\text{batt}}$ [Whr/kg]  & 350 \\
-% %                      &        & $\rho_{\text{solar}}$ [kg/m$^2$] & 0.3 \\
-% %                      &        & Day of the Year      & 355 (Dec 21st)\\
-% % \bottomrule
-% % \label{t:gassolarparams}
-% %  \end{longtable}
-% 
-% 
-% \subsection{Sizing for Different Latitude Bands}
-% 
-% An analysis of the max take-off weight required to operate at different latitudes reveals that is it is more difficult for solar-electric powered aircraft to operate at higher wind speeds than it is for gas powered aircraft.
-% This analysis is achieved by minimizing the max take-off weight in both the gas and solar-electric powered optimization models at different latitudes. 
-% For this study, the gas-powered aircraft is constrained to fly for 8 days. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
-%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{ Trade study of max take-off weight and latitude.
-Plots show required maximum take-off weight required to operate between a band of latitudes.
-A comparison of the solar and gas powered latitude analysis shows that gas powered aircraft are more robust to operating in higher wind speeds.
-}}
-%  \label{f:latvsmtowtrade}
-% \end{figure}
-% 
-% One way to interpret Figure~\ref{f:latvsmtowtrade} is that a solar-powered aircraft weighing 175 lbs is able to operate between $\pm$40 degrees North latitude in 90th percentile wind speeds.  
-% This analysis shows that gas powered architectures are able to operate in more locations than solar-electric powered aircraft.  
-% Figure~\ref{f:mtowvslatgas} levels off around 35 degrees latitude because wind speeds are highest at that latitude at 15,000 ft. 
-% Thus, if the gas powered aircraft can fly at 35 degrees latitude it can fly in any band of latitudes.  
-% On the other hand, solar-electric powered aircraft design becomes less feasible at higher latitudes because even though wind speeds peak around 42 degrees latitude at 60,000 ft, solar flux continues to decrease at higher latitudes making it difficult to reach latitude bands greater than $\pm$50 degrees.
-% 
-% \subsection{Gas Powered Endurance}
-% 
-% While, gas powered aircraft may be more robust to high wind speeds, solar-electric powered aircraft can achieve higher endurance. 
-% The gas powered optimization model was set to minimize max take-off weight for different endurance requirements to understand what endurance limits are possible. 
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}
-%     \caption{ \textbf{ Trade study between endurance and max take-off weight for a gas powered architecture. }}
-% 	\label{f:mtowvsendsimple}
-% 	\end{center}
-% \end{figure}
-% 
-% For a gas-powered architecture, achieving endurances above 10 days is difficult.  Solar-electric powered aircraft, however, can theoretically fly indefinitely.
-
-% \subsection{Solar-electric Powered Trade Space Exploration}
-% 
-% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.
-(Figure~\ref{f:solarcontours})
-% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
-% 
-%  \begin{figure}[H]
-%  \begin{subfigmatrix}{3}% number of columns
-%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
-%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
-%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
-%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
-%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
-%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
-%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
-%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
-%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
-%  \end{subfigmatrix}
-%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft.
-Each point on the graph corresponds to a unique design.} }
-%  \label{f:solarcontours}
-% \end{figure}
-
-\subsubsection{Wing Spar Model}
-
-A wing spar is one of the primary structural elements of both aircraft. 
-A conservative approach to calculating the size of the spar is to assume that the spar carries all of the bending loads caused by the lifting loads and weight of the aircraft.  
-It is assumed that there are two out of plane bending cases to which both the gas powered and the solar-electric powered aircraft are subject: standard wing bending and gust loads.
-
-%\begin{itemize}
-
-\paragraph{\textbf{Standard Wing Bending Case}} 
-The distributed load along the wing is a combination of the lifting and weight distributions along the wing.
-The wing loading distribution $q(y)$, can be approximated as a scaling of the local chord,\cite{bending}
-
-\begin{equation}
-    \label{e:wingloading}
-    q(y) \approx K_q c(y) 
-\end{equation}
-
-where $y$ is the distance from the root wing location. The loading constant $K_q$\cite{bending}, is defined as
-
-\begin{equation}
-    \label{e:kq}
-    K_q = \frac{N_{\text{max}}W_{\text{cent}}}{S}
-\end{equation}
-
-where $W_{\text{cent}}$ is the sum of loads acting at the center of the aircraft and the safety load factor is an input $N_{\text{max}}=5$ ($N_{\text{max}}=1$ corresponds to steady level flight).
-Using the equation for the local chord of a constant tapered wing\cite{bending} with a taper ratio $\lambda=0.5$, the pre-computed distributed load
-
-% \begin{equation}
-%     \label{e:localchord}
-%     \bar{c}(y) \equiv \frac{c(y)}{S/b} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right)
-% \end{equation}
-
-% where the taper ratio is an input, $\lambda=0.5$.  Thus, the pre-computed distributed load
-
-\begin{equation}
-    \label{e:qbar}
-    \bar{q}(y) \equiv \frac{q(y)b}{N_{\text{max}}W_{\text{cent}}} = \frac{2}{1+\lambda} \left( 1 + (\lambda - 1) \frac{2y}{b} \right),
-\end{equation}
-
-is inputted into a standard beam model to predict the bending moments and deflections. The center weight $W_{\text{cent}}$ for the gas powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{fuel}} + W_{\text{fuselage}} + W_{\text{engine}} + W_{\text{payload}} + W_{\text{emp}}.
-\end{equation}
-
-The center weight for the solar powered aircraft is
-
-\begin{equation}
-    W_{\text{cent}} \geq W_{\text{payload}} + W_{\text{motor}} + W_{\text{emp}}.
-\end{equation}
-
-\paragraph{\textbf{Gust Loading Case}}
-
-Because long-endurance aircraft typically have high aspect ratio wings, they are considered flexible aircraft. 
-To account for this flexibility and size a structural wing spar that will ensure enough rigidity in the event of a gust load, a distributed gust lifting load is added to the steady level flight loading distribution
-
-\begin{align}
-    q(y) &= \frac{N_{\text{max}}W_{\text{cent}}}{b}\bar{c}(y) + c_{l_{\alpha}} \alpha_{\text{gust}} \frac{1}{2} \rho V^2 \frac{S}{b}\bar{c}(y) \\
-    \bar{q}(y) = \frac{q(y)b}{W_{\text{cent}}N_{\text{max}}} &\geq \bar{c}(y) \left[1 + \frac{c_{l_{\alpha}}}{C_L} \alpha_{\text{gust}} (y) \left(1 + \frac{W_{\text{wing}}}{W_{\text{cent}}} \right) \right]
-\end{align}
-
-where $\bar{c}(y)$ is pre-computed prior to the optimization solve for a taper ratio $\lambda = 0.5$.
-The safety load factor is $N_{\text{max}}=2$ for the gust loading case.
-The weight of the wing for the gas powered aircraft is assumed to be the weight of the spar cap plus the weight of the skin
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}}.
-\end{equation}
-
-For the solar-electric powered aircraft the batteries and solar cells are also included in the wing weight
-
-\begin{equation}
-    W_{\text{wing}} \geq W_{\text{spar}} + W_{\text{skin}} + W_{\text{batt}} + W_{\text{solar}}.
-\end{equation}
-
-The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by
-
-\begin{equation}
-    \label{e:gustalpha}
-    \alpha_{\text{gust}}(y)  = \tan^{-1}\left(\frac{V_{\text{gust}}(y)}{V} \right).
-\end{equation}
-
-Because the $\arctan$ function is not GP compatible, a monomial approximation was calculated using techniques described by Hoburg et.~al\cite{fitting},
-
-\begin{equation}
-    \alpha_{\text{gust}}(y)  = 0.946 \left(\frac{V_{\text{gust}}(y)}{V} \right)^{0.996}
-\end{equation}
-
-and has an RMS log-space error of 0.039 for $V_{\text{gust}}/V \in [0, 0.7]$.
-The gust velocity has an assumed profile along the wing\cite{acgust},
-
-\begin{equation}
-    \label{e:gustwind}
-    V_{\text{gust}}(y) = V_{\text{ref}} \left(1-\cos\left(\frac{2y}{b} \frac{\pi}{2} \right) \right)
-\end{equation}
-
-where the reference velocity is an assumed conservative value\cite{acgust}, $V_{\text{ref}} = 10$ [m/s]. The gust velocity profile is computed prior to solve to preserve GP-compatibility.
-
-\paragraph{\textbf{Discretized Beam Model}}
-
-Using a standard Bernoulli-Euler discretized beam model with $n=5$ nodes\cite{bending}, the shear forces, moments, angles, and deflections can be expressed in a GP-compatible form using the distributed loads $q(y)$, as an input,
-
-% \begin{align}
-%     \label{e:shear}
-%     \mathcal{S}_i &= \mathcal{S}_{i+1} - \frac{q_{i+1} + q_i}{2}\Delta y \\
-%     \label{e:moment}
-%     M_i &= M_{i+1} - \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2}\Delta y \\
-%     \label{e:shearboundary}
-%     \mathcal{S}_n &= 0 \\
-%     \label{e:momentboundary}
-%     M_n &= 0
-% \end{align}
-% 
-% Similarly, the angle deflection and deflection can be calculated with boundary conditions of zero angle and deflection and the wing root.\cite{bending}
-% 
-% \begin{align}
-%     \label{e:angle}
-%     \Theta_{i} &= \Theta_{i+1} + \frac{1}{2} \left(\frac{M_i}{EI_i} + \frac{M_{i-1}}{EI_{i-1}} \right) \Delta y \\
-%     \label{e:deflection}
-%     w_{i} &= w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y \\
-%     \label{e:angleboundary}
-%     \Theta_0 &= 0 \\
-%     \label{e:defboundary}
-%     w_0 &= 0 
-% \end{align}
-%  
-% Equations~\eqref{e:shear}-\eqref{e:defboundary} are GP-compatible if expressed as
-
-\begin{align}
-    \label{e:sheargp}
-    \mathcal{S}_{i+1} &\geq \mathcal{S}_i + \frac{q_{i+1} + q_i}{2} \Delta y \\
-    \label{e:momentgp}
-    \mathcal{M}_{i+1} &\geq \mathcal{M}_i + \frac{\mathcal{S}_{i+1} + \mathcal{S}_i}{2} \Delta y \\
-    \label{e:anglegp}
-    \Theta_{i} &\geq \Theta_{i+1} + \frac{1}{2} \left(\frac{\mathcal{M}_i}{EI_i} + \frac{\mathcal{M}_{i-1}}{EI_{i-1}} \right) \Delta y \\
-    \label{e:deflection}
-    w_{i} &\geq w_{i+1} + \frac{1}{2} (\Theta_i + \Theta_{i-1}) \Delta y .
-\end{align}
-
-where the Young's Modulus of carbon fiber is $E = 20$ [MPa]. Boundary conditions are zero shear forces and moments at the wing tips and zero angle and deflection at the wing root.\cite{bending} \\
-
-\textbf{Cap Spar for Bending Loads} 
-A cap spar is considered for the solar-electric and gas powered aircraft.
-A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.
-A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.9\textwidth,natwidth=547,natheight=74]{capsparcross.eps}
-    \caption{\textbf{Cross sectional view of a cap spar.}}
-	\label{f:capspar}
-	\end{center}
-\end{figure}
-
-The moment of inertia of the cap spar is modeled by only considering the spar caps, not the foam interior.  
-This conservative assumption is made because the contribution of the foam core is much less than that of the spar caps.  
-The equation for the moment of inertia\cite{bending} of a cap spar is
-
-\begin{equation}
-    \label{e:moispar}
-    I = \frac{w_{\text{cap}}t_{\text{cap}}^3}{6} + 2w_{\text{cap}}t_{\text{cap}}\left( \frac{h_{\text{cap}}}{2} + \frac{t_{\text{cap}}}{2} \right)^2.
-\end{equation}
-
-This equation is not GP compatible.  However, using a first order conservative approximation, the moment of inertia can be simplified to be written in a GP-compatible form,
-
-\begin{equation}
-    \label{e:moispar}
-    I \leq 2w_{\text{cap}}t_{\text{cap}}\left(\frac{h_{\text{cap}}}{2}\right)^2.
-\end{equation}
-
-There are also geometric constraints imposed on the width and thickness.
-The total spar cap thickness cannot be greater than the thickness of the airfoil cross section, $\tau_t = 0.115$.
-The width of the spar cap is assumed no greater than 30\% of chord, $\tau_w = 0.3$.
-
-\begin{align}
-    \label{e:thickness}
-    c(y)\tau_t &\geq h_{\text{cap}} + 2t_{\text{cap}} \\
-    \label{e:width}
-    c(y)\tau_w &\geq w_{\text{cap}} 
-    \end{align}
-
-To match the discretized beam model, the spar cross section can also be written in a discretized form such that each section has a unique width and thickness.
-
-\begin{align}
-    I_i &\leq 2w_{\text{cap}_i}t_{\text{cap}_i}\left(\frac{h_{\text{cap}_i}}{2}\right)^2 \\
-    c(y)\tau_t &\geq h_{\text{cap}_i} + 2t_{\text{cap}_i} \\
-    c(y)\tau_w &\geq w_{\text{cap}_i} 
-\end{align}
-
-The wing spar at each section root must be strong enough to withstand the bending moment and stiff enough to not exceed some deflection limit.
-Both constraints are imposed in the optimization model as
-
-\begin{align}
-    \label{e:stresscont}
-    \sigma_{\text{cfrp}} &\geq \mathcal{M}_i \frac{h_{\text{cap}_i}+t_{\text{cap}_i}}{I_i}\\
-    \label{e:defcont}
-    w_n &\leq w_{\text{max}}
-\end{align}
-
-where the ultimate tensile strength for unidirectional carbon fiber is $\sigma_{\text{cfrp}} = 1700$ [MPa].\cite{cfprop}
-The tip deflection is constrained to be less than 20\% of the half span, $\frac{w_{\text{max}}}{b/2} = 0.2$.
-
-Finally, the weight of the spar cap is computed as
-
-\begin{align}
-    \label{e:sparmass}
-    \Delta W_i &\geq \rho_{\text{cfrp}} w_{\text{cap}_i}t_{\text{cap}_i} \frac{b/2}{n-1}g \\
-    \label{e:sparmasssum}
-    W_{\text{spar}} &\geq 2 \sum\limits_{1}^{n-1} \Delta W_i
-\end{align}
-
-where $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$].\cite{cfply}
-
-% \end{itemize}
-
-\subsubsection{Additional Wing Weight}
-
-It is assumed that the wing skin is made of carbon fiber.  The weight of the wing skin is
-
-\begin{equation}
-    \label{e:wingskinweight}
-    W_{\text{skin}} \geq 2 \rho_{A_{\text{cfrp}}} S g.
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], or approximately the area density of one ply of carbon fiber.\cite{cfply} The wing skin is asssumed not to contribute to the bending stiffness.
-
-Additional wing weight $W_{\text{fadd}}$, accounts for additional structural weight (ribs, rear spar, actuators, etc.),
-
-\begin{equation}
-    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
-\end{equation}
-
-where $m_{\text{fac}} = 1.2$.
-
-\subsubsection{Empennage}
-
-An empennage model is added to both the solar-electric and gas powered aircraft models.  The empennage model consists of a single tail boom, horizontal tail and vertical tail.  
-The empennage adds both weight and drag to each aircraft.
-
-The tail boom has an optimized diameter $d$, root wall thickness $t_0$, root moment of inertia $I_0$, modulus $E=150$ [GPa]\cite{cfprop}, density $\rho_{\text{cfrp}} = 1.6$ [g/cm$^3$]\cite{cfprop}, and length $l_{\text{h}}$.
-The total mass and root bending inertia are imposed in the optimization model as
-
-\begin{align}
-    m &\geq \pi \rho_{\text{cfrp}} t_0 d l_{\text{h}} \left( 1 - \frac{1}{2} k\right) \\
-    I_0 &\leq \pi t_0 d^3/8
-\end{align}
-
-where the index $k=0$ corresponds to a uniform wall thickness and stiffness, and $k=1$ corresponds to a linear drop-off to zero.
-For both the solar-electric and gas powered aircraft $k=0.8$ is assumed.
-When the tail boom is loaded at the endpoint $x=l_{\text{h}}$, by the horizontal tail lift $L_{\text{h}}$, the end deflection angle follows from standard beam analysis
-
-\begin{align}
-    \label{e:boomdefl}
-    \theta &\geq \frac{L_{\text{h}} l_{\text{h}}^2}{EI_0} \frac{1+k}{2} \\
-    L_{\text{h}} &= \frac{1}{2} C_{L_{\text{h}}} \rho V^2 S_{\text{h}}.
-\end{align}
-
-The horizontal tail is sized to satisfy a horizontal tail volume coefficient condition $V_{\text{h}} = 0.45$,\cite{aircraftrules}
-
-\begin{equation}
-    V_{\text{h}} = \frac{S_{\text{h}}l_{\text{h}}}{Sc}
-\end{equation}
-
-% meet a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
-% 
-% \begin{equation}
-%     \label{e:deltacg}
-%     \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2
-% \end{equation}
-% 
-% A single constraint, whose derivation is explained in Appendix C, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
-travel range,
-% 
-% \begin{align}
-%     \label{e:smcorr}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:htv}
-%     V_{\text{w}} &= \frac{S_{\text{h}}}{S} \frac{l_{\text{h}}}{c}
-% \end{align}
-% 
-% Because this equation is not posynomial, it was reformulated as conservative monomial approximation
-% 
-% \begin{align}
-%     1 &\geq z_{\text{SM}_1} \left(\frac{m_{\text{w}} \mathcal{F}_{NE}}{m_{\text{h}} V_{\text{w}}})\right) + \frac{d\epsilon}{d\alpha} \\
-%     z_{\text{SM}_2} &\leq V_{\text{w}}\frac{(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-%     \label{e:smcorrmon}
-%     z_{\text{SM}_1}^{0.48} z_{\text{SM}_1}^{0.52} &\geq \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}}
-% \end{align}
-% 
-% where the values, 0.48 and 0.52, are chosen as an approximate percentage that each term is of the right hand side of Equation~\eqref{e:smcorrmon}.
-
-The vertical tail is sized to meet a conservative tail volume coefficient $V_{\text{v}}= 0.04$,\cite{aircraftrules}
-
-\begin{equation}
-    \label{e:vtv}
-    V_{\text{v}} = \frac{S_{\text{v}}}{S} \frac{l_{\text{v}}}{b}
-\end{equation}
-
-where $l_{\text{v}}$ is the vertical tail moment arm, assumed to be equal to the horizontal tail moment arm, $l_{\text{v}} = l_{\text{h}}$.
-
-Both the horizontal and vertical tails are assumed to have a carbon fiber skin and solid foam interior where their respective densities are $\rho_{A_{\text{cfrp}}} = 0.049$ [g/cm$^2$], $\rho_{\text{foam}} = 1.5$ [lbf/ft$^3$].
-The weight of the tails is
-
-\begin{equation}
-    W_{\text{(v,h)}}/m_{\text{fac}} = \rho_{\text{foam}} \frac{S_{\text{(v,h)}}^2}{b_{\text{(v,h)}}} \bar{A} + g\rho_{A_{\text{cfrp}}} S_{\text{(v,h)}} \\
-\end{equation}
-
-where $b_{\text{h}}$ and $b_{\text{v}}$ are the spans of the horizontal and vertical tails respectively and $\bar{A}$ is the cross sectional area of the NACA 0008 airfoil.
-The margin factor $m_{\text{fac}}=1.1$, is included to account for control surfaces, attachment joints, actuators, etc.
-
-The drag of the empennage was modeled as three separate parts with no interference drag.  The drag of the tail boom is calculated using a turbulent flat plate model,
-
-\begin{align}
-    \label{e:boomdrag}
-    D_{\text{boom}} &\geq \frac{1}{2} C_f \rho V^2 l_{\text{h}}\pi d \\
-    C_f &\geq \frac{0.445}{Re_{\text{boom}}^{0.3}} \\
-    Re_{\text{boom}} &= \frac{V\rho l_{\text{h}}}{\mu}
-\end{align}
-
-The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
-where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
-The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.
-
-% \begin{align}
-%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
-%     \label{e:taildrag}
-%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
-%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
-%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
-% \end{align}
-
-% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.
-
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
-%     \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles). (Log-space RMS error = 0.0173)}}
-% 	\label{f:taildragpolar}
-% 	\end{center}
-% \end{figure}
-
-% \subsection{Weight Breakdown}
-% 
-% The weight of each component of both the gas and solar platforms is summed to constrain the max take-off weight.  
-% The structural weight comprises the weight of the wing, fuselage, horizontal and vertical stabilizers, and engine. 
-% The weight breakdown for the gas powered aircraft is a summation of the structural weight, the total fuel weight, and the payload weight. 
-% 
-% \begin{align}
-%     \label{e:weightmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}}  + W_{\text{payload}} + W_{\text{fuel}} + W_{\text{engine}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} + W_{\text{fuse}}
-% \end{align}
-% 
-% A similar weight breakdown is done for the solar-electric powered aircraft, 
-% 
-% \begin{align}
-%     \label{e:weightsmtow}
-%     W_{\text{MTO}} &\geq W_{\text{structural}} + W_{\text{payload}} + W_{\text{solar}} + W_{\text{batt}} + W_{\text{motor}} \\
-%     W_{\text{structural}} &\geq W_{\text{wing}} + W_{\text{boom}} + W_{\text{h}}+ W_{\text{v}} \\
-%     W_{\text{solar}} &\geq \rho_{\text{solar}} S_{\text{solar}} g \\
-%     W_{\text{batt}} &\geq \frac{E_{\text{batt}}}{h_{\text{batt}}} g
-% \end{align}
-% 
-% where the solar cell density is $\rho_{\text{solar}} = 0.27$ [kg/m$^2$].\cite{solartech}\cite{solarparam}
-% Lithim-sulfur batteries are selected with a battery specific energy $h_{\text{batt}} = 350$ [Whr/kg].\cite{lithiumsul}
-% The payload weight is assumed to be $W_{\text{payload}} = 10$ [lbs].
-
-\subsection{Solar-Electric Aircraft Physical Models}
-
-\subsubsection{Wind Speeds}
-
-There exists an optimum flight altitude for solar-electric powered aircraft.  
-The local minimum in wind speed around 19,000 m and the variation of air density with altitude suggest that the altitude should be an output of the optimization. 
-To accomplish this, constraints relating wind speed to altitude are imposed. 
-This approach assumes that the solar-electric powered aircraft will not be confined to fly at a single altitude, but will fly at the altitude where conditions are most favorable.
-
-Because wind speeds do not increase monotonically with latitude, it is possible that the design for a given latitude is constrained by winds at a latitude inside the required latitude band. 
-To handle this, a multipoint set of latitude constraints is imposed. 
-The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
-In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
-% The constraint on December wind speeds and air density at the 30 degree latitude is 
-% 
-% \begin{align}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
-%     \label{e:windfit35}
-%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
-% \end{align}
-% 
-% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed.
-The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.
-% 
-% \begin{figure}[H]
-% 	\begin{center}
-% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
-%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
-% 	\label{f:windfitl35}
-% 	\end{center}
-% \end{figure}
-
-For the gas powered aircraft, the altitude is not optimized.  
-Gas-powered aircraft, which are naturally aspirated, will be less efficient at higher altitudes.  
-Additionally, wind speeds increase monotonically from sea level up to $\sim$9,100 m.  
-Therefore, the altitude for gas-powered aircraft is determined as the minimum altitude necessary for the payload to be effective.\cite{orion}
-For this sizing study, the altitude requirement for the gas powered aircraft is $h_{\text{min}}=4,500$ m, corresponding to a 100 km diameter footprint with a 5 degree lookup angle.
-
-\subsubsection{Solar-Electric Power}
-
-At a given time of year DOY, and latitude $\phi$, there exists an available solar irradiated energy $(E/S)_{\text{sun}}$ per unit area during one day,
-
-\begin{equation}
-    \label{e:solarfunc}
-    (E/S)_{\text{sun}} = f(\phi, \text{DOY}).
-\end{equation}
-
-To operate continually for multiple days the total available solar energy $(E/S)_{\text{sun}}$, must be greater than the required pre-conversion-efficiency solar energy per unit area to power the aircraft during the day $(E/S)_{\text{day}}$, and the energy required to power the aircraft during the night via batteries $E_{\text{batt}}$,\cite{solartech}
-
-\begin{align}
-    \label{e:solarreq}
-    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
-    \label{e:solarbatt}
-    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
-\end{align}
-
-where the charge, discharge and solar cell efficiencies are assumed to be $\eta_{\text{charge}} = \eta_{\text{discharge}} = 0.98$ and $\eta_{\text{solar}}= 0.22$. 
-An additional energy term $(E/S)_{\text{twilight}}$ is included in Equation~\eqref{e:solarbatt} to account for hours of low solar irradiance, during mornings and evenings, when the aircraft must continue to fly on partial battery power.
-Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt} are graphically represented in Figure~\ref{f:lat30}, which represents the solar energy at 30 degrees latitude.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
-    \caption{\textbf{Solar power on Dec 21st at 30$^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
-	\label{f:lat30}
-	\end{center}
-\end{figure}
-
-In order for the aircraft to begin charging batteries there must be a minimum solar irradiance power
-
-\begin{align}
-    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
-    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
-\end{align}
-
-where $\eta_{\text{motor}} = 0.95$. 
-Thus, for a given latitude and day, and therefore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
-The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude.
-$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
-% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
-%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
-%  \end{subfigmatrix}
-%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
-%  \label{f:energyapprox}
-% \end{figure}
-
-For the purposes of this design study, it is assumed that the solar cells are placed on the wing and the horizontal tail but not on the vertical tail. 
-The fractional solar cell area index $f_{\text{solar}}$ is one when the solar cells completely cover the main wing and greater than one if solar cells are also placed on the horizontal tail,
-
-\begin{equation}
-    \label{e:solarssolar}
-    S_{\text{solar}} \leq f_{\text{solar}}S.
-\end{equation}
-
-\subsubsection{Motor Weight}
-
-The solar-electric powered aircraft has a motor whose weight is based on the approximation\cite{electricengine}
-
-\begin{equation}
-    \label{e:electricengine}
-    P_{\text{max}} = B_{\text{PM}} m_{\text{motor}}
-\end{equation}
-
-where $m_{\text{motor}}$ is the motor mass, $P_{\text{max}} \geq P_{\text{oper}}$ is the maximum operating power, and the assumed power to mass ratio is $B_{PM} = 4140.8$ [W/kg]. \\
-
-\subsection{Gas Powered Aircraft Physical Models}
-
-\subsubsection{Breguet Endurance}
-
-A key sizing equation for a long endurance gas powered aircraft is the Breguet Range equation.  
-For GP-compatibility and to optimize endurance, not range, a variation of the Breguet Range equation is used,
-
-\begin{equation}
-    \label{e:breguetendurance}
-    t = \frac{W_{\text{ave}}}{P_{\text{shaft}}\text{BSFC}g} \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}}\right)}.
-\end{equation}
-
-% The derivation begins with the differential form of Breguet Range,\cite{br2}
-%
-% \begin{equation}
-%     \label{e:breguetdiff}
-%     -\frac{dW}{dt} = g\dot{m}_{\text{fuel}}.
-% \end{equation}
-% 
-% Using the definition of $\text{BSFC}$
-% 
-% \begin{equation}
-%     \label{e:brBSFC}
-%     \text{BSFC} = \frac{\dot{m}_{\text{fuel}}}{P_{\text{shaft}}},
-% \end{equation}
-% 
-% Equation~\eqref{e:breguetdiff} can be written as
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -dW = g P_{\text{shaft}} \text{BSFC} dt.
-% \end{equation}
-
-This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
-One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
-The segment weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as
-
-% Dividing by $W$,
-% 
-% \begin{equation}
-%     \label{e:brdiff2}
-%     -\frac{dW}{W} = \frac{g P_{\text{shaft}}\text{BSFC} }{W} dt,
-% \end{equation}
-% 
-% and integrating, the Breguet Range equation can be expressed as
-% 
-% \begin{equation}
-%     \label{e:be1}
-%     \ln{\left( \frac{W_{\text{initial}}}{W_{\text{final}}} \right)} = \frac{gP_{\text{shaft}}\text{BSFC}}{W_{\text{ave}}} t,
-% \end{equation}
-% 
-% where $W_{\text{ave}}$ is the average weight of the aircraft during the flight segment.
-
-\begin{equation}
-    \label{e:gpmean}
-    W_{\text{ave}} = \sqrt{W_{\text{initial}}W_{\text{final}}}.
-\end{equation}
-
-    To make Equation~\eqref{e:breguetendurance} GP compatible, a Taylor expansion is used,\cite{hoburgthesis}
-
-\begin{align}
-    \label{e:brzbre}
-    z_{\text{bre}} &\geq \frac{P_{\text{shaft}}t \text{BSFC} g}{W}\\
-    \label{e:brtaylor}
-    \frac{W_{\text{fuel}}}{W_\text{final}} &\geq z_{\text{bre}} + \frac{z_{\text{bre}}^2}{2} + \frac{z_{\text{bre}}^3}{6} + \frac{z_{\text{bre}}^4}{24} + \dots
-\end{align}
-
-    Equations~\eqref{e:brzbre} and~\eqref{e:brtaylor} are monomial and posynomial respectively and therefore GP compatible.
-For long-endurance aircraft, missions can last days, causing the power to weight ratio $(P_{\text{shaft}}/W)$ to vary significantly during the course of the flight.
-    Equations~\eqref{e:brzbre},~\eqref{e:brtaylor}, and~\eqref{e:slfweight} can be discretized to account for this,
-
-\begin{align}
-    \label{e:slfweightd}
-    \sqrt{W_i W_{i+1}} &= \frac{1}{2} \rho_i V_i^2 C_{L_i} S \\
-    \label{e:brzbred}
-    z_{bre_i} &\geq \frac{P_{\text{shaft}_i}t_i \text{BSFC} g}{\sqrt{W_i W_{i+1}}}\\
-    \label{e:brtaylord}
-    \frac{W_{\text{fuel}_i}}{W_{i+1}} &\geq z_{bre_i} + \frac{z_{bre_i}^2}{2} + \frac{z_{bre_i}^3}{6} + \frac{z_{bre_i}^3}{24}.
-    \end{align}
-
-    For evaluation of long-endurance, gas-powered aircraft a discretization of $N=5$ was used.
-
-\subsubsection{Elliptical Fuselage}
-
-For the gas powered aircraft it is assumed that the fuel is carried in an elliptically shaped fuselage.
-The fuselage will increase the overall weight and drag of the aircraft.
-The solar-electric powered aircraft is assumed to carry the batteries in the wings will therefore have a small fuselage whose effects will be ignored.
-
-The driving constraint for the size of the fuselage is to ensure that all of the fuel required for the mission can fit inside the fuselage,
-
-\begin{equation}
-    \label{e:fusevol}
-    \mathcal{V}_{\text{fuse}} \geq \frac{W_\text{fuel}}{\rho_\text{fuel}}
-\end{equation}
-
-where the fuel is assumed to have a density $\rho_\text{fuel} = 6.01$ [lbf/gallon].  The dimensions of the fuselage are constrained by
-
-\begin{equation}
-    \label{e:fusevol2}
-    \mathcal{V}_{\text{fuse}} \leq \frac{4}{3}\pi \frac{l_{\text{fuse}}}{2}R_{\text{fuse}}^2
-\end{equation}
-
-where $l_{\text{fuse}}$ is the length of the fuselage and $R_{\text{fuse}}$ is the radius.
-Using the length and radius, the surface area can be calculated using Thomsen's approximation,\cite{ellipsoidSA}
-
-\begin{equation}
-    \label{e:fusesa}
-    3 \left( \frac{S_{\text{fuse}}}{\pi} \right)^{1.6075} \geq 2(2l_{\text{fuse}}R_{\text{fuse}})^{1.6075} + (4R_{\text{fuse}}^2)^{1.6075}.
-\end{equation}
-
-The weight of the fuselage is constrained by
-
-\begin{equation}
-    \label{e:fuseweight}
-    W_{\text{fuse}} \geq S_{\text{fuse}} \rho_{A_{\text{cfrp}}} g
-\end{equation}
-
-where $\rho_{A_{\text{cfrp}}} = 0.0975$ [g/cm$^2$], or the area density of two plies of carbon fiber.\cite{cfply}  The surface area is also used to calculate the drag assuming a skin friction based drag model,
-
-\begin{align}
-    \label{e:fusedrag}
-    D_{\text{fuse}} &\geq C_f k_{\text{fuse}} \frac{1}{2} \rho V^2 S_{\text{fuse}} \\
-    C_f &\geq \frac{0.455}{Re^{0.3}}
-\end{align}
-
-where $k_{\text{fuse}}$ is the form factor approximated by\cite{raymer}
-
-\begin{equation}
-    \label{e:fuseform}
-    k_{\text{fuse}} \geq 1 + \frac{60}{(l_{\text{fuse}}/2R_{\text{fuse}})^3} + \frac{(l_{\text{fuse}}/2R_{\text{fuse}})}{400}.
-\end{equation}
-
-\subsubsection{Engine Weight}
-
-The engine weight of the gas powered aircraft is governed by a simple power law derived from existing two-stroke and four-stroke engines\cite{gasengine}
-
-\begin{equation}
-    \label{e:powerlaw}
-    \frac{W_{\text{engine}}}{W_{\text{engine-ref}}} = 1.27847 \left(\frac{P_{\text{SL-max}}}{P_{\text{ref}}} \right)^{0.772392}
-\end{equation}
-
-where $W_{\text{engine-ref}} = 10$ [lbs] and $P_{\text{ref}} = 10$ [hp].   
-Equation~\eqref{e:powerlaw} is compared to the data in Figure~\ref{f:powervsweightfit}.
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.eps}
-    \caption{\textbf{Power law fit to University of North Dakota engine weight data\cite{gasengine}. (Log-space RMS error = 0.34)}}
-	\label{f:powervsweightfit}
-	\end{center}
-\end{figure}
-
-\subsubsection{Gas Powered Engine Performance}
-
-Two characteristics of gas engines affect the performance of long-endurance aircraft.  
-The first is brake specific fuel consumption, $\text{BSFC}$; a lower $\text{BSFC}$ will result in increased endurance.  
-The second is the lapse rate.  
-Assuming a propeller driven aircraft and a naturally aspirated engine, as the aircraft reaches higher altitudes, the engine will have decreased available power. 
-To account for these two effects a two-stroke, double cylinder engine, the DF70 from RCV Engines Ltd, England, was selected as a representative engine.  
-RCV Engines Ltd provided manufacturing data that was used to generate representative performance curves.\cite{rcvengines}
-It is assumed that engines of a similar size will perform similarly to the DF70 engine.  
-This engine performance model is only valid for internal combustion engines.
-
-To capture throttling effects as aircraft weight decreases, a GP-compatible curve was generated from the DF70 data to relate $\text{BSFC}$ and shaft power. (Figure~\ref{f:powertobsfc})
-Using this approximation the required shaft power determines the $\text{BSFC}$.
-A comparison of the DF70 data provided by RCV Engine Ltd in Figure~\ref{f:powertobsfc} to the $\text{BSFC}$ to power curves in Goering\cite{bsfcperf} verify that this performance curve is representative of engines of similar sizes.
-
-% \begin{equation}
-%     \label{e:rpmtobsfc}
-%     \left( \frac{\text{BSFC}}{\text{BSFC}_{100\%}} \right)^{18.5563} \geq \num{8.66321d-3}\left( \frac{P_{\text{shaft}}}{P_{\text{max}}}\right)^{-7.70161} + 1.38628\left( \frac{P_{\text{shaft}}}{P_{\text{max}}} \right)^{1.12921} \\
-% \end{equation}
-
-% where $\text{BSFC}_{100\%} = 0.3162$ [kg/kW/hr].\cite{rcvengines}
-
-\begin{figure}[h!]
-	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.eps}
-    \caption{\textbf{Representative engine performance fit based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
-	\label{f:powertobsfc}
-	\end{center}
-\end{figure}
-
-The lapse rate $L_{\text{eng}}(h) \leq 1$, is assumed to affect the maximum power output,
-
-\begin{equation}
-    \label{e:lapse}
-    L_{\text{eng}}(h) \equiv \frac{P_{\text{max}}}{P_{\text{SL-max}}}.
-\end{equation}
-
-The lapse is calculated from the required flight altitude $h=4,500$ [m], prior to the optimization solve using an approximate engine loss rate for normally aspirated engines of 3.5\% hp per 300 m,\cite{enginelapse}
-
-\begin{equation}
-    \label{e:lapsefit}
-    L_{\text{eng}}(h) = 1 - \frac{0.035}{300 \text{ [m]}} h.
-\end{equation}
-
-\subsubsection{Climb Constraints}
-
-Because the gas engine is naturally aspirated, as the aircraft climbs there will be less available power.  
-The climb constraint ultimately sizes the engine because at the top of climb, when the least amount of power is available, the engine must provide the necessary power to meet a minimum climb rate,
-
-\begin{equation}
-    \label{e:climbrate}
-    \dot{h}_{\text{min}} \geq 30 \text{ [m/min]}.
-\end{equation}
-
-The climb rate affects the required thrust, and therefore the required power during climb,
-
-\begin{equation}
-    \label{e:climb}
-    T \geq \frac{1}{2} C_D \rho V^2 S + W \frac{\dot{h}_{\text{min}}}{V}
-\end{equation}
-
-where $W$ is the weight of the aircraft during climb.
-
-\section{Results}
+\chapter{Results}
 
@@
 The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
-This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.
+Highlighted design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements.
+The rest of this chapter explores how these results may vary for different requirements, input parameter values, and physics modeling assumptions.
 
-\newpage
+\footnotesize
 \begin{longtable}{llll}
@@
 
-\subsection{Changing Requirements}
+\input{../svals.generated.tex}
+\input{../gvals.generated.tex}
+
+\normalsize
+
+\section{Changing Requirements}
 
 Important trade offs between the gas powered and solar-electric powered architectures are highlighted and quantified by changing the latitude and endurance requirements.
+
+\subsection{Latitude Requirement Analysis}
+
 Both optimization models were solved by minimizing the max take-off weight across different latitude requirements. 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
     \caption{\textbf{Gas architecture feasible for all latitudes. Next integer latitude for each solar-electric curve is infeasible.}}
@@
 
+\subsection{Endurance Requirement Analysis}
 While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
     \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
@@
 
-% Changing the objective function can be informative to the design space. 
-% High efficiency solar cells are expensive and from a manufacturing perspective, a larger aircraft that requires fewer solar cells could be advantageous.
-% By altering the objective function to solar cell area it is observed that when the objective function is solar cell area the airplane is slightly bigger but requires fewer solar cells.
-(Figure~\ref{f:solarobjcomp})
-% 
-% \begin{figure}[H]
-%  \begin{subfigmatrix}{2}% number of columns
-%      \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
-%      \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
-%  \end{subfigmatrix}
-%     \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
-%     \label{f:solarobjcomp}
-% \end{figure}
+\subsection{Season Requirement Analysis}
 
-\newpage
-\subsection{Changing Parameter Values}
+The season requirement for the solar-electric powered aicraft if met, allows for year round coverage.  
+However, there may be applications that require flying for only 10 months of the year or even half the year.  
+Figure~\ref{f:season} shows which latitudes are feasible for different season requirements and what the aircraft weight would be. 
+Interestingly, flying for only 10 months of the year does not increase the ability to reach higher latitudes.  
+However, there is considerable benefit to only flying for 6 months of the year.
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{season.pdf}
+    \caption{\textbf{Flexible season requirements allow for smaller aircraft and increased latitude.}}
+	\label{f:season}
+	\end{center}
+\end{figure}
+
+\section{Changing Parameter Values}
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{battsolarcontour.pdf}
     \caption{\textbf{Contours of latitude. Reaching higher latitudes requires better solar cells and batteries.}}
@@
  \begin{subfigmatrix}{3}% number of columns
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.eps}}
-  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.eps}}
-  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.eps}}
-  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.eps}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.pdf}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.pdf}}
+  \subfigure[35th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.pdf}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.pdf}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.pdf}}
+  \subfigure[30th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.pdf}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.80$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.pdf}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.85$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.pdf}}
+  \subfigure[25th Latitude, $p_{\text{wind}}=0.90$]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.pdf}}
  \end{subfigmatrix}
@@
 
-\subsection{Changing Physical Modeling Assumptions}
+\section{Changing Physical Modeling Assumptions}
 
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper2.pdf}
  \caption{\textbf{Comparison of simplified and detailed structural models highlights trade between wing weight and air density.}}
@@
 	\begin{center}
-	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.eps}
+	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission2.pdf}
     \caption{\textbf{Wind speed constraint moves lift to drag ratio off maximum $(C_L^{1.5}/C_D)$ point (plus signs). Solid lines are drag polars.}}
@@
 
-\subsection{Sensitivities}
+\section{Sensitivities}
 
@@
 
-% Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
-% For example, both batteries and solar cells are expensive and important to the design.  
-% The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
-% The ratio of their magnitudes is 1.76.  
-% Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
+Sensitivities can also be used to give insight into the break-even point (in terms of cost) for investing in various technologies.  
+For example, both batteries and solar cells are expensive and important to the design.  
+The sensitivity to the battery specific energy is -2.27 and the sensitivity to solar cell efficiency is -1.29 for the 25th latitude and 85th percentile winds. 
+The ratio of their magnitudes is 1.76.  
+Therefore, the break-even point for investing in these technologies occurs when a given percentage improvement in battery specific energy costs 1.76 times as much to achieve as the same percentage improvement in solar cell efficiency.
 
-\input{sens.generated.tex}
+\footnotesize
+\input{../sens.generated.tex}
 
-\newpage
-
-\input{svals.generated.tex}
-
-\input{gassens.generated.tex}
-
-\input{gvals.generated.tex}
-
-\section{Conclusion}
-
-Using geometric programming as a means of evaluation of the design space, high level intuition and understanding about design trade studies for long-endurance aircraft and their driving requirements is achieved.
-One high level discovery is the importance of wind speed.  
-If station keeping during any season or latitude is critical to the mission of a proposed aircraft, then that requirement is more likely to be met by a gas-powered architecture.
-However, if that constraint can be relaxed such that the aircraft only needs to station keep for 80\% of the time, then for certain latitudes, a solar powered aircraft could achieve greater endurance.
-An unsurprising discovery is the sensitivity of the solar-powered aircraft to the battery specific energy and solar cell efficiency.
-Using higher energy density batteries can result in significant weight and performance savings.
-The rapid solve time of geometric programming is able to present these trade studies and give intuition into design trades by quantifying the disparity in performance between the two options.
-
-% produces the bibliography section when processed by BibTeX
-
-% \section*{Appendix A}
-% 
-% \subsection{Altitude and relevant atmosphereic quantities}
-% 
-% The altitude also determines atmospheric properties such as air density and viscosity.
-For the gas powered aircraft air density is calculated prior to solving the optimization using the altitude requirement and atmospheric equations specific to the troposphere,\cite{isaatm}
-% 
-% \begin{align}
-%     \label{e:Talt}
-%     T_{\text{atm}} = T_{\text{sl}} - Lh \\
-%     \label{e:rhot}
-%     \rho = \frac{T_{\text{sl}}T_{\text{atm}}^{\left( \frac{g}{R_{\text{spec}}L} -1 \right)}}{R T_{\text{sl}}^{\frac{g}{R_{\text{spec}}L}}}.
-% \end{align}
-% 
-% Air viscosity is also computed prior to solving and is assumed to vary according to Sutherland's law,\cite{fluiddyhandbook}
-% 
-% \begin{equation}
-%     \label{e:sutherland}
-%     \mu = \mu_0 \frac{T_0 + C}{T+C} \left( \frac{T}{T_0} \right)^{\frac{3}{2}}
-% \end{equation}
-% 
-% where $\mu_0 = 1.827\text{e}-5$ [Ns/m$^2$], $T_0 = 291.15$ [K], and $C = 120$ [K]. \\
-% 
-% For the solar aircraft altitude is calculated based off of the optimized air density using the relations,\cite{isaatm} 
-% 
-% \begin{align}
-%     \label{e:tropopress}
-%     p &= \rho R_{\text{spec}}T_{11} \\
-%     \label{e:tropoalt}
-%     h &= 11,000 \text{ [m] } + \frac{R_{\text{spec}}T_{11}}{g}\frac{p}{p_{11}}.
-% \end{align}
-% 
-% According to Sutherland's law\cite{fluiddyhandbook}, viscosity varies with temperature.
-Because temperature is assumed constant 11,000-20,000 meters\cite{isaatm}, viscosity is also assumed to be constant, $\mu=\num{1.42d-5}$.
-\\
-
-\section{Appendix}
-
-\subsection{Discussion on the use of the \emph{JH01} airfoil}
-
-The \emph{sd7032} airfoil was redesigned to prevent drag creep by weakening the pressure spike associated with premature separation at higher Reynolds numbers.  
-Figure~\ref{f:jhcps} shows the pressure distributions generated in XFOIL of the \emph{JH01} airfoil at $C_L=0.0$ and $C_L=1.35$ with $Re=\num{3d5}$.
-The redesigned airfoil was named \emph{JH01}. We would like to thank Mark Drela for the redesign.
-
-\begin{figure}[h!]
- \begin{subfigmatrix}{2}% number of columns
-     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmin.eps}}
-     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.45\textwidth,natwidth=792,natheight=612]{cpmax.eps}}
- \end{subfigmatrix}
- \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number $Re=\num{3d5}$.}}
- \label{f:jhcps}
-\end{figure}
-
-\subsection{Convex fitted functions}
-
-Some constraints in this optimization study are convex equations that approximate data or other functions.  
-These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
-Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
-All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,
-
-\begin{align}
-    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
-    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
-\end{align}
-
-The log-space RMS error,
-
-\begin{equation}
-\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
-\end{equation}
-
- is given throughout the paper for each equation, where $m$ is the number of data points.
-
-Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.
-In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.
-
-% The each wind speed constraint follows the following form
-% 
-% \begin{equation}
-%     \label{e:windspeedgen}
-%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
-% \end{equation}
-% 
-% Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60.
-Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.
-% 
-% \tiny
-% \begin{longtable}{lccccccccccccc} \\
-% \label{t:windvals} \\
-%     \caption{Wind Speed Function Coefficients and Exponents} \\
-%     \toprule
-%     \toprule
-%     Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
-%     \midrule
-% 20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
-% 21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
-% 22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
-% 23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
-% 24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
-% 25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
-% 26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
-% 27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
-% 28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
-% 29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
-% 30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
-% 31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
-% 32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
-% 33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
-% 34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
-% 35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\
-% 36 & 456 & 5.77 & 45 & 3.4e-15 & -9.35 & 12.8 & 834 & 6.79 & 8.98 & 1.23e-09 & -5.52 & 58.8 & 6.92\\
-% 37 & 5.15e-14 & -8.63 & 12.1 & 225 & 5.94 & 8.34 & 8.45e-09 & -5.03 & 57.5 & 196 & 5.19 & 43.2 & 6.52\\
-% 38 & 84.8 & 4.63 & 41.6 & 4.69e-08 & -4.61 & 55.3 & 67.8 & 5.13 & 7.68 & 8.55e-13 & -7.87 & 11.2 & 6.1\\
-% 39 & 1.65e-11 & -7.05 & 10.1 & 35.7 & 4.06 & 40.1 & 22.4 & 4.33 & 6.95 & 2.55e-07 & -4.18 & 51.7 & 5.6\\
-% 40 & 16.7 & 3.57 & 38.4 & 2.62e-10 & -6.28 & 9.05 & 1.32e-06 & -3.75 & 48 & 8.73 & 3.63 & 6.25 & 5.11\\
-% 41 & 3.42e-09 & -5.56 & 7.84 & 3.93 & 2.99 & 5.53 & 8.86e-06 & -3.19 & 43.4 & 11.2 & 3.27 & 37.3 & 4.6\\
-% 42 & 2.06 & 2.42 & 4.8 & 4.04e-08 & -4.87 & 6.63 & 10.3 & 3.16 & 35.8 & 6.03e-05 & -2.61 & 39.2 & 4.07\\
-% 43 & 13.8 & 3.32 & 34.9 & 0.000428 & -2.01 & 36.7 & 1.27 & 1.93 & 4.09 & 4.21e-07 & -4.22 & 5.58 & 3.55\\
-% 44 & 0.929 & 1.54 & 3.39 & 19.9 & 3.54 & 34.2 & 3.77e-06 & -3.6 & 4.62 & 0.00191 & -1.54 & 34.3 & 3.03\\
-% 45 & 0.0061 & -1.16 & 32.6 & 2.75e-05 & -3.04 & 3.83 & 0.783 & 1.24 & 2.76 & 27.5 & 3.75 & 34.2 & 2.55\\
-% 46 & 35.2 & 3.97 & 32.9 & 0.000197 & -2.47 & 3.02 & 0.758 & 1.01 & 2.13 & 0.0176 & -0.791 & 30.5 & 2.04\\
-% 47 & 0.00143 & -1.89 & 2.2 & 42.2 & 4.28 & 28.5 & 0.816 & 0.811 & 1.47 & 0.046 & -0.43 & 28.9 & 1.49\\
-% 48 & 0.926 & 0.622 & 0.814 & 0.0108 & -1.29 & 1.34 & 0.106 & -0.0538 & 28.2 & 57.2 & 4.99 & 20 & 0.895\\
-% 49 & 56.5 & 5.67 & 11.3 & 0.128 & 0.183 & 27.4 & 0.95 & 0.455 & 0.349 & 0.0533 & -0.793 & 0.695 & 0.445\\
-% 50 & 0.869 & 0.355 & 0.148 & 24.1 & 5.47 & 7.62 & 0.0838 & 0.217 & 27.1 & 0.139 & -0.487 & 0.358 & 0.229\\
-% 51 & 4.88 & 4.99 & 5.72 & 0.657 & 0.238 & 0.0196 & 0.0293 & 0.214 & 27.7 & 0.345 & -0.21 & 0.116 & 0.0703\\
-% 52 & 0.00842 & 0.166 & 28 & 0.834 & 4.31 & 4.86 & 0.615 & -0.0754 & 0.0355 & 0.383 & 0.186 & -0.0168 & 0.0212\\
-% 53 & 0.307 & 0.168 & -0.0266 & 0.335 & 3.6 & 4.1 & 0.687 & -0.0538 & 0.0264 & 0.00564 & 0.12 & 27.7 & 0.0155\\
-% 54 & 0.00497 & 0.0916 & 27.5 & 0.223 & 3.09 & 3.42 & 0.54 & -0.067 & 0.0346 & 0.449 & 0.0992 & -0.0245 & 0.0144\\
-% 55 & 0.429 & -0.0809 & 0.0543 & 0.00527 & 0.0771 & 27.7 & 0.206 & 2.77 & 3.25 & 0.553 & 0.0722 & -0.0305 & 0.0158\\
-% 56 & 0.134 & 2.62 & 3.28 & 0.628 & 0.0434 & -0.0319 & 0.00348 & 0.0804 & 28 & 0.358 & -0.072 & 0.0644 & 0.0106\\
-% 57 & 0.00421 & 0.0943 & 28 & 0.0734 & -0.168 & 0.207 & 0.905 & 0.0123 & -0.013 & 0.155 & 2.47 & 3.06 & 0.0128\\
-% 58 & 0.00709 & 0.107 & 28.8 & 0.0366 & -0.258 & 0.474 & 0.253 & 2.33 & 2.98 & 0.921 & 0.00373 & -0.013 & 0.0225\\
-% 59 & 0.0242 & 0.892 & 28.4 & 0.000155 & -1.21 & 16.1 & 0.945 & -0.0147 & 0.0161 & 0.188 & 1.98 & 1.32 & 0.0233\\
-% 60 & 0.0173 & 0.691 & 28.1 & 0.948 & -0.0136 & 0.015 & 0.209 & 2.15 & 1.6 & 7.54e-05 & -1.41 & 14.3 & 0.0222 \\
-% \bottomrule
-% \end{longtable}
-% \normalsize
-
-% \section*{Appendix C}
-% 
-% To derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass
-% 
-% \begin{align}
-%     \label{e:mcenter}
-%     M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
-%     \label{e:eq6}
-%     \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
-% \end{align}
-% 
-% It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible. 
-% 
-% Using moment and lift approximations
-% 
-% \begin{align}
-%     C_{m_{\text{w}}} &= constant \\
-%     C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
-%     \label{e:eq10}
-%     C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
-%     m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
-%     m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
-% \end{align}
-% 
-% where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate
-% 
-% \begin{equation}
-%     \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
-% \end{equation}
-% 
-% so that \eqref{e:eq10} can be written as
-% 
-% \begin{equation}
-%     C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})
-% \end{equation}
-% 
-% This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.
-% 
-% \begin{align}
-%     C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
-%     \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
-% \end{align}
-% 
-% Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.
-% 
-% \begin{align}
-%     C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
-%     \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
-% \end{align}
-% 
-% Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin
-% 
-% \begin{equation}
-%     -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
-% \end{equation}
-% 
-% The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.
-% 
-% \begin{equation}
-%     \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
-% \end{equation}
-% 
-% Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$
-% 
-% \begin{equation}
-%     0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
-% \end{equation}
-% 
-% The pitch authority requirement is that at the forward-most c.g.
-position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$.
-Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation
-% 
-% \begin{align}
-%     \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
-% \end{align}
-% 
-% where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.
-
-\subsection{Solar energy calculations}
-
-The total available solar energy per day is an integral of the available solar power during the course of the day,
-
-    \begin{equation}
-        \label{e:solares}
-        (E/S)_{\text{sun}} = \int_{t_{\text{sun rise}}}^{t_{\text{sun set}}} (P/S)_{\text{sun}} dt.
-    \end{equation}
-
-For simplicity, energy and power will refer to energy and power per unit area throughout the rest of this section (i.e. $(P/S) = P$). 
-The solar irradiated power $P_{\text{sun}}$, is a function of $\theta$, the angle between the normal to the flat surface, or aircraft wing, and the sun beam.\cite{solar}
-
-\begin{equation}
-    \label{e:solarp}
-    P_{\text{sun}} = P_0 \cos{\theta}
-\end{equation}
-
-The angle $\theta$, depends on the time of day $t$, latitude $\phi$, and declination angle $\Delta$,\cite{solar}
-
-    \begin{equation}
-        \label{e:solartheta}
-        \cos{\theta} = \sin{\Delta} \sin{\phi} + \cos{\Delta} \cos{\phi} \cos{2\pi t/24}.
-    \end{equation}
-
- The declination angle $\Delta$, can be found using the relation\cite{solar}
-
-    \begin{align}
-        \label{e:solardelta}
-        \Delta = &0.006918 - 0.399912 \cos{\beta} + 0.070257\sin{\beta} - 0.006758\cos{2\beta} + \nonumber \\
-        & 0.000907\sin{2\beta} - 0.002697\cos{3\beta} + 0.00148\sin{3\beta},
-    \end{align}
-
-    where $\beta = 2\pi (\text{DOY}-1)/365$.
-    The time of day $t_{\text{day}}$, and the time of night $t_{\text{night}}$, can be calculated using a derivation of Equation~\eqref{e:solartheta}, \cite{solar}
-
-    \begin{align}
-        \label{e:solartday}
-        \cos{(\pi t_{\text{sun rise}}/12)} &= -\tan{\Delta} \tan{\phi} \\
-        \label{e:solarsunrise}
-        t_{\text{sun rise}} &= -t_{\text{sun set}} \\
-        \label{e:solartday2}
-        t_{\text{day}} &= 2t_{\text{sun rise}} \\
-        \label{e:solartnight}
-        t_{\text{night}} &= 24 - t_{\text{day}}
-    \end{align}
-
-    where noon is $t=0$.
-Both $t_{\text{day}}$ and $t_{\text{night}}$ affect the battery size as defined in Equations~\eqref{e:solarreq} and~\eqref{e:solarbatt}.
-The solar power available assuming no inclination angle $P_0$, is found using the eccentricity of the earth's orbit,
-
-    \begin{align}
-        \label{e:solarp0}
-        P_0 & = P_{\text{sun surface}} \frac{R_{\text{sun}}^2}{R_{\text{earth orbit}}^2}, \\
-        \label{e:solareo}
-        R_{\text{earth orbit}} & = r_0 \left[ 1 + 0.017 \sin{\left( 2\pi \frac{\text{DOY}-93}{365}\right)} \right],
-    \end{align}
-
-    where
-
-    \[ \begin{array}{lcl}
-        P_{\text{sun surface}} & : & \text{Power emitted at the sun's surface} \\
-        R_{\text{sun}} & : & \text{Radius of the sun} \\
-        R_{\text{earth orbit}} & : & \text{Distance from earth to sun} \\
-        r_0 & : & \text{Average distance from earth to sun} \\
-        0.017 & : & \text{Eccentricity of earth-sun orbit}.
-    \end{array} \]
-
-    Using a trapezoidal integration of Equation~\eqref{e:solares}, the total available solar energy per unit area can be obtained for a given latitude and day of the year.
-Because the equations in this appendix are not GP compatible, the total solar energy per unit area per day, the length of the day and the length of the night are calculated from the latitude and the day of the year prior to an optimization solve.
-
-\bibliography{biblibrary}
-\bibliographystyle{aiaa}
-
-\end{document}
-
-% - Release $Name:  $ -
+\input{../gassens.generated.tex}

### 2017-05-19 docs/thesis/cover.tex
add thesis folder

% -*-latex-*-
% 
% For questions, comments, concerns or complaints:
% thesis@mit.edu
% 
%
% $Log: cover.tex,v $
% Revision 1.8  2008/05/13 15:02:15  jdreed
% Degree month is June, not May.  Added note about prevdegrees.
% Arthur Smith's title updated
%
% Revision 1.7  2001/02/08 18:53:16  boojum
% changed some \newpages to \cleardoublepages
%
% Revision 1.6  1999/10/21 14:49:31  boojum
% changed comment referring to documentstyle
%
% Revision 1.5  1999/10/21 14:39:04  boojum
% *** empty log message ***
%
% Revision 1.4  1997/04/18  17:54:10  othomas
% added page numbers on abstract and cover, and made 1 abstract
% page the default rather than 2.  (anne hunter tells me this
% is the new institute standard.)
%
% Revision 1.4  1997/04/18  17:54:10  othomas
% added page numbers on abstract and cover, and made 1 abstract
% page the default rather than 2.  (anne hunter tells me this
% is the new institute standard.)
%
% Revision 1.3  93/05/17  17:06:29  starflt
% Added acknowledgements section (suggested by tompalka)
% 
% Revision 1.2  92/04/22  13:13:13  epeisach
% Fixes for 1991 course 6 requirements
% Phrase "and to grant others the right to do so" has been added to 
% permission clause
% Second copy of abstract is not counted as separate pages so numbering works
% out
% 
% Revision 1.1  92/04/22  13:08:20  epeisach

% NOTE:
% These templates make an effort to conform to the MIT Thesis specifications,
% however the specifications can change.  We recommend that you verify the
% layout of your title page with your thesis advisor and/or the MIT 
% Libraries before printing your final copy.
\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}

\author{Michael Burton}
% If you wish to list your previous degrees on the cover page, use the 
% previous degrees command:
%       \prevdegrees{A.A., Harvard University (1985)}
% You can use the \\ command to list multiple previous degrees
%       \prevdegrees{B.S., University of California (1978) \\
%                    S.M., Massachusetts Institute of Technology (1981)}
\prevdegrees{B.S., Massachusetts Institute of Technology (2016)}
\department{Department of Aeronautics and Astronautics}

% If the thesis is for two degrees simultaneously, list them both
% separated by \and like this:
% \degree{Doctor of Philosophy \and Master of Science}
\degree{Master of Science in Aeronautics and Astronautics}

% As of the 2007-08 academic year, valid degree months are September, 
% February, or June.  The default is June.
\degreemonth{June}
\degreeyear{2017}
\thesisdate{May 25, 2017}

%% By default, the thesis will be copyrighted to MIT.  If you need to copyright
%% the thesis to yourself, just specify the `vi' documentclass option.  If for
%% some reason you want to exactly specify the copyright notice text, you can
%% use the \copyrightnoticetext command.  
%\copyrightnoticetext{\copyright IBM, 1990.  Do not open till Xmas.}

% If there is more than one supervisor, use the \supervisor command
% once for each.
\supervisor{Warren Hoburg}{Associate Professor}

% This is the department committee chairman, not the thesis committee
% chairman.  You should replace this with your Department's Committee
% Chairman.
\chairman{Youssef M. Marzouk}{Chairman, Department Committee on Graduate Theses}

% Make the titlepage based on the above information.  If you need
% something special and can't use the standard form, you can specify
% the exact text of the titlepage yourself.  Put it in a titlepage
% environment and leave blank lines where you want vertical space.
% The spaces will be adjusted to fill the entire page.  The dotted
% lines for the signatures are made with the \signature command.
\maketitle

% The abstractpage environment sets up everything on the page except
% the text itself.  The title and other header material are put at the
% top of the page, and the supervisors are listed at the bottom.  A
% new page is begun both before and after.  Of course, an abstract may
% be more than one page itself.  If you need more control over the
% format of the page, you can use the abstract environment, which puts
% the word "Abstract" at the beginning and single spaces its text.

%% You can either \input (*not* \include) your abstract file, or you can put
%% the text of the abstract directly between the \begin{abstractpage} and
%% \end{abstractpage} commands.

% First copy: start a new page, and save the page number.
\cleardoublepage
% Uncomment the next line if you do NOT want a page number on your
% abstract and acknowledgments pages.
% \pagestyle{empty}
\setcounter{savepage}{\thepage}
\begin{abstractpage}
\input{abstract}
\end{abstractpage}

% Additional copy: start a new page, and reset the page number.  This way,
% the second copy of the abstract is not counted as separate pages.
% Uncomment the next 6 lines if you need two copies of the abstract
% page.
% \setcounter{page}{\thesavepage}
% \begin{abstractpage}
% \input{abstract}
% \end{abstractpage}

\cleardoublepage

\section*{Acknowledgments}

This is the acknowledgements section.  You should replace this with your
own acknowledgements.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% -*-latex-*-

### 2017-05-21 docs/thesis/app_tbflex.tex
modify gas to include sp bool; print tail boom table for gas; create appendix and tbflex section in thesis

\chapter{Tail Boom Flexibility Forumlation}

I would like to thank Mark Drela for the formulation of the tail boom flexibility model.
The derivation of Equation~\eqref{e:smcorr} begins with the moment about the aircraft's center of mass

\begin{align}
    \label{e:mcenter}
    M_{cg} &= M_{\text{w}} + (x_{cg} - x_{ac})L_{\text{w}} - l_{\text{h}} L_{\text{h}} \\
    \label{e:eq6}
    \frac{M_{cg}}{qSc} = C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} C_{L_W} - V_{\text{w}} C_{L_{\text{h}}}
\end{align}

It is assumed that the tail boom's effective root location is at the wing's aerodynamic center $x_{ac}$, and that the tail's pitching moment about its own aerodynamic center is negligible.

Using moment and lift approximations

\begin{align}
    C_{m_{\text{w}}} &= constant \\
    C_{L_W} &= C_{L_{W_0}} + m_{\text{w}} \alpha \\
    \label{e:eq10}
    C_{L_{\text{h}}} &= C_{L_{h_0}} + m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \theta \right] + C_{L_{h_{\delta}}}\delta_e \\
    m_{\text{w}} &= \frac{2\pi}{1 + 2/AR_{\text{w}}} \\
    m_{\text{h}} &= \frac{2\pi}{1 + 2/AR_{\text{h}}}
\end{align}

where $\epsilon$ is the wing's downwash angle seen at the tail. Using a vortex approximation and neglecting taper effects, we can estimate

\begin{equation}
    \frac{d\epsilon}{d\alpha} = \frac{m_{\text{w}}}{4\pi} \frac{c}{l_{\text{h}}}
\end{equation}

so that \eqref{e:eq10} can be written as

\begin{equation}
    C_{L_{\text{h}}} = m_{\text{h}} \left[\left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) C_{L_{\text{h}}}\right] + C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) 
\end{equation}

This can be further simplified by defining a tail boom flexibility factor $\mathcal{F}$.

\begin{align}
    C_{L_{\text{h}}} &= \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0}) \\
    \mathcal{F} &= 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
\end{align}

Using the wing lift coefficient and the recast tail lift coefficient, the pitching moment is equation and derivative are given as follows.

\begin{align}
    C_m &= C_{m_{\text{w}}} + \frac{x_{cg} - x_{ac}}{c} (C_{L_{W_0}} + m_{\text{w}} \alpha) - V_{\text{w}} \left[ \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) \alpha + \mathcal{F}^{-1} C_{L_{h_{\delta}}}(\delta_e - \delta_{e_0})\right] \\
    \frac{dC_m}{d\alpha} & = \frac{x_{cg} - x_{ac}}{c} m_{\text{w}}  - V_{\text{w}} \mathcal{F}^{-1} m_{\text{h}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) 
\end{align}

Now by dividing by $dC_{L_W}/d\alpha = m_{\text{h}}$, which defines the static margin

\begin{equation}
    -\frac{dC_m/d\alpha}{dC_{L_W}/d\alpha} = \text{SM} = V_{\text{w}} \mathcal{F}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{x_{cg} - x_{ac}}{c}
\end{equation}

The case for meeting the minimum static margin requirement is at the never-exceed dynamic pressure $q_{NE}$ and when the c.g is at its aft-most position.

\begin{equation}
    \text{SM}_{\text{min}} = V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) - \frac{(x_{cg})_{\text{aft}} - x_{ac}}{c} 
\end{equation}

Dividing Equation~\eqref{e:eq6} by $C_{L_W}$, gives a requirement on the tail lift coefficient required to achieve a pitch trim condition $C_m=1$

\begin{equation}
    0 = \frac{C_{m_{\text{w}}}}{C_{L_W}} + \frac{x_{cg} - x_{ac}}{c} - V_{\text{w}} \frac{C_{L_{\text{h}}}}{C_{L_W}}
\end{equation}

The pitch authority requirement is that at the forward-most c.g. position and maximum lift with the tail lift coefficient equal to the most-negative allowable value $(C_{L_{\text{h}}})_{\text{min}}$. Combining the static margin and pitch authority requirement results in the horizontal tail sizing equation

\begin{align}
    \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{\text{w}}}}{C_{L_{\text{max}}}} &\leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
\end{align}

where $\Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}}$.

### 2017-05-21 docs/thesis/chap_results.tex
modify gas to include sp bool; print tail boom table for gas; create appendix and tbflex section in thesis

@@
 
+\subsection{Representing Structural Weight as Fraction of Total}
+
 Insight into the design space can also be gained by changing the physical modeling assumptions.
@@
 
+\subsection{Lift to Drag Ratio Shift Due to Wind Speed Constraint}
+
 Another interesting result is the operating lift to drag ratio for the gas powered aircraft.  
@@
 If it is assumed that wind speeds are negligible or that station-keeping is not important then velocity will be optimized such that the lift to drag ratio is at the maximum $C_L^{1.5}/C_D$ as shown in Figure~\ref{f:polarmission}.
+
+\subsection{Tail Boom Flexibility}
+
+By changing using higher fidelity physical models, more accurate results of the sizing analysis can be achieved.  
+For example, sizing the horizontal tail using a volume coefficient does not account for the fexibility of the tail boom. 
+A flexible tail boom will cause a slight delay in the responsiveness to pitch control inputs, effectively lowering the control authority of the horizontal tail. 
+However, using a minimum static margin $\text{SM}_{\text{min}} = 0.35$ and minimum desirable c.g. travel range,
+
+\begin{equation}
+    \label{e:deltacg}
+    \Delta x_{cg} = (x_{cg})_{\text{aft}} - (x_{cg})_{\text{fwd}} = 0.2,
+\end{equation}
+
+a single constraint, whose derivation is explained in Appendix~\ref{app:tbflex}, minimizes the horizontal tail volume coefficient $V_{\text{w}}$, while meeting the minimum static margin and minimum c.g.
+travel range,
+
+\begin{equation}
+    \label{e:smcorr}
+    \text{SM}_{\text{min}} + \frac{\Delta x_{cg}}{c} - \frac{C_{M_{	ext{w}}}}{C_{L_{\text{max}}}} \leq V_{\text{w}} \mathcal{F}_{NE}^{-1} \frac{m_{\text{h}}}{m_{\text{w}}} \left( 1 - \frac{d\epsilon}{d\alpha}\right) + V_{\text{w}} \frac{-(C_{L_{\text{h}}})_{\text{min}}}{C_{L_{\text{max}}}} \\
+\end{equation}
+
+Because this equation is not GP-compatible a signomial program is used to solve the optimization model for both the solar-electric and gas powered models.\cite{sp}
+The flexibility factor,
+
+\begin{equation}
+     \mathcal{F} = 1 + m_{\text{h}} \frac{qS_{\text{h}}l_{\text{h}}^2}{EI_0}(1-\frac{1}{2}k) 
+\end{equation}
+
+is a measure of how the pitch control authority is degraded by the flexibility of the tail boom. 
+Thus, it is expected that the horizontal tail volume coefficient will increase from an increased horizontal tail surface area and tail boom length. 
+Tables~\ref{t:tbftable} and~\ref{t:tbfgtable} confirm the increased size of the horizontal tail and tail boom when the tail boom flexibility model is activiated.
+
+\input{../tbftable.generated.tex}
+\input{../tbfgtable.generated.tex}
 

### 2017-05-21 gassolar/gas/disp_sens.py

@@ -21,19 +21,64 @@ if __name__ == "__main__":
         sol = M.solve("mosek")
         sols.append(sol)
 
-    varnames = ["V_{wind}_Mission/Loiter/FlightSegment", "W_{pay}", "\\eta_{prop}", "BSFC_{min}", "t_Mission/Loiter", "N_{max}_Mission/AircraftLoading/WingLoading/ChordSparL"]
-    latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$", "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$", "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
+    varnames = ["V_{wind}_Mission/Loiter/FlightSegment", "W_{pay}",
+                "\\eta_{prop}", "BSFC_{min}", "t_Mission/Loiter",
+                "N_{max}_Mission/AircraftLoading/WingLoading/ChordSparL"]
+    latns = ["$V_{\\mathrm{wind}}$", "$W_{\\mathrm{pay}}$",
+             "$\\eta_{\\mathrm{prop}}$", "$BSFC_{\\mathrm{min}}$",
+             "$t_{\\mathrm{loiter}}$", "$N_{\\mathrm{max}}$"]
     fig, ax = plot_sens(M, sols[2], varnames)
-    
-    dvarns = ["MTOW", "b_Mission/Aircraft/Wing", "AR_Mission/Aircraft/Wing", "W_{fuel-tot}", "W_Mission/Aircraft/Wing", "W_Mission/Aircraft/Engine", "BSFC", "C_L", "C_D"]
-    dlatns = ["MTOW", "$b$", "$A$", "$W_{\\mathrm{fuel}}$", "$W_{\\mathrm{wing}}$", "$W_{\\mathrm{engine}}$", "BSFC", "$C_L$", "$C_D$"]
+
+    dvarns = ["MTOW", "b_Mission/Aircraft/Wing", "AR_Mission/Aircraft/Wing",
+              "W_{fuel-tot}", "W_Mission/Aircraft/Wing",
+              "W_Mission/Aircraft/Engine", "BSFC", "C_L", "C_D"]
+    dlatns = ["MTOW", "$b$", "$A$", "$W_{\\mathrm{fuel}}$",
+              "$W_{\\mathrm{wing}}$", "$W_{\\mathrm{engine}}$", "BSFC", "$C_L$",
[…]

### 2017-05-21 gassolar/gas/gas.py

@@ -1,7 +1,8 @@
 """Jungle Hawk Owl"""
 import numpy as np
 from gpkitmodels.GP.aircraft.engine.gas_engine import Engine
-from gpkitmodels.GP.aircraft.wing.wing import Wing
+from gpkitmodels.GP.aircraft.wing.wing import Wing as WingGP
+from gpkitmodels.SP.aircraft.wing.wing import Wing as WingSP
 from gpkitmodels.GP.aircraft.fuselage.elliptical_fuselage import Fuselage
 from gpkitmodels.GP.aircraft.tail.empennage import Empennage
 from gpkitmodels.GP.aircraft.tail.tail_boom import TailBoomState
@@ -15,10 +16,15 @@ from loiter import Loiter
 
 class Aircraft(Model):
     "the JHO vehicle"
-    def setup(self, Wfueltot):
+    def setup(self, Wfueltot, sp=False):
+
+        self.sp = sp
 
         self.fuselage = Fuselage(Wfueltot)
-        self.wing = Wing()
+        if sp:
+            self.wing = WingSP()
+        else:
+            self.wing = WingGP()
         self.engine = Engine()
         self.empennage = Empennage()
 
@@ -31,9 +37,13 @@ class Aircraft(Model):
         Wwing = Variable("W_{wing}", "lbf", "wing weight for loading")
         etaprop = Variable("\\eta_{prop}", 0.8, "-", "propulsive efficiency")
 
-        self.empennage.substitutions["V_h"] = 0.45
         self.empennage.substitutions["V_v"] = 0.04
-        self.empennage.substitutions["m_h"] = 0.4
+
+        if not sp:
+            self.empennage.substitutions["V_h"] = 0.45
+            self.empennage.substitutions["AR_h"] = 5
+            self.empennage.substitutions["m_h"] = 0.1
+
         constraints = [
             Wzfw >= sum(summing_vars(components, "W")) + Wpay + Wavn,
             Wwing >= sum(summing_vars([self.wing], "W")),
@@ -45,10 +55,6 @@ class Aircraft(Model):
                 self.empennage.verticaltail["S"]
                 * self.empennage.verticaltail["l_v"]/self.wing["S"]
                 / self.wing["b"]),
-            # self.wing["C_{L_{max}}"]/self.wing["m_w"] <= (
-            #     self.empennage.horizontaltail["C_{L_{max}}"]
-            #     / self.empennage.horizontaltail["m_h"])
-            self.empennage.horizontaltail["C_{L_{max}}"] == 1.5,
[…]

### 2017-05-21 gassolar/solar/print_sens.py

@@ -23,10 +23,16 @@ def sol_table(sols, models, varnames, filename, ttype, latns=[],
                     "Percentile Winds & 90th Percentile Winds \\\\\n")
         elif ttype == "tbflex":
             f.write("\\multirow{2}{*}{Variable} & "
-                    "\\multicol{2}{c}{Without Tail Boom Flex} & "
-                    "\\multicol{2}{c}{With Tail Boom Flex} \\\\\n")
+                    "\\multicolumn{2}{c}{Without Tail Boom Flex} & "
+                    "\\multicolumn{2}{c}{With Tail Boom Flex} \\\\\n")
             f.write("& 25$^{\circ}$ Latitude & 30$^{\circ}$ Latitude & "
                     "25$^{\circ}$ Latitude & 30$^{\circ}$ Latitude \\\\\n")
+        elif ttype == "tbflexg":
+            f.write("\\multirow{2}{*}{Variable} & "
+                    "\\multicolumn{2}{c}{Without Tail Boom Flex} & "
+                    "\\multicolumn{2}{c}{With Tail Boom Flex} \\\\\n")
+            f.write("& 7 Day Endurance & 9 Day Endurance "
+                    "& 7 Day Endurance & 9 Day Endurance \\\\\n")
         elif ttype == "gas":
             f.write("Variable & 5 Day Endurance & 7 Day Endurance & 9 Day "
                     "Endurance\\\\\n")
@@ -192,14 +198,19 @@ if __name__ == "__main__":
               "$W_{\\mathrm{batt}}$", "$W_{\\mathrm{solar}}$", "$C_L$", "$C_D$",
               "$h$"]
 
-    tbfvar = ["W_{total}", "b_Mission/Aircraft/Wing", "d_0", "l_h", "V_h"]
-    tbflat = ["MTOW", "$b$", "$d_0$", "l_{\\mathrm{h}}", "V_{\\mathrm{h}}"]
+    tbfvar = ["W_{total}", "b_Mission/Aircraft/Wing",
+              "W_Mission/Aircraft/Empennage/HorizontalTail",
+              "W_Mission/Aircraft/Empennage/TailBoom", "d_0", "l_h",
+              "S_Mission/Aircraft/Empennage/HorizontalTail", "V_h"]
+    tbflat = ["MTOW", "$b$", "$W_{\\mathrm{h}}$", "$W_{\\mathrm{boom}}$",
[…]

### 2017-05-22 docs/thesis/chap_results.tex
change naming for tables; update intro conclusion and result sections; add sp reference

@@
 
-The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs}. 
-Key design parameters for both aircraft optimization models are listed in Table~\ref{t:gassolarparams}.
-The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 84.0 lbs.  
+The gas and solar-electric powered aircraft optimization models were solved by minimizing the take-off weight for the requirements listed in Table~\ref{t:mreqs} and the assumptions and key design parameters described herein.
 The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take-off weight of 194.7 lbs.
-Highlighted design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements.
-The rest of this chapter explores how these results may vary for different requirements, input parameter values, and physics modeling assumptions.
+The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take-off weight of 72.3 lbs.  
+Important design parameters are listed in Table~\ref{t:gassolarparams}.
+Key design variables for both architectures are listed in Tables~\ref{t:svals} and~\ref{t:gvals} for various mission requirements. 
+The rest of this section explores how these result may vary for different requirements, input parameter values, and physics modeling assumptions.
 
@@
 
- \begin{figure}[H]
+ \begin{figure}[h!]
  \begin{subfigmatrix}{3}% number of columns
@@
 
-Because this equation is not GP-compatible a signomial program is used to solve the optimization model for both the solar-electric and gas powered models.\cite{sp}
+Because this equation is not GP-compatible a signomial program is used to solve the optimization model for both the solar-electric and gas powered models.\cite{sp}\cite{gp}
 The flexibility factor,
@@
 Tables~\ref{t:tbftable} and~\ref{t:tbfgtable} confirm the increased size of the horizontal tail and tail boom when the tail boom flexibility model is activiated. 
+Because tail boom flexibility is strictly worse for aircraft weight because it requires more horizontal tail surface area or more tail boom length it is surprising to see that for the 30$^{\circ}$ latitude case the horizontal tail weight actually decreases on the solar aircraft.
+Adding the additional constraints to account for the tail boom flexibility allows for the horizontal tail aspect ratio to be optimized, whereas before it was a required input. 
+The high aspect ratio of the horizontal tails means low interior foam weight even though the surface area is bigger.
 

### 2017-05-22 docs/thesis/cover.tex
change naming for tables; update intro conclusion and result sections; add sp reference

@@
 %                    S.M., Massachusetts Institute of Technology (1981)}
-\prevdegrees{B.S., Massachusetts Institute of Technology (2016)}
+\prevdegrees{B.S., Massachusetts Institute of Technology, 2016}
 \department{Department of Aeronautics and Astronautics}
@@
 % once for each.
-\supervisor{Warren Hoburg}{Associate Professor}
+\supervisor{Warren Hoburg}{Associate Professor of Aeronautics and Astronautics}
 
@@
 
-This is the acknowledgements section.  You should replace this with your
-own acknowledgements.
+The success I achieved in completing this thesis was only possible because of the incredible support from family, friends, mentors, and MIT faculty.
+I was able to complete this thesis a full year ealier than I expected and I must thank the many people who made that possible.
+
+I first need to thank my family beginning with my beautiful wife, Jasmine.  
+Her support and encouragement have been unwavering.  
+When I was up late struggling with a problem, she never believed I wouldn't figure it out.  
+She has always been enthusiastic and supportive of my successes and given sound advice and encouragment when facing difficult problems or situations. 
+I could not have done this without her.
+
+My parents and sibilings were also constant sources of support and encouragement. 
+My parents are the wisest people I know and I am immensely grateful for their advice.  
+They have always pushed me to be the best me and I would not have ever made it to MIT without them.
+
+I want to greatly thank my advisor, Warren Hoburg (Woody), for his guidance, friendship, and belief in my abilities.  
+Woody is a demanding advisor, who pushed me to learn, grow, and develop at a pace and intensity that would not have been possible otherwise. 
+While rigorous in his pursuit of high quality work, Woody also gave me the flexibility to explore my own interests and taught me to think critically for myself.
+He is truly a brillant professor and I was fortunate that he selected me to join his lab this year.
+
+I would not have made it through the rigors and stress of MIT without my best friend and officemate Martin York. 
+I could not have shared an office with a more brillant, passionate and fun person. 
+Our common interests, beliefs, and mentality gave me freedom to express my views, opinions, challenges and successes without fear of judgement or offense.
+
+I also need to thank the project managers, professors, and sponsors of the Jungle Hawk Owl, John Hansmen, Navid Yazdani, Mark Drela, Warren Hoburg and Tony Tao. 
+The original problem developed by this team of professionals, of which architecture would was best for the Jungle Hawk Owl was the basis for my thesis work.
+They all acted as mentors to me, teaching me techincally, guiding me in the industry and giving me opportunites to present my work. 
+It was a privilege to work with and for these people to build the Jungle Hawk Owl.
+
+Finally, I want to thank the many fellow students I worked with over the past year and a half building and designing the Jungle Hawk Owl.  
+We went through an extremely intense and rapid design and build cycle. 
+The late nights, long, arduous layups, and successful test flights have forged a camaraderie between us that I will not forget.
 

### 2017-05-22 gassolar/gas/disp_sens.py

@@ -39,9 +39,10 @@ if __name__ == "__main__":
     tbfvar = ["MTOW", "b_Mission/Aircraft/Wing",
               "W_Mission/Aircraft/Empennage/HorizontalTail",
               "W_Mission/Aircraft/Empennage/TailBoom", "d_0", "l_h",
-              "S_Mission/Aircraft/Empennage/HorizontalTail", "V_h"]
+              "S_Mission/Aircraft/Empennage/HorizontalTail", "AR_h", "V_h"]
     tbflat = ["MTOW", "$b$", "$W_{\\mathrm{h}}$", "$W_{\\mathrm{boom}}$",
-              "$d_0$", "$l_{\\mathrm{h}}$", "$S_h$", "$V_{\\mathrm{h}}$"]
+              "$d_0$", "$l_{\\mathrm{h}}$", "$S_h$", "$A_{\\mathrm{h}}$",
+              "$V_{\\mathrm{h}}$"]
 
     tbsols = []
     tbMs = []

### 2017-05-22 gassolar/gas/gas.py

@@ -41,7 +41,7 @@ class Aircraft(Model):
 
         if not sp:
             self.empennage.substitutions["V_h"] = 0.45
-            self.empennage.substitutions["AR_h"] = 5
+            self.empennage.substitutions["AR_h"] = 5.0
             self.empennage.substitutions["m_h"] = 0.1
 
         constraints = [

### 2017-05-22 gassolar/solar/print_sens.py

@@ -69,6 +69,9 @@ def sol_table(sols, models, varnames, filename, ttype, latns=[],
                 else:
                     val = s(vnm).magnitude
                     units = " [" + unitstr(s(vnm)) + "]"
+                if "**" in units:
+                    sp = units.split("**")
+                    units = sp[0] + "$^{" + sp[-1].replace("]", "") + "}$]"
                 vals.append(val)
             row = ltnm + units + "&" + " & ".join(["%.3g" % x for x in vals])
             f.write(row + "\\\\\n")
@@ -201,9 +204,10 @@ if __name__ == "__main__":
     tbfvar = ["W_{total}", "b_Mission/Aircraft/Wing",
               "W_Mission/Aircraft/Empennage/HorizontalTail",
               "W_Mission/Aircraft/Empennage/TailBoom", "d_0", "l_h",
-              "S_Mission/Aircraft/Empennage/HorizontalTail", "V_h"]
+              "S_Mission/Aircraft/Empennage/HorizontalTail", "AR_h", "V_h"]
     tbflat = ["MTOW", "$b$", "$W_{\\mathrm{h}}$", "$W_{\\mathrm{boom}}$",
-              "$d_0$", "$l_{\\mathrm{h}}$", "$S_h$", "$V_{\\mathrm{h}}$"]
+              "$d_0$", "$l_{\\mathrm{h}}$", "$S_h$", "$A_{\\mathrm{h}}$",
+              "$V_{\\mathrm{h}}$"]
 
     tbsols = []
     tbMs = []

### 2017-05-22 gassolar/solar/solar.py

@@ -47,7 +47,7 @@ class Aircraft(Model):
 
         if not sp:
             self.empennage.substitutions["V_h"] = 0.45
-            self.empennage.substitutions["AR_h"] = 5
+            self.empennage.substitutions["AR_h"] = 5.0
             self.empennage.substitutions["m_h"] = 0.1
 
         constraints = [

### 2017-05-22 docs/thesis/cover.tex
thesis minor edits

@@
 % Chairman.
-\chairman{Youssef M. Marzouk}{Chairman, Department Committee on Graduate Theses}
+\chairman{Youssef M. Marzouk}{Associate Professor of Aeronautics and Astonautics\\Chairman, Graduate Program Committee}
 
@@
 
-The success I achieved in completing this thesis was only possible because of the incredible support from family, friends, mentors, and MIT faculty.
-I was able to complete this thesis a full year ealier than I expected and I must thank the many people who made that possible.
+\begin{singlespace}
+The success I achieved in completing this thesis a year earlier than expected was only possible because of the incredible support from family, friends, mentors, and MIT faculty.
 
-I first need to thank my family beginning with my beautiful wife, Jasmine.  
+I first need to thank my family, beginning with my beautiful wife, Jasmine.  
 Her support and encouragement have been unwavering.  
-When I was up late struggling with a problem, she never believed I wouldn't figure it out.  
-She has always been enthusiastic and supportive of my successes and given sound advice and encouragment when facing difficult problems or situations. 
+When I was up late struggling with a problem, she always believed I would figure it out.  
+She has been enthusiastic and supportive of my successes and given sound advice and encouragement when facing difficult problems or situations. 
 I could not have done this without her.
 
-My parents and sibilings were also constant sources of support and encouragement. 
+My parents and siblings were also constant sources of support and encouragement. 
 My parents are the wisest people I know and I am immensely grateful for their advice.  
-They have always pushed me to be the best me and I would not have ever made it to MIT without them.
+They have pushed me to be the best me and I would not have ever made it to MIT without them.
 
-I want to greatly thank my advisor, Warren Hoburg (Woody), for his guidance, friendship, and belief in my abilities.  
-Woody is a demanding advisor, who pushed me to learn, grow, and develop at a pace and intensity that would not have been possible otherwise. 
+I want to thank my outstanding advisor, Warren Hoburg (Woody), for his guidance, friendship, and belief in my abilities.  
+Woody is a demanding advisor who pushed me to learn, grow, and develop at a pace and intensity that I did not believe possible. 
 While rigorous in his pursuit of high quality work, Woody also gave me the flexibility to explore my own interests and taught me to think critically for myself.
-He is truly a brillant professor and I was fortunate that he selected me to join his lab this year.
+He is truly a brilliant professor and a wise mentor and I was fortunate that he selected me to join his lab this year.
 
 I would not have made it through the rigors and stress of MIT without my best friend and officemate Martin York. 
-I could not have shared an office with a more brillant, passionate and fun person. 
-Our common interests, beliefs, and mentality gave me freedom to express my views, opinions, challenges and successes without fear of judgement or offense.
+I could not have shared an office with a more brilliant, passionate and fun person. 
+Our common interests, beliefs, and convictions gave me freedom to express my views, opinions, challenges and successes without fear of judgement or offense.
 
@@
 The original problem developed by this team of professionals, of which architecture would was best for the Jungle Hawk Owl was the basis for my thesis work.
-They all acted as mentors to me, teaching me techincally, guiding me in the industry and giving me opportunites to present my work. 
+They all acted as mentors to me, teaching me technically, guiding me in the industry and giving me opportunities to present my work. 
 It was a privilege to work with and for these people to build the Jungle Hawk Owl.
@@
 The late nights, long, arduous layups, and successful test flights have forged a camaraderie between us that I will not forget. 
+\end{singlespace}
 

### 2017-05-22 gassolar/gas/disp_sens.py

@@ -47,16 +47,15 @@ if __name__ == "__main__":
     tbsols = []
     tbMs = []
     for sp in [False, True]:
-        for t in [7, 9]:
-            M = Mission(sp=sp)
-            M.cost = M["MTOW"]
-            M.substitutions.update({"t_Mission/Loiter": t})
-            tbMs.append(M)
-            if sp:
-                sol = M.localsolve("mosek")
-            else:
-                sol = M.solve("mosek")
-            tbsols.append(sol)
+        M = Mission(sp=sp)
+        M.cost = M["MTOW"]
+        M.substitutions.update({"t_Mission/Loiter": 9})
+        tbMs.append(M)
+        if sp:
+            sol = M.localsolve("mosek")
+        else:
+            sol = M.solve("mosek")
+        tbsols.append(sol)
 
     if len(sys.argv) > 1:
         path = sys.argv[1]

### 2017-05-22 gassolar/solar/print_sens.py

@@ -19,8 +19,9 @@ def sol_table(sols, models, varnames, filename, ttype, latns=[],
             f.write("\\multirow{2}{*}{Variable} & 25$^{\circ}$ Latitude & "
                     "30$^{\circ}$ Latitude & 25$^{\circ}$ Latitude & "
                     "30$^{\circ}$ Latitude \\\\\n")
-            f.write("& 85th Percentile Winds & 85th Percentile Winds & 90th "
-                    "Percentile Winds & 90th Percentile Winds \\\\\n")
+            f.write("& $p_{\\mathrm{wind}}=0.85$& $p_{\\mathrm{wind}}=0.85$ & "
+                    "$p_{\\mathrm{wind}}=0.90$ & "
+                    "$p_{\\mathrm{wind}}=0.90$ \\\\\n")
         elif ttype == "tbflex":
             f.write("\\multirow{2}{*}{Variable} & "
                     "\\multicolumn{2}{c}{Without Tail Boom Flex} & "
@@ -28,11 +29,9 @@ def sol_table(sols, models, varnames, filename, ttype, latns=[],
             f.write("& 25$^{\circ}$ Latitude & 30$^{\circ}$ Latitude & "
                     "25$^{\circ}$ Latitude & 30$^{\circ}$ Latitude \\\\\n")
         elif ttype == "tbflexg":
-            f.write("\\multirow{2}{*}{Variable} & "
-                    "\\multicolumn{2}{c}{Without Tail Boom Flex} & "
-                    "\\multicolumn{2}{c}{With Tail Boom Flex} \\\\\n")
-            f.write("& 7 Day Endurance & 9 Day Endurance "
-                    "& 7 Day Endurance & 9 Day Endurance \\\\\n")
+            f.write("\\multirow{2}{*}{Variable} & Without Tail Boom Flex & "
+                    "With Tail Boom Flex \\\\\n")
+            f.write("& 9 Day Endurance & 9 Day Endurance \\\\\n")
         elif ttype == "gas":
             f.write("Variable & 5 Day Endurance & 7 Day Endurance & 9 Day "
                     "Endurance\\\\\n")
@@ -65,7 +64,7 @@ def sol_table(sols, models, varnames, filename, ttype, latns=[],
                         units = ""
                     if vnm == "\\rho":
                         val = altitude(val)*0.3048
-                        units = "[m]"
+                        units = " [m]"
                 else:
                     val = s(vnm).magnitude
                     units = " [" + unitstr(s(vnm)) + "]"

### 2017-05-24 gassolar/environment/wind_fit.py
fits for all lats all months

@@ -8,7 +8,6 @@ from gpfit.fit import fit
 plt.rc("text", usetex=True)
 plt.rcParams.update({'font.size':15})
 
-GENERATE = False
 PERCT_NORM = 100.0
 WIND_NORM = 100.0
 RHO_NORM = 1.0
@@ -58,7 +57,7 @@ def fit_setup(altitude=(40000, 80000), latitude=45, day=355):
 
     return x, y
 
-def plot_fits(xdata, ydata, yfit, latitude):
+def plot_fits(xdata, ydata, yfit, latitude, plot=False):
 
     x1 = np.flipud(np.unique(xdata[0]))
     x2 = np.unique(xdata[1])
@@ -81,7 +80,7 @@ def plot_fits(xdata, ydata, yfit, latitude):
 
     ax.fill_between(np.exp(x1), yfits[0], yfits[-1], alpha=0.2,
                     facecolor="#3E31AE", edgecolor="None")
-    if not GENERATE:
+    if plot:
         for i, p in enumerate(["75\%", "85\%", "95\%"]):
             ax.text(np.exp(x1)[0]+0.005, yfits[i][0]-1.0, p)
     ax.set_xlabel("Air Density [kg/m$^3$]")
@@ -90,14 +89,14 @@ def plot_fits(xdata, ydata, yfit, latitude):
     ax.grid()
     return fig, ax
 
-def make_fits(day, latrange, month, gen=False, path=""):
+def make_fits(day, latrange, month, gen=False, path="", plot=False):
 
     for l in latrange:
         print "Fitting for %d latitude" % l
-        altitudestart = range(40000, 50500, 500)
+        altitudestart = range(43000, 50500, 500)
         for j, a in enumerate(altitudestart):
-            print "Trying Altitude Range: %d-80000" % a
-            X, Y = fit_setup(altitude=(a, 80000), latitude=l, day=day)
+            print "Trying Altitude Range: %d-72000" % a
+            X, Y = fit_setup(altitude=(a, 72000), latitude=l, day=day)
             rms_best = 1
             cn_best = None
             df_best = None
@@ -111,7 +110,7 @@ def make_fits(day, latrange, month, gen=False, path=""):
                     except ValueError:
                         print "Fit failed: Lat %d; K = %d; ftype = %s;" % (
                             l, K, ftype)
-                        err = [0.9, 0.9]
+                        err = 0.9
                     if err < rms_best:
                         df = cns.get_dataframe(X)
                         if "0.0" in df.values or "inf" in df.values:
[…]
